### 2017-12-25 Report/BigDataHomework1.tex
Initial commit

\documentclass[12pt]{article}

\usepackage{hyperref}
\hypersetup{
	colorlinks,
	citecolor=blue,
	%filecolor=black,
	linkcolor=black,
	urlcolor=blue
}
\usepackage{float}
\usepackage[english]{babel}
\usepackage[square,numbers]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{url}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\graphicspath{{images/}}
\usepackage{parskip}
\usepackage{fancyhdr}
\usepackage{vmargin}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\setmarginsrb{3 cm}{2.5 cm}{3 cm}{2.5 cm}{1 cm}{1.5 cm}{1 cm}{1.5 cm}
\usepackage[table]{xcolor}
\usepackage{todonotes}
\usepackage{menukeys}

\title{Spatiotemporal Data Classification}   % Title
\author{ Nikiforos Pittaras, M1422 \\ 
		 Chrysoula Themeli, M1423}                               % Authors
\date{\today}  % Date

\makeatletter
\let\thetitle\@title
\let\theauthor\@author
\let\thedate\@date
\makeatother

\pagestyle{fancy}
\fancyhf{}
%\rhead{\theauthor}
\lhead{\thetitle}
\cfoot{\thepage}

\begin{document}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\begin{titlepage}
		\centering
		\vspace*{0.5 cm}
		\includegraphics[scale = 0.75]{ekpalogo.png}\\[1.0 cm]   % University Logo
		\textsc{\LARGE National and Kapodistrian University of Athens}\\[2.0 cm]   % University Name
		\textsc{\Large Department of Informatics and Telecommunications}\\[0.5 cm]               % Department
		\textsc{\large Large-Scale Data Analysis Techniques}\\[0.5 cm] % Course Name
		\rule{\linewidth}{0.2 mm} \\[0.4 cm]
		{ \huge \bfseries \thetitle}\\
		\rule{\linewidth}{0.2 mm} \\[1.5 cm]

		\begin{minipage}{0.4\textwidth}
			\begin{center} \large
				\theauthor
			\end{center}
		\end{minipage}~
		\begin{minipage}{0.4\textwidth}
		\end{minipage}\\[2 cm]

		{\large \thedate}\\[2 cm]

		\vfill

	\end{titlepage}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\tableofcontents
    \newpage
	\listoffigures
	\newpage

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\section{Introduction}
    This is the report related to the Python project for the course "Large-Scale Data Analysis Techniques". In the below sections, the implemented logic will be explicitly analyzed. The project is divided in two parts: in the first part the goals are: to preprocess and clean the given data as well as to visualize five of the given trips in the train\_set file. In the second part, the tasks are: to find for all test trips the k nearest neabours from the given train trips, to find the train trips with the longest similar subroute for each test trip, to use 10-fold cross validation to the training data and classify them with three different methods (k-nearest neighbours, logistic regression and random forest) and finally to choose one of the above classification methods and improve the result.

    In the Main.py file, it is required to provide the path to the output folder as well as the input file with the training data. Afterwards, three folders are created to keep the images with the trips' visualizations for questions 1c, 2a1 and 2a2 and then the functions implementing the project's tasks are called.

	\section{Exercise 1}
	In this section, further details are provided related to the preprocessing and cleaning of the training data as well as to the trip's visualization. All files implementing the requirements of this part are in the folder FirstQuestion.

	\subsection{Question A}
	The first task of the project was to pre-process the given data, which is implemented in the file Preprocessing.py. In the read\_file\_contents, the train\_set.csv file is provided as input and for each line the necessary information is stored in a dictionary (journey\_id, vehicle\_id, timestamp, longitude and latitude).

	Since now the data from the input file are stored in the dictionary, the next step is to save the data based on the vehicle id. The function get\_data\_per\_vehicle\_id gets this dictionary as input and for every object in the dictionary, creates a list with key the vehicle\_id and then stores the remaining data connected with this id. The same process is followed for the sorting of data based on the journey\_id, therefore the structure where data are stored is a dictionary with key the vehicle\_id and with value another dictionary with key the journey\_id and with value a list containing several lists with the timestamp, the longitude and the latitude information.

	Finally, in the function get\_trips, the above structure is stored in the file trips.csv. The output format follows the requested pattern giving each trip an id. The final data structure is a list where each item has the trip\_id, the journey\_id and lists containing the timestamp, the longitude and the latitude.

	\subsection{Question B}
	The second goal was to clean the given data, based on the total trip distance and the max distance between two successive points. Total distance should not be less than 2km and max distance should not exceed 2km.

	The above logic is implemented in the file CleanData.py in function filter\_trips. The two lists (trips\_too\_small and trips\_too\_big) will keep the excluded trips due to either not valid total distance or to invalid max distance between successive points. For each trip in the list (returned from question A) the total distance is calculate, using the function calculate\_lonlat\_distance, which returns the distance between two successive points using the haversine formula. Following the same logic, the max distance between successive points of a trip is calculated and a distance is found more than 2km, the trip is excluded.

	Finally, the cleaned trips are kept in a list under the same format that was used in Question A and they are extracted in the file cleanTrips.csv, but also in the file tripsClean.pickle which keeps the data structure used. The functions for the data export are in the file utils.py (write\_trips\_to\_file and write\_trips\_using\_pickle).

	\subsection{Question C}
	This task is implemented in the file DataVisualization.py and the created output is stored in the folder Question1C. The requirement is to visualize five of the given train trips. For this purporse, a new trips list is creating containing five trips, excluding all trips with null journey\_id. Afterwards, the data\_visualization function is called, and a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).

	Finally, the function write\_group\_gml from the utils.py file is called. In this function, it is calculated the mean for both longitudes and latitudes and then gmplot is uded so as to visualize the selected trips. Below, there is an example of a trip shown on map is presented:

	\begin{figure} [H]
		\begin{center}
			\includegraphics [scale = 0.75] {questionCexample.jpg}
			\caption{Example of trip visualization}
		\end{center}
	\end{figure}

	\section{Exercise 2}
	The files implementing the requirements of this part are in the folder SecondQuestion.

	\subsection{Question A1}
	For this part a test file is given (test\_set\_a1.csv) containing additional trips. For each of these trips, the k nearest neighbours are targeted (in this exercise k=5) and an image file is produced, showing all six trips (the test trip and the trips from the training file) on six different maps. Under each map, the neighbour index, the journey\_id, the dynamic time warping and the time needed to process this neighbour in millisec.

	The code for this question is in the file NearestNeighbours.py. Firstly, all test trips in the test file are extracted and stored in the same data structure as the trip list that contains the training dataset. Subsequently, for each test trip all nearest neighbours are calculated in the function find\_nearest\_neighbours\_for\_test\_trip, keeping the time needed in a variable. For the test trip and all the training trips, the longitudes and latitudes are stored in tuples which are provided to the function calculate\_dynamic\_time\_warping in order to find the distance between two points. The calculated distances are stored in a list, which finally is sorted in ascending order and only the five nearest neighbours are kept.

	Since now the nearest neighbour trips are found, the trips list is updated (function: get\_updated\_trips\_list) containing now only those five trips. The final step now is to visualize the six trips. For this purpose, three list are created: a list with all the points (longitudes and latitudes), a list with the labels and a list with the colors (in this question the only color is the blue). All this data is provided to the function visualize\_paths in the utils.py file.

	In order to be able to create an image containing all the six maps each time, the rasterize.js file is required. In this function, both .html and .jpg files are created. Their names are kept in a list and when the final image file is created these files are deleted. 
	\todo{describe how the visualization function works}

	An example of this question image output can be found below:

	\begin{figure} [H]
		\begin{center}
			\includegraphics [scale = 0.50] {question2a1example.jpg}
			\caption{Example of nearest neighbours trips visualization}
		\end{center}
	\end{figure}

	\subsection{Question A2}
	Following the same logic as above, the given test file(test\_set\_a2.csv) is parsed and the data are stored in the usual format and the function find\_similar\_subroots\_per\_test\_trip is used to determine the nearest sub-routes for each test trip.

	The points are stored as tuples, similarly to the above question, both for the test and the training datasets, and the function datetime.datetime.now() is used so as to calculate the start and end time of each nearest subroute trip processing. The latest is performed in the method calc\_lcss.
	\todo{describe calc and update lcss}

	Finally, a similar data preprocessing is followed before visualize the data. This time the points are separated so as to show which of them belong to the nearest subroute and color them with red color. The function for the visualization is the same that was used for the question 2A1.

	\subsection{Question B}
	The purpose of this task was to export features for classification for the training dataset. In short, a grid should be created and the points of each trip trace should be replaced by a specific square of the grid. The output file where the exported features are stored is the tripFeatures.csv and the tripFeatures.pickle, which was used in order to keep the final data structure with the features.

	In order to be able to draw the grid, it was necessary to find the min and max longitudes and latitudes (function: find\_min\_max\_latlong). Having now this information, the next function called is the create\_grid giving as input the number\_of\_cells and the result of the previous method. The distance per cell in the grid is calculated as the difference between the max and the min lat/long divided by the total number of cells. The next step is to create the lines of the rows and the columns and then to provide a name for each cell.
	\todo{explain the replacepoints}

	Finally, the method visualize\_grid is called.
	\todo{explain the function}

	\subsection{Question C}

	\subsubsection{Beat ​ ​the ​ ​Benchmark}

	\newpage
	\addcontentsline{toc}{section}{References}
	\begin{thebibliography}{30}
		\bibitem{first} 
    \end{thebibliography}

\end{document}

### 2017-12-25 FirstQuestion/CleanData.py

@@ -1,0 +1,64 @@
+from math import sin, cos, sqrt, atan2, radians
+import os
+import utils
+
+
+def question_1b(output_folder, trips_list):
+    print("Total num of trips in trips.csv file: %d" % len(trips_list))
+    trips_list = filter_trips(trips_list)
+    print("Final num of trips to be writtern in tripsClean.csv: %d" % len(trips_list))
+    output_file = os.path.join(output_folder, "tripsClean.csv")
+    pickle_file = os.path.join(output_folder, "tripsClean.pickle")
+    utils.write_trips_to_file(output_file, trips_list)
+    utils.write_trips_using_pickle(pickle_file, trips_list)
+    return trips_list
+
+
+def filter_trips(trips_list):
+    trips_too_small, trips_too_big = [], []
+    for trip in trips_list:
+        total_dist = caluclate_total_distance_per_trip(trip)
+        # too small journey check
+        if total_dist < 2:
+            trips_too_small.append(trip)
+            trips_list.remove(trip)
+        max_dist = calculate_max_dist(trip)
+        if max_dist > 2:
+            trips_too_big.append(trip)
+            trips_list.remove(trip)
+    print("Total trips deleted due to total distance less than 2km: %d" % len(trips_too_small))
+    print("Total trips deleted due to max distance between two points more than 2km: %d" % len(trips_too_big))
+    return trips_list
+
+
+def caluclate_total_distance_per_trip(trip):
+    total_distance = 0
+    lonlatlist = trip[2:]
+    for i in range(len(lonlatlist)-1):
+        dist =  calculate_lonlat_distance(lonlatlist[i][1:], lonlatlist[i+1][1:])
+        total_distance += dist
+    return total_distance
+
+
+def calculate_lonlat_distance(point1, point2):
+    """
+    Uses the haversine formula
+    """
+    # print(point1, point2)
+    point1 = [radians(point1[0]), radians(point1[1])]
+    point2 = [radians(point2[0]), radians(point2[1])]
+    R = 6371.0
+    a = sin((point2[1]-point1[1])/2)**2 + cos(point1[1]) * cos(point2[1]) * sin((point2[0]-point1[0])/2)**2
+    c = 2 * atan2(sqrt(a), sqrt(1-a))
+    d = R * c
+    return d
+
+
+def calculate_max_dist(trip):
+    lonlatlist = trip[2:]
+    max_dist=2
[…]

### 2017-12-25 FirstQuestion/DataVisualization.py

@@ -1,0 +1,30 @@
+import random
+import os
+import utils
+
+
+def question_1c(output_folder, trips_list):
+    output_folder = os.path.join(output_folder, "Question1C")
+    output_file_base = os.path.join(output_folder, "mymap")
+    trips_visualization(output_file_base, trips_list, 5)
+
+
+def trips_visualization(base_file_name, trips_list, num_of_trips):
+    non_null_trips = get_trip_list_not_null_jid(trips_list)
+
+    for i in range(num_of_trips):
+        trip = non_null_trips[i]
+        file_name = base_file_name + str(i) + ".html"
+        data_visualization(file_name, trip)
+
+
+def get_trip_list_not_null_jid(trips_list):
+    non_null_trips = [t for t in trips_list if t[1] != 'null']
+    random.shuffle(non_null_trips)
+    return non_null_trips
+
+
+def data_visualization(file_name, trip):
+    points = [utils.idx_to_lonlat(list(range(len(trip[2:]))), trip)]
+    colors = ['b' for _ in range(len(points) + 1)]
+    utils.write_group_gml(points, file_name, colors)

### 2017-12-25 FirstQuestion/Preprocessing.py

@@ -1,0 +1,94 @@
+def question_1a(input_file, output_file):
+    '''
+    Runs the first subquestion
+    :param input_file: The training file
+    :param output_file: The processed file with the trips
+    :return:
+    '''
+    line_objects = read_train_file(input_file)
+    map_per_vid = get_data_per_vehicle_id(line_objects)
+    map_per_vid = get_data_per_vehicle_and_journey_id(map_per_vid)
+    trips_list = write_trips(output_file, map_per_vid)
+    return trips_list
+
+
+def read_train_file(train_file):
+    '''
+    Reads the training file
+    :param train_file:
+    :return:
+    '''
+    line_objects = []
+    with open(train_file, "r") as f:
+        next(f)
+        for line in f:
+            line_contents = line.strip().split(',')
+            obj = {}
+            obj["jid"] = line_contents[0]
+            obj["vid"] = line_contents[1]
+            obj["ts"] = line_contents[2]
+            obj["lon"] = line_contents[3]
+            obj["lat"] = line_contents[4]
+            line_objects.append(obj)
+    return line_objects
+
+
+def get_data_per_vehicle_id(line_objects):
+    '''
+    Aggregates data wrt the vehicle ID
+    :param line_objects:
+    :return:
+    '''
+    map_per_vid = {}
+    for obj in line_objects:
+        vid = obj["vid"]
+        if vid not in map_per_vid:
+            map_per_vid[vid] = []
+        obj_without_vid = {key: obj[key] for key in obj if key != "vid"}
+        map_per_vid[vid].append(obj_without_vid)
+    return map_per_vid
+
+
+def get_data_per_vehicle_and_journey_id(map_per_vid):
+    '''
+    Aggregates data wrt the journey ID
+    :param map_per_vid:
+    :return:
+    '''
+    for vehicle_id in map_per_vid:
+        map_per_jid = {}
+        vid_list = map_per_vid[vehicle_id]
+        for item in vid_list:
+            jid = item["jid"]
+            if jid not in map_per_jid:
+                map_per_jid[jid] = []
+            obj_without_jid = {key: item[key] for key in item if key != "jid"}
+            map_per_jid[jid].append(obj_without_jid)
+        for jid in map_per_jid:
+            jid_list = map_per_jid[jid]
+            sorted_jid_list = sorted(jid_list, key=lambda k: k['ts'])
+            map_per_jid[jid] = sorted_jid_list
[…]

### 2017-12-25 Main.py

@@ -1,0 +1,79 @@
+import sys, os, pkgutil
+import FirstQuestion.Preprocessing as prep
+import FirstQuestion.CleanData as clean
+import FirstQuestion.DataVisualization as visual
+
+import SecondQuestion.NearestNeighbours as nn
+import SecondQuestion.NearestSubroutes as ns
+import SecondQuestion.MapInGridView as gv
+import SecondQuestion.JourneyClassification as jc
+
+
+def main(output_folder, input_file):
+    # Prepare files and folders
+    output_file, test_file1, test_file2 = create_files_folders(output_folder, input_file)
+    trips_list = question_1(input_file, output_file, output_folder)
+    question_2(output_folder, test_file1, test_file2, trips_list)
+
+
+def create_files_folders(output_folder, input_file):
+    # question 1
+    output_file = "%s/trips.csv" % output_folder
+    os.makedirs("%s/Question1C" % output_folder, exist_ok=True)
+    # question 2
+    # test_file1 = "test_set_a1.csv"
+    # test_file2 = "test_set_a2.csv"
+    # TODO correct the test files
+    test_file1 = input_file
+    test_file2 = input_file
+    os.makedirs("%s/Question2A1" % output_folder, exist_ok=True)
+    os.makedirs("%s/Question2A2" % output_folder, exist_ok=True)
+
+    return output_file, test_file1, test_file2
+
+
+def question_1(input_file, output_file, output_folder):
+    # Question 1
+    print(">>> Running question 1a - parsing the training data")
+    trips_list = prep.question_1a(input_file, output_file)
+    print(">>> Running question 1b - cleaning the training data")
+    trips_list = clean.question_1b(output_folder, trips_list)
+    print(">>> Running question 1b - visualizing the training data")
+    visual.question_1c(output_folder, trips_list)
+    return trips_list
+
+
+def question_2(output_folder, test_file1, test_file2, trips_list):
+    # Question 2
+    print(">>> Running question 2a1 - Nearest neighbours computation")
+    nn.question_a1(output_folder, test_file1, trips_list)
+    print(">>> Running question 2a2 - Nearest subroutes computation")
+    ns.question_a2(output_folder, test_file2, trips_list)
+    print(">>> Running question 2b - Cell grid quantization")
+    cellgrid = (4, 3)
+    print("Using cell grid:", cellgrid)
+    features_file = gv.subquestion_b(trips_list, cellgrid, output_folder)
[…]

### 2017-12-25 SecondQuestion/JourneyClassification.py

@@ -1,0 +1,98 @@
+import pickle, os
+import numpy as np
+from sklearn.model_selection import KFold
+from sklearn.neighbors import KNeighborsClassifier
+from sklearn.linear_model import LogisticRegression
+from sklearn.ensemble import RandomForestClassifier
+from sklearn.metrics import accuracy_score
+import utils
+
+
+def preprocess_data(data):
+    targets = [d[1] for d in data]
+    data = [d[2] for d in data]
+
+    maxlen = len(max(data, key=lambda x : len(x)))
+    data = [[int(d[1:]) for d in dlist] for dlist in data]
+
+    for i, datum in enumerate(data):
+        if len(datum) < maxlen:
+            data[i] = datum + [3 for _ in range(maxlen - len(datum))]
+
+    # map journey ids to numbers
+    num_ids = {}
+    targets_nums = []
+    for t in targets:
+        if t not in num_ids:
+            num_ids[t] = len(num_ids)
+        targets_nums.append(num_ids[t])
+    return data, targets_nums
+
+def question_c(file):
+    # get the data ready
+    trips_list = get_grid_trips(file)
+    kf = KFold(n_splits=10)
+    grid_features, targets = preprocess_data(trips_list)
+
+    folds_idxs = list(kf.split(grid_features))
+    trips_array = np.asarray(grid_features)
+    targets = np.asarray(targets)
+
+    classifiers = ["knn", "logreg", "randfor"]
+    accuracies = {}
+    # classify
+    output_dir = "classification_charts"
+    os.makedirs(output_dir, exist_ok=True)
+    for classifier in classifiers:
+        print("Testing classifier [%s]" % classifier)
+        # train & test each classifier
+        # for each fold
+        accuracies[classifier] = []
+        for i, (train_idx, val_idx) in enumerate(folds_idxs):
+            print("\tClassifing fold %d/%d" % (i+1, len(folds_idxs)))
+            train = (trips_array[train_idx], targets[train_idx])
+            val = (trips_array[val_idx], targets[val_idx])
+            if classifier == "knn":
+                k = 5
+                res = knn_classification(train, val, targets, k)
+            elif classifier == "logreg":
+                res = logreg_classification(train,val)
+            elif classifier == "randfor":
+                res = randfor_classification(train, val)
+            accuracies[classifier].append(res)
[…]

### 2017-12-25 SecondQuestion/MapInGridView.py

@@ -1,0 +1,121 @@
+import matplotlib.pyplot as plt
+import utils
+import os
+
+
+def subquestion_b(trips_list, number_of_cells, output_folder):
+    output_file = os.path.join(output_folder, "tripFeatures.csv")
+    pickle_file = os.path.join(output_folder, "tripFeatures.pickle")
+    max_lat, max_lon, min_lat, min_lon = find_min_max_latlong(trips_list)
+    rows,columns,cell_names = create_grid(number_of_cells,max_lat,max_lon,min_lat,min_lon)
+    replace_points(trips_list, rows, columns, cell_names, output_file, pickle_file)
+    return pickle_file
+
+
+def find_min_max_latlong(trips_list):
+    max_lat=trips_list[0][2][2]
+    max_lon=trips_list[0][2][1]
+    min_lat=trips_list[0][2][2]
+    min_lon=trips_list[0][2][1]
+    for trip in trips_list:
+        if trip[2][1]>max_lon:
+            max_lon=trip[2][1]
+        if trip[2][1]<min_lon:
+            min_lon=trip[2][1]
+        if trip[2][2]>max_lat:
+            max_lat=trip[2][2]
+        if trip[2][2]<min_lat:
+            min_lat=trip[2][2]
+    return max_lat, max_lon, min_lat, min_lon
+
+
+def create_grid(number_of_cells, max_lat,max_lon,min_lat,min_lon):
+    cell_lat_dist, cell_lon_dist = get_distance_per_cell(number_of_cells, min_lat, min_lon, max_lat, max_lon)
+    rows = create_grid_lines(number_of_cells[0], min_lat, cell_lat_dist)
+    columns=create_grid_lines(number_of_cells[1], min_lon, cell_lon_dist)
+    cell_names = create_cell_names(number_of_cells)
+
+    visualize_grid(rows,columns,min_lat,min_lon,max_lat,max_lon)
+    return rows,columns,cell_names
+
+
+def get_distance_per_cell(number_of_cells, min_lat, min_lon, max_lat, max_lon):
+    total_lat_dist = max_lat - min_lat
+    total_lon_dist = max_lon - min_lon
+    cell_lat_dist = total_lat_dist / number_of_cells[0]
+    cell_lon_dist = total_lon_dist / number_of_cells[1]
+    return cell_lat_dist, cell_lon_dist
+
+
+def create_grid_lines(number_of_cells, min_point, dist):
+    lines_list = []
+    new_min = min_point
[…]

### 2017-12-25 SecondQuestion/NearestNeighbours.py

@@ -1,0 +1,66 @@
+import FirstQuestion.CleanData as clean
+import datetime
+import os
+import utils
+
+
+def question_a1(output_folder, file, trips_list):
+    test_trip_list = utils.create_test_trip_list(file)
+    for i,test_trip in enumerate(test_trip_list):
+        print("Examining test element %d/%d" % (i+1, len(test_trip_list)))
+        outfile_name = os.path.join(output_folder, "Question2A1", "output_%d" % (i+1))
+        millis_start = utils.tic()
+        nearest_neighbours = find_nearest_neighbours_for_test_trip(test_trip, trips_list)
+        elapsed = utils.tictoc(millis_start)
+        trips_list = get_updated_trips_list(trips_list, nearest_neighbours)
+        preprocessing_for_visualization(test_trip, nearest_neighbours, trips_list, outfile_name, elapsed, i)
+
+
+def find_nearest_neighbours_for_test_trip(test_trip, trips_list):
+    test_lonlat = utils.create_list_tuples(test_trip)
+    nearest_neighbours = []
+    #print("knn for ",test_lonlat)
+    for trip in trips_list:
+        trip_lonlat = utils.create_list_tuples(trip)
+        distance = calculate_dynamic_time_warping(test_lonlat, trip_lonlat)
+        nearest_neighbours.append((trip[1], distance))
+    nearest_neighbours = sorted(nearest_neighbours, key=lambda k: k[1])
+    nearest_neighbours = nearest_neighbours[:5]
+    return nearest_neighbours
+
+
+def calculate_dynamic_time_warping(latlons1, latlons2):
+    dtw = [ [float('Inf') for _ in range(len(latlons2)+1)] for _ in range(len(latlons1)+1)]
+    dtw[0][0] = 0
+    for i in range(1,1+len(latlons1)):
+        for j in range(1,1+len(latlons2)):
+            cost = clean.calculate_lonlat_distance(latlons1[i-1], latlons2[j-1])
+            dtw[i][j] = cost + min(dtw[i-1][j], dtw[i][j-1], dtw[i-1][j-1])
+    return dtw[-1][-1]
+
+
+def get_updated_trips_list(trips_list, nearest_neighbours):
+    """
+    gets a trips_list as input, in the format of tripsClean.csv file and returns a list of tuples with lats and lons
+
+    :param trips_list:
+    :param nearest_neighbours:
+    :return:
+    """
+    new_trips_list=[]
[…]

### 2017-12-25 SecondQuestion/NearestSubroutes.py

@@ -1,0 +1,112 @@
+import FirstQuestion.CleanData as clean
+import os
+import datetime
+import utils
+
+
+def question_a2(output_folder, test_file, trips_list):
+    test_list = utils.create_test_trip_list(test_file)
+    k=5
+    for i,test_trip in enumerate(test_list):
+        print("Extracting neighbours for test trip %d/%d" % (i+1,len(test_list)))
+        file_name = os.path.join(output_folder, "Question2A2", "subroutes_%d" % (i + 1))
+        max_subseqs = find_similar_subroots_per_test_trip(test_trip, trips_list, k)
+        preprocessing_for_visualisation(test_trip, max_subseqs, trips_list, file_name)
+
+
+def find_similar_subroots_per_test_trip(test_trip, trips_list, k):
+    test_lonlat = utils.create_list_tuples(test_trip)
+    max_subseqs = []
+
+    for tripidx, trip in enumerate(trips_list):
+        trip_lonlat = utils.create_list_tuples(trip)
+        millis_start = datetime.datetime.now()
+        subseqs, pp = calc_lcss(test_lonlat, trip_lonlat)
+        millis_end = datetime.datetime.now()
+        elapsed = millis_end - millis_start
+        # sort by decr. length
+        pp = sorted(pp, key = lambda x : len(x), reverse=True)
+        max_subseqs = update_lcss(max_subseqs, pp, k, elapsed, tripidx)
+    return max_subseqs
+
+
+def calc_lcss(t1, t2):
+    L = [ [0 for _ in t2] for _ in t1]
+    seqs = []
+    idxs = []
+    z = 0
+    for i, p1 in enumerate(t1):
+        for j, p2 in enumerate(t2):
+            # for testing
+            if type(p1) == str:
+                equal = p1 == p2
+            else:
+                dist = clean.calculate_lonlat_distance(p1, p2)
+                equal = dist < 200
+            if equal:
+                if i == 0 or j == 0:
+                    L[i][j] =1
+                    seqs.append(t2[j])
+                    idxs.append([j])
+                else:
+                    L[i][j] =  L[i-1][j-1] + 1
+                    if L[i][j] > z:
+                        z = L[i][j]
+                        seqs.append(t2[j-z+1:j+1])
+                        idxs.append(list(range(j-z+1,j+1)))
+            else:
+                print(i,j,"-",p1,p2)
+                for ll in L: print(ll)
+                L[i][j] = 0
[…]

### 2017-12-25 rasterize.js

@@ -1,0 +1,49 @@
+"use strict";
+var page = require('webpage').create(),
+    system = require('system'),
+    address, output, size;
+
+if (system.args.length < 3 || system.args.length > 5) {
+    console.log('Usage: rasterize.js URL filename [paperwidth*paperheight|paperformat] [zoom]');
+    console.log('  paper (pdf output) examples: "5in*7.5in", "10cm*20cm", "A4", "Letter"');
+    console.log('  image (png/jpg output) examples: "1920px" entire page, window width 1920px');
+    console.log('                                   "800px*600px" window, clipped to 800x600');
+    phantom.exit(1);
+} else {
+    address = system.args[1];
+    output = system.args[2];
+    page.viewportSize = { width: 1200, height: 1200 };
+    if (system.args.length > 3 && system.args[2].substr(-4) === ".pdf") {
+        size = system.args[3].split('*');
+        page.paperSize = size.length === 2 ? { width: size[0], height: size[1], margin: '0px' }
+                                           : { format: system.args[3], orientation: 'portrait', margin: '1cm' };
+    } else if (system.args.length > 3 && system.args[3].substr(-2) === "px") {
+        size = system.args[3].split('*');
+        if (size.length === 2) {
+            pageWidth = parseInt(size[0], 10);
+            pageHeight = parseInt(size[1], 10);
+            page.viewportSize = { width: pageWidth, height: pageHeight };
+            page.clipRect = { top: 0, left: 0, width: pageWidth, height: pageHeight };
+        } else {
+            console.log("size:", system.args[3]);
+            pageWidth = parseInt(system.args[3], 10);
+            pageHeight = parseInt(pageWidth * 3/4, 10); // it's as good an assumption as any
+            console.log ("pageHeight:",pageHeight);
+            page.viewportSize = { width: pageWidth, height: pageHeight };
+        }
+    }
+    if (system.args.length > 4) {
+        page.zoomFactor = system.args[4];
+    }
+    page.open(address, function (status) {
+        if (status !== 'success') {
+            console.log('Unable to load the address!');
+            phantom.exit(1);
+        } else {
+            window.setTimeout(function () {
+                page.render(output);
+                phantom.exit();
+            }, 200);
+        }
+    });
+}

### 2017-12-25 run.sh

@@ -1,0 +1,3 @@
+#! /usr/bin/env bash
+[ -z $(which phantomjs) ] && echo "Need phantomjs to create output report files!" && exit 
+python Main.py "$(pwd)/results_folder" ./input_folder/train_set_dev.csv

### 2017-12-25 utils.py

@@ -1,0 +1,177 @@
+import gmplot
+import subprocess
+import os
+import matplotlib.pyplot as plt
+import pylab
+from scipy.misc import imread
+import datetime
+import FirstQuestion.Preprocessing as prep
+import pickle
+
+
+def tic():
+    return datetime.datetime.now()
+
+
+def tictoc(previous_tic):
+    # get duration in usec
+    diff = tic() - previous_tic
+    days = diff.days
+    secs = diff.seconds
+    usec = diff.microseconds
+    msec, usec = divmod(usec, 1000)
+    sec_, msec = divmod(msec, 1000)
+    secs += sec_
+    mins, sec = divmod(secs, 60)
+    hrs, mins = divmod(mins,60)
+    msg = "%02d:%02d:%02d:%03d" % (hrs, mins, secs, msec)
+    if days:
+        msg = "%d days! " % days + msg
+    return msg
+
+
+def write_trips_to_file(output_file, trips_list):
+    with open(output_file, "w") as f:
+        for trip in trips_list:
+            trip_str = str(trip)
+            f.write(trip_str+"\n")
+
+
+def write_trips_using_pickle(output_file, trips_list):
+    with open(output_file, "wb") as f:
+        pickle.dump(trips_list, f)
+
+
+def create_test_trip_list(file):
+    '''
+    Test files processing
+    '''
+    # test_file = path_to_files + file
+    # with open(test_file, "r") as f:
+    #     test_trip_list=[]
+    #     for line in f:
+    #         test_trip_list.append(list(line))
+    # TODO read from test file
+    print("\tNote: Will use the training file from question 1a for now, limited to 5 elements")
+    tl =  prep.question_1a(file, "test_clean_trips.csv")
+    return tl[:5]
+
+
+# Functions used for trips visualization
+
+def create_list_tuples(trips_list):
+    '''
+    get lats and lots as list of tuples
+    '''
+    trips_list = trips_list[2:]
+    new_list_tuples = []
+    for item in trips_list:
+        new_list_tuples.append((item[1], item[2]))
+    return new_list_tuples
+
+
+def idx_to_lonlat(idx, trip):
+    if type(idx) != list:
+        idx = [idx]
+    lats = []
+    lons = []
+    for i in idx:
+        lats.append(trip[2+i][2])
+        lons.append(trip[2+i][1])
+    return (lons, lats)
+
+
+def write_group_gml(points, outpath, colors):
[…]

### 2017-12-27 SecondQuestion/NearestNeighbours.py
update report

@@ -1,5 +1,4 @@
 import FirstQuestion.CleanData as clean
-import datetime
 import os
 import utils
 
@@ -17,9 +16,16 @@ def question_a1(output_folder, file, trips_list):
 
 
 def find_nearest_neighbours_for_test_trip(test_trip, trips_list):
+
+    '''
+    :param test_trip: a give trip from the test file
+    :param trips_list: all trips from the training_set file
+    :return: returns the 5 nearest neighbours for the given test trip
+    '''
+
     test_lonlat = utils.create_list_tuples(test_trip)
     nearest_neighbours = []
-    #print("knn for ",test_lonlat)
+    # print("knn for ",test_lonlat)
     for trip in trips_list:
         trip_lonlat = utils.create_list_tuples(trip)
         distance = calculate_dynamic_time_warping(test_lonlat, trip_lonlat)
@@ -30,6 +36,12 @@ def find_nearest_neighbours_for_test_trip(test_trip, trips_list):
 
 
 def calculate_dynamic_time_warping(latlons1, latlons2):
+    '''
+    Calculate the DTW of two points, used in order find the nearest neighbours
+    :param latlons1:
+    :param latlons2:
+    :return:
+    '''
     dtw = [ [float('Inf') for _ in range(len(latlons2)+1)] for _ in range(len(latlons1)+1)]
     dtw[0][0] = 0
     for i in range(1,1+len(latlons1)):
@@ -40,13 +52,11 @@ def calculate_dynamic_time_warping(latlons1, latlons2):
 
 
 def get_updated_trips_list(trips_list, nearest_neighbours):
-    """
-    gets a trips_list as input, in the format of tripsClean.csv file and returns a list of tuples with lats and lons
-
-    :param trips_list:
-    :param nearest_neighbours:
-    :return:
-    """
+    '''
+    :param trips_list: all the trips from the training set
+    :param nearest_neighbours: the 5 nearest neighbours
+    :return: trips_list containing only the 5 nearest neighbours
+    '''
     new_trips_list=[]
     for trip in trips_list:
         for neighbour in nearest_neighbours:
@@ -56,6 +66,15 @@ def get_updated_trips_list(trips_list, nearest_neighbours):
 
 
 def preprocessing_for_visualization(test_trip, nearest_neighbours, trips_list, outfile_name, elapsed, i):
+    '''
+    :param test_trip: the given test trip from the test file
+    :param nearest_neighbours: the 5 nearest neighbours of this test trip
[…]

### 2017-12-28 Report/BigDataHomework1.tex
update report
changes to report

@@
 	\newpage
+	\listoftodos
+	\newpage
 
@@
 For each of these trips, the k nearest neighbours are targeted (in this exercise k=5) and an image file is produced, showing all six trips (the test trip and the trips from the training file) on six different maps.
-Under each map, the neighbour index, the journey\_id, the dynamic time warping and the time needed to process this neighbour in millisec.
+Under each map, the neighbour index, the journey\_id, the dynamic time warping and the time needed to process the neighbours of this test trip in millisec.
 
@@
 Subsequently, for each test trip all nearest neighbours are calculated in the function find\_nearest\_neighbours\_for\_test\_trip, keeping the time needed in a variable.
-For the test trip and all the training trips, the longitudes and latitudes are stored in tuples which are provided to the function calculate\_dynamic\_time\_warping in order to find the distance between two points.
+At the beginning of the trips processing, the tic method from utils.py file is used (finds the current time in millisec using the datetime.datetime.now() method).
+After having found the nearest neighbours for the given test trip, the method tictoc() is called, returning the required time.
+In order to find the nearest neighbours, the following process is implemented: For the test trip and all the training trips, the longitudes and latitudes are stored in tuples which are provided to the function calculate\_dynamic\_time\_warping in order to find the distance between two points.
 The calculated distances are stored in a list, which finally is sorted in ascending order and only the five nearest neighbours are kept.
@@
 For this purpose, three list are created: a list with all the points (longitudes and latitudes), a list with the labels and a list with the colors (in this question the only color is the blue).
-All this data is provided to the function visualize\_paths in the utils.py file.
+All this data is provided to the function visualize\_paths in the utils.py file in order to export the trips as map images.
 
-	In order to be able to create an image containing all the six maps each time, the rasterize.js file is required.
+	In order to be able to create an image containing all the six maps each time, the rasterize.js file is required as well as phantomjs.
 In this function, both .html and .jpg files are created.
-Their names are kept in a list and when the final image file is created these files are deleted.
-	\todo{describe how the visualization function works}
+First of all, the function write\_group\_gml is used so as to create the html files (this function is described above in question c).
+These html files are converted to images using the function html\_to\_png.
+File names are kept in a list and when the final image file is created (containing all six maps) these files are deleted.
+Afterwards, images as displayed as collection of plots and exported in the .jpg format.
 
@@
 	\subsection{Question A2}
-	Following the same logic as above, the given test file(test\_set\_a2.csv) is parsed and the data are stored in the usual format and the function find\_similar\_subroots\_per\_test\_trip is used to determine the nearest sub-routes for each test trip.
+	Following the same logic as above, the given test file(test\_set\_a2.csv) is parsed storing the data in the usual format and the function find\_similar\_subroots\_per\_test\_trip is used to determine the nearest sub-routes for each test trip.
 
-	The points are stored as tuples, similarly to the above question, both for the test and the training datasets, and the function datetime.datetime.now() is used so as to calculate the start and end time of each nearest subroute trip processing.
-The latest is performed in the method calc\_lcss.
+	Similarly to the above question, the points are stored as tuples and the required time is calculated using the relevant methods from the utils.py file. To begin with, the method calc\_lcss is called
 	\todo{describe calc and update lcss}

### 2017-12-28 Report/BigDataHomework1.tex
Update report until last question, 2 todos remain

@@
 
-	Similarly to the above question, the points are stored as tuples and the required time is calculated using the relevant methods from the utils.py file. To begin with, the method calc\_lcss is called
-	\todo{describe calc and update lcss}
+	Similarly to the above question, the points are stored as tuples and the required time is calculated using the relevant methods from the utils.py file.
+To begin with, the method calc\_lcss is called.
+In this function, the list of lonlat coordinate tuples are given as input.
+We keep all similar point values and indexes in two lists while iterating the two given lists.
+The distance between two points is calculated using the method calculate\_lonlat\_distance in the CleanData.py file.
+Based on the exercise requirements, two points are considered as equals if their distance is less than 200 metres.
+Therefore, if the points are equal, and the indexes are the first point of one of the given lists the cell is marked to unit cost and a new similar sequence is started.
+Otherwise, we continue to an existing similar sequence, increasing by 1 its length.
+\todo{explain z variable} On the other hand, if points are not equal, then a new subsequence should start, and therefore L[i][j] is set to zero.
+
+	Having now found the longest subsequences, we sort the indexes based on the length of the subsequences and we call the update\_current\_maxsubseq in order to update our result comparing previous subsequences.
+First of all, the new sequences found are iterated, checking if the length of the subsequence list has reached the given number (in our case 5).
+If there is still space to add the subsequence, then it is added, otherwise we check if the new subsequence is longer from those kept in the list and therefore it should replace it.
 
 	Finally, a similar data preprocessing is followed before visualize the data.
-This time the points are separated so as to show which of them belong to the nearest subroute and color them with red color.
+This time the points are separated so as to show which of them belong to the nearest sub-route and color them with red color.
 The function for the visualization is the same that was used for the question 2A1.
@@
 The next step is to create the lines of the rows and the columns and then to provide a name for each cell.
-	\todo{explain the replacepoints}
 
-	Finally, the method visualize\_grid is called.
-	\todo{explain the function}
+	Afterwards, all points in the trip list should be replaced by the cell names that each point corresponds to.
+This logic is implemented in the method replace\_points.
+For each trip in the list, all coordinates are transformed in tuples format and using the find\_index function a name is mapped to this point.
+After having replaced all points with their new names, this list is exported in .csv and .pickle format.
+In order to be able to easily check the created grid, the method visualize\_grid is implemented so as to plot the grid.
 
 	\subsection{Question C}
+	The file corresponding to this question is the JourneyClassification.py.
+First of all, the data structure is loaded from the .pickle file, created in above question as we have already mentioned.
+In the requirements of this question it is mentioned to use 10-fold cross validation, therefore the KFold class is used providing the required folds number.
+Then it is used in order to split our list to the training and test datasets.
+	\todo{explain preprocessdata function}
+
+	It worths to be mentioned here, that method asarrays from numpy was used to the necessary for this question lists.
+In addition, a list with the names of the three classifiers is created and for each of them, the relevant function is called in order to train the classifier and then count the accuracy.
+
+	In this question, the scikit learn library was used.
+Firstly, for the knn classifier the class KNeighborsClassifier was used, and more specifically the fit and predict methods.
+In order to calculate the accuracy of the classifiers we have used the accuracy\_score method (this result is return from each classification function in the JourneyClassification.py).
+The same logic is followed for the other two classifiers(logistic regression and random forest).
+What should be mentioned, though, is that the LogisticRegression class implements the predict\_proba method, which returns a list with probabilities and argmax from numpy was used in order to determine the result with the max probability.
 

### 2017-12-28 question2/JourneyClassification.py

@@ -28,6 +28,7 @@ def preprocess_data(data):
         targets_nums.append(num_ids[t])
     return data, targets_nums
 
+
 def question_c(file):
     # get the data ready
     trips_list = get_grid_trips(file)

### 2017-12-28 question2/NearestSubroutes.py

@@ -1,6 +1,5 @@
 import question1.CleanData as clean
 import os
-import datetime
 import utils
 
 
@@ -72,8 +71,8 @@ def calc_lcss(t1, t2):
                         seqs.append(t2[j-z+1:j+1])
                         idxs.append(list(range(j-z+1,j+1)))
             else:
-                #print(i,j,"-",p1,p2)
-                #for ll in L: print(ll)
+                # print(i,j,"-",p1,p2)
+                # for ll in L: print(ll)
                 L[i][j] = 0
     return seqs, list(idxs)
 

### 2017-12-28 utils.py

@@ -90,6 +90,8 @@ def read_trips(filepath):
                     trip.append(vals)
                 trips.append(trip)
     return trips
+
+
 # Functions used for trips visualization
 #######################################
 def read_train_set(filepath):
@@ -112,6 +114,7 @@ def read_train_set(filepath):
             line_objects.append(obj)
     return line_objects
 
+
 def idx_to_lonlat(trip, idx = None, format = "tuple"):
     '''
     Get the longitudes and latitudes coresponding to the input points of the input trip
@@ -133,6 +136,7 @@ def idx_to_lonlat(trip, idx = None, format = "tuple"):
     elif format == "tuples":
         return list(zip(lons,lats))
 
+
 def write_group_gml(lonlats_tuplelists, outpath, colors=None):
     '''
      Make a color plot a collection of points.
@@ -148,8 +152,8 @@ def write_group_gml(lonlats_tuplelists, outpath, colors=None):
     delta_lonlat = [mx-mn for (mx,mn) in zip(max_lonlat, min_lonlat)]
     center_lonlat = [min_lonlat[i] + delta_lonlat[i] for i in range(2)]
     zoom = 14
-    #print("points:",lonlats_tuplelists)
-    #print("center:",center_lonlat,"delta:",delta_lonlat,"zoom",zoom)
+    # print("points:",lonlats_tuplelists)
+    # print("center:",center_lonlat,"delta:",delta_lonlat,"zoom",zoom)
     gmap = gmplot.GoogleMapPlotter(center_lonlat[1], center_lonlat[0], zoom)
     if colors is None:
         colors = ['b' for _ in lonlats_tuplelists]
@@ -158,6 +162,7 @@ def write_group_gml(lonlats_tuplelists, outpath, colors=None):
         gmap.plot(pts[1], pts[0], colors[idx], edge_width=5)
     print("Writing plot", outpath)
     gmap.draw(outpath)
+
 
 def html_to_png(html_path, png_path):
     '''
@@ -173,6 +178,7 @@ def html_to_png(html_path, png_path):
         print("Failed to run command %s" % " ".join(cmd))
         print(ex)
         exit(1)
+
 
 def visualize_paths(all_pts, colors, labels, file_name):
     '''
@@ -229,6 +235,7 @@ def visualize_paths(all_pts, colors, labels, file_name):
     pylab.savefig(base_file_name + ".nnplot.jpg", dpi=300)
     plt.close()
 
+
[…]

### 2017-12-28 Report/BigDataHomework1.tex
changes in report: main.py, output folders, add references

@@
 
-    In the Main.py file, it is required to provide the path to the output folder as well as the input file with the training data.
-Afterwards, three folders are created to keep the images with the trips' visualizations for questions 1c, 2a1 and 2a2 and then the functions implementing the project's tasks are called.
+    In the Main.py file, it is required to provide the path to the input folder (containing the training and test files) and the output folder where all results will be stored.
+In this folder an additional directory is created, in order to keep all maps extracted from this application.
+There are two ways in order to run this project, either add these folders in the edit configurations choice from the menu, or use the run.sh in order to provide these args.
+Finally, the dependencies are checked and the folders and files are prepared before starting to run the application.
 
@@
 	\subsection{Question C}
-	This task is implemented in the file DataVisualization.py and the created output is stored in the folder Question1C.
+	This task is implemented in the file DataVisualization.py and the created output is stored in the output folder in the gmplots directory.
 The requirement is to visualize five of the given train trips.
@@
 	\subsection{Question A1}
-	For this part a test file is given (test\_set\_a1.csv) containing additional trips.
+	For this part a test file is given (test\_set\_a1.csv, included in the input folder) containing additional trips.
 For each of these trips, the k nearest neighbours are targeted (in this exercise k=5) and an image file is produced, showing all six trips (the test trip and the trips from the training file) on six different maps.
@@
 	\begin{thebibliography}{30}
-		\bibitem{first} 
+		\bibitem{knn}SciKit Learn: Nearest Neighbors \\
+		 \url{http://scikit-learn.org/stable/modules/neighbors.html}
+		 \bibitem{logreg} SciKit Learn: Logistic Regression \\
+		 \url{http://scikit-learn.org/stable/modules/linear_model.html#logistic-regression}
+		 \bibitem{ranfor} SciKit Learn: Random Forest \\
+		 \url{http://scikit-learn.org/stable/modules/generated/sklearn.ensemble.RandomForestClassifier.html#sklearn.ensemble.RandomForestClassifier}
+		 \bibitem{accuracy} SciKit Learn: Accuracy \\
+		 \url{http://scikit-learn.org/stable/modules/generated/sklearn.metrics.accuracy_score.html#sklearn.metrics.accuracy_score}
+		 \bibitem{knn2}A Complete Guide to K-Nearest-Neighbors with Applications in Python and R \\
+		 \url{https://kevinzakka.github.io/2016/07/13/k-nearest-neighbor/}
+		 \bibitem{dwt} Dynamic time warping \\
+		 \url{https://en.wikipedia.org/wiki/Dynamic_time_warping}
+		 \bibitem{lcss} Longest common subsequence problem \\
+		 \url{https://en.wikipedia.org/wiki/Longest_common_subsequence_problem}
     \end{thebibliography}

### 2017-12-28 Main.py

@@ -10,7 +10,6 @@ import question2.JourneyClassification as jc
 import utils
 
 
-
 def question_1(input_file, output_file, output_folder, maps_folder):
     # Question 1
     print(">>> Running question 1a - parsing the training data")
@@ -20,14 +19,15 @@ def question_1(input_file, output_file, output_folder, maps_folder):
     print(">>> Running question 1b - visualizing the training data")
     visual.question_1c(maps_folder, trips_list)
 
+
 def question_2(train_file, test_files, output_folder):
     # Question 2
     # Read the training data
     trips_list = utils.read_trips(train_file)
     print(">>> Running question 2a1 - Nearest neighbours computation")
-    #nn.question_a1(output_folder, test_files[0], trips_list)
+    nn.question_a1(output_folder, test_files[0], trips_list)
     print(">>> Running question 2a2 - Nearest subroutes computation")
-    #ns.question_a2(output_folder, test_files[1], trips_list)
+    ns.question_a2(output_folder, test_files[1], trips_list)
     print(">>> Running question 2b - Cell grid quantization")
     cellgrid = (4, 3)
     print("Using cell grid:", cellgrid)

### 2017-12-29 Main.py
add threads to questions 2a1 and 2a2

@@ -26,9 +26,9 @@ def question_2(train_file, test_files, output_folder, maps_folder):
     # Read the training data
     trips_list = utils.read_trips(train_file)
     print(">>> Running question 2a1 - Nearest neighbours computation")
-    #nn.question_a1(maps_folder, test_files[0], trips_list)
+    nn.question_a1(maps_folder, test_files[0], trips_list)
     print(">>> Running question 2a2 - Nearest subroutes computation")
-    #ns.question_a2(output_folder, test_files[1], trips_list)
+    ns.question_a2(output_folder, test_files[1], trips_list)
     print(">>> Running question 2b - Cell grid quantization")
     cellgrid = (4, 3)
     print("Using cell grid:", cellgrid)

### 2017-12-29 question2/NearestNeighbours.py

@@ -1,6 +1,7 @@
 import question1.CleanData as clean
 import os
 import utils
+from multiprocessing.pool import ThreadPool
 
 
 def question_a1(output_folder, test_file, trips_list):
@@ -32,17 +33,21 @@ def calculate_nns(test_trip, trips_list):
     # get test coordinate list
     test_lonlat = test_trip["points"]
     nearest_neighbours = []
+    pool = ThreadPool(processes=10)
     for i, trip in enumerate(trips_list):
         # get candidate coordinate list
         trip_lonlat = trip["points"]
         # calcluate distance
-        distance = calculate_dynamic_time_warping(test_lonlat, trip_lonlat)
+        async_result = pool.apply_async(calculate_dynamic_time_warping, (test_lonlat, trip_lonlat))
+        distance = async_result.get()
+        # distance = calculate_dynamic_time_warping(test_lonlat, trip_lonlat)
         print("Calculated distance: %.2f for trip: %d/%d : %s" % (distance, i+1, len(trips_list), str(trip['id'])))
         nearest_neighbours.append((trip['id'], distance))
     # sort the list to increasing distance
     nearest_neighbours = sorted(nearest_neighbours, key=lambda k: k[1])
     # return the top 5
     return nearest_neighbours[:5]
+
 
 def calculate_dynamic_time_warping(latlons1, latlons2):
     '''
@@ -53,9 +58,12 @@ def calculate_dynamic_time_warping(latlons1, latlons2):
     '''
     dtw = [ [float('Inf') for _ in range(len(latlons2)+1)] for _ in range(len(latlons1)+1)]
     dtw[0][0] = 0
+    pool = ThreadPool(processes=10)
     for i in range(1,1+len(latlons1)):
         for j in range(1,1+len(latlons2)):
-            cost = clean.calculate_lonlat_distance(latlons1[i-1], latlons2[j-1])
+            async_result = pool.apply_async(clean.calculate_lonlat_distance, (latlons1[i-1], latlons2[j-1]))
+            cost = async_result.get()
+            # cost = clean.calculate_lonlat_distance(latlons1[i-1], latlons2[j-1])
             dtw[i][j] = cost + min(dtw[i-1][j], dtw[i][j-1], dtw[i-1][j-1])
     return dtw[-1][-1]
 

### 2017-12-29 question2/NearestSubroutes.py

@@ -1,6 +1,7 @@
 import question1.CleanData as clean
 import os
 import utils
+from multiprocessing.pool import ThreadPool
 
 
 def question_a2(output_folder, test_file, trips_list):
@@ -18,13 +19,16 @@ def find_similar_subroots_per_test_trip(test_trip, trips_list, k):
     test_lonlat = test_trip['points']
     # keep track of the current common subsequences
     max_subseqs = []
+    pool = ThreadPool(processes=10)
     # iterate over the candidates
     for tripidx, trip in enumerate(trips_list):
         # get coordinates
         trip_lonlat = trip['points']
         timestart = utils.tic()
         # compute common subsequences between the test trip and the current candidate
-        subseqs, subseqs_idx = calc_lcss(test_lonlat, trip_lonlat)
+        async_result = pool.apply_async(calc_lcss, (test_lonlat, trip_lonlat))
+        subseqs, subseqs_idx = async_result.get()
+        # subseqs, subseqs_idx = calc_lcss(test_lonlat, trip_lonlat)
         elapsed = utils.tictoc(timestart)
         # sort by decr. length
         subseqs_idx = sorted(subseqs_idx, key = lambda x : len(x), reverse=True)

### 2017-12-29 Report/BigDataHomework1.tex
changes to report

@@
 The two lists (trips\_too\_small and trips\_too\_big) will keep the excluded trips due to either not valid total distance or to invalid max distance between successive points.
-For each trip in the list (returned from question A) the total distance is calculate, using the function calculate\_lonlat\_distance, which returns the distance between two successive points using the haversine formula.
+For each trip in the training dataset (in the dictionary format from question A) the total distance is calculate, using the function calculate\_lonlat\_distance, which returns the distance between two successive points using the haversine formula.
 Following the same logic, the max distance between successive points of a trip is calculated and a distance is found more than 2km, the trip is excluded.
@@
 The requirement is to visualize five of the given train trips.
-For this purporse, a new trips list is creating containing five trips, excluding all trips with null journey\_id.
-Afterwards, the data\_visualization function is called, and a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).
+For this purporse, after having excluded all trips with null journey\_id and with less than 1 lists with coordinates, the trips list is shuffled and then iterated only for 5 trips.
+Afterwards, a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).
 

### 2017-12-29 Main.py
add todos

@@ -28,7 +28,7 @@ def question_2(train_file, test_files, output_folder, maps_folder):
     print(">>> Running question 2a1 - Nearest neighbours computation")
     nn.question_a1(maps_folder, test_files[0], trips_list)
     print(">>> Running question 2a2 - Nearest subroutes computation")
-    ns.question_a2(output_folder, test_files[1], trips_list)
+    ns.question_a2(maps_folder, test_files[1], trips_list)
     print(">>> Running question 2b - Cell grid quantization")
     cellgrid = (4, 3)
     print("Using cell grid:", cellgrid)

### 2017-12-29 question2/JourneyClassification.py

@@ -5,6 +5,7 @@ from sklearn.neighbors import KNeighborsClassifier
 from sklearn.linear_model import LogisticRegression
 from sklearn.ensemble import RandomForestClassifier
 from sklearn.metrics import accuracy_score
+from sklearn.metrics import classification_report
 import utils
 
 
@@ -93,6 +94,8 @@ def knn_classification(train, val, targets, k):
     knn_classifier = KNeighborsClassifier(n_neighbors=k)
     knn_classifier.fit(train[0], train[1])
     res = knn_classifier.predict(val[0])
+    # TODO fix the below
+    # print(classification_report(res, np.array(twenty_train.target)[test_index], target_names=twenty_train.target_names))
     return accuracy_score(res, val[1])
 
 

### 2017-12-29 utils.py

@@ -9,13 +9,13 @@ import pickle
 import json
 
 
-
 def tic():
     '''
     Get timestamp
     :return:
     '''
     return datetime.datetime.now()
+
 
 def tictoc(previous_tic):
     '''
@@ -39,6 +39,7 @@ def tictoc(previous_tic):
     if days:
         msg = "%d days! " % days + msg
     return msg
+
 
 def write_trips(output_file, trips_list):
     '''
@@ -74,6 +75,7 @@ def serialize_trips(output_file, trips_list):
     with open(output_file, "wb") as f:
         pickle.dump(trips_list, f)
 
+
 def read_trips(filepath):
     '''
     Read a trips file, csv or pickle
@@ -105,6 +107,7 @@ def read_trips(filepath):
                 obj["points"] = points
                 trips.append(obj)
     return trips
+
 
 # Functions used for trips visualization
 #######################################
@@ -196,6 +199,7 @@ def html_to_png(html_path, png_path):
         print(ex)
         exit(1)
 
+
 def get_lonlat_tuple(points):
     '''
     :param points: [(lon1,lat1), (lon2,lat2),...]
@@ -203,6 +207,8 @@ def get_lonlat_tuple(points):
     '''
     return ([l[0] for l in points], [l[1] for l in points])
 
+
+# TODO check, problem with maps in result, wrong zoom
 def visualize_point_sequences(all_pts, colors, labels, file_name):
     '''
     Generic geocoordinate multi-plot visualization function with gmplot.

### 2017-12-29 question2/NearestNeighbours.py
add pool.close and references to report

@@ -45,6 +45,9 @@ def calculate_nns(test_trip, trips_list):
         # distance = calculate_dynamic_time_warping(test_lonlat, trip_lonlat)
         print("Calculated distance: %.2f for trip: %d/%d : %s" % (distance, i+1, len(trips_list), str(trip['id'])))
         nearest_neighbours.append((trip['id'], distance))
+
+    pool.close()
+    pool.join()
     # sort the list to increasing distance
     nearest_neighbours = sorted(nearest_neighbours, key=lambda k: k[1])
     # return the top 5
@@ -67,6 +70,8 @@ def calculate_dynamic_time_warping(latlons1, latlons2):
             cost = async_result.get()
             # cost = clean.calculate_lonlat_distance(latlons1[i-1], latlons2[j-1])
             dtw[i][j] = cost + min(dtw[i-1][j], dtw[i][j-1], dtw[i-1][j-1])
+    pool.close()
+    pool.join()
     return dtw[-1][-1]
 
 def get_trip_from_id(trips_list, nearest_neighbours):

### 2017-12-29 question2/NearestSubroutes.py

@@ -35,6 +35,8 @@ def find_similar_subroots_per_test_trip(test_trip, trips_list, k):
         # update the list of the longest subsequences
         max_subseqs = update_current_maxsubseq(max_subseqs, subseqs_idx, k, elapsed, tripidx)
     print("Got %d common subsequences" % len(max_subseqs))
+    pool.close()
+    pool.join()
     if len(max_subseqs) != k:
         print("WARNING: Specified %d subseqs!" % k)
     return max_subseqs

### 2018-01-03 Main.py
Various fixes

@@ -8,6 +8,7 @@ import question2.NearestSubroutes as ns
 import question2.MapInGridView as gv
 import question2.JourneyClassification as jc
 import utils
+import random
 
 
 def question_1(input_file, output_file, output_folder, maps_folder):
@@ -21,12 +22,12 @@ def question_1(input_file, output_file, output_folder, maps_folder):
     print("Finished question1")
 
 
-def question_2(train_file, test_files, output_folder, maps_folder):
+def question_2(train_file, test_files, output_folder, maps_folder, paropts):
     # Question 2
     # Read the training data
     trips_list = utils.read_trips(train_file)
     print(">>> Running question 2a1 - Nearest neighbours computation")
-    nn.question_a1(maps_folder, test_files[0], trips_list)
+    nn.question_a1(maps_folder, test_files[0], trips_list, paropts)
     print(">>> Running question 2a2 - Nearest subroutes computation")
     ns.question_a2(maps_folder, test_files[1], trips_list)
     print(">>> Running question 2b - Cell grid quantization")
@@ -58,6 +59,11 @@ if __name__ == '__main__':
     input_folder  = os.path.abspath(sys.argv[1])
     output_folder = os.path.abspath(sys.argv[2])
 
+    rand_seed = 123123
+    random.seed(rand_seed)
+
+    paropts = ("processes",8)
+
     # question 1
     ############
 
@@ -79,4 +85,4 @@ if __name__ == '__main__':
     test_files = [ os.path.join(input_folder, "test_set_a%d.csv" % t) for t in [1,2]]
 
     # run
-    question_2(output_file_clean, test_files, output_folder, maps_folder)
+    question_2(output_file_clean, test_files, output_folder, maps_folder, paropts)

### 2018-01-03 question1/CleanData.py

@@ -24,23 +24,26 @@ def filter_trips(trips_list):
     :param trips_list:
     :return:
     '''
-    trips_too_small, trips_too_big = [], []
+
+    trips_too_small, trips_too_big, nulls = [], [], []
     for trip in trips_list:
-        total_dist = caluclate_total_distance_per_trip(trip)
+        total_dist = calculate_total_distance_per_trip(trip)
         # too small journey check
         if total_dist < 2:
             trips_too_small.append(trip)
             trips_list.remove(trip)
+            continue
         max_dist = calculate_max_dist(trip)
         if max_dist > 2:
             trips_too_big.append(trip)
             trips_list.remove(trip)
+            continue
     print("Total trips deleted due to total distance less than 2km: %d" % len(trips_too_small))
     print("Total trips deleted due to max distance between two points more than 2km: %d" % len(trips_too_big))
     return trips_list
 
 
-def caluclate_total_distance_per_trip(trip):
+def calculate_total_distance_per_trip(trip):
     '''
     Calculates the total distance of a trip
     :param trip:

### 2018-01-03 question1/Preprocessing.py

@@ -9,8 +9,9 @@ def question_1a(input_file, output_file):
     '''
     line_objects = utils.read_train_set(input_file)
     map_per_vid = get_data_per_vehicle_id(line_objects)
-    map_per_vid = get_data_per_vehicle_and_journey_id(map_per_vid)
-    trips_list = write_trips(output_file, map_per_vid)
+    map_per_vidjid = get_data_per_vehicle_and_journey_id(map_per_vid)
+    trips_list = write_trips(output_file, map_per_vidjid)
+    print("Done with question 1!")
     return trips_list
 
 def get_data_per_vehicle_id(line_objects):
@@ -19,6 +20,7 @@ def get_data_per_vehicle_id(line_objects):
     :param line_objects:
     :return:
     '''
+    print("Aggregating per vehicle id")
     map_per_vid = {}
     for obj in line_objects:
         vid = obj["vid"]
@@ -28,29 +30,35 @@ def get_data_per_vehicle_id(line_objects):
         map_per_vid[vid].append(obj_without_vid)
     return map_per_vid
 
-
 def get_data_per_vehicle_and_journey_id(map_per_vid):
     '''
     Aggregates data wrt the journey ID
     :param map_per_vid:
-    :return:
+    :return: vid1:{jid1:[pt1,pt2...], jid2:[pt1,pt2],...},vid2:{...}
     '''
+    print("Aggregating per journey id")
     for vehicle_id in map_per_vid:
+        # get elements for this vehicle id
         map_per_jid = {}
         vid_list = map_per_vid[vehicle_id]
+        # make a map wrt to the journey id
         for item in vid_list:
             jid = item["jid"]
             if jid not in map_per_jid:
                 map_per_jid[jid] = []
+            # assign the points of each journeyid, minus that id
             obj_without_jid = {key: item[key] for key in item if key != "jid"}
             map_per_jid[jid].append(obj_without_jid)
+        # for each jid for this vehicle
         for jid in map_per_jid:
+            # get poitns
             jid_list = map_per_jid[jid]
+            # sort per timestamp
             sorted_jid_list = sorted(jid_list, key=lambda k: k['ts'])
             map_per_jid[jid] = sorted_jid_list
+        # assign jid-points map to this vehicle id
         map_per_vid[vehicle_id] = map_per_jid
     return map_per_vid
-
 
 def write_trips(filename, map_per_vid):
     '''
@@ -59,6 +67,7 @@ def write_trips(filename, map_per_vid):
[…]

### 2018-01-03 question2/NearestNeighbours.py

@@ -1,12 +1,16 @@
 import question1.CleanData as clean
 import os
 import utils
-from multiprocessing.pool import ThreadPool
+from multiprocessing.pool import Pool, ThreadPool
+import threading
+import itertools
+from dtw import dtw as libdtw
+
 
 
 # TODO check time
 # TODO check for threads in plots
-def question_a1(output_folder, test_file, trips_list):
+def question_a1(output_folder, test_file, trips_list, paropts):
     print("WARNING: Reading training file part because test file is not supplied.")
     test_trip_list = trips_list[:5]
     for i,test_trip in enumerate(test_trip_list):
@@ -15,7 +19,8 @@ def question_a1(output_folder, test_file, trips_list):
         # prepare to count time
         millis_start = utils.tic()
         # compute nearest neighbours
-        nns_ids_distances = calculate_nns(test_trip, trips_list)
+        #parallelize = None
+        nns_ids_distances = calculate_nns(test_trip, trips_list, paropts = paropts)
         # get time elapsed
         elapsed = utils.tictoc(millis_start)
         # get the whole data of the returned trips, add distances
@@ -24,7 +29,8 @@ def question_a1(output_folder, test_file, trips_list):
         preprocessing_for_visualization(test_trip, nns_ids_distances, neighbour_trips, outfile_name, elapsed, i)
 
 
-def calculate_nns(test_trip, trips_list):
+
+def calculate_nns(test_trip, trips_list, paropts = None, ret_container = None):
 
     '''
     :param test_trip: a trip row
@@ -32,47 +38,158 @@ def calculate_nns(test_trip, trips_list):
     :return: returns a list of the 5 nearest neighbours from trips_list, for the test_trip. Format is (trip_id, distance)
     '''
 
+    # parallelization type
+    if paropts:
+        print("Parallelizing with",paropts)
+        partype, numpar = paropts
+    else:
+        partype, numpar = None, None
+
+    tic = utils.tic()
     # get test coordinate list
     test_lonlat = test_trip["points"]
-    nearest_neighbours = []
-    pool = ThreadPool(processes=10)
-    for i, trip in enumerate(trips_list):
-        # get candidate coordinate list
-        trip_lonlat = trip["points"]
-        # calcluate distance
-        async_result = pool.apply_async(calculate_dynamic_time_warping, (test_lonlat, trip_lonlat))
-        distance = async_result.get()
[…]

### 2018-01-03 utils.py

@@ -64,6 +64,16 @@ def write_trips(output_file, trips_list):
         with open(output_file, "w") as f:
             json.dump(trips_list, f)
 
+def sublist(llist, slen):
+    '''
+    return llist divided to sublists of at most slen
+    :param llist:
+    :param sublist_length:
+    :param only_num:
+    :return:
+    '''
+    divs = range(0, len(llist), slen)
+    return [ llist[i:i+slen] for i in divs]
 
 def serialize_trips(output_file, trips_list):
     '''
@@ -74,7 +84,6 @@ def serialize_trips(output_file, trips_list):
     '''
     with open(output_file, "wb") as f:
         pickle.dump(trips_list, f)
-
 
 def read_trips(filepath):
     '''
@@ -117,7 +126,8 @@ def read_train_set(filepath):
     :param filepath:
     :return:
     '''
-    print("Reading file", filepath)
+    print("Reading training file ignoring null jids", filepath)
+    num_nulls, num_empties = 0, 0
     line_objects = []
     with open(filepath, "r") as f:
         next(f)
@@ -125,11 +135,21 @@ def read_train_set(filepath):
             line_contents = line.strip().split(',')
             obj = {}
             obj["jid"] = line_contents[0]
+            # drop nulls and empties
+            if obj["jid"] == "null":
+                num_nulls += 1
+                continue
+            # drop nulls and empties
+            if obj["jid"] == "":
+                num_empties += 1
+                continue
             obj["vid"] = line_contents[1]
             obj["ts"] = line_contents[2]
             obj["lon"] = line_contents[3]
             obj["lat"] = line_contents[4]
             line_objects.append(obj)
+    print("Discarded %d null journey id entries" % num_nulls)
+    print("Discarded %d empty journey id entries" % num_empties)
     return line_objects
 
 

### 2018-01-06 Report/BigDataHomework1.tex
add pool.close and references to report
Various fixes
little change to pandas files

@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+	\todo{Remove the null jids from the start}
 	\section{Introduction}
@@
 	In this section, further details are provided related to the preprocessing and cleaning of the training data as well as to the trip's visualization.
-All files implementing the requirements of this part are in the folder FirstQuestion.
+All files implementing the requirements of this part are in the folder question1.
 
@@
 		 \url{https://en.wikipedia.org/wiki/Longest_common_subsequence_problem}
+		 \bibitem{report} SciKit Learn: Classification report \\
+		 \url{http://scikit-learn.org/stable/modules/generated/sklearn.metrics.classification_report.html}
+		 \bibitem{threads} ThreadPool \\
+		 \url{https://stackoverflow.com/questions/8533318/multiprocessing-pool-when-to-use-apply-apply-async-or-map}
     \end{thebibliography}

### 2018-01-06 ReportPandas/BigDataHomeworkPandas.tex
little change to pandas files

--- 2018-01-06 Report/BigDataHomework1.tex
@@
     In the Main.py file, it is required to provide the path to the input folder (containing the training and test files) and the output folder where all results will be stored.
-In this folder an additional directory is created, in order to keep all maps extracted from this application.
+In this folder an additional directories are created, in order to keep all maps extracted from this application and also to store the charts showing the classification accuracy.
 There are two ways in order to run this project, either add these folders in the edit configurations choice from the menu, or use the run.sh in order to provide these args.
@@
 	\subsection{Question A}
-	The first task of the project was to pre-process the given data, which is implemented in the file Preprocessing.py.
-In the read\_file\_contents, the train\_set.csv file is provided as input and for each line the necessary information is stored in a dictionary (journey\_id, vehicle\_id, timestamp, longitude and latitude).
+	The requirement in this part is to preprocess the input training data and create a new file where will be in 3 columns.
+The first column is the trip\_id incrementing by 1 for each trip, the second is the journey\_id and the last one contains a list of lists, where each item contains a timestamp, a longitude and a latitude.
+In order to read the given csv file, pandas library was used transforming the csv file to a dataframe.
+The first necessary thing to do, was to remove all data with null journeyPatternId and then using the groupby method on the columns vehicleID and journeyPatternId, three different dataframes were created.
+By concatenating them, the final dataframe contains five columns, where the three last columns contain a list of timestamps, a list of longitudes and a list of latitudes respectively.
 
-	Since now the data from the input file are stored in the dictionary, the next step is to save the data based on the vehicle id.
-The function get\_data\_per\_vehicle\_id gets this dictionary as input and for every object in the dictionary, creates a list with key the vehicle\_id and then stores the remaining data connected with this id.
-The same process is followed for the sorting of data based on the journey\_id, therefore the structure where data are stored is a dictionary with key the vehicle\_id and with value another dictionary with key the journey\_id and with value a list containing several lists with the timestamp, the longitude and the latitude information.
-
-	Finally, in the function get\_trips, the above structure is stored in the file trips.csv.
-The output format follows the requested pattern giving each trip an id.
-The final data structure is a list where each item has the trip\_id, the journey\_id and lists containing the timestamp, the longitude and the latitude.
+	By iterating to each dataframe's row, the three lists are now concatenated in one column and therefore the other two are dropped from the dataframe.
+Then, the column vehicleID is replaced by the tripId column which is the index of this dataframe, starting from 0 and incrementing by 1.
+Finally, in the column points, which now contains the list of lists with the timestamps and the poins, is sorted based on the timestamp of each list item.
+The final result is exported in a file, keeping the requested format.
 

### 2018-01-06 Main.py

@@ -1,19 +1,14 @@
 import sys, os, pkgutil
 import question1.Preprocessing as prep
-import question1.PandasPreprocessing as pdp
 import question1.CleanData as clean
-import question1.CleanDataPandas as cdp
 import question1.DataVisualization as visual
-import question1.DataVisualizationPandas as dvp
+import question1_pandas as qp1
 
+import question2_pandas as qp2
 import question2.NearestNeighbours as nn
-import question2.NearestNeighboursPandas as nnp
 import question2.NearestSubroutes as ns
-import question2.NearestSubroutsPandas as nsp
 import question2.MapInGridView as gv
-import question2.MapInGridViewPandas as gvp
 import question2.JourneyClassification as jc
-import question2.JourneyClassificationPandas as jcp
 import utils
 import random
 
@@ -22,11 +17,11 @@ def question_1(input_file, output_file, output_file_clean, maps_folder, pandas_o
     # Question 1
     if pandas_option:
         print(">>> Running question 1a - parsing the training data")
-        trips_list, df = pdp.create_trips_file(input_file, output_file)
+        trips_list, df = qp1.create_trips_file(input_file, output_file)
         print(">>> Running question 1b - cleaning the training data")
-        trips_list, df = cdp.filter_trips_pandas(output_file_clean, df)
+        trips_list, df = qp1.filter_trips_pandas(output_file_clean, df)
         print(">>> Running question 1c - visualizing the training data")
-        dvp.visualize_trips(maps_folder, df)
+        qp1.visualize_trips(maps_folder, df)
     else:
         print(">>> Running question 1a - parsing the training data")
         trips_list = prep.question_1a(input_file, output_file)
@@ -41,15 +36,15 @@ def question_2(train_file, test_files, output_folder, maps_folder, class_folder,
     # Question 2
     if pandas_option:
         print(">>> Running question 2a1 - Nearest neighbours computation")
-        # nnp.question_a1(output_folder, train_file, test_files[0], paropts)
+        qp2.question_a1(output_folder, train_file, test_files[0], paropts)
         print(">>> Running question 2a2 - Nearest subroutes computation")
-        # nsp.question_a2(maps_folder, test_files[1], train_file)
+        qp2.question_a2(maps_folder, test_files[1], train_file)
         print(">>> Running question 2b - Cell grid quantization")
         cellgrid = (4, 3)
         print("Using cell grid:", cellgrid)
-        features_file = gvp.question_b(train_file, cellgrid, output_folder)
[…]

### 2018-01-06 UtilsPandas.py

@@ -1,229 +1,0 @@
-import gmplot
-import subprocess
-import os
-import matplotlib.pyplot as plt
-import pylab
-from scipy.misc import imread
-import datetime
-import pickle
-import json
-
-
-def tic():
-    '''
-    Get timestamp
-    :return:
-    '''
-    return datetime.datetime.now()
-
-
-def tictoc(previous_tic):
-    '''
-    Get duration from previous timestamp
-    :param previous_tic:
-    :return:
-    '''
-    # get duration in usec
-    diff = tic() - previous_tic
-    # get durations of datetime.now() members
-    days = diff.days
-    secs = diff.seconds
-    usec = diff.microseconds
-    # aggregate to larger time units
-    msec, usec = divmod(usec, 1000)
-    sec_, msec = divmod(msec, 1000)
-    secs += sec_
-    mins, sec = divmod(secs, 60)
-    hrs, mins = divmod(mins,60)
-    msg = "%02d:%02d:%02d:%03d" % (hrs, mins, secs, msec)
-    if days:
-        msg = "%d days! " % days + msg
-    return msg
-
-
-def idx_to_lonlat(points, idx=None, format="tuple"):
-    if idx is None:
-        idx = list(range(len(points)))
-    if type(idx) != list:
-        idx = [idx]
-    lats, lons = [], []
-    for i in idx:
-        lons.append(points[i][1])
-        lats.append(points[i][2])
-    if format == "tuple":
-        return (lons, lats)
-    elif format == "tuples":
-        return list(zip(lons,lats))
-
-
-def sublist(llist, slen):
-    '''
-    return llist divided to sublists of at most slen
-    :param llist:
-    :param sublist_length:
-    :param only_num:
-    :return:
-    '''
-    divs = range(0, len(llist), slen)
-    return [ llist[i:i+slen] for i in divs]
-
-
-def get_total_points(df):
-    total_points = []
-    for index, row in df.iterrows():
-        points = row["points"]
-        total_points.append(points)
-    return total_points
-
-
-def get_lonlat_tuple(points):
-    '''
-    :param points: [(lon1,lat1), (lon2,lat2),...]
-    :return: ([lon1,lon2,...],[lat1,lat2,...])
-    '''
-    return ([l[1] for l in points], [l[2
-                                     ] for l in points])
-
-
-# Visualization
-##################
-
-def write_group_gml(lonlats_tuplelists, outpath, colors=None):
-    '''
[…]

### 2018-01-06 main_pandas.py

@@ -1,0 +1,80 @@
+import sys, os, pkgutil
+import question1_pandas as qp1
+import question2_pandas as qp2
+import random
+
+
+def question_1(input_file, output_file, output_file_clean, maps_folder):
+    # Question 1
+    print(">>> Running question 1a - parsing the training data")
+    trips_list, df = qp1.create_trips_file(input_file, output_file)
+    print(">>> Running question 1b - cleaning the training data")
+    trips_list, df = qp1.filter_trips_pandas(output_file_clean, df)
+    print(">>> Running question 1c - visualizing the training data")
+    qp1.visualize_trips(maps_folder, df)
+    print("Finished question1")
+
+
+def question_2(train_file, test_files, output_folder, maps_folder, class_folder, paropts):
+    # Question 2
+    print(">>> Running question 2a1 - Nearest neighbours computation")
+    qp2.question_a1(output_folder, train_file, test_files[0], paropts)
+    print(">>> Running question 2a2 - Nearest subroutes computation")
+    qp2.question_a2(maps_folder, test_files[1], train_file)
+    print(">>> Running question 2b - Cell grid quantization")
+    cellgrid = (4, 3)
+    print("Using cell grid:", cellgrid)
+    features_file = qp2.question_b(train_file, cellgrid, output_folder)
+    print(">>> Running question 2c - Classification")
+    qp2.question_c(features_file, class_folder)
+
+
+def check_dependencies():
+    deps = ['numpy', 'scipy']
+    for dep in deps:
+        res = pkgutil.find_loader(dep)
+        if res is None:
+            print("The python3 package [%s] is required to run." % dep)
+            exit(1)
+    if not os.path.exists('rasterize.js'):
+        print("The rasterize.js tool from the phantomjs framework is required to run.")
+        exit(1)
+
+
+if __name__ == '__main__':
+    if len(sys.argv) < 2:
+        print("Usage: %s inputfolder outputfolder" % sys.argv[0])
+        exit(1)
+    check_dependencies()
+    print("Running %s with arguments: %s" % (os.path.basename(sys.argv[0]), sys.argv))
+    input_folder  = os.path.abspath(sys.argv[1])
+    output_folder = os.path.abspath(sys.argv[2])
+
+    rand_seed = 123123
+    random.seed(rand_seed)
+
+    paropts = None
+
+    # question 1
+    ############
+
+    # prepare files
[…]

### 2018-01-06 question1/DataVisualizationPandas.py

@@ -1,6 +1,6 @@
 import os
 import random
-import UtilsPandas as up
+import utils_pandas as up
 
 
 def visualize_trips(output_folder, df):

### 2018-01-06 question1_pandas.py

@@ -1,0 +1,136 @@
+import pandas as pd
+from math import sin, cos, sqrt, atan2, radians
+import os
+import random
+import utils_pandas as up
+
+
+# Question 1a
+#################
+
+def create_trips_file(input_file, output_file):
+    print("Reading training file with pandas ignoring null jids", input_file)
+    df = pd.read_csv(input_file)
+    df = df[pd.notnull(df['journeyPatternId'])].reindex()
+    timestamps = df.groupby(["vehicleID", "journeyPatternId"])["timestamp"].apply(list)
+    lons = df.groupby(["vehicleID", "journeyPatternId"])["longitude"].apply(list)
+    lats = df.groupby(["vehicleID", "journeyPatternId"])["latitude"].apply(list)
+    df = pd.concat([timestamps, lons, lats], axis=1, ignore_index=False)
+    print("File read")
+    print("Start processing data")
+
+    for index, row in df.iterrows():
+        tslist, lonslist, latslist = [], [], []
+        tslist = row["timestamp"]
+        lonslist = row["longitude"]
+        latslist = row["latitude"]
+        data_list = []
+        for i,timestamp in enumerate(tslist):
+            new_list = []
+            new_list.append(timestamp)
+            new_list.append(lonslist[i])
+            new_list.append(latslist[i])
+            data_list.append(new_list)
+        row["timestamp"] = data_list
+    new_df = create_new_dataframe()
+    new_df = pd.concat([new_df, df["timestamp"]], axis=1, ignore_index=False)
+    new_df.to_csv(output_file)
+    df = pd.read_csv(output_file)
+    df = df.drop(["vehicleID", "tripId"], axis=1)
+    df.index.name = "tripId"
+    df.columns = ["journeyId", "points"]
+    df = sort_timestamps(df)
+    df.to_csv(output_file)
+    trips_list = df.to_dict(orient='dict')
+    return trips_list, df
+
+
+def create_new_dataframe():
+    columns = []
+    columns.append("tripId")
+    new_df = pd.DataFrame(columns=columns)
+    return new_df
+
+
+def sort_timestamps(df):
+    for index, row in df.iterrows():
+        points = row["points"]
+        points = eval(points)
+        points.sort(key=lambda x: x[0])
+        df.at[index, "points"] = points
+    return df
+
+
+# Question 1b
+#############
+
+def filter_trips_pandas(output_file, df):
+    trips_too_small, trips_too_big = [], []
+    for index, row in df.iterrows():
[…]

### 2018-01-06 question2/JourneyClassificationPandas.py

@@ -1,47 +1,9 @@
-import pandas as pd
-import os
 import numpy as np
 from sklearn.model_selection import KFold
 from sklearn.neighbors import KNeighborsClassifier
 from sklearn.linear_model import LogisticRegression
 from sklearn.ensemble import RandomForestClassifier
 from sklearn.metrics import accuracy_score
-from sklearn.metrics import classification_report
-import UtilsPandas as up
-
-
-def question_c(features_file, output_folder):
-    feature_df = pd.read_csv(features_file)
-    kf = KFold(n_splits=10)
-    grid_features, targets = preprocess_data(feature_df)
-    folds_idxs = list(kf.split(grid_features))
-    folds_idxs = list(kf.split(grid_features))
-    trips_array = np.asarray(grid_features)
-    targets = np.asarray(targets)
-
-    classifiers = ["knn", "logreg", "randfor"]
-    accuracies = {}
-    # classify
-    for classifier in classifiers:
-        print("Testing classifier [%s]" % classifier)
-        # train & test each classifier
-        # for each fold
-        accuracies[classifier] = []
-        for i, (train_idx, val_idx) in enumerate(folds_idxs):
-            print("\tClassifing fold %d/%d" % (i + 1, len(folds_idxs)))
-            train = (trips_array[train_idx], targets[train_idx])
-            val = (trips_array[val_idx], targets[val_idx])
-            if classifier == "knn":
-                k = 5
-                res = knn_classification(train, val, targets, k)
-            elif classifier == "logreg":
-                res = logreg_classification(train, val)
-            elif classifier == "randfor":
-                res = randfor_classification(train, val)
-            accuracies[classifier].append(res)
-        titlestr = "%s, overall accuracy: %2.4f" % (classifier, np.mean(accuracies[classifier]))
-        up.barchart(list(range(1, 11)), accuracies[classifier], title=titlestr, ylabel="accuracy",
-                       save=os.path.join(output_folder, classifier))
 
 
 def preprocess_data(feature_df):

### 2018-01-06 question2/MapInGridViewPandas.py

@@ -1,17 +1,6 @@
 import os
-import pandas as pd
 import matplotlib.pyplot as plt
-import UtilsPandas as up
-
-
-def question_b(train_file, number_of_cells, output_folder):
-    # specify files
-    output_file = os.path.join(output_folder, "tripFeatures.csv")
-    train_df = pd.read_csv(train_file)
-    max_lat, max_lon, min_lat, min_lon = find_min_max_latlong(train_df)
-    rows, columns, cell_names = create_grid(number_of_cells, max_lat, max_lon, min_lat, min_lon, output_folder=output_folder)
-    replace_points(train_df, rows, columns, cell_names, output_file)
-    return output_file
+import utils_pandas as up
 
 
 def find_min_max_latlong(train_df):

### 2018-01-06 question2/NearestNeighboursPandas.py

@@ -1,28 +1,10 @@
 import pandas as pd
 import os
-import UtilsPandas as up
-from multiprocessing.pool import Pool, ThreadPool
+import utils_pandas as up
+from multiprocessing.pool import ThreadPool
 import threading
 from dtw import dtw as libdtw
-import question1.CleanDataPandas as cdp
-
-
-def question_a1(output_folder, clean_file, test_file, paropts):
-    test_df = pd.read_csv(test_file)
-    train_df = pd.read_csv(clean_file)
-    for index, row in test_df.iterrows():
-        print("Examining test element %d" % (index + 1))
-        outfile_name = os.path.join(output_folder, "nn_%d_" % (index + 1))
-        # prepare to count time
-        millis_start = up.tic()
-        # compute nearest neighbours
-        test_points = row["points"]
-        test_points = eval(test_points)
-        nns_ids_distances = calculate_nns(test_points, train_df, paropts=paropts)
-        # get time elapsed
-        elapsed = up.tictoc(millis_start)
-        # visualize
-        preprocessing_for_visualization(test_points, nns_ids_distances, outfile_name, elapsed, index)
+import question1_pandas as qp1
 
 
 def calculate_nns(test_points, train_df, paropts=None):
@@ -155,13 +137,13 @@ def calculate_dynamic_time_warping(latlons1, latlons2, paropts = None, impl = "d
         dtw[0][0] = 0
         for i in range(1, 1 + len(latlons1)):
             for j in range(1, 1 + len(latlons2)):
-                cost = cdp.calculate_lonlat_distance(latlons1[i-1], latlons2[j-1])
+                cost = qp1.calculate_lonlat_distance(latlons1[i-1], latlons2[j-1])
                 dtw[i][j] = cost + min(dtw[i - 1][j], dtw[i][j - 1], dtw[i - 1][j - 1])
         return dtw[-1][-1]
 
     elif impl == "lib":
         # https://github.com/pierre-rouanet/dtw/blob/master/examples/simple%20example.ipynb
-        ret = libdtw(latlons1, latlons2, lambda x,y : cdp.calculate_lonlat_distance(x,y))
+        ret = libdtw(latlons1, latlons2, lambda x,y : qp1.calculate_lonlat_distance(x,y))
         return ret[0]
 
 
@@ -174,11 +156,11 @@ def compute_dists(points_list, paropts):
     if partype == 'process':
         pool = ThreadPool(processes=numpar)
[…]

### 2018-01-06 question2/NearestSubroutsPandas.py

@@ -1,20 +1,6 @@
-import pandas as pd
-import os
-import UtilsPandas as up
+import utils_pandas as up
 from multiprocessing.pool import ThreadPool
-import question1.CleanDataPandas as cdp
-
-
-def question_a2(output_folder, test_file, train_file):
-    test_df = pd.read_csv(test_file)
-    train_df = pd.read_csv(train_file)
-    for index, row in test_df.iterrows():
-        print("Extracting subroutes for test trip %d/%d" % (index + 1, len(test_df.index)))
-        file_name = os.path.join(output_folder, "subroutes_%d_" % (index + 1))
-        test_points = row["points"]
-        test_points = eval(test_points)
-        max_subseqs = find_similar_subroutes_per_test_trip(test_points, train_df)
-        preprocessing_for_visualisation(test_points, max_subseqs, file_name, index)
+import question1_pandas as qp1
 
 
 def find_similar_subroutes_per_test_trip(test_points, train_df, k=5):
@@ -57,7 +43,7 @@ def calc_lcss(t1, t2):
     for i, p1 in enumerate(t1):
         for j, p2 in enumerate(t2):
             # calculate the dist
-            dist = cdp.calculate_lonlat_distance(p1, p2)
+            dist = qp1.calculate_lonlat_distance(p1, p2)
             equal = dist < 200
             if equal:
                 # the points are equal enough

### 2018-01-06 question2_pandas.py

@@ -1,0 +1,83 @@
+import question2.NearestNeighboursPandas as nn
+import question2.NearestSubroutsPandas as ns
+import question2.MapInGridViewPandas as gvp
+import question2.JourneyClassificationPandas as jcp
+from sklearn.model_selection import KFold
+import pandas as pd
+import numpy as np
+import utils_pandas as up
+import os
+
+
+def question_a1(output_folder, clean_file, test_file, paropts):
+    test_df = pd.read_csv(test_file)
+    train_df = pd.read_csv(clean_file)
+    for index, row in test_df.iterrows():
+        print("Examining test element %d" % (index + 1))
+        outfile_name = os.path.join(output_folder, "nn_%d_" % (index + 1))
+        # prepare to count time
+        millis_start = up.tic()
+        # compute nearest neighbours
+        test_points = row["points"]
+        test_points = eval(test_points)
+        nns_ids_distances = nn.calculate_nns(test_points, train_df, paropts=paropts)
+        # get time elapsed
+        elapsed = up.tictoc(millis_start)
+        # visualize
+        nn.preprocessing_for_visualization(test_points, nns_ids_distances, outfile_name, elapsed, index)
+
+
+def question_a2(output_folder, test_file, train_file):
+    test_df = pd.read_csv(test_file)
+    train_df = pd.read_csv(train_file)
+    for index, row in test_df.iterrows():
+        print("Extracting subroutes for test trip %d/%d" % (index + 1, len(test_df.index)))
+        file_name = os.path.join(output_folder, "subroutes_%d_" % (index + 1))
+        test_points = row["points"]
+        test_points = eval(test_points)
+        max_subseqs = ns.find_similar_subroutes_per_test_trip(test_points, train_df)
+        ns.preprocessing_for_visualisation(test_points, max_subseqs, file_name, index)
+
+
+def question_b(train_file, number_of_cells, output_folder):
+    # specify files
+    output_file = os.path.join(output_folder, "tripFeatures.csv")
+    train_df = pd.read_csv(train_file)
+    max_lat, max_lon, min_lat, min_lon = gvp.find_min_max_latlong(train_df)
+    rows, columns, cell_names = gvp.create_grid(number_of_cells, max_lat, max_lon, min_lat, min_lon, output_folder=output_folder)
+    gvp.replace_points(train_df, rows, columns, cell_names, output_file)
+    return output_file
+
+
+def question_c(features_file, output_folder):
+    feature_df = pd.read_csv(features_file)
[…]

### 2018-01-06 utils_pandas.py

@@ -1,0 +1,229 @@
+import gmplot
+import subprocess
+import os
+import matplotlib.pyplot as plt
+import pylab
+from scipy.misc import imread
+import datetime
+import pickle
+import json
+
+
+def tic():
+    '''
+    Get timestamp
+    :return:
+    '''
+    return datetime.datetime.now()
+
+
+def tictoc(previous_tic):
+    '''
+    Get duration from previous timestamp
+    :param previous_tic:
+    :return:
+    '''
+    # get duration in usec
+    diff = tic() - previous_tic
+    # get durations of datetime.now() members
+    days = diff.days
+    secs = diff.seconds
+    usec = diff.microseconds
+    # aggregate to larger time units
+    msec, usec = divmod(usec, 1000)
+    sec_, msec = divmod(msec, 1000)
+    secs += sec_
+    mins, sec = divmod(secs, 60)
+    hrs, mins = divmod(mins,60)
+    msg = "%02d:%02d:%02d:%03d" % (hrs, mins, secs, msec)
+    if days:
+        msg = "%d days! " % days + msg
+    return msg
+
+
+def idx_to_lonlat(points, idx=None, format="tuple"):
+    if idx is None:
+        idx = list(range(len(points)))
+    if type(idx) != list:
+        idx = [idx]
+    lats, lons = [], []
+    for i in idx:
+        lons.append(points[i][1])
+        lats.append(points[i][2])
+    if format == "tuple":
+        return (lons, lats)
+    elif format == "tuples":
+        return list(zip(lons,lats))
+
+
+def sublist(llist, slen):
+    '''
+    return llist divided to sublists of at most slen
+    :param llist:
+    :param sublist_length:
+    :param only_num:
+    :return:
+    '''
+    divs = range(0, len(llist), slen)
+    return [ llist[i:i+slen] for i in divs]
+
+
+def get_total_points(df):
+    total_points = []
+    for index, row in df.iterrows():
+        points = row["points"]
+        total_points.append(points)
+    return total_points
+
+
+def get_lonlat_tuple(points):
+    '''
+    :param points: [(lon1,lat1), (lon2,lat2),...]
+    :return: ([lon1,lon2,...],[lat1,lat2,...])
+    '''
+    return ([l[1] for l in points], [l[2
+                                     ] for l in points])
+
+
+# Visualization
+##################
+
+def write_group_gml(lonlats_tuplelists, outpath, colors=None):
+    '''
[…]

### 2018-01-06 Report/BigDataHomeworkPandas.tex
Merge reports

--- 2018-01-06 ReportPandas/BigDataHomeworkPandas.tex
@@
 	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	\todo{Remove the null jids from the start}
+
 	\section{Introduction}
@@
 	In this section, further details are provided related to the preprocessing and cleaning of the training data as well as to the trip's visualization.
-All files implementing the requirements of this part are in the folder question1.
+All files implementing the requirements of this part are in the file question1\_pandas.py.
 
@@
 
-	The above logic is implemented in the file CleanData.py in function filter\_trips.
+	The above logic is implemented in function filter\_trips\_pandas.
 The two lists (trips\_too\_small and trips\_too\_big) will keep the excluded trips due to either not valid total distance or to invalid max distance between successive points.
-For each trip in the training dataset (in the dictionary format from question A) the total distance is calculate, using the function calculate\_lonlat\_distance, which returns the distance between two successive points using the haversine formula.
+For each trip in the training dataset the total distance is calculated, using the function calculate\_lonlat\_distance, which returns the distance between two successive points using the haversine formula.
 Following the same logic, the max distance between successive points of a trip is calculated and a distance is found more than 2km, the trip is excluded.
 
-	Finally, the cleaned trips are kept in a list under the same format that was used in Question A and they are extracted in the file cleanTrips.csv, but also in the file tripsClean.pickle which keeps the data structure used.
-The functions for the data export are in the file utils.py (write\_trips\_to\_file and write\_trips\_using\_pickle).
+	Finally, the distances are checked and rows with invalid trips are removed from the dataframe, which is afterwards exported in a new file (trips\_clean.csv).
 

### 2018-01-06 Report/BigDataHomeworkPandas.tex
Update report with pandas

@@
 	\subsection{Question C}
-	This task is implemented in the file DataVisualization.py and the created output is stored in the output folder in the gmplots directory.
+	This task is implemented in the function visualize\_trips and the created output is stored in the output folder in the gmplots directory.
 The requirement is to visualize five of the given train trips.
-For this purporse, after having excluded all trips with null journey\_id and with less than 1 lists with coordinates, the trips list is shuffled and then iterated only for 5 trips.
+For this purporse, the trips list is shuffled in order to visualize random trips and then iterated only for 5 trips.
 Afterwards, a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).
 
-	Finally, the function write\_group\_gml from the utils.py file is called.
-In this function, it is calculated the mean for both longitudes and latitudes and then gmplot is uded so as to visualize the selected trips.
+	Finally, the function write\_group\_gml from the utils\_pandas.py file is called.
+In this function, it is calculated the mean for both longitudes and latitudes and then gmplot is used so as to visualize the selected trips.
 Below, there is an example of a trip shown on map is presented:
@@
 	\section{Exercise 2}
-	The files implementing the requirements of this part are in the folder SecondQuestion.
+	The files implementing the requirements of this part are in the file question2\_pandas.py and some functions implementing the logic for this question are in the files in the folder question 2.
 
@@
 
-	The code for this question is in the file NearestNeighbours.py.
-Firstly, all test trips in the test file are extracted and stored in the same data structure as the trip list that contains the training dataset.
-Subsequently, for each test trip all nearest neighbours are calculated in the function find\_nearest\_neighbours\_for\_test\_trip, keeping the time needed in a variable.
-At the beginning of the trips processing, the tic method from utils.py file is used (finds the current time in millisec using the datetime.datetime.now() method).
-After having found the nearest neighbours for the given test trip, the method tictoc() is called, returning the required time.
-In order to find the nearest neighbours, the following process is implemented: For the test trip and all the training trips, the longitudes and latitudes are stored in tuples which are provided to the function calculate\_dynamic\_time\_warping in order to find the distance between two points.
-The calculated distances are stored in a list, which finally is sorted in ascending order and only the five nearest neighbours are kept.
+	In the function question\_a1, two dataframes are created for both the test and the training trips.
+The test dataframe's rows are iterated and for each row the function calculate\_nns from the file NearestNeighboursPandas.py is called in order to get the nearest neighbours, calculating also the time needed for this action.
 
-	Since now the nearest neighbour trips are found, the trips list is updated (function: get\_updated\_trips\_list) containing now only those five trips.
-The final step now is to visualize the six trips.
-For this purpose, three list are created: a list with all the points (longitudes and latitudes), a list with the labels and a list with the colors (in this question the only color is the blue).
-All this data is provided to the function visualize\_paths in the utils.py file in order to export the trips as map images.
+	\todo{paralization}
+
+	In this function, the 3rd column of the dataframe is providing and it is transformed in tuples with the latitudes and longitudes.
+The list nearest neighbours will keep in tuples for each train trip the trip\_id, the calculated distance (which is calculated using the dynamic time wraping algorithm), the journeyId and finally all the points of the journey.
+	\todo{custom algorithm or library}
+
+	The final list is sorted and only the 5 nearest neighbours are kept for visualization.
+preprocessing\_for\_visualization, three list are created: a list with all the points (longitudes and latitudes) of the test trip and its neighbours, a list with the labels and a list with the colors (in this question the only color is the blue).
+All this data is provided to the function visualize\_paths in the utils\_pandas.py file in order to export the trips as map images.
 
@@
 	\subsection{Question A2}
-	Following the same logic as above, the given test file(test\_set\_a2.csv) is parsed storing the data in the usual format and the function find\_similar\_subroots\_per\_test\_trip is used to determine the nearest sub-routes for each test trip.
+	Following the same logic as above, in the file question2\_pandas and in the function question\_a2, two dataframes are created for the test and the training dataset using pandas library.
+The test dataframe is iterated row by row, in order to find the 5 trips from the training dataset which have the longest common subsequence with the specific test trip.
 
@@
 We keep all similar point values and indexes in two lists while iterating the two given lists.
-The distance between two points is calculated using the method calculate\_lonlat\_distance in the CleanData.py file.
+The distance between two points is calculated using the method calculate\_lonlat\_distance in the question1\_pandas.py file.
 Based on the exercise requirements, two points are considered as equals if their distance is less than 200 metres.
@@
 This time the points are separated so as to show which of them belong to the nearest sub-route and color them with red color.
-The function for the visualization is the same that was used for the question 2A1.
+The function for the visualization is the same that was used for the question 2a1 in the utils\_pandas.py file.
 
@@
 In short, a grid should be created and the points of each trip trace should be replaced by a specific square of the grid.
-The output file where the exported features are stored is the tripFeatures.csv and the tripFeatures.pickle, which was used in order to keep the final data structure with the features.
+The output file where the exported features are stored is the tripFeatures.csv.
 
@@
 
-	Afterwards, all points in the trip list should be replaced by the cell names that each point corresponds to.
+	Afterwards, all points in the trip dataframe should be replaced by the cell names that each point corresponds to.
 This logic is implemented in the method replace\_points.
-For each trip in the list, all coordinates are transformed in tuples format and using the find\_index function a name is mapped to this point.
-After having replaced all points with their new names, this list is exported in .csv and .pickle format.
+For each row in the dataframe, all coordinates are transformed in tuples format and using the find\_index function a name is mapped to this point.
+After having replaced all points with their new names, the dataframe is exported in .csv format.
 In order to be able to easily check the created grid, the method visualize\_grid is implemented so as to plot the grid.
@@
 	\subsection{Question C}
-	The file corresponding to this question is the JourneyClassification.py.
-First of all, the data structure is loaded from the .pickle file, created in above question as we have already mentioned.
-In the requirements of this question it is mentioned to use 10-fold cross validation, therefore the KFold class is used providing the required folds number.
-Then it is used in order to split our list to the training and test datasets.
-	\todo{explain preprocessdata function}
+	In the method question\_c in question2\_pandas.py file, a new dataframe is created from the features file, containing the grid cell names.
+All functions used in this question are in the file JourneyClassificationPandas.py.
+Before starting the classification, the method preprocess\_data was used in order to get the journey\_id of all trips as numbers and also to remove the letter C from the cell names.
 
+	In the requirements of this question it is mentioned to use 10-fold cross validation, therefore the KFold class is used providing the required folds number and then the training dataset is splitted in training and test folds.
 It worths to be mentioned here, that method asarrays from numpy was used to the necessary for this question lists.

### 2018-01-06 main_pandas.py
remove output folder

@@ -18,7 +18,7 @@ def question_1(input_file, output_file, output_file_clean, maps_folder):
 def question_2(train_file, test_files, output_folder, maps_folder, class_folder, paropts):
     # Question 2
     print(">>> Running question 2a1 - Nearest neighbours computation")
-    qp2.question_a1(output_folder, train_file, test_files[0], paropts)
+    qp2.question_a1(maps_folder, train_file, test_files[0], paropts)
     print(">>> Running question 2a2 - Nearest subroutes computation")
     qp2.question_a2(maps_folder, test_files[1], train_file)
     print(">>> Running question 2b - Cell grid quantization")

### 2018-01-06 Report/BigDataHomeworkPandas.tex
update report, update dependecies

@@
 
-	\todo{paralization}
+	There are three ways to execute this task: using processes, threads or serial execution.
+This information is provided as input giving the kind of execution and the number of processes/threads as well.
+In the function run\_with\_processes an async task is executed for each process and then the result is stored in a list, while in the function run\_with\_threads the dataframe is splitted in sub dataframes (the number is specified by the number of threads) and then distance is calculated for each subdataframe.
+After thread join, the result is also stored in the nearest\_neighbours list.
 
-	In this function, the 3rd column of the dataframe is providing and it is transformed in tuples with the latitudes and longitudes.
+	In the function calculate\_dists, the 3rd column of the dataframe is provided and it is transformed in tuples with the latitudes and longitudes.
 The list nearest neighbours will keep in tuples for each train trip the trip\_id, the calculated distance (which is calculated using the dynamic time wraping algorithm), the journeyId and finally all the points of the journey.
+
 	\todo{custom algorithm or library}
@@
 The function for the visualization is the same that was used for the question 2a1 in the utils\_pandas.py file.
+
+	Below, there is an example of the 6 produced maps, showing the nearest sub-routes related to a test trip:
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.50] {question2a1example.jpg}
+			\caption{Example of nearest sub-routes trips visualization}
+		\end{center}
+	\end{figure}
 

### 2018-01-06 Main.py

@@ -105,7 +105,7 @@ if __name__ == '__main__':
     ############
 
     # prepare files
-    test_files = [ os.path.join(input_folder, "test_set_a%d.csv" % t) for t in [1,2]]
+    test_files = [os.path.join(input_folder, "test_set_a%d.csv" % t) for t in [1,2]]
 
     # run
     question_2(output_file_clean, test_files, output_folder, maps_folder, class_folder, paropts, pandas_option=False)

### 2018-01-06 main_pandas.py

@@ -18,7 +18,7 @@ def question_1(input_file, output_file, output_file_clean, maps_folder):
 def question_2(train_file, test_files, test_file, output_folder, maps_folder, class_folder, paropts):
     # Question 2
     print(">>> Running question 2a1 - Nearest neighbours computation")
-    qp2.question_a1(maps_folder, train_file, test_files[0], paropts)
+    # qp2.question_a1(maps_folder, train_file, test_files[0], paropts)
     print(">>> Running question 2a2 - Nearest subroutes computation")
     qp2.question_a2(maps_folder, test_files[1], train_file)
     print(">>> Running question 2b - Cell grid quantization")
@@ -30,7 +30,7 @@ def question_2(train_file, test_files, test_file, output_folder, maps_folder, cl
 
 
 def check_dependencies():
-    deps = ['numpy', 'scipy']
+    deps = ['numpy', 'scipy', 'pandas']
     for dep in deps:
         res = pkgutil.find_loader(dep)
         if res is None:

### 2018-01-06 question2/NearestSubroutsPandas.py

@@ -3,6 +3,7 @@ from multiprocessing.pool import ThreadPool
 import question1_pandas as qp1
 
 
+#TODO include only one time a trip
 def find_similar_subroutes_per_test_trip(test_points, train_df, k=5):
     test_lonlat = up.idx_to_lonlat(test_points, format="tuples")
     max_subseqs = []

### 2018-01-07 Report/BigDataHomework.tex
clean files, update report

--- 2018-01-06 Report/BigDataHomeworkPandas.tex
@@
 
-    In the Main.py file, it is required to provide the path to the input folder (containing the training and test files) and the output folder where all results will be stored.
+    In the main.py file, it is required to provide the path to the input folder (containing the training and test files) and the output folder where all results will be stored.
 In this folder an additional directories are created, in order to keep all maps extracted from this application and also to store the charts showing the classification accuracy.
@@
 	In this section, further details are provided related to the preprocessing and cleaning of the training data as well as to the trip's visualization.
-All files implementing the requirements of this part are in the file question1\_pandas.py.
+All files implementing the requirements of this part are in the file question1.py.
 
 	\subsection{Question A}
-	The requirement in this part is to preprocess the input training data and create a new file where will be in 3 columns.
-The first column is the trip\_id incrementing by 1 for each trip, the second is the journey\_id and the last one contains a list of lists, where each item contains a timestamp, a longitude and a latitude.
-In order to read the given csv file, pandas library was used transforming the csv file to a dataframe.
-The first necessary thing to do, was to remove all data with null journeyPatternId and then using the groupby method on the columns vehicleID and journeyPatternId, three different dataframes were created.
+	The requirement in this part is to preprocess the input training data and create a new file with 3 columns, containingthe trip\_id (a number incrementing by 1 for each trip), the journey\_id(excluding trips with null id) and all points and timestamps in the format: [[timestamp1, longitude1, latitude1], [timestamp2, longitude2, latitude2],...].
+The given csv file with the training data was transformed to a dataframe using the relevant pandas library.
+As already mentioned, the data with null journeyPatternId should be removed from the dataframe, therefore the lines are checked for NaN and "null" journey id.
+Then, the columns vehicleID and journeyPatternId are groupedby three times in order to create three different dataframes.
+Foe example, the first dataframe contains the columns: vehicleID, journeyPatternId and timpestamp.
+In a similar way, the two remaining dataframes were formatted.
 By concatenating them, the final dataframe contains five columns, where the three last columns contain a list of timestamps, a list of longitudes and a list of latitudes respectively.
@@
 Then, the column vehicleID is replaced by the tripId column which is the index of this dataframe, starting from 0 and incrementing by 1.
-Finally, in the column points, which now contains the list of lists with the timestamps and the poins, is sorted based on the timestamp of each list item.
-The final result is exported in a file, keeping the requested format.
+Finally, the column points, which now contains the list of lists with the timestamps and the points, is sorted based on the timestamp of each list item.
+The final result is exported in a file with the requested format.
 
 	\subsection{Question B}
-	The second goal was to clean the given data, based on the total trip distance and the max distance between two successive points.
-Total distance should not be less than 2km and max distance should not exceed 2km.
+	The second goal was to clean the given data, based on two criteria: total trip distance should not be less than 2km while max distance between two successive points should not exceed 2km.
 
-	The above logic is implemented in function filter\_trips\_pandas.
+	The above logic is implemented in function filter\_trips.
 The two lists (trips\_too\_small and trips\_too\_big) will keep the excluded trips due to either not valid total distance or to invalid max distance between successive points.
@@
 
-	Finally, the distances are checked and rows with invalid trips are removed from the dataframe, which is afterwards exported in a new file (trips\_clean.csv).
+	Finally, the distances are checked and rows with invalid trips are removed from the dataframe, which is, afterwards, exported in a new file (trips\_clean.csv).
 
@@
 For this purporse, the trips list is shuffled in order to visualize random trips and then iterated only for 5 trips.
-Afterwards, a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).
+Then, a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).
 
-	Finally, the function write\_group\_gml from the utils\_pandas.py file is called.
+	Finally, the function write\_group\_gml from the utils.py file is called.
 In this function, it is calculated the mean for both longitudes and latitudes and then gmplot is used so as to visualize the selected trips.
@@
 	\section{Exercise 2}
-	The files implementing the requirements of this part are in the file question2\_pandas.py and some functions implementing the logic for this question are in the files in the folder question 2.
+	The files implementing the requirements of this part are in the file question2.py and some functions implementing the logic for this question are in the files: NearestNeighbours.py, NearestSubroutes.py, MapInGridView, JourneyClassification.py.
 
@@
 For each of these trips, the k nearest neighbours are targeted (in this exercise k=5) and an image file is produced, showing all six trips (the test trip and the trips from the training file) on six different maps.
-Under each map, the neighbour index, the journey\_id, the dynamic time warping and the time needed to process the neighbours of this test trip in millisec.
+Under each map, the test or neighbour index, the journey\_id, the dynamic time warping and the time needed to process the neighbours of this test trip in millisec.
 
 	In the function question\_a1, two dataframes are created for both the test and the training trips.
-The test dataframe's rows are iterated and for each row the function calculate\_nns from the file NearestNeighboursPandas.py is called in order to get the nearest neighbours, calculating also the time needed for this action.
+The test dataframe's rows are iterated and for each row the function calculate\_nns from the file NearestNeighbours.py is called in order to get the nearest neighbours, calculating also the time needed for this action.
 
@@
 This information is provided as input giving the kind of execution and the number of processes/threads as well.
-In the function run\_with\_processes an async task is executed for each process and then the result is stored in a list, while in the function run\_with\_threads the dataframe is splitted in sub dataframes (the number is specified by the number of threads) and then distance is calculated for each subdataframe.
+In the function run\_with\_processes an async task is executed for each process and then the result is stored in a list, while in the function run\_with\_threads, the dataframe is splitted in sub dataframes (the number is specified by the number of threads) and then distance is calculated for each subdataframe.
 After thread join, the result is also stored in the nearest\_neighbours list.
@@
 	The final list is sorted and only the 5 nearest neighbours are kept for visualization.
-preprocessing\_for\_visualization, three list are created: a list with all the points (longitudes and latitudes) of the test trip and its neighbours, a list with the labels and a list with the colors (in this question the only color is the blue).
-All this data is provided to the function visualize\_paths in the utils\_pandas.py file in order to export the trips as map images.
+However, before the visualization, the preprocessing\_for\_visualization function is used, where three list are created: a list with all the points (longitudes and latitudes) of the test trip and its neighbours, a list with the labels and a list with the colors (in this question the only color is the blue).
+All this data is provided to the function visualize\_paths in the utils.py file in order to export the trips as map images.
 
@@
 	\subsection{Question A2}
-	Following the same logic as above, in the file question2\_pandas and in the function question\_a2, two dataframes are created for the test and the training dataset using pandas library.
+	Following the same logic as above, in the function question\_a2, two dataframes are created for the test and the training dataset using pandas library.
 The test dataframe is iterated row by row, in order to find the 5 trips from the training dataset which have the longest common subsequence with the specific test trip.
@@
 We keep all similar point values and indexes in two lists while iterating the two given lists.
-The distance between two points is calculated using the method calculate\_lonlat\_distance in the question1\_pandas.py file.
+The distance between two points is calculated using the method calculate\_lonlat\_distance in the question1.py file.
 Based on the exercise requirements, two points are considered as equals if their distance is less than 200 metres.
@@
 This time the points are separated so as to show which of them belong to the nearest sub-route and color them with red color.
-The function for the visualization is the same that was used for the question 2a1 in the utils\_pandas.py file.
+The function for the visualization is the same that was used for the question 2a1 in the utils.py file.
 
@@
 
-	In order to be able to draw the grid, it was necessary to find the min and max longitudes and latitudes (function: find\_min\_max\_latlong).
+	In order to be able to draw the grid, it was necessary to find the min and max longitudes and latitudes (function: find\_min\_max\_latlong, file:MapInGridView.py).
 Having now this information, the next function called is the create\_grid giving as input the number\_of\_cells and the result of the previous method.
@@
 	\subsection{Question C}
-	In the method question\_c in question2\_pandas.py file, a new dataframe is created from the features file, containing the grid cell names.
-All functions used in this question are in the file JourneyClassificationPandas.py.
+	In the method question\_c, a new dataframe is created from the features file, containing the grid cell names.
+All functions used in this question are in the file JourneyClassification.py.
 Before starting the classification, the method preprocess\_data was used in order to get the journey\_id of all trips as numbers and also to remove the letter C from the cell names.
@@
 	In the requirements of this question it is mentioned to use 10-fold cross validation, therefore the KFold class is used providing the required folds number and then the training dataset is splitted in training and test folds.
-It worths to be mentioned here, that method asarrays from numpy was used to the necessary for this question lists.
-In addition, a list with the names of the three classifiers is created and for each of them, the relevant function is called in order to train the classifier and then count the accuracy.
+It worths to be mentioned here, that method asarrays from numpy was used to the, necessary for this question, lists.
+In addition, a list with the names of the three classifiers is created and for each of them, the relevant function is called in order to train the classifier and then count the predictions' accuracy on the test dataset.
 
@@
 Firstly, for the knn classifier the class KNeighborsClassifier was used, and more specifically the fit and predict methods.
-In order to calculate the accuracy of the classifiers we have used the accuracy\_score method (this result is return from each classification function in the JourneyClassification.py).
+In order to calculate the accuracy of the classifiers we have used the accuracy\_score method (this result is return from each classification function in the JourneyClassification.py), but also the classification\_report in order to print the precision and recall values for each prediction of the classifier.
 The same logic is followed for the other two classifiers(logistic regression and random forest).

### 2018-01-07 JourneyClassification.py

@@ -89,11 +89,8 @@ def randfor_classification(train, val, targets):
     return accTrain, accVal
 
 
-
-
 # test file in the same format as the features file
 def improve_classification(features_file, test_file, output_folder, classifier):
-    print("IXTA todo")
     return
     train_df = pd.read_csv(features_file)
     test_df = pd.read_csv(test_file)

### 2018-01-07 Main.py

@@ -1,111 +1,0 @@
-import sys, os, pkgutil
-import question1.Preprocessing as prep
-import question1.CleanData as clean
-import question1.DataVisualization as visual
-import question1_pandas as qp1
-
-import question2_pandas as qp2
-import question2.NearestNeighbours as nn
-import question2.NearestSubroutes as ns
-import question2.MapInGridView as gv
-import question2.JourneyClassification as jc
-import utils
-import random
-
-
-def question_1(input_file, output_file, output_file_clean, maps_folder, pandas_option=True):
-    # Question 1
-    if pandas_option:
-        print(">>> Running question 1a - parsing the training data")
-        trips_list, df = qp1.create_trips_file(input_file, output_file)
-        print(">>> Running question 1b - cleaning the training data")
-        trips_list, df = qp1.filter_trips_pandas(output_file_clean, df)
-        print(">>> Running question 1c - visualizing the training data")
-        qp1.visualize_trips(maps_folder, df)
-    else:
-        print(">>> Running question 1a - parsing the training data")
-        trips_list = prep.question_1a(input_file, output_file)
-        print(">>> Running question 1b - cleaning the training data")
-        trips_list = clean.question_1b(output_file_clean, trips_list)
-        print(">>> Running question 1c - visualizing the training data")
-        visual.question_1c(maps_folder, trips_list)
-    print("Finished question1")
-
-
-def question_2(train_file, test_files, output_folder, maps_folder, class_folder, paropts, pandas_option=True):
-    # Question 2
-    if pandas_option:
-        print(">>> Running question 2a1 - Nearest neighbours computation")
-        qp2.question_a1(output_folder, train_file, test_files[0], paropts)
-        print(">>> Running question 2a2 - Nearest subroutes computation")
-        qp2.question_a2(maps_folder, test_files[1], train_file)
-        print(">>> Running question 2b - Cell grid quantization")
-        cellgrid = (4, 3)
-        print("Using cell grid:", cellgrid)
-        features_file = qp2.question_b(train_file, cellgrid, output_folder)
-        print(">>> Running question 2c - Classification")
-        qp2.question_c(features_file, class_folder)
-    else:
-        # Read the training data
-        trips_list = utils.read_trips(train_file)
-        print(">>> Running question 2a1 - Nearest neighbours computation")
[…]

### 2018-01-07 MapInGridView.py

@@ -1,6 +1,6 @@
 import os
 import matplotlib.pyplot as plt
-import utils_pandas as up
+import utils as up
 
 
 def find_min_max_latlong(train_df):

### 2018-01-07 NearestNeighbours.py

@@ -1,10 +1,10 @@
 import pandas as pd
 import os
-import utils_pandas as up
+import utils as up
 from multiprocessing.pool import ThreadPool
 import threading
 from dtw import dtw as libdtw
-import question1_pandas as qp1
+import question1 as qp1
 
 
 def calculate_nns(test_points, train_df, paropts=None):

### 2018-01-07 NearestSubrouts.py

@@ -1,9 +1,8 @@
-import utils_pandas as up
+import utils as up
 from multiprocessing.pool import ThreadPool
-import question1_pandas as qp1
+import question1 as qp1
 
 
-#TODO include only one time a trip
 def find_similar_subroutes_per_test_trip(test_points, train_df, k=5):
     test_lonlat = up.idx_to_lonlat(test_points, format="tuples")
     max_subseqs = []

### 2018-01-07 main.py

@@ -1,6 +1,6 @@
 import sys, os, pkgutil
-import question1_pandas as qp1
-import question2_pandas as qp2
+import question1 as qp1
+import question2 as qp2
 import random
 
 
@@ -9,7 +9,7 @@ def question_1(input_file, output_file, output_file_clean, maps_folder):
     print(">>> Running question 1a - parsing the training data")
     trips_list, df = qp1.create_trips_file(input_file, output_file)
     print(">>> Running question 1b - cleaning the training data")
-    trips_list, df = qp1.filter_trips_pandas(output_file_clean, df)
+    trips_list, df = qp1.filter_trips(output_file_clean, df)
     print(">>> Running question 1c - visualizing the training data")
     qp1.visualize_trips(maps_folder, df)
     print("Finished question1")

### 2018-01-07 question1.py

@@ -2,7 +2,7 @@ import pandas as pd
 from math import sin, cos, sqrt, atan2, radians
 import os
 import random
-import utils_pandas as up
+import utils as up
 
 
 # Question 1a
@@ -65,17 +65,17 @@ def sort_timestamps(df):
 # Question 1b
 #############
 
-def filter_trips_pandas(output_file, df):
+def filter_trips(output_file, df):
     trips_too_small, trips_too_big = [], []
     for index, row in df.iterrows():
         jounreyId = row["journeyId"]
         points = row["points"]
-        total_dist = calculate_total_distance_per_trip_pandas(points)
+        total_dist = calculate_total_distance_per_trip(points)
         if total_dist < 2000:
             trips_too_small.append(jounreyId)
             df.drop(index, inplace=True)
             continue
-        max_dist = calculate_max_dist_pandas(points)
+        max_dist = calculate_max_dist(points)
         if max_dist > 2000:
             trips_too_big.append(jounreyId)
             df.drop(index, inplace=True)
@@ -88,7 +88,7 @@ def filter_trips_pandas(output_file, df):
     return trips_list, df
 
 
-def calculate_total_distance_per_trip_pandas(points):
+def calculate_total_distance_per_trip(points):
     total_distance = 0
     for i in range(len(points) - 1):
         dist = calculate_lonlat_distance(points[i][1:], points[i + 1][1:])
@@ -96,7 +96,7 @@ def calculate_total_distance_per_trip_pandas(points):
     return total_distance
 
 
-def calculate_max_dist_pandas(points):
+def calculate_max_dist(points):
     max_dist = 2000
     for i in range(len(points) - 1):
         dist = calculate_lonlat_distance(points[i][1:], points[i + 1][1:])

### 2018-01-07 question1/CleanData.py

@@ -1,89 +1,0 @@
-from math import sin, cos, sqrt, atan2, radians
-import utils
-
-
-def question_1b(output_file, trips_list):
-    '''
-    Runs the second subquestion
-    :param output_folder:
-    :param trips_list:
-    :return:
-    '''
-
-    print("Number of trips in the input file: %d" % len(trips_list))
-    trips_list = filter_trips(trips_list)
-    print("Number of clean trips: %d" % len(trips_list))
-    utils.write_trips(output_file, trips_list)
-    utils.serialize_trips(output_file + ".pickle", trips_list)
-    return trips_list
-
-
-def filter_trips(trips_list):
-    '''
-    Removes too long or too short trips
-    :param trips_list:
-    :return:
-    '''
-
-    trips_too_small, trips_too_big, nulls = [], [], []
-    for trip in trips_list:
-        total_dist = calculate_total_distance_per_trip(trip)
-        # too small journey check
-        if total_dist < 2:
-            trips_too_small.append(trip)
-            trips_list.remove(trip)
-            continue
-        max_dist = calculate_max_dist(trip)
-        if max_dist > 2:
-            trips_too_big.append(trip)
-            trips_list.remove(trip)
-            continue
-    print("Total trips deleted due to total distance less than 2km: %d" % len(trips_too_small))
-    print("Total trips deleted due to max distance between two points more than 2km: %d" % len(trips_too_big))
-    return trips_list
-
-
-def calculate_total_distance_per_trip(trip):
-    '''
-    Calculates the total distance of a trip
-    :param trip:
-    :return:
-    '''
-    total_distance = 0
-    lonlatlist = trip[2:]
-    for i in range(len(lonlatlist)-1):
-        dist =  calculate_lonlat_distance(lonlatlist[i][1:], lonlatlist[i+1][1:])
-        total_distance += dist
-    return total_distance
-
-
-def calculate_lonlat_distance(point1, point2):
-    """
-    Calculates the distance in meters between two points (lon, lat) using the haversine formula.
-    """
-    # to rad
-    point1 = [radians(point1[0]), radians(point1[1])]
-    point2 = [radians(point2[0]), radians(point2[1])]
-    delta_lat = point2[1] - point1[1]
-    delta_lon = point2[0] - point1[0]
-    earth_radius = 6371.0
[…]

### 2018-01-07 question1/CleanDataPandas.py

@@ -1,57 +1,0 @@
-from math import sin, cos, sqrt, atan2, radians
-
-
-def filter_trips_pandas(output_file, df):
-    trips_too_small, trips_too_big = [], []
-    for index, row in df.iterrows():
-        jounreyId = row["journeyId"]
-        points = row["points"]
-        total_dist = calculate_total_distance_per_trip_pandas(points)
-        if total_dist < 2000:
-            trips_too_small.append(jounreyId)
-            df.drop(index, inplace=True)
-            continue
-        max_dist = calculate_max_dist_pandas(points)
-        if max_dist > 2000:
-            trips_too_big.append(jounreyId)
-            df.drop(index, inplace=True)
-            continue
-    print("Total trips deleted due to total distance less than 2km: %d" % len(trips_too_small))
-    print("Total trips deleted due to max distance between two points more than 2km: %d" % len(trips_too_big))
-    df.to_csv(output_file)
-    trips_list = df.to_dict(orient='dict')
-    return trips_list, df
-
-
-def calculate_total_distance_per_trip_pandas(points):
-    total_distance = 0
-    for i in range(len(points) - 1):
-        dist = calculate_lonlat_distance(points[i][1:], points[i + 1][1:])
-        total_distance += dist
-    return total_distance
-
-
-def calculate_max_dist_pandas(points):
-    max_dist = 2000
-    for i in range(len(points) - 1):
-        dist = calculate_lonlat_distance(points[i][1:], points[i + 1][1:])
-        if (dist > max_dist):
-            max_dist = dist
-    return max_dist
-
-
-def calculate_lonlat_distance(point1, point2):
-    """
-    Calculates the distance in meters between two points (lon, lat) using the haversine formula.
-    """
-    # to rad
-    point1 = [radians(point1[0]), radians(point1[1])]
-    point2 = [radians(point2[0]), radians(point2[1])]
-    delta_lat = point2[1] - point1[1]
-    delta_lon = point2[0] - point1[0]
-    earth_radius = 6371.0
-    a = sin(delta_lat/2)**2 + cos(point1[1]) * cos(point2[1]) * sin(delta_lon/2)**2
-    c = 2 * atan2(sqrt(a), sqrt(1-a))
-    d = earth_radius * c
-    # return to meters
-    return d * 1000

### 2018-01-07 question1/DataVisualization.py

@@ -1,26 +1,0 @@
-import random
-import os
-import utils
-
-
-def question_1c(output_folder, trips_list):
-    '''
-    Runs the 3rd subquestion. Visualize 5 random different journeys.
-    :param output_folder:
-    :param trips_list:
-    :return:
-    '''
-    output_file_base = os.path.join(output_folder, "mapplot")
-    num_visualize = 5
-    # Apply some restrictions to visualized candidates
-    print("Will restrict visualization to a) non-null journey ids and b) having at least 2 points")
-    trips_list = [t for t in trips_list if t[1] != "null" and len(t[2:]) > 1]
-    # shuffle them up
-    random.shuffle(trips_list)
-    for i in range(num_visualize):
-        trip = trips_list[i]
-        file_name = output_file_base + str(i) + ".html"
-        # get point coordinates lon-lat
-        points_lonlat = [utils.idx_to_lonlat(trip)]
-        # produce output htmls
-        utils.write_group_gml(points_lonlat, file_name)

### 2018-01-07 question1/DataVisualizationPandas.py

@@ -1,19 +1,0 @@
-import os
-import random
-import utils_pandas as up
-
-
-def visualize_trips(output_folder, df):
-    output_file_base = os.path.join(output_folder, "mapplot")
-    num_visualize = 5
-    total_points = up.get_total_points(df)
-    random.shuffle(total_points)
-    for i in range(num_visualize):
-        file_name = output_file_base + str(i) + ".html"
-        # get point coordinates lon-lat
-        points_lonlat = [up.idx_to_lonlat(total_points[i])]
-        # produce output htmls
-        up.write_group_gml(points_lonlat, file_name)
-
-
-

### 2018-01-07 question1/PandasPreprocessing.py

@@ -1,53 +1,0 @@
-import pandas as pd
-
-
-def create_trips_file(input_file, output_file):
-    print("Reading training file with pandas ignoring null jids", input_file)
-    df = pd.read_csv(input_file)
-    df = df[pd.notnull(df['journeyPatternId'])].reindex()
-    timestamps = df.groupby(["vehicleID", "journeyPatternId"])["timestamp"].apply(list)
-    lons = df.groupby(["vehicleID", "journeyPatternId"])["longitude"].apply(list)
-    lats = df.groupby(["vehicleID", "journeyPatternId"])["latitude"].apply(list)
-    df = pd.concat([timestamps, lons, lats], axis=1, ignore_index=False)
-    print("Processing data.")
-
-    for index, row in df.iterrows():
-        tslist, lonslist, latslist = [], [], []
-        tslist = row["timestamp"]
-        lonslist = row["longitude"]
-        latslist = row["latitude"]
-        data_list = []
-        for i,timestamp in enumerate(tslist):
-            new_list = []
-            new_list.append(timestamp)
-            new_list.append(lonslist[i])
-            new_list.append(latslist[i])
-            data_list.append(new_list)
-        row["timestamp"] = data_list
-    new_df = create_new_dataframe()
-    new_df = pd.concat([new_df, df["timestamp"]], axis=1, ignore_index=False)
-    new_df.to_csv(output_file)
-    df = pd.read_csv(output_file)
-    df = df.drop(["vehicleID", "tripId"], axis=1)
-    df.index.name = "tripId"
-    df.columns = ["journeyId", "points"]
-    df = sort_timestamps(df)
-    df.to_csv(output_file)
-    trips_list = df.to_dict(orient='dict')
-    return trips_list, df
-
-
-def create_new_dataframe():
-    columns = []
-    columns.append("tripId")
-    new_df = pd.DataFrame(columns=columns)
-    return new_df
-
-
-def sort_timestamps(df):
-    for index, row in df.iterrows():
-        points = row["points"]
-        points = eval(points)
-        points.sort(key=lambda x: x[0])
-        df.at[index, "points"] = points
-    return df

### 2018-01-07 question1/Preprocessing.py

@@ -1,91 +1,0 @@
-import utils
-
-
-def question_1a(input_file, output_file):
-    '''
-    Runs the first subquestion
-    :param input_file: The training file
-    :param output_file: The processed file with the trips
-    :param read_option: file will be read using pandas or not
-    :return:
-    '''
-
-    line_objects = utils.read_train_set(input_file)
-    map_per_vid = get_data_per_vehicle_id(line_objects)
-    map_per_vidjid = get_data_per_vehicle_and_journey_id(map_per_vid)
-    trips_list = write_trips(output_file, map_per_vidjid)
-    print("Done with question 1!")
-    return trips_list
-
-
-def get_data_per_vehicle_id(line_objects):
-    '''
-    Aggregates data wrt the vehicle ID
-    :param line_objects:
-    :return:
-    '''
-    print("Aggregating per vehicle id")
-    map_per_vid = {}
-    for obj in line_objects:
-        vid = obj["vid"]
-        if vid not in map_per_vid:
-            map_per_vid[vid] = []
-        obj_without_vid = {key: obj[key] for key in obj if key != "vid"}
-        map_per_vid[vid].append(obj_without_vid)
-    return map_per_vid
-
-
-def get_data_per_vehicle_and_journey_id(map_per_vid):
-    '''
-    Aggregates data wrt the journey ID
-    :param map_per_vid:
-    :return: vid1:{jid1:[pt1,pt2...], jid2:[pt1,pt2],...},vid2:{...}
-    '''
-    print("Aggregating per journey id")
-    for vehicle_id in map_per_vid:
-        # get elements for this vehicle id
-        map_per_jid = {}
-        vid_list = map_per_vid[vehicle_id]
-        # make a map wrt to the journey id
-        for item in vid_list:
-            jid = item["jid"]
-            if jid not in map_per_jid:
-                map_per_jid[jid] = []
-            # assign the points of each journeyid, minus that id
-            obj_without_jid = {key: item[key] for key in item if key != "jid"}
-            map_per_jid[jid].append(obj_without_jid)
-        # for each jid for this vehicle
-        for jid in map_per_jid:
-            # get poitns
-            jid_list = map_per_jid[jid]
-            # sort per timestamp
-            sorted_jid_list = sorted(jid_list, key=lambda k: k['ts'])
-            map_per_jid[jid] = sorted_jid_list
-        # assign jid-points map to this vehicle id
-        map_per_vid[vehicle_id] = map_per_jid
-    return map_per_vid
-
-
[…]

### 2018-01-07 question2.py

@@ -1,11 +1,11 @@
-import question2.NearestNeighboursPandas as nn
-import question2.NearestSubroutsPandas as ns
-import question2.MapInGridViewPandas as gvp
-import question2.JourneyClassificationPandas as jcp
+import NearestNeighbours as nn
+import NearestSubrouts as ns
+import MapInGridView as gvp
+import JourneyClassification as jcp
 from sklearn.model_selection import KFold
 import pandas as pd
 import numpy as np
-import utils_pandas as up
+import utils as up
 import os
 
 

### 2018-01-07 question2/JourneyClassification.py

@@ -1,120 +1,0 @@
-import pickle, os
-import numpy as np
-from sklearn.model_selection import KFold
-from sklearn.neighbors import KNeighborsClassifier
-from sklearn.linear_model import LogisticRegression
-from sklearn.ensemble import RandomForestClassifier
-from sklearn.metrics import accuracy_score
-from sklearn.metrics import classification_report
-import utils
-
-
-def preprocess_data(data_orig):
-    '''
-    Prepare data for classification
-    :param data:
-    :return:
-    '''
-    # get journey ids
-    targets = [d['jid'] for d in data_orig]
-    # get features
-    data = [d['points'] for d in data_orig]
-
-    # get maximum length of feature lists
-    maxlen = len(max(data, key=lambda x : len(x)))
-    # convert string to numerics
-    data = [[int(d[1:]) for d in dlist] for dlist in data]
-
-    # pad to the maximum length
-    for i, datum in enumerate(data):
-        if len(datum) < maxlen:
-            data[i] = datum + [3 for _ in range(maxlen - len(datum))]
-
-    # convert journey ids to numbers
-    num_ids = {}
-    targets_nums = []
-    for t in targets:
-        if t not in num_ids:
-            num_ids[t] = len(num_ids)
-        targets_nums.append(num_ids[t])
-    # count occurences
-    hist = []
-    for jid in num_ids:
-        occurences = sum([1 if t == jid else 0 for t in targets])
-        hist.append((jid, occurences))
-
-    sorted(hist, key = lambda x : x[1])
-    print("Most frequent 5 jids:")
-    for (jid, occ) in hist[:5]:
-        print(jid,":",occ)
-    return data, targets_nums
-
-
-def question_c(file, output_folder):
-    # get the data ready
-    trips_list = get_grid_trips(file)
-    kf = KFold(n_splits=10)
-    grid_features, targets = preprocess_data(trips_list)
-
-    folds_idxs = list(kf.split(grid_features))
-    trips_array = np.asarray(grid_features)
-    targets = np.asarray(targets)
-
-    classifiers = ["knn", "logreg", "randfor"]
-    accuracies = {}
-    # classify
-    for classifier in classifiers:
-        print("Testing classifier [%s]" % classifier)
-        # train & test each classifier
-        # for each fold
-        accuracies[classifier] = []
-        for i, (train_idx, val_idx) in enumerate(folds_idxs):
-            print("\tClassifing fold %d/%d" % (i+1, len(folds_idxs)))
[…]

### 2018-01-07 question2/MapInGridView.py

@@ -1,118 +1,0 @@
-import matplotlib.pyplot as plt
-import utils
-import os
-
-
-def subquestion_b(trips_list, number_of_cells, output_folder):
-    # specify files
-    output_file = os.path.join(output_folder, "tripFeatures.csv")
-    pickle_file = os.path.join(output_folder, "tripFeatures.pickle")
-    # use a given data collection to judge min/max data points
-    max_lat, max_lon, min_lat, min_lon = find_min_max_latlong(trips_list)
-    rows,columns,cell_names = create_grid(number_of_cells,max_lat,max_lon,min_lat,min_lon, output_folder=output_folder)
-    replace_points(trips_list, rows, columns, cell_names, output_file, pickle_file)
-    return pickle_file
-
-
-def find_min_max_latlong(trips_list):
-    max_lat, max_lon = -1000, -1000
-    min_lat, min_lon = 1000, 1000
-    for trip in trips_list:
-        max_lon = max([t[0] for t in trip['points']] + [max_lon])
-        min_lon = min([t[0] for t in trip['points']] + [min_lon])
-        max_lat = max([t[1] for t in trip['points']] + [max_lat])
-        min_lat = min([t[1] for t in trip['points']] + [min_lat])
-    return max_lat, max_lon, min_lat, min_lon
-
-
-def create_grid(number_of_cells, max_lat,max_lon,min_lat,min_lon, output_folder):
-    cell_lat_dist, cell_lon_dist = get_distance_per_cell(number_of_cells, min_lat, min_lon, max_lat, max_lon)
-    rows = create_grid_lines(number_of_cells[0], min_lat, cell_lat_dist)
-    columns=create_grid_lines(number_of_cells[1], min_lon, cell_lon_dist)
-    cell_names = create_cell_names(number_of_cells)
-
-    visualize_grid(rows,columns,min_lat,min_lon,max_lat,max_lon, output_folder=output_folder)
-    return rows,columns,cell_names
-
-
-def get_distance_per_cell(number_of_cells, min_lat, min_lon, max_lat, max_lon):
-    total_lat_dist = max_lat - min_lat
-    total_lon_dist = max_lon - min_lon
-    cell_lat_dist = total_lat_dist / number_of_cells[0]
-    cell_lon_dist = total_lon_dist / number_of_cells[1]
-    return cell_lat_dist, cell_lon_dist
-
-
-def create_grid_lines(number_of_cells, min_point, dist):
-    lines_list = []
-    new_min = min_point
[…]

### 2018-01-07 question2/NearestNeighbours.py

@@ -1,228 +1,0 @@
-import question1.CleanData as clean
-import os
-import utils
-from multiprocessing.pool import Pool, ThreadPool
-import threading
-from dtw import dtw as libdtw
-
-
-def question_a1(output_folder, test_file, trips_list, paropts):
-    print("WARNING: Reading training file part because test file is not supplied.")
-    test_trip_list = trips_list[:5]
-    for i,test_trip in enumerate(test_trip_list):
-        print("Examining test element %d/%d" % (i+1, len(test_trip_list)))
-        outfile_name = os.path.join(output_folder, "nn_%d_" % (i+1))
-        # prepare to count time
-        millis_start = utils.tic()
-        # compute nearest neighbours
-        # parallelize = None
-        nns_ids_distances = calculate_nns(test_trip, trips_list, paropts = paropts)
-        # get time elapsed
-        elapsed = utils.tictoc(millis_start)
-        # get the whole data of the returned trips, add distances
-        neighbour_trips = [n for n in trips_list if n['id'] in [t[0] for t in nns_ids_distances]]
-        # visualize
-        preprocessing_for_visualization(test_trip, nns_ids_distances, neighbour_trips, outfile_name, elapsed, i)
-
-
-def calculate_nns(test_trip, trips_list, paropts = None, ret_container = None):
-
-    '''
-    :param test_trip: a trip row
-    :param trips_list: a list of trip rows to find NNs of test_trip
-    :return: returns a list of the 5 nearest neighbours from trips_list, for the test_trip. Format is (trip_id, distance)
-    '''
-
-    # parallelization type
-    if paropts:
-        print("Parallelizing with",paropts)
-        partype, numpar = paropts
-    else:
-        partype, numpar = None, None
-
-    tic = utils.tic()
-    # get test coordinate list
-    test_lonlat = test_trip["points"]
-    # initialize res
-    nearest_neighbours = [-1 for _ in trips_list]
-
-    if partype:
-        # num threads or processes
-        if partype == "processes":
-            # K elements at a time - only works for K=1. else, computation is done multiple times
-            K = 1
-            trips = utils.sublist(trips_list, K)
-            # results here
-            rres = [[] for _ in trips]
-            pool = ThreadPool(processes=numpar)
-            tasks = [[] for _ in trips]
-            for i,tlist in enumerate(trips):
-                # print(len(tlist),tlist)
[…]

### 2018-01-07 question2/NearestSubroutes.py

@@ -1,164 +1,0 @@
-import question1.CleanData as clean
-import os
-import utils
-from multiprocessing.pool import ThreadPool
-
-
-def question_a2(output_folder, test_file, trips_list):
-    print("WARNING: Reading training file part because test file is not supplied.")
-    test_list = trips_list[:5]
-    k=5
-    for i,test_trip in enumerate(test_list):
-        print("Extracting subroutes for test trip %d/%d" % (i+1,len(test_list)))
-        file_name = os.path.join(output_folder, "subroutes_%d_" % (i + 1))
-        max_subseqs = find_similar_subroots_per_test_trip(test_trip, trips_list, k)
-        preprocessing_for_visualisation(test_trip, max_subseqs, trips_list, file_name)
-
-
-def find_similar_subroots_per_test_trip(test_trip, trips_list, k):
-    test_lonlat = test_trip['points']
-    # keep track of the current common subsequences
-    max_subseqs = []
-    pool = ThreadPool(processes=10)
-    # iterate over the candidates
-    for tripidx, trip in enumerate(trips_list):
-        # get coordinates
-        trip_lonlat = trip['points']
-        timestart = utils.tic()
-        # compute common subsequences between the test trip and the current candidate
-        async_result = pool.apply_async(calc_lcss, (test_lonlat, trip_lonlat))
-        subseqs, subseqs_idx = async_result.get()
-        # subseqs, subseqs_idx = calc_lcss(test_lonlat, trip_lonlat)
-        elapsed = utils.tictoc(timestart)
-        # sort by decr. length
-        subseqs_idx = sorted(subseqs_idx, key = lambda x : len(x), reverse=True)
-        # update the list of the longest subsequences
-        max_subseqs = update_current_maxsubseq(max_subseqs, subseqs_idx, k, elapsed, tripidx)
-    print("Got %d common subsequences" % len(max_subseqs))
-    pool.close()
-    pool.join()
-    if len(max_subseqs) != k:
-        print("WARNING: Specified %d subseqs!" % k)
-    return max_subseqs
-
-
-def calc_lcss(t1, t2):
-    print("Have to check the lcss")
-    '''
-
-    :param t1: list of lonlat coordinate tuples
-    :param t2: same
-    :return:
-    '''
-    L = [ [0 for _ in t2] for _ in t1]
-    # store the sequence of similar point values and indexes
-    seqs, idxs = [], []
-    z = 0
-    for i, p1 in enumerate(t1):
[…]

### 2018-01-07 utils.py

@@ -5,8 +5,7 @@ import matplotlib.pyplot as plt
 import pylab
 from scipy.misc import imread
 import datetime
-import pickle
-import json
+import pandas as pd
 
 
 def tic():
@@ -41,138 +40,56 @@ def tictoc(previous_tic):
     return msg
 
 
-def write_trips(output_file, trips_list):
-    '''
-    Self explanatory
-    :param output_file:
-    :param trips_list:
-    :return:
-    '''
-    # write list primitives
-    if type(trips_list[0]) == list:
-        with open(output_file, "w") as f:
-            for trip in trips_list:
-                # make csv
-                header = [str(t) for t in trip[:2]]
-                coords = [str(t) for coord_tuple in trip[2:] for t in coord_tuple]
-                trip_str = ",".join(header + coords)
-                # write it
-                f.write(trip_str+"\n")
-    # write list of objects
-    if type(trips_list[0]) == dict:
-        # use json
-        with open(output_file, "w") as f:
-            json.dump(trips_list, f)
-
-def sublist(llist, slen):
-    '''
-    return llist divided to sublists of at most slen
-    :param llist:
-    :param sublist_length:
-    :param only_num:
-    :return:
-    '''
-    divs = range(0, len(llist), slen)
-    return [ llist[i:i+slen] for i in divs]
-
-def serialize_trips(output_file, trips_list):
-    '''
-    Serialize to python pickle
-    :param output_file:
-    :param trips_list:
-    :return:
-    '''
-    with open(output_file, "wb") as f:
-        pickle.dump(trips_list, f)
-
-def read_trips(filepath):
-    '''
-    Read a trips file, csv or pickle
-    :param filepath:
-    :return:
-    '''
-    if filepath.endswith(".pickle"):
-        with open(filepath, "rb") as f:
-            trips = pickle.load(f)
-    else:
-        # read the lines
-        trips=[]
-        with open(filepath, "r") as f:
-            for line in f:
-                line = line.strip()
-                if not line:
-                    continue
-                elems = line.split(",")
-                obj={}
-                obj["id"] = elems[0]
-                obj["jid"] = elems[1]
-                elems = elems[2:]
-                timestamps, points = [], []
-                for i in range(0, len(elems), 3):
[…]

### 2018-01-07 utils_pandas.py

@@ -1,231 +1,0 @@
-import gmplot
-import subprocess
-import os
-import matplotlib.pyplot as plt
-import pylab
-from scipy.misc import imread
-import datetime
-import pandas as pd
-
-
-def tic():
-    '''
-    Get timestamp
-    :return:
-    '''
-    return datetime.datetime.now()
-
-
-def tictoc(previous_tic):
-    '''
-    Get duration from previous timestamp
-    :param previous_tic:
-    :return:
-    '''
-    # get duration in usec
-    diff = tic() - previous_tic
-    # get durations of datetime.now() members
-    days = diff.days
-    secs = diff.seconds
-    usec = diff.microseconds
-    # aggregate to larger time units
-    msec, usec = divmod(usec, 1000)
-    sec_, msec = divmod(msec, 1000)
-    secs += sec_
-    mins, sec = divmod(secs, 60)
-    hrs, mins = divmod(mins,60)
-    msg = "%02d:%02d:%02d:%03d" % (hrs, mins, secs, msec)
-    if days:
-        msg = "%d days! " % days + msg
-    return msg
-
-
-def idx_to_lonlat(points, idx=None, format="tuple"):
-    if idx is None:
-        idx = list(range(len(points)))
-    if type(idx) != list:
-        idx = [idx]
-    lats, lons = [], []
-    for i in idx:
-        lons.append(points[i][1])
-        lats.append(points[i][2])
-    if format == "tuple":
-        return lons, lats
-    elif format == "tuples":
-        return list(zip(lons,lats))
-
-
-def get_sub_dataframes(df, num):
-    print("number of tasks: ", str(num))
-    num_data, rem = divmod(len(df.index), num)
-    sudframes = []
-    start = 0
-    end = num_data
-    while end <= len(df.index):
-        new_df = df[start:end]
-        sudframes.append(new_df)
-        start = end
-        end = end + num_data
-    if rem:
-        sudframes = sudframes[:num]
-        sudframes[-1] += df[-rem:]
-    return sudframes
-
-
-def get_total_points(df):
-    total_points = []
-    for index, row in df.iterrows():
-        points = row["points"]
-        total_points.append(points)
-    return total_points
-
-
-def get_lonlat_tuple(points):
-    '''
-    :param points: [(lon1,lat1), (lon2,lat2),...]
-    :return: ([lon1,lon2,...],[lat1,lat2,...])
-    '''
-    return [l[1] for l in points], [l[2] for l in points]
-
-
[…]

### 2018-01-07 Report/BigDataHomework.tex
update report

@@
 
-	\todo{custom algorithm or library}
+	The algorithm dynamic time wraping is implemented with two ways based on the input (diy or diy\_initial). \todo{diy and diyinitial}
 
@@
 Therefore, if the points are equal, and the indexes are the first point of one of the given lists the cell is marked to unit cost and a new similar sequence is started.
-Otherwise, we continue to an existing similar sequence, increasing by 1 its length.
-\todo{explain z variable} On the other hand, if points are not equal, then a new subsequence should start, and therefore L[i][j] is set to zero.
+Otherwise, we continue to an existing similar sequence, increasing by 1 its length, keeping the current cost in the z variable.
+On the other hand, if points are not equal, then a new subsequence should start, and therefore L[i][j] is set to zero.
 

### 2018-01-07 question2.py

@@ -1,5 +1,5 @@
 import NearestNeighbours as nn
-import NearestSubrouts as ns
+import NearestSubroutes as ns
 import MapInGridView as gvp
 import JourneyClassification as jcp
 from sklearn.model_selection import KFold

### 2018-01-13 ProjectReport/BigDataHomework.tex
reorganize folders

--- 2018-01-07 Report/BigDataHomework.tex
@@
 
-	\subsection{Question A}
+	\subsection{Data Preprocessing (question A)}
 	The requirement in this part is to preprocess the input training data and create a new file with 3 columns, containingthe trip\_id (a number incrementing by 1 for each trip), the journey\_id(excluding trips with null id) and all points and timestamps in the format: [[timestamp1, longitude1, latitude1], [timestamp2, longitude2, latitude2],...].
@@
 
-	\subsection{Question B}
+	\subsection{Clean Data (question B)}
 	The second goal was to clean the given data, based on two criteria: total trip distance should not be less than 2km while max distance between two successive points should not exceed 2km.
@@
 
-	\subsection{Question C}
+	\subsection{Trip Visualization (question C)}
 	This task is implemented in the function visualize\_trips and the created output is stored in the output folder in the gmplots directory.
@@
 
-	\subsection{Question A1}
+	\subsection{Nearest Neighbours (question A1)}
 	For this part a test file is given (test\_set\_a1.csv, included in the input folder) containing additional trips.
@@
 
-	\subsection{Question A2}
+	\subsection{Nearest Sub-routes (question A2)}
 	Following the same logic as above, in the function question\_a2, two dataframes are created for the test and the training dataset using pandas library.
@@
 		\begin{center}
-			\includegraphics [scale = 0.50] {question2a1example.jpg}
+			\includegraphics [scale = 0.50] {question2a2example.jpg}
 			\caption{Example of nearest sub-routes trips visualization}
@@
 
-	\subsection{Question B}
+	\subsection{Features extraction (question B)}
 	The purpose of this task was to export features for classification for the training dataset.
@@
 
-	\subsection{Question C}
+	\subsection{Classification (question C)}
 	In the method question\_c, a new dataframe is created from the features file, containing the grid cell names.

### 2018-01-13 Paper/Presentation/mll_review.tex
Update presentation gitignore

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}
\setbeamercovered{transparent}
\setbeamercolor{local structure}{fg=black}
\setbeamertemplate{caption}{\raggedright\insertcaption\par}
\usefonttheme[onlymath]{serif}

\mode<presentation> {

	% The Beamer class comes with a number of default slide themes
	% which change the colors and layouts of slides. Below this is a list
	% of all the themes, uncomment each in turn to see what they look like.

	%\usetheme{default}
	%\usetheme{AnnArbor}
	%\usetheme{Antibes}
	%\usetheme{Bergen}
	%\usetheme{Berkeley}
	%\usetheme{Berlin}
	%\usetheme{Boadilla}
	\usetheme{CambridgeUS}
	%\usetheme{Copenhagen}
	%\usetheme{Darmstadt}
	%\usetheme{Dresden}
	%\usetheme{Frankfurt}
	%\usetheme{Goettingen}
	%\usetheme{Hannover}
	%\usetheme{Ilmenau}
	%\usetheme{JuanLesPins}
	%\usetheme{Luebeck}
	%\usetheme{Madrid}
	%\usetheme{Malmoe}
	%\usetheme{Marburg}
	%\usetheme{Montpellier}
	%\usetheme{PaloAlto}
	%\usetheme{Pittsburgh}
	%\usetheme{Rochester}
	%\usetheme{Singapore}
	%\usetheme{Szeged}
	%\usetheme{Warsaw}

	% As well as themes, the Beamer class has a number of color themes
	% for any slide theme. Uncomment each of these in turn to see how it
	% changes the colors of your current slide theme.

	%\usecolortheme{albatross}
	%\usecolortheme{beaver}
	%\usecolortheme{beetle}
	%\usecolortheme{crane}
	%\usecolortheme{dolphin}
	%\usecolortheme{dove}
	%\usecolortheme{fly}
	%\usecolortheme{lily}
	\usecolortheme{orchid}
	%\usecolortheme{rose}
	%\usecolortheme{seagull}
	%\usecolortheme{seahorse}
	%\usecolortheme{whale}
	%\usecolortheme{wolverine}

	%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
	%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

	%\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables
\usepackage{caption}
\captionsetup{font=scriptsize,labelfont=scriptsize}
\usepackage{amssymb}
\usepackage{bm}
\usepackage[colorinlistoftodos,prependcaption,textsize=small]{todonotes}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Large-Scale Data Analysis Techniques]{A Review on Multi-Label Learning Algorithms} % The short title appears at the bottom of every slide, the full title is only on the title page

\author[Sissy Themeli, Nikiforos Pittaras]{Min-Ling Zhang and Zhi-Hua Zhou} % Your name
\institute[DI-UOA] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
	IEEE TRANSACTIONS ON KNOWLEDGE AND DATA ENGINEERING\\ % Your institution for the title page
	\medskip
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

	\begin{frame}
	\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}
\frametitle{Overview} % Table of contents slide, comment this block out to remove it
\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
%\setbeamercolor{section in toc}{fg=black}
%\setbeamercolor{subsection in toc}{fg=black}
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
\section{Introduction} % Sections can be created in order to organize your presentation into discrete blocks, all sections and subsections are automatically printed in the table of contents as an overview of the talk
\subsection{Problem Definition}
\begin{frame}
\frametitle{\insertsection : \insertsubsection}
Single-label supervised learning
\begin{itemize}
	\item Applied to classification
	\item Supervised: Given dataset $\{ (x_i, y_i)\}, i = 1, \dots N, x \in X, y \in Y$
	\item Goal is to learn a model $f: X \rightarrow Y$
\end{itemize}

Multi-label supervised learning
\begin{itemize}
	\item Multiple labels per instance: $\{ (x_i, \bm{y}_i)\}, i = 1, \dots N, x \in X, \bm{y} \in \mathbb{P}(Y)$
	\item Learn a model $f: X \times Y \rightarrow r$ where $r\in \mathbb{R}$ is the confidence that $y_i$ characterizes $x_i$
	\item For classification, assume that $x_i$ belongs to $y_i$ if $r$ is greater than some threshold $t(x_i)$

		\begin{itemize}
			\item $t(\cdot)$ can be a predetermined constant function or learned from $X$
		\end{itemize}

\end{itemize}
TODO: Include stuff about multi-labeled dataset characterization?
\end{frame}

%------------------------------------------------
\subsection{Algorithm Strategies}
\begin{frame}
\frametitle{\insertsection : \insertsubsection}
Label search space $\mathbb{S_Y}$ grows exponentially as a function of $|Y|$: 
\begin{itemize}
	\item e.g. for $|Y|=20, |\mathbb{S_Y}| = 2 ^ {|\mathbb{P(Y)}|} = 2^{20} \ge 10^6$
\end{itemize}
Solution: integrate in the learning process potential label correlations.
This work examines algorithms grouped in three broad categories:

\begin{enumerate}
	\item First-order strategies
		\begin{itemize}
			\item Transform multi-labeled problem to multiple, single-label problems 
			\item Ignore label correlations 
			\item Simple, scalable, suboptimal
		\end{itemize}
	\item Second-order strategies
		\begin{itemize}
			\item Consider \emph{pairwise} label relations
			\item Good trade-off between generalization performance and scalability
			\item Lacking in some real-world applications
		\end{itemize} 
	\item Higher-order strategies
		\begin{itemize}
			\item Capture more complicated label interdependencies
			\item Strong modeling capabilities
			\item Computationally demanding, less scalable
		\end{itemize}

\end{enumerate}
\end{frame}

%------------------------------------------------
\subsection{Evaluation Metrics}
\begin{frame}
\frametitle{\insertsection : \insertsubsection}
Need for more complicated evaluation than traditional supervised learning(accuracy, F-measure etc). There are 2 metrics categories 
	\begin{itemize}
		\item \emph{Example-based: }Evaluates system's performance on each example and returns the mean value. 
		\begin{itemize}
			\item \emph{From classification perspective:} $Subset Accuracy$, $Hamming Loss$, $accuracy_{exam}$, $precision_{exam}$, $recall_{exam}$, $F^\beta_{exam}$
			\item \emph{From ranking perspective:} one-error, coverage, ranking loss, average precision
		\end{itemize}
		\item \emph{Label-based: }Evaluates system's performance on each class label separately and returns macro/micro averaged value across all class labels
		\begin{itemize}
			\item From classification perspective: $B_{macro}$, $B_{micro}$
			\item From ranking perspective: $AUC_{macro}$, $AUC_{micro}$
		\end{itemize}
	\end{itemize}
TODO: theoretical results?
\end{frame}
%------------------------------------------------
\section{Multi-label Learning Algorithms}
%------------------------------------------------

\subsection{Simple Categorization}

\begin{frame}
\frametitle{\insertsection : \insertsubsection}
Algorithms in two categories:
\begin{itemize}
	\item \emph{Problem Transformation Methods: }Transform the learning problem into other well-established learning scenarios (fit data to algorithm philosophy)
	\item \emph{Algorithm Adaptation Methods: }Adapt popular learning techniques to deal with multi-label data directly (fit algorithm to data philosophy)
\end{itemize}
\begin{figure}
	\begin{center}
		\includegraphics[scale = 0.25]{images/algorithmsclass.png}
		\caption{8 representative algorithms}
	\end{center}
\end{figure}
\end{frame}

%------------------------------------------------
\subsection{Problem Transformation Methods}

\begin{frame}
\frametitle{\insertsection : \insertsubsection}
\emph{Binary Relevance}
\begin{itemize}
	\item Decompose problem into q independent binary classification problems per possible label in the label space
\end{itemize}

\end{frame}
%------------------------------------------------

\begin{frame}
\frametitle{Classifier chains}
\begin{itemize}
\item <2-> 
\begin{itemize}
\item 
\item <4-> 
\end{itemize}
\end{itemize}
\begin{itemize}
\item<5-> 
\end{itemize}

\begin{itemize}
\item <7-> 
\begin{itemize}
\item <8-> 
\end{itemize}
\end{itemize}
\end{frame}
%------------------------------------------------

\begin{frame}
\frametitle{Calibrated Label Ranking}
\begin{itemize}
\item <2-> 
\begin{itemize}
\item 
\item <4-> 
\end{itemize}
\end{itemize}
\begin{itemize}
\item<5-> 
\end{itemize}

\begin{itemize}
\item <7-> 
\begin{itemize}
\item <8-> 
\end{itemize}
\end{itemize}
\end{frame}
%------------------------------------------------
\begin{frame}
\frametitle{Random k-Labelsets}
\begin{itemize}
\item <2-> 
\begin{itemize}
\item 
\item <4-> 
\end{itemize}
\end{itemize}
\begin{itemize}
\item<5-> 
\end{itemize}

\begin{itemize}
\item <7-> 
\begin{itemize}
\item <8-> 
\end{itemize}
\end{itemize}
\end{frame}
%------------------------------------------------

\subsection{Alogrithm Adaptation Methods}

\begin{frame}
\frametitle{Multi-Label k-Nearest Neighbor}
\begin{itemize}
\item <2-> 
\item <3-> 
\item <4-> 
\item<5-> 
\item <6-> 
% \item<7->
\end{itemize}
\end{frame}

%------------------------------------------------
\begin{frame}
\frametitle{Multi-label Decision Tree}

%\begin{figure}
%	\includegraphics[width = \textwidth, height = \textheight, keepaspectratio]{modelcharacters.png}
%	\caption{Character enhanced word representations}
%	\end{figure}
\end{frame}
%------------------------------------------------
\begin{frame}
\frametitle{Ranking Support Vector Machine}

%\begin{figure}
%	\includegraphics[width = \textwidth, height = \textheight, keepaspectratio]{modelcharacters.png}
%	\caption{Character enhanced word representations}
%	\end{figure}
\end{frame}
%------------------------------------------------
\begin{frame}
\frametitle{Collective Multi-Label Classifier}

%\begin{figure}
%	\includegraphics[width = \textwidth, height = \textheight, keepaspectratio]{modelcharacters.png}
%	\caption{Character enhanced word representations}
%	\end{figure}
\end{frame}
%------------------------------------------------
\section{Related Learning Settings}
%------------------------------------------------

\begin{frame}
\frametitle{Related Learning Settings}
\begin{itemize}
\item 
\item 
\item 
\item 
\end{itemize}
\end{frame}

%------------------------------------------------
\section{Conclusion}
%------------------------------------------------
\begin{frame}
\frametitle{Conclusion}
\begin{itemize}
\item 
\item 
\end{itemize}
\onslide<4->{}
\begin{itemize}
\item <5-> 
\item <6->
\item <7-> 
\end{itemize}
\end{frame}

%------------------------------------------------
\section{The end}
\begin{frame}
\Huge{\centerline{Thank you}}
\end{frame}

%------------------------------------------------

\end{document}

### 2018-01-13 Paper/Presentation/mll_review.tex
Update presentation

@@
 	\item Learn a model $f: X \times Y \rightarrow r$ where $r\in \mathbb{R}$ is the confidence that $y_i$ characterizes $x_i$
-	\item For classification, assume that $x_i$ belongs to $y_i$ if $r$ is greater than some threshold $t(x_i)$
+	\item For classification, assume that $x_i$ belongs to $y_i$ if $r \ge $ $t(x_i)$
 
@@
 \frametitle{\insertsection : \insertsubsection}
-Algorithms in two categories:
+
+Assume multi-label classification problem with $q$ possible class labels $Y=\{y_1, \dots y_q\}$
+
+Group algorithms in two categories:
 \begin{itemize}
@@
 \end{itemize}
+\end{frame}
+
+%------------------------------------------------
+
+\begin{frame}
+\frametitle{\insertsection : \insertsubsection}
 \begin{figure}
 	\begin{center}
-		\includegraphics[scale = 0.25]{images/algorithmsclass.png}
-		\caption{8 representative algorithms}
+		\includegraphics[scale = 0.75]{images/aa.png}
+	\end{center}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\frametitle{\insertsection : \insertsubsection}
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale = 0.7]{images/pt.png}
 	\end{center}
@@
 \begin{frame}
+
 \frametitle{\insertsection : \insertsubsection}
-\emph{Binary Relevance}
+\textbf{Binary Relevance}
 \begin{itemize}
-	\item Decompose problem into q independent binary classification problems per possible label in the label space
+	\item Decompose to $q$ independent binary classification problems, $h_i(x)$
+	\item Cross-train on $q$ binary training sets (one for each $y_i$)
+	\item Predict labels of unseen $x$ by evaluating each classifier $h_i(x)$
+	\item Output labels where  $h_i(x) > 0$ else $\underset{1 \geq i \geq q}{argmax} \{h_i(x))\}$ 
+\end{itemize}
+
+Pros \& cons:
+\begin{itemize}
+	\item Simple, parallelizable
+	\item Sensitive to class-imbalanced data
+	\item Ignores label correlations (first-order method)
 \end{itemize}

### 2018-01-13 Paper/Presentation/mll_review.tex
update presentation

@@
 \subsection{Problem Transformation Methods}
+%------------------------------------------------
+\begin{frame}
+\Huge{\centerline{Problem Transformation Methods}}
+\end{frame}
+%------------------------------------------------
+\begin{frame}
+\frametitle{Binary Relevance}
 
-\begin{frame}
-
-\frametitle{\insertsection : \insertsubsection}
-\textbf{Binary Relevance}
 \begin{itemize}
@@
 \begin{itemize}
-\item <2-> 
-\begin{itemize}
-\item 
-\item <4-> 
-\end{itemize}
-\end{itemize}
-\begin{itemize}
-\item<5-> 
+	\item High-order approach algorithm
+	\item Transform the problem into a chain of binary classification problems
+	\item Build subsequent binary classifiers upon the predictions of preceding ones
+	\item A permutation function is used to specify an order to the labels
+	\item Binary training set
+constructed by appending each instance with its relevance
+to those labels
+	\item Predict relevant labels for unknown instances by iteratively traversing the classifier chain
 \end{itemize}
 
+Pros \& cons:
 \begin{itemize}
-\item <7-> 
-\begin{itemize}
-\item <8-> 
-\end{itemize}
+	\item Effectiveness largely affected by the ordering
+	\item Exploitation of label correlations
+	\item Does not support parallel implementation
 \end{itemize}
@@
 \begin{itemize}
-\item <2-> 
+	\item Second-order approach algorithm
+	\item Transform the problem into a label ranking problem
+	\item Ranking implemented by pairwise comparison techniques
+	\item For $q$ labels, generate $q(q-1)/2$ binary classifiers by pairwise comparison
+	\item Construct training set per label pairs considering the example's relevance
+	\item For unknown instance, get the overall classifiers' votes
+	\item Virtual label $y_v$ as thresholding parameter
+\end{itemize}
+Pros \& cons:
 \begin{itemize}
-\item 
-\item <4-> 
-\end{itemize}
-\end{itemize}
+	\item Smooth the class-imbalance
+	\item Classifier's number grows from linear to quadratic scale
 	\begin{itemize}
-\item<5-> 
-\end{itemize}
-
-\begin{itemize}
-\item <7-> 
-\begin{itemize}
-\item <8-> 
+		\item Improvement: Exact or approximate pruning to reduce classifier's number to be queried
 	\end{itemize}
@@
 \begin{itemize}
-\item <2-> 
+	\item High-order approach algorithm
+	\item Decompose to an ensemble of multi-class classification problems
+	\item Each component targets a random subset of $Y$, inducing a multi-class classifier by Label Powerset (LP) techniques
+	\item Convert training set into a multi-class set, treating each label set as a new class
+	\item Each example is reassigned with the mapped single-label participating in multi-class classifier induction
+	\item For unknown $x$, counts the maximum and actual number of votes
+	\item Label set is relevant when the actual number of votes exceeds
+half of the maximum ones
+\end{itemize}
+Pros \& cons:
 \begin{itemize}
-\item 
-\item <4-> 
-\end{itemize}
-\end{itemize}
-\begin{itemize}
-\item<5-> 
-\end{itemize}
-
-\begin{itemize}
-\item <7-> 
-\begin{itemize}
-\item <8-> 
-\end{itemize}
+	\item Incompleteness: predicts only label sets included in the training set
+	\item Inefficiency: large $Y$ implies high training complexity.Improvement by combining LP and ensemble learning
 \end{itemize}
@@
 %------------------------------------------------
+\subsection{Alogrithm Adaptation Methods}
+%------------------------------------------------
+\begin{frame}
+\Huge{\centerline{Alogrithm Adaptation Methods}}
+\end{frame}
+%------------------------------------------------
+\begin{frame}
+\frametitle{Multi-Label k-Nearest Neighbour}
+\begin{itemize}
+	\item First-order approach algorithm, adapting knn techniques
+	\item Maximum a posteriori (MAP) rule to make prediction by reasoning and embody label info in the neighbours
+	\item For new instances, $N(x)$ represents the set of its knn (uses Euclidean distance)
 
-\subsection{Alogrithm Adaptation Methods}
-
-\begin{frame}
-\frametitle{Multi-Label k-Nearest Neighbor}
-\begin{itemize}
-\item <2-> 
-\item <3-> 
-\item <4-> 
-\item<5-> 
-\item <6-> 
-% \item<7->
 \end{itemize}
@@
 
-%\begin{figure}
-%	\includegraphics[width = \textwidth, height = \textheight, keepaspectratio]{modelcharacters.png}
-%	\caption{Character enhanced word representations}
-%	\end{figure}
 \end{frame}
@@
 	\item 
-\item 
-\end{itemize}
-\onslide<4->{}
-\begin{itemize}
-\item <5-> 
-\item <6->
-\item <7-> 
 \end{itemize}
@@
 \end{frame}
-
 %------------------------------------------------
-
 \end{document}

### 2018-01-13 Paper/Presentation/mll_review.tex
update presentation

@@
 	\item For unknown instance, get the overall classifiers' votes
-	\item Virtual label $y_v$ as thresholding parameter
+	\item Virtual label $y_v$ as artificial splitting point between relevant and irrelevant labels
 \end{itemize}
@@
 	\item Incompleteness: predicts only label sets included in the training set
-	\item Inefficiency: large $Y$ implies high training complexity.Improvement by combining LP and ensemble learning
+	\item Inefficiency: large $Y$ implies high training complexity.Improve by invoking LP only on random k-labelsets and ensemble learning to LP classifiers
 \end{itemize}
@@
 \begin{frame}
-\Huge{\centerline{Alogrithm Adaptation Methods}}
+\Huge{\centerline{Algorithm Adaptation Methods}}
 \end{frame}
@@
 	\item First-order approach algorithm, adapting knn techniques
+	\item Lazy learning and Bayesian reasoning
 	\item Maximum a posteriori (MAP) rule to make prediction by reasoning and embody label info in the neighbours
 	\item For new instances, $N(x)$ represents the set of its knn (uses Euclidean distance)
+	\item Posterior probability $\mathbb{P}(H_j|C_j)$ and $\mathbb{P}(\neg H_j|C_j)$, where $C_j$ is the knn of $x$ and $H_j$ is the event that $x$ has the label $y_j$
+	\item Predict relevant label set based on which probability is higher(MAP rule)
+	\item Frequency counting strategy:
+	\begin{enumerate}
+		\item Estimation of prior probabilities ($\mathbb{P}(H_j)$ \& $\mathbb{P}(\neg H_j)$) using a smoothing parameter $s$ (usually $s=1$)
+		\item Estimation of likelyhoods (($\mathbb{P}(C_j|H_j)$ \& $\mathbb{P}(C_j|\neg H_j)$)
+	\end{enumerate}
 
@@
 \frametitle{Multi-label Decision Tree}
-
+\begin{itemize}
+	\item First-order approach algorithm, that adopts decision tree techniques
+	\item ML entropy is used to
+build the decision tree recursively
+	\item ML dataset $T$ is divided based on a splitting value $\theta$ ($T^-$ \& $T^+$)
+	\item Starting from the root, ML-DT identifies the feature and the splitting value which
+maximizes the information gain and creates the two child nodes
+	\item ML-DT assumes independence among labels and computes the ML
+entropy in a decomposable way
+	\item Any test instance allocated in a leaf node is associated with the most frequent label $y$ present in that node
+\end{itemize}
+Pros \& cons:
+\begin{itemize}
+	\item High efficiency
+	\item Improvement: pruning strategy  or ensemble learning techniques
+\end{itemize}
 \end{frame}
@@
 \frametitle{Ranking Support Vector Machine}
-
+\begin{itemize}
+	\item Second-order approach algorithm, that adopts maximum margin strategy 
+	\item $q$ linear classifiers, optimized to minimize the empirical
+ranking loss and enabled to handle nonlinear cases with kernel tricks 
+	\item $W = {(w_j , b_j ) | 1 ≤ j ≤ q}$,where $w_j \in \mathbb{R}^d$ and
+$b_j \in \mathbb{R}$ are the weight vector and bias
+	\item Rank-SVM defines the learning system's
+margin on ($x_i$ , $Y_i$), considering its ranking ability on the example’s relevant/irrelevant labels
+\end{itemize}
 \end{frame}
@@
 \frametitle{Collective Multi-Label Classifier}
-
+\begin{itemize}
+	\item 
+	\item 
+	\item 
+	\item 
+\end{itemize}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Related Learning Settings}
+\frametitle{\insertsection}
 \begin{itemize}
@@
 \begin{frame}
-\frametitle{Conclusion}
+\frametitle{\insertsection}
 \begin{itemize}

### 2018-01-13 Project/NearestSubroutes.py
Add non consequtive lcss, options bundle. Various misc updates.

@@ -4,7 +4,7 @@ import threading
 import question1 as qp1
 
 
-def find_similar_subroutes_per_test_trip(test_points, train_df, k=5, paropts=None):
+def find_similar_subroutes_per_test_trip(test_points, train_df, k, paropts=None, conseq_lcss = True):
     if paropts:
         print("Parallelizing with", paropts)
         partype, numpar = paropts
@@ -16,17 +16,17 @@ def find_similar_subroutes_per_test_trip(test_points, train_df, k=5, paropts=Non
     if partype:
         # num threads or processes
         if partype == "processes":
-            max_subseqs = exec_with_processes(train_df, numpar, test_lonlat, k)
+            max_subseqs = exec_with_processes(train_df, numpar, test_lonlat, k, conseq_lcss)
         elif partype == "threads":
-            max_subseqs = exec_with_threads(train_df, numpar, test_lonlat, k)
+            max_subseqs = exec_with_threads(train_df, numpar, test_lonlat, k, conseq_lcss)
     else:
-        max_subseqs = serial_execution(train_df, test_lonlat, k)
+        max_subseqs = serial_execution(train_df, test_lonlat, k, conseq_lcss)
     if len(max_subseqs) != k:
         print("WARNING: Specified %d subseqs!" % k)
     return max_subseqs
 
 
-def serial_execution(df, test_lonlat, k):
+def serial_execution(df, test_lonlat, k, conseq_lcss):
     max_subseqs = []
     for index, row in df.iterrows():
         train_points = row["points"]
@@ -34,12 +34,14 @@ def serial_execution(df, test_lonlat, k):
         train_lonlat = utils.idx_to_lonlat(train_points, format="tuples")
         timestart = utils.tic()
         # compute common subsequences between the test trip and the current candidate
-        subseqs, subseqs_idx = calc_lcss(test_lonlat, train_lonlat)
+        subseqs, subseqs_idx = calc_lcss(test_lonlat, train_lonlat, conseq_lcss= conseq_lcss)
         elapsed = utils.tictoc(timestart)
         # sort by decr. length
         subseqs_idx = sorted(subseqs_idx, key=lambda x: len(x), reverse=True)
         # update the list of the longest subsequences
-        max_subseqs = update_current_maxsubseq(max_subseqs, subseqs_idx, k, elapsed, row)
+        if subseqs:
+            max_subseqs = update_current_maxsubseq(max_subseqs, subseqs_idx, k, elapsed, row)
[…]

### 2018-01-13 Paper/Presentation/mll_review.tex
update report

@@
 	\item Applied to classification
-	\item Supervised: Given dataset $\{ (x_i, y_i)\}, i = 1, \dots N, x \in X, y \in Y$
+	\item Supervised: Dataset $\{ (x_i, y_i)\}, i = 1, \dots N, x \in X, y \in Y$
 	\item Goal is to learn a model $f: X \rightarrow Y$
@@
 \frametitle{\insertsection : \insertsubsection}
-Label search space $\mathbb{S_Y}$ grows exponentially as a function of $|Y|$: 
+Label search space $\mathbb{S_Y}$ grows exponentially as a function of $|Y|=q$
 \begin{itemize}
-	\item e.g. for $|Y|=20, |\mathbb{S_Y}| = 2 ^ {|\mathbb{P(Y)}|} = 2^{20} \ge 10^6$
+	\item e.g. for $q=20, |\mathbb{S_Y}| = 2 ^ {|\mathbb{P(Y)}|} = 2^{20} \ge 10^6$
 \end{itemize}
@@
 \end{frame}
-
 %------------------------------------------------
-
+\subsection{Problem Transformation Methods}
+%------------------------------------------------
 \begin{frame}
-\frametitle{\insertsection : \insertsubsection}
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale = 0.75]{images/aa.png}
-	\end{center}
-\end{figure}
+\Huge{\centerline{Problem Transformation Methods}}
 \end{frame}
-
+%------------------------------------------------
 \begin{frame}
@@
 \end{frame}
-
-%------------------------------------------------
-\subsection{Problem Transformation Methods}
-%------------------------------------------------
-\begin{frame}
-\Huge{\centerline{Problem Transformation Methods}}
-\end{frame}
 %------------------------------------------------
@@
 \begin{itemize}
-	\item Decompose to $q$ independent binary classification problems, $h_i(x)$
-	\item Cross-train on $q$ binary training sets (one for each $y_i$)
-	\item Predict labels of unseen $x$ by evaluating each classifier $h_i(x)$
-	\item Output labels where  $h_i(x) > 0$ else $\underset{1 \geq i \geq q}{argmax} \{h_i(x))\}$ 
+	\item Decompose multi-label problem to $q$ independent binary classification problems
+	\item Construct $q$ binary (one-vs-all) training sets (one for each $y_i$)
+	\item Cross-train each binary classifier $h_i(x)$ on its respective dataset
+
+	\item Predict labels of an unseen $x$ by evaluating each classifier $h_i(x)$
+	\item Output labels where  $h_i(x) > 0$ or $\underset{1 \geq i \geq q}{argmax} \{h_i(x))\}$, if none
 \end{itemize}
@@
 \Huge{\centerline{Algorithm Adaptation Methods}}
+\end{frame}
+%------------------------------------------------
+\begin{frame}
+\frametitle{\insertsection : \insertsubsection}
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale = 0.75]{images/aa.png}
+	\end{center}
+\end{figure}
 \end{frame}

### 2018-01-14 Paper/Presentation/mll_review.tex
update presentation, check the last 2 algorithms

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Beamer Presentation
-% LaTeX Template
-% Version 1.0 (10/11/12)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
+%-------------------------------------------------------------------------------------
 %	PACKAGES AND THEMES
-%----------------------------------------------------------------------------------------
-
+%-------------------------------------------------------------------------------------
 \documentclass{beamer}
@@
 \usepackage[colorinlistoftodos,prependcaption,textsize=small]{todonotes}                                                                                                    
-
-%----------------------------------------------------------------------------------------
+%-------------------------------------------------------------------------------------
 %	TITLE PAGE
-%----------------------------------------------------------------------------------------
+%-------------------------------------------------------------------------------------
 
@@
 {
-	IEEE TRANSACTIONS ON KNOWLEDGE AND DATA ENGINEERING\\ % Your institution for the title page
+	IEEE Transactions On Knowledge And Data Engineering\\ % Your institution for the title page
 	\medskip
@@
 
-%----------------------------------------------------------------------------------------
+%-------------------------------------------------------------------------------------
 %	PRESENTATION SLIDES
-%----------------------------------------------------------------------------------------
+%-------------------------------------------------------------------------------------
 
-%------------------------------------------------
 \section{Introduction} % Sections can be created in order to organize your presentation into discrete blocks, all sections and subsections are automatically printed in the table of contents as an overview of the talk
@@
 \end{frame}
-
 %------------------------------------------------
@@
 margin on ($x_i$ , $Y_i$), considering its ranking ability on the example’s relevant/irrelevant labels
+
+TODO!!!
 \end{itemize}
+Pros \& cons:
+\begin{itemize}
+	\item Benefits from kernels to handle non-linear classification problems
+	\item Multiple kernel learning techniques can be
+employed to learn from multi-label data
+\end{itemize}
+
 \end{frame}
@@
 \begin{itemize}
-	\item 
+	\item Second-order approach algorithm, that adapts maximum entropy
+principle \item Correlations
+among labels are encoded as constraints that the resulting
+	distribution must satisfy
 	\item 
@@
 \end{itemize}
+Pros \& cons:
+\begin{itemize}
+	\item More general than Rank-SVM, that only considers relevant-irrelevant label pairs
+\end{itemize}
+\end{frame}
+%------------------------------------------------
+\begin{frame}
+\frametitle{\insertsection : Summary}
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale = 0.45]{images/summary.png}
+		\caption{Summary of Representative Multi-Label Learning Algorithms}
+	\end{center}
+\end{figure}
 \end{frame}
@@
 %------------------------------------------------
-
 \begin{frame}
@@
 \begin{itemize}
-	\item 
-	\item 
-	\item 
-	\item 
+	\item Multi-instance learning
+	\begin{itemize}
+		\item Each example described by a bag of instances while associated with a single/binary label
+		\item Models the object’s ambiguities in input space
+	\end{itemize}
+	\item Ordinal classification
+	\begin{itemize}
+		\item Natural ordering of class labels and their relevance
+		\item Vague (ordinal) instead
+of definite judgment on the label relevance
+	\end{itemize}
+	\item Multi-task learning
+	\begin{itemize}
+		\item Multiple tasks are trained in parallel, usually small number of tasks
+		\item Training
+information of related tasks used as an inductive bias to improve the generalization performance of other tasks
+		\item Tasks
+can be in the same or different feature space
+	\end{itemize}
+	\item Data streams classification
+	\begin{itemize}
+		\item Real-world objects are generated online and processed in a real-time manner
+	\end{itemize}
 \end{itemize}
@@
 \begin{frame}
+\Huge{\centerline{Conclusion}}
+\end{frame}
+%------------------------------------------------
+\begin{frame}
+\frametitle{\insertsection : Online resources}
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale = 0.47]{images/online.png}
+		\caption{Online Resources for Multi-Label Learning}
+	\end{center}
+\end{figure}
+\end{frame}
+%------------------------------------------------
+\begin{frame}
 \frametitle{\insertsection}
+Summary
 \begin{itemize}
-	\item 
+	\item ML learning problem definition 
+	\item ML learning representative algorithms
+	\item Related learning settings
+\end{itemize}
+
+Open research problems
+\begin{itemize}
+	\item Full understanding and appropriate usage of label correlations (might be asymetric or local)
+	\item Conduct thorough experimental studies to
+get insights on the pros and cons of different multi-label
+learning algorithms
 \end{itemize}
@@
 \Huge{\centerline{Thank you}}
+\Huge{\centerline{Questions?}}
 \end{frame}

### 2018-01-14 Project/JourneyClassification.py
Fixes on various stuff

@@ -12,7 +12,7 @@ import utils
 import os
 
 
-def classify(features, targets, num_folds, classifiers, output_folder, filename_tag = "", classifier_obj = None):
+def train(features, targets, num_folds, classifiers, output_folder, filename_tag ="", classifier_obj = None):
     kf = KFold(n_splits = num_folds)
     folds_idxs = list(kf.split(features))
     trips_array = np.asarray(features)
@@ -54,12 +54,11 @@ def classify(features, targets, num_folds, classifiers, output_folder, filename_
     return mean_accuracies
 
 # TODO improve classification
-def preprocess_data(feature_df):
-    print("Preprocessing data for classification.")
+def preprocess_train_data(feature_df, seed):
+    print("Preprocessing training data")
     targets = []
     data = []
-    # TODO make an options dict bundle cardundle
-    seed = 123
+    # shuffle data
     feature_df = feature_df.sample(frac=1, random_state = seed).reset_index(drop=True)
     for index, row in feature_df.iterrows():
         targets.append(row["journeyId"])
@@ -69,8 +68,11 @@ def preprocess_data(feature_df):
         for point in train_points:
             points.append(point[1])
         data.append(points)
+
     # get maximum length of feature lists
     maxlen = len(max(data, key=lambda x: len(x)))
+    print("Padding training features to a max length of", maxlen)
+    # keep the numeric part of the text features
     data = [[int(d[1:]) for d in dlist] for dlist in data]
     # pad to the maximum length
     for i, datum in enumerate(data):
@@ -85,19 +87,35 @@ def preprocess_data(feature_df):
             num_ids[t] = len(num_ids)
         targets_nums.append(num_ids[t])
 
-    # count occurences
-    hist = []
-    for jid in num_ids:
-        occurences = sum([1 if t == jid else 0 for t in targets])
-        hist.append((jid, occurences))
+    # count instances per jid
+    jids = [j for j in num_ids]
+    counts = [targets_nums.count(num_ids[j]) for j in jids]
+    jids_sorted = sorted(zip(jids,counts), key = lambda x : x[1], reverse = True)
+    print("10 Jids with most data:")
+    for i in range(10):
+        print(i,":",jids_sorted[i])
 
-    # hist = sorted(hist, key=lambda x: x[1], reverse= True)
-    # print("Most frequent 5 jids:")
-    # for (jid, occ) in hist[:5]:
-    #     print(jid, ":", occ)
[…]

### 2018-01-14 Project/MapInGridView.py

@@ -1,29 +1,41 @@
 import os
 import matplotlib.pyplot as plt
 import utils
+from collections import OrderedDict
 
 
-def find_min_max_latlong(train_df):
+def find_min_max_latlon(train_df, output_folder):
     max_lat, max_lon = -1000, -1000
     min_lat, min_lon = 1000, 1000
+    llats, llons = [],[]
     for index, row in train_df.iterrows():
-        train_points = row["points"]
-        train_points = eval(train_points)
-        max_lon = max([t[1] for t in train_points] + [max_lon])
-        min_lon = min([t[1] for t in train_points] + [min_lon])
-        max_lat = max([t[2] for t in train_points] + [max_lat])
-        min_lat = min([t[2] for t in train_points] + [min_lat])
-    return max_lat, max_lon, min_lat, min_lon
+        train_points = eval(row["points"])
+        lats = [t[2] for t in train_points]
+        lons = [t[1] for t in train_points]
+        llats.extend(lats)
+        llons.extend(lons)
+        max_lon = max(lons + [max_lon])
+        min_lon = min(lons + [min_lon])
+        max_lat = max(lats + [max_lat])
+        min_lat = min(lats + [min_lat])
+    plt.plot(llats,llons,"b.")
+    plt.savefig(os.path.join(output_folder,"data_minmax_grid_extent.png"))
+    plt.close()
+    return (max_lon,max_lat),(min_lon, min_lat),llats, llons #max_lat, max_lon, min_lat, min_lon
 
 
-def create_grid(number_of_cells, max_lat,max_lon,min_lat,min_lon, output_folder):
-    cell_lat_dist, cell_lon_dist = get_distance_per_cell(number_of_cells, min_lat, min_lon, max_lat, max_lon)
-    rows = create_grid_lines(number_of_cells[0], min_lat, cell_lat_dist)
-    columns=create_grid_lines(number_of_cells[1], min_lon, cell_lon_dist)
+def create_grid(number_of_cells, max_lonlat,min_lonlat,all_lats, all_lons, output_folder):
+    total_dists = [ m-n for (m,n) in zip(max_lonlat, min_lonlat)]
+    cell_dists = [m/n for (m,n) in zip(total_dists, number_of_cells)]
+    cell_lat_dist, cell_lon_dist = cell_dists
+
+    rows = create_grid_lines(number_of_cells[0], min_lonlat[1], cell_lat_dist)
[…]

### 2018-01-14 Project/NearestNeighbours.py

@@ -30,10 +30,8 @@ def calculate_nns(test_points, train_df, paropts=None, k=5):
     nearest_neighbours = sorted(nearest_neighbours, key=lambda k: k[1])
     # return the top 5
     print("Elapsed for parallelization: ", str(partype), numpar, " is:", utils.tictoc(tic))
-    print("Neighbours:")
     nearest_neighbours = nearest_neighbours[:k]
-    for neigh in nearest_neighbours:
-        print(neigh)
+    print("Neighbours:",[n[0] for n in nearest_neighbours])
     return nearest_neighbours
 
 

### 2018-01-14 Project/main.py

@@ -24,18 +24,18 @@ def question_2(opts):
     print("=====================")
     # Question 2
     print("\n>>> Running question 2a1 - Nearest neighbours computation")
-    qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"])
+    # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"])
 
     print("\n>>> Running question 2a2 - Nearest subroutes computation")
-    qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],opts["conseq_lcss"], opts["k"], opts["paropts"])
+    # qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],opts["conseq_lcss"], opts["k"], opts["paropts"])
 
     print("\n>>> Running question 2b - Cell grid quantization")
-    cellgrid = (10, 10)
+    cellgrid = opts["grid"]
     print("Using cell grid with dimensions", cellgrid)
-    features_file = qp2.question_b(opts["cleanfile"], cellgrid, opts["outdir"])
+    features_file, grid_file = qp2.question_b(opts["cleanfile"], cellgrid, opts["outdir"])
 
     print("\n>>> Running question 2c - Classification")
-    # qp2.question_c(features_file, test_file, class_folder)
+    qp2.question_c(features_file, grid_file, opts["testfiles"][2], opts["classifdir"], opts["seed"], opts["classiffile"], opts["folds"])
 
 
 def check_dependencies():
@@ -81,6 +81,9 @@ if __name__ == '__main__':
     options["cleanfile"]  = join(options["outdir"], "trips_clean.csv")
     options["mapsdir"]    = join(options["outdir"], "gmplots")
     options["classifdir"] = join(options["outdir"], "classification_charts")
+    options["classiffile"] = join(options["outdir"],"​testSet_JourneyPatternIDs.csv")
+    options["folds"] = 2
+    options["grid"] = (5,5)
 
     os.makedirs(options["outdir"], exist_ok=True)
     os.makedirs(options["mapsdir"], exist_ok=True)

### 2018-01-14 Project/question2.py

@@ -5,6 +5,7 @@ import JourneyClassification as jcp
 import pandas as pd
 import utils
 import os
+import pickle
 
 
 def question_a1(output_folder, clean_file, test_file, paropts, k):
@@ -39,32 +40,58 @@ def question_a2(output_folder, test_file, train_file, conseq_lcss, k, paropts):
 
 def question_b(train_file, number_of_cells, output_folder):
     # specify files
-    output_file = os.path.join(output_folder, "tripFeatures.csv")
+    grid_file = os.path.join(output_folder,"grid.pickle")
+    feature_file = os.path.join(output_folder, "tripFeatures.csv")
+    # read data and make the grid
     train_df = pd.read_csv(train_file)
-    max_lat, max_lon, min_lat, min_lon = gvp.find_min_max_latlong(train_df)
-    rows, columns, cell_names = gvp.create_grid(number_of_cells, max_lat, max_lon, min_lat, min_lon, output_folder=output_folder)
-    gvp.replace_points(train_df, rows, columns, cell_names, output_file)
-    return output_file
+    max_lonlat, min_lonlat, all_lats, all_lons = gvp.find_min_max_latlon(train_df, output_folder)
+    grid = gvp.create_grid(number_of_cells, max_lonlat, min_lonlat, all_lats, all_lons, output_folder=output_folder)
+    # save grid and transform data
+    with open(grid_file, "wb") as f:
+        pickle.dump(grid, f)
+    gvp.map_to_features(train_df, grid, feature_file)
+    return feature_file, grid_file
 
 
-def question_c(features_file, test_file, output_folder):
+def question_c(features_file, grid_file, test_file, output_folder, seed, classif_file, num_folds):
     df_features = pd.read_csv(features_file)
-    features, targets = jcp.preprocess_data(df_features)
-    num_folds = 5
+    features, jid_mapping, targets = jcp.preprocess_train_data(df_features, seed)
     classifiers = ["knn", "logreg", "randfor"]
     # classifiers = ["logreg"]
-    mean_accuracies = jcp.classify(features, targets, num_folds, classifiers, output_folder)
+    mean_accuracies = jcp.train(features, targets, num_folds, classifiers, output_folder)
 
     # print mean accuracy per classifier
+    print()
     for classifier in mean_accuracies:
         print(classifier, "accuracy train/val:", mean_accuracies[classifier])
-    impr_classifier = "logreg"
+
+    # select the logistic regression algorithm to beat the benchmark
+    impr_classifier_name = "logreg"
[…]

### 2018-01-14 Project/run.sh

@@ -1,3 +1,3 @@
 #! /usr/bin/env bash
 [ -z $(which phantomjs) ] && echo "Need phantomjs to create output report files!" && exit 
-python Main.py "$(pwd)/inputfolder" "$(pwd)/outputfolder"
+python main.py "$(pwd)/inputfolder" "$(pwd)/outputfolder"

### 2018-01-14 Paper/PaperReport/PaperReport.tex

\documentclass[12pt]{report}
\usepackage[a4paper]{geometry}
\usepackage[myheadings]{fullpage}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{graphicx, wrapfig, subcaption, setspace, booktabs}
\usepackage[T1]{fontenc}
\usepackage[font=small, labelfont=bf]{caption}
\usepackage{fourier}
\usepackage[protrusion=true, expansion=true]{microtype}
\usepackage[english]{babel}
\usepackage{sectsty}
\usepackage{url, lipsum}
\usepackage{listings}
\usepackage{float}
\usepackage{todonotes}
\graphicspath{{images/}}

\lstset{
	basicstyle=\ttfamily,
	columns=fullflexible,
	frame=single,
	breaklines=true,
	postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
}

\newcommand{\HRule}[1]{\rule{\linewidth}{#1}}
\onehalfspacing
\setcounter{tocdepth}{5}
\setcounter{secnumdepth}{5}

%-------------------------------------------------------------------------------
% HEADER & FOOTER
%-------------------------------------------------------------------------------
\pagestyle{fancy}
\fancyhf{}
\setlength\headheight{15pt}
\fancyfoot[R]{Page \thepage\ of \pageref{LastPage}}
%-------------------------------------------------------------------------------
% TITLE PAGE
%-------------------------------------------------------------------------------

\begin{document}

	\title{ \normalsize \textsc{Large-Scale Data Analysis Techniques}
		\\ [2.0cm]
		\HRule{0.5pt} \\
		\LARGE \textbf{\uppercase{A Review on Multi-Label Learning Algorithms}}
		\HRule{2pt} \\ [0.5cm]
		\normalsize \today \vspace*{5\baselineskip}}

	\date{}
	\author{
		Nikiforos Pittaras, M1422\\
		Chrysoula Themeli, M1423 \\ 
		National and Kapodistrian University of Athens\\
		Department of Informatics and Telecommunications }

	\maketitle
	\tableofcontents
	\listoffigures
	\newpage

	%-------------------------------------------------------------------------------
	% Section title formatting
	\sectionfont{\scshape}
	%-------------------------------------------------------------------------------

	%-------------------------------------------------------------------------------
	% Summary of the paper
	%-------------------------------------------------------------------------------

	\section*{Summary}
	\addcontentsline{toc}{section}{Summary}

	%\begin{figure}[H]
	%	\centering
	%	\includegraphics[width=0.8\textwidth]{semsearch.png}
	%	\caption{An overview of the SemSearch architecture}
	%	\centering
	%\end{figure}

	%-------------------------------------------------------------------------------
	% REFERENCES
	%-------------------------------------------------------------------------------
	\newpage
	\addcontentsline{toc}{section}{Bibliography}
	\begin{thebibliography}{10}
		\bibitem{intelligent} 
		Madhu, G., Dr A. Govardhan, and Dr TV Rajinikanth. "Intelligent semantic web search engines: a brief survey." \url{https://arxiv.org/pdf/1102.0831.pdf}
	\end{thebibliography}
\end{document}

### 2018-01-14 Paper/Presentation/mll_review.tex
Update presentation

--- 2018-01-13 Paper/Presentation/mll_review.tex
@@
 \frametitle{\insertsection : \insertsubsection}
-Need for more complicated evaluation than traditional supervised learning(accuracy, F-measure etc). There are 2 metrics categories 
+Two categories and perspectives:
 	\begin{itemize}
-		\item \emph{Example-based: }Evaluates system's performance on each example and returns the mean value. 
+		\item \emph{Example-based: }Evaluates multi-labelled performance on each example, extrapolate to whole dataset
 		\begin{itemize}
-			\item \emph{From classification perspective:} $Subset Accuracy$, $Hamming Loss$, $accuracy_{exam}$, $precision_{exam}$, $recall_{exam}$, $F^\beta_{exam}$
-			\item \emph{From ranking perspective:} one-error, coverage, ranking loss, average precision
+			\item \emph{Classification perspective:} 
+				\begin{itemize}
+					\item Subset Accuracy, Hamming Loss
+					\item Precision, Recall, F-measure, F$^\beta$ -measure
 				\end{itemize}
-		\item \emph{Label-based: }Evaluates system's performance on each class label separately and returns macro/micro averaged value across all class labels
+			\item \emph{Ranking perspective:} one-error, coverage, ranking loss, average precision
+		\end{itemize}
+		\item \emph{Label-based: }Evaluates performance on each label separately, extrapolate to whole label set
 		\begin{itemize}
-			\item From classification perspective: $B_{macro}$, $B_{micro}$
-			\item From ranking perspective: $AUC_{macro}$, $AUC_{micro}$
+			\item Classification perspective: Macro/Micro averaging for single-label example-based, classification-based measures
+			\item Ranking perspective: Macro/Micro averaging for AUC 
 		\end{itemize}
 	\end{itemize}
-TODO: theoretical results?
+	Classifiers should aim to optimize \emph{multiple} metrics
 \end{frame}
@@
 
-\subsection{Simple Categorization}
+\subsection{Categorization}
 
@@
 
-Assume multi-label classification problem with $q$ possible class labels $Y=\{y_1, \dots y_q\}$
-
 Group algorithms in two categories:
 \begin{itemize}
-	\item \emph{Problem Transformation Methods: }Transform the learning problem into other well-established learning scenarios (fit data to algorithm philosophy)
-	\item \emph{Algorithm Adaptation Methods: }Adapt popular learning techniques to deal with multi-label data directly (fit algorithm to data philosophy)
+	\item \emph{Problem Transformation Methods:}
+
+		\begin{itemize}
+			\item Transform the learning problem into other well-established learning scenarios
+			\item ``Fit data to algorithm'' philosophy
 		\end{itemize}
+
+	\item \emph{Algorithm Adaptation Methods:}
+		\begin{itemize}
+			\item Adapt popular learning techniques to deal with multi-label data directly
+			\item ``Fit algorithm to data'' philosophy
+		\end{itemize}
+\end{itemize}
+
+Include algorithms that:
+\begin{itemize}
+	\item Has broad, unique characteristics
+	\item Has primitive impact, i.e. leads to a number follow-up related methods
+	\item Is influential and highly-cited in multi-label learning
+\end{itemize}
+
 \end{frame}
@@
 \begin{itemize}
-	\item Simple, parallelizable
+	\item Simple, one-vs-rest scheme, parallelizable
 	\item Sensitive to class-imbalanced data
@@
 \begin{itemize}
-	\item High-order approach algorithm
-	\item Transform the problem into a chain of binary classification problems
-	\item Build subsequent binary classifiers upon the predictions of preceding ones
-	\item A permutation function is used to specify an order to the labels
-	\item Binary training set
-constructed by appending each instance with its relevance
-to those labels
+	\item A permutation function $f_p$ is used to order $Y$
+	\item Transform into a \emph{chain} of binary classification problems
+	\item Enrich representations at each step $j$ by concatenating each $x_i$ with the confidence of the $j$-th classifier
 	\item Predict relevant labels for unknown instances by iteratively traversing the classifier chain
@@
 \begin{itemize}
-	\item Effectiveness largely affected by the ordering
-	\item Exploitation of label correlations
-	\item Does not support parallel implementation
+	\item Higher-order method: exploitation of label correlations, but in a random manner
+	\item Highly sensitive to $f_p$. Ensemble schemes attempt to overcome it. 
+	\item Iterative operation prevents parallel implementation
 \end{itemize}
@@
 \begin{itemize}
-	\item Second-order approach algorithm
-	\item Transform the problem into a label ranking problem
-	\item Ranking implemented by pairwise comparison techniques
+	\item Transform into a label pairwise comparison ranking problem
 	\item For $q$ labels, generate $q(q-1)/2$ binary classifiers by pairwise comparison
-	\item Construct training set per label pairs considering the example's relevance
-	\item For unknown instance, get the overall classifiers' votes
-	\item Virtual label $y_v$ as artificial splitting point between relevant and irrelevant labels
+	\item Construct training sets $D_{jk} : \{x_i | y_j \in Y_i \oplus y_k \in Y_i\}$
+	\item For an unknown instance, all classifiers' votes are aggregated and ranked 
+	\item A virtual label $y_v$ serves as the artificial splitting point between relevant and irrelevant labels
 \end{itemize}
@@
 \begin{itemize}
-	\item Smooth the class-imbalance
-	\item Classifier's number grows from linear to quadratic scale
+	\item Second-order approach algorithm. One-vs-one scheme.
+	\item Smooths out the class-imbalance problem
+	\item Number of classifiers s quadratic to $|Y|$ (compared to linear for BR) 
 	\begin{itemize}
-		\item Improvement: Exact or approximate pruning to reduce classifier's number to be queried
+		\item Pruning methods to reduce the search space 
 	\end{itemize}
@@
 \begin{itemize}
-	\item High-order approach algorithm
 	\item Decompose to an ensemble of multi-class classification problems
-	\item Each component targets a random subset of $Y$, inducing a multi-class classifier by Label Powerset (LP) techniques
-	\item Convert training set into a multi-class set, treating each label set as a new class
+	\item Each component targets a random subset of $Y$, classified with Label Powerset (LP) techniques:
+	\begin{itemize}
+	\item Transform to single-label data by treating each distinct labelset as a new class
 	\item Each example is reassigned with the mapped single-label participating in multi-class classifier induction
 	\item For unknown $x$, counts the maximum and actual number of votes
-	\item Label set is relevant when the actual number of votes exceeds
-half of the maximum ones
+	\item Label set is relevant when the number of votes exceeds half of the max
+	\end{itemize}
 \end{itemize}
@@
 \begin{itemize}
-	\item Incompleteness: predicts only label sets included in the training set
-	\item Inefficiency: large $Y$ implies high training complexity.Improve by invoking LP only on random k-labelsets and ensemble learning to LP classifiers
+	\item High-order approach algorithm
+	\item Data-sensitive: Cannot generalize to labelsets not in the training set, too few examples for some labelsets
+	\item Large $Y$ implies high training complexity.
+	\item Improve by invoking an ensemble on random k-sized labelsets
 \end{itemize}
@@
 	\item First-order approach algorithm, adapting knn techniques
-	\item Lazy learning and Bayesian reasoning
-	\item Maximum a posteriori (MAP) rule to make prediction by reasoning and embody label info in the neighbours
-	\item For new instances, $N(x)$ represents the set of its knn (uses Euclidean distance)
-	\item Posterior probability $\mathbb{P}(H_j|C_j)$ and $\mathbb{P}(\neg H_j|C_j)$, where $C_j$ is the knn of $x$ and $H_j$ is the event that $x$ has the label $y_j$
-	\item Predict relevant label set based on which probability is higher(MAP rule)
-	\item Frequency counting strategy:
-	\begin{enumerate}
-		\item Estimation of prior probabilities ($\mathbb{P}(H_j)$ \& $\mathbb{P}(\neg H_j)$) using a smoothing parameter $s$ (usually $s=1$)
-		\item Estimation of likelyhoods (($\mathbb{P}(C_j|H_j)$ \& $\mathbb{P}(C_j|\neg H_j)$)
-	\end{enumerate}
+	\item Compares the MAP probabilities: $P(K|C_j), K \in \{H_j, \neg H_j\}$ 
+	\item Priors $P(K)$ computed by smoothed frequency estimation in the training data 
+	\item Likelihoods $P(C_j|K)$ are computed as a function of:
+	\begin{itemize}
+		\item $\kappa(r)$, the number of examples labelled $y_j$ with $r$ neighbours labelled $y_j$
+		\item $\overset{\sim}{\kappa}(r)$, the number of examples not labelled $y_j$ with $r$ neighbours labelled $y_j$
+	\end{itemize}
+\end{itemize}
 
+Pros \& cons:
+\begin{itemize}
+	\item First-order approach, Bayesian and Lazy reasoning
+	\item Decision boundary can be modified on-line via new instances
+	\item Class imbalance can be overcome by calculating priors
+	\item Extensions proposed for label correlation exploitations
 \end{itemize}
+
 \end{frame}
-
 %------------------------------------------------
@@
 \begin{itemize}
-	\item First-order approach algorithm, that adopts decision tree techniques
-	\item ML entropy is used to
-build the decision tree recursively
-	\item ML dataset $T$ is divided based on a splitting value $\theta$ ($T^-$ \& $T^+$)
-	\item Starting from the root, ML-DT identifies the feature and the splitting value which
-maximizes the information gain and creates the two child nodes
-	\item ML-DT assumes independence among labels and computes the ML
-entropy in a decomposable way
-	\item Any test instance allocated in a leaf node is associated with the most frequent label $y$ present in that node
+	\item Multi-label entropy is used to build a decision tree recursively
+	\item Split position is such that the information gain criterion is maximized
+	\item Dataset $T$ is split on the $l$-th feature valued $\theta$ to branches ($T^-$ \& $T^+$), where $ x_{il} \leq \theta, x_i \in T^-$ and $ x_{il} > \theta, x_i \in T^+$
+	\item Recurse on subtrees until a stopping criterion is met (e.g. child size)
+	\item Single-label entropy is computed by considering labelsets as new classes
+	\item Assumes independence among labels to ensure low computational cost
+	\item Unseen instances are assigned the label of the majority of members of the leaf they arrive
 \end{itemize}
@@
 \begin{itemize}
-	\item High efficiency
-	\item Improvement: pruning strategy  or ensemble learning techniques
+	\item First-order algorithm, highly efficient
+	\item Assumes label independence
+	\item Pruning and ensemble strategies proposed
 \end{itemize}
@@
 \begin{itemize}
-	\item Second-order approach algorithm, that adopts maximum margin strategy 
-	\item $q$ linear classifiers, optimized to minimize the empirical
-ranking loss and enabled to handle nonlinear cases with kernel tricks 
-	\item $W = {(w_j , b_j ) | 1 ≤ j ≤ q}$,where $w_j \in \mathbb{R}^d$ and
-$b_j \in \mathbb{R}$ are the weight vector and bias
-	\item Rank-SVM defines the learning system's
-margin on ($x_i$ , $Y_i$), considering its ranking ability on the example’s relevant/irrelevant labels
+	\item $q$ linear classifiers, optimized to minimize the empirical ranking loss
+	\item For each relevant-nonrelevant label pair $(y_j,y_k) \in Y_j \times \bar Y$ their decision boundary is marked by $\left<w_j,w_k\right> + b_j - b_k=0$
+	\item Retrieves a ranked label pairs list for each unseen instance
+	\item Non-linearity achievable through feature mapping and the kernel trick
+	\item Convex quadratic programming problem, solved by any QP solver
 \end{itemize}
+Pros \& cons:
+\begin{itemize}
+	\item Second-order approach, maximum margin strategy 
+	\item Extensible by kernel learning, kernel selection can be overcome by MKL techniques 
+	\item Adaptable learning by selecting an appropriate loss function
+\end{itemize}
+
 \end{frame}
@@
 \begin{itemize}
-	\item 
-	\item 
-	\item 
-	\item 
+	\item Labels encoded as binary random vectors into a joint pdf
+	\item Label correlations encoded into the joint pdf as constraints
+	\item Model via the maximum information entropy criterion: $H_p(x, y)$, subject to constraints $\mathbb{E_p}[f_k(x,y)] = F_k$, where $F_k$ are estimated from the training set
+	\item Objective arrived by Lagrange multipliers, solvable by any constraint optimization solver
+	\item Unseen instances labelled with $Y = \underset{y}{\text{argmax }} P(y | x)$
 \end{itemize}
+Pros \& cons:
+\begin{itemize}
+	\item Second-order approach. All label pairs considered, not just relevant-irrelevant ones.
+	\item Intractable $\text{argmax}$ operation for a large label space without pruning.
+\end{itemize}
+\end{frame}
+%------------------------------------------------
+\begin{frame}
+\frametitle{\insertsection : Summary}
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale = 0.45]{images/summary.png}
+		\caption{Summary of Representative Multi-Label Learning Algorithms}
+	\end{center}
+\end{figure}
 \end{frame}
@@
 \begin{itemize}
-	\item 
-	\item 
-	\item 
-	\item 
+	\item Multi-instance learning
+	\begin{itemize}
+		\item Each ML example described by a bag of instances while associated with a single/binary label
+		\item Models the object’s complex semantics in input space, rather than its output
+	\end{itemize}
+	\item Ordinal classification
+	\begin{itemize}
+		\item Model class co-relevance in a (vague) graded membership vector
+		\item Transform ML problem to a set of ordinal set of problems
+	\end{itemize}
+	\item Multi-task learning
+	\begin{itemize}
+		\item Multiple tasks trained in parallel, sharing information
+		\item Knowledge from related tasks used as an inductive bias to improve generalization 
+		\item Shared or different feature space, small task workload
+	\end{itemize}
+	\item Data streams classification
+	\begin{itemize}
+		\item Real-world objects are generated online and processed in  real-time 
+		\item Concept drift problem
+	\end{itemize}
 \end{itemize}
@@
 \begin{frame}
+\Huge{\centerline{Conclusion}}
+\end{frame}
+%------------------------------------------------
+\begin{frame}
+\frametitle{\insertsection : Online resources}
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale = 0.47]{images/online.png}
+		\caption{Online Resources for Multi-Label Learning}
+	\end{center}
+\end{figure}
+\end{frame}
+%------------------------------------------------
+\begin{frame}
 \frametitle{\insertsection}
+Summary:
 \begin{itemize}
-	\item 
+	\item Multi-label learning problem definition 
+	\item Multi-label learning representative algorithms
+	\item Related learning settings
+\end{itemize}
+
+Future goals:
+\begin{itemize}
+	\item Holy grail: Formal characterization on the underlying concept / mechanism on the appropriate usage of label correlations, especially on large output spaces
+	\item Thorough experimental comparative study to discover to pros and cons of different multi-label learning algorithms
 \end{itemize}

### 2018-01-15 Paper/Presentation/mll_review.tex
update presentation with related work

@@
 \subsection{Problem Definition}
+%------------------------------------------------
+\begin{frame}
+\Huge{\centerline{Introduction}}
+\end{frame}
+%------------------------------------------------
 \begin{frame}
@@
 \end{frame}
-
+%------------------------------------------------
+\subsection{Related Work}
+\begin{frame}
+\frametitle{\insertsection : \insertsubsection}
+\begin{itemize}
+\item Madjarov, Gjorgji, et al. "An extensive experimental comparison of methods for multi-label learning." Pattern Recognition 45.9 (2012): 3084-3104
+\item Tsoumakas, Grigorios, and Ioannis Katakis. "Multi-label classification: An overview." International Journal of Data Warehousing and Mining 3.3 (2006)
+\item Zhang, Min-Ling, and Kun Zhang.
+"Multi-label learning by exploiting label dependency." Proceedings of the 16th ACM SIGKDD international conference on Knowledge discovery and data mining.
+ACM, 2010
+\end{itemize}
+\end{frame}
 %------------------------------------------------
@@
 %------------------------------------------------
-
 \subsection{Categorization}
-
+%------------------------------------------------
+\begin{frame}
+\Huge{\centerline{Multi-label Learning Algorithms}}
+\end{frame}
+%------------------------------------------------
 \begin{frame}
@@
 \item Holy grail: Formal characterization on the underlying concept / mechanism on the appropriate usage of label correlations, especially on large output spaces
-	\item Thorough experimental comparative study to discover to pros and cons of different multi-label learning algorithms
+\item Thorough experimental comparative study to discover pros and cons of different multi-label learning algorithms
 \end{itemize}

### 2018-01-15 Paper/Presentation/mll_review.tex
Minor updates

@@
 \section{Introduction} % Sections can be created in order to organize your presentation into discrete blocks, all sections and subsections are automatically printed in the table of contents as an overview of the talk
-\subsection{Problem Definition}
+\subsection{Multi-label learning}
 %------------------------------------------------
@@
 \frametitle{\insertsection : \insertsubsection}
-Single-label supervised learning
+Applied to supervised classification
 \begin{itemize}
-\item Applied to classification
-\item Supervised: Dataset $\{ (x_i, y_i)\}, i = 1, \dots N, x \in X, y \in Y$
-\item Goal is to learn a model $f: X \rightarrow Y$
-\end{itemize}
-
-Multi-label supervised learning
-\begin{itemize}
-\item Multiple labels per instance: $\{ (x_i, \bm{y}_i)\}, i = 1, \dots N, x \in X, \bm{y} \in \mathbb{P}(Y)$
-\item Learn a model $f: X \times Y \rightarrow r$ where $r\in \mathbb{R}$ is the confidence that $y_i$ characterizes $x_i$
-\item For classification, assume that $x_i$ belongs to $y_i$ if $r \ge $ $t(x_i)$
-
+\item Given input data and labels, goal is to learn a model $f: X \rightarrow Y$ 
+\item $f(x_i, y_i) = r$ where $r\in \mathbb{R}$ is the confidence that $y_i$ characterizes $x_i$
+\item Assume that $x_i$ belongs to $y_i$ if $r \ge $ $t(x_i)$
 \begin{itemize}
@@
 \end{itemize}
+\end{itemize}
+
+Single-label learning
+\begin{itemize}
+	\item Dataset $\{ (x_i, y_i)\}, i = 1, \dots N, x \in X, y \in Y=\{y_1,\dots,y_q\}$
+\end{itemize}
+
+Multi-label learning
+\begin{itemize}
+\item Multi-label dataset: Multiple labels per instance. $\{ (x_i, \bm{y}_i)\}, i = 1, \dots N, x \in X, \bm{y} \in \mathbb{P}(Y)$
+% \item Learn a model $f: X \times Y \rightarrow r$
 
 \end{itemize}
-TODO: Include stuff about multi-labeled dataset characterization?
+%TODO: Include stuff about multi-labeled dataset characterization?
 \end{frame}
 %------------------------------------------------
-\subsection{Related Work}
-\begin{frame}
-\frametitle{\insertsection : \insertsubsection}
-\begin{itemize}
-\item Madjarov, Gjorgji, et al. "An extensive experimental comparison of methods for multi-label learning." Pattern Recognition 45.9 (2012): 3084-3104
-\item Tsoumakas, Grigorios, and Ioannis Katakis. "Multi-label classification: An overview." International Journal of Data Warehousing and Mining 3.3 (2006)
-\item Zhang, Min-Ling, and Kun Zhang.
-"Multi-label learning by exploiting label dependency." Proceedings of the 16th ACM SIGKDD international conference on Knowledge discovery and data mining.
-ACM, 2010
-\end{itemize}
-\end{frame}
 %------------------------------------------------
@@
 \begin{itemize}
-\item Transform multi-labeled problem to multiple, single-label problems
 \item Ignore label correlations
+\item Often transform M-L problem to multiple, single-label problems
 \item Simple, scalable, suboptimal
@@
 \frametitle{\insertsection : \insertsubsection}
+Extension of single-label metrics to the M-L case.
+
 Two categories and perspectives:
@@
 \end{itemize}
+\item[\rightarrow] Classifiers should aim to optimize \emph{multiple} metrics
 \end{itemize}
-Classifiers should aim to optimize \emph{multiple} metrics
 \end{frame}
@@
 \begin{itemize}
-\item Has broad, unique characteristics
+\item Has broad, noteworthy or unique characteristics
 \item Has primitive impact, i.e. leads to a number follow-up related methods
@@
 \item Predict labels of an unseen $x$ by evaluating each classifier $h_i(x)$
-\item Output labels where  $h_i(x) > 0$ or $\underset{1 \geq i \geq q}{argmax} \{h_i(x))\}$, if none
+\item Assign label $y_i$ according to $h_i(x)$ and the thresholding/ranking scheme 
 \end{itemize}
@@
 \item Transform into a \emph{chain} of binary classification problems
-\item Enrich representations at each step $j$ by concatenating each $x_i$ with the confidence of the $j$-th classifier
+\item Enrich representations at step $j$ by concatenating each $x_i$ with the confidence of preceeding $1,\dots,j-1$-th classifiers
 \item Predict relevant labels for unknown instances by iteratively traversing the classifier chain
 \end{itemize}
+todo prediction
 
@@
 \begin{itemize}
-\item Higher-order method: exploitation of label correlations, but in a random manner
+\item Higher-order method: exploitation of label correlations to a degree, but in a random manner
 \item Highly sensitive to $f_p$. Ensemble schemes attempt to overcome it.
@@
 \item For $q$ labels, generate $q(q-1)/2$ binary classifiers by pairwise comparison
-\item Construct training sets $D_{jk} : \{x_i | y_j \in Y_i \oplus y_k \in Y_i\}$
+\item Construct training sets $D_{jk} : \{x_i, Y_i | y_j \in Y_i \oplus y_k \in Y_i\}$
 \item For an unknown instance, all classifiers' votes are aggregated and ranked
@@
 \end{itemize}
+todo how it works exactly cardum
 Pros \& cons:
@@
 \item Decompose to an ensemble of multi-class classification problems
-\item Each component targets a random subset of $Y$, classified with Label Powerset (LP) techniques:
+\item Each component targets a random subset of $\mathbb{P}(Y)$ that appears in $X$, classified with Label Powerset (LP) techniques:
 \begin{itemize}
 \item Transform to single-label data by treating each distinct labelset as a new class
-\item Each example is reassigned with the mapped single-label participating in multi-class classifier induction
-\item For unknown $x$, counts the maximum and actual number of votes
-\item Label set is relevant when the number of votes exceeds half of the max
+\item Each example is reassigned with the new mapped class and classified through regular single-label classification
+\item M-L classify $x$ with $y_i$ when the votes received for $y_i$ from the ensemble exceed half the max possible it can get
 \end{itemize}
+todo how does each LP classify, k-labelsets
 \end{itemize}
@@
 %------------------------------------------------
+\subsection{Related Work}
+\begin{frame}
+\frametitle{\insertsection : \insertsubsection}
+\begin{itemize}
+\item Madjarov, Gjorgji, et al. "An extensive experimental comparison of methods for multi-label learning." Pattern Recognition 45.9 (2012): 3084-3104
+\item Tsoumakas, Grigorios, and Ioannis Katakis. "Multi-label classification: An overview." International Journal of Data Warehousing and Mining 3.3 (2006)
+\item Zhang, Min-Ling, and Kun Zhang.
+"Multi-label learning by exploiting label dependency." Proceedings of the 16th ACM SIGKDD international conference on Knowledge discovery and data mining.
+ACM, 2010
+\end{itemize}
+\end{frame}
+
+%------------------------------------------------
 \section{The end}

### 2018-01-16 Paper/Presentation/mll_review.tex
Update presentation and some todos

@@
 \usepackage{bm}
+%\usepackage{enumitem}
+\usepackage{amsfonts}
 \usepackage[colorinlistoftodos,prependcaption,textsize=small]{todonotes}
@@
 \begin{itemize}
-\item Given input data and labels, goal is to learn a model $f: X \rightarrow Y$ 
-\item $f(x_i, y_i) = r$ where $r\in \mathbb{R}$ is the confidence that $y_i$ characterizes $x_i$
-\item Assume that $x_i$ belongs to $y_i$ if $r \ge $ $t(x_i)$
-\begin{itemize}
-  \item $t(\cdot)$ can be a predetermined constant function or learned from $X$
+\item[$\bullet$] Given input data and labels, goal is to learn a model $f: X \rightarrow Y$ 
+\item[$\bullet$] $f(x_i, y_i) = r$ where $r\in \mathbb{R}$ is the confidence that $y_i$ characterizes $x_i$
+\item[$\bullet$] Assume that $x_i$ belongs to $y_i$ if $r \ge $ $t(x_i)$
+\begin{itemize}%[label={$\checkmark$}]
+\item[$\checkmark$] $t(\cdot)$ can be a predetermined constant function or learned from $X$
 \end{itemize}
@@
 \begin{itemize}
-	\item Dataset $\{ (x_i, y_i)\}, i = 1, \dots N, x \in X, y \in Y=\{y_1,\dots,y_q\}$
+\item[$\bullet$] Dataset $\{ (x_i, y_i)\}, i = 1, \dots N, x \in X, y \in Y=\{y_1,\dots,y_q\}$
 \end{itemize}
@@
 \begin{itemize}
-\item Multi-label dataset: Multiple labels per instance. $\{ (x_i, \bm{y}_i)\}, i = 1, \dots N, x \in X, \bm{y} \in \mathbb{P}(Y)$
+\item[$\bullet$] Multi-label dataset: Multiple labels per instance. $\{ (x_i, \bm{y}_i)\}, i = 1, \dots N, x \in X, \bm{y} \in \mathbb{P}(Y)$
 % \item Learn a model $f: X \times Y \rightarrow r$
@@
 \begin{itemize}
-\item e.g. for $q=20, |\mathbb{S_Y}| = 2 ^ {|\mathbb{P(Y)}|} = 2^{20} \ge 10^6$
+\item[$\triangleright$] e.g. for $q=20, |\mathbb{S_Y}| = 2 ^ {|\mathbb{P(Y)}|} = 2^{20} \ge 10^6$
 \end{itemize}
@@
 \begin{itemize}
-\item Ignore label correlations
-\item Often transform M-L problem to multiple, single-label problems
-\item Simple, scalable, suboptimal
+\item[$\circ$] Ignore label correlations
+\item[$\circ$] Often transform M-L problem to multiple, single-label problems
+\item[$\circ$] Simple, scalable, suboptimal
 \end{itemize}
@@
 \begin{itemize}
-\item Consider \emph{pairwise} label relations
-\item Good trade-off between generalization performance and scalability
-\item Lacking in some real-world applications
+\item[$\circ$] Consider \emph{pairwise} label relations
+\item[$\circ$] Good trade-off between generalization performance and scalability
+\item[$\circ$] Lacking in some real-world applications
 \end{itemize}
@@
 \begin{itemize}
-\item Capture more complicated label interdependencies
-\item Strong modeling capabilities
-\item Computationally demanding, less scalable
+\item[$\circ$] Capture more complicated label interdependencies
+\item[$\circ$] Strong modeling capabilities
+\item[$\circ$] Computationally demanding, less scalable
 \end{itemize}
@@
 \begin{itemize}
-\item \emph{Example-based: }Evaluates multi-labelled performance on each example, extrapolate to whole dataset
+\item[$\bullet$] \emph{Example-based: }Evaluates multi-labeled performance on each example, extrapolate to whole dataset
 \begin{itemize}
-\item \emph{Classification perspective:}
+\item[$\circ$] \emph{Classification perspective:}
 \begin{itemize}
-\item Subset Accuracy, Hamming Loss
-\item Precision, Recall, F-measure, F$^\beta$ -measure
+\item[$\star$] Subset Accuracy, Hamming Loss
+\item[$\star$] Precision, Recall, F-measure, F$^\beta$ -measure
 \end{itemize}
-\item \emph{Ranking perspective:} one-error, coverage, ranking loss, average precision
+\item[$\circ$] \emph{Ranking perspective:} one-error, coverage, ranking loss, average precision
 \end{itemize}
-\item \emph{Label-based: }Evaluates performance on each label separately, extrapolate to whole label set
+\item[$\bullet$] \emph{Label-based: }Evaluates performance on each label separately, extrapolate to whole label set
 \begin{itemize}
-\item Classification perspective: Macro/Micro averaging for single-label example-based, classification-based measures
-\item Ranking perspective: Macro/Micro averaging for AUC
+\item[$\circ$] Classification perspective: Macro/Micro averaging for single-label example-based, classification-based measures
+\item[$\circ$] Ranking perspective: Macro/Micro averaging for AUC
 \end{itemize}
-\item[\rightarrow] Classifiers should aim to optimize \emph{multiple} metrics
 \end{itemize}
+$\ast$ Classifiers should aim to optimize \emph{multiple} metrics
 \end{frame}
@@
 \begin{itemize}
-\item \emph{Problem Transformation Methods:}
+\item[$\bullet$] \emph{Problem Transformation Methods:}
 
 \begin{itemize}
-\item Transform the learning problem into other well-established learning scenarios
-\item ``Fit data to algorithm'' philosophy
+\item[$\circ$] Transform the learning problem into other well-established learning scenarios
+\item[$\circ$] ``Fit data to algorithm'' philosophy
 \end{itemize}
 
-\item \emph{Algorithm Adaptation Methods:}
+\item[$\bullet$] \emph{Algorithm Adaptation Methods:}
 \begin{itemize}
-\item Adapt popular learning techniques to deal with multi-label data directly
-\item ``Fit algorithm to data'' philosophy
+\item[$\circ$] Adapt popular learning techniques to deal with multi-label data directly
+\item[$\circ$] ``Fit algorithm to data'' philosophy
 \end{itemize}
@@
 \begin{itemize}
-\item Has broad, noteworthy or unique characteristics
-\item Has primitive impact, i.e. leads to a number follow-up related methods
-\item Is influential and highly-cited in multi-label learning
+\item[$\checkmark$] Has broad, noteworthy or unique characteristics
+\item[$\checkmark$] Has primitive impact, i.e. leads to a number follow-up related methods
+\item[$\checkmark$] Is influential and highly-cited in multi-label learning
 \end{itemize}
@@
 \begin{itemize}
-\item Decompose multi-label problem to $q$ independent binary classification problems
-\item Construct $q$ binary (one-vs-all) training sets (one for each $y_i$)
-\item Cross-train each binary classifier $h_i(x)$ on its respective dataset
+\item[$\bullet$] Decompose multi-label problem to $q$ independent binary classification problems
+\item[$\bullet$] Construct $q$ binary (one-vs-all) training sets (one for each $y_i$)
+\item[$\bullet$] Cross-train each binary classifier $h_i(x)$ on its respective dataset
 
-\item Predict labels of an unseen $x$ by evaluating each classifier $h_i(x)$
-\item Assign label $y_i$ according to $h_i(x)$ and the thresholding/ranking scheme 
+\item[$\bullet$] Predict labels of an unseen $x$ by evaluating each classifier $h_i(x)$
+\item[$\bullet$] Assign label $y_i$ according to $h_i(x)$ and the thresholding/ranking scheme 
 \end{itemize}
@@
 \begin{itemize}
-\item Simple, one-vs-rest scheme, parallelizable
-\item Sensitive to class-imbalanced data
-\item Ignores label correlations (first-order method)
+\item[$\bullet$] Simple, one-vs-rest scheme, parallelizable
+\item[$\bullet$] Sensitive to class-imbalanced data
+\item[$\bullet$] Ignores label correlations (first-order method)
 \end{itemize}
@@
 \begin{itemize}
-\item A permutation function $f_p$ is used to order $Y$
-\item Transform into a \emph{chain} of binary classification problems
-\item Enrich representations at step $j$ by concatenating each $x_i$ with the confidence of preceeding $1,\dots,j-1$-th classifiers
-\item Predict relevant labels for unknown instances by iteratively traversing the classifier chain
+\item[$\bullet$] A permutation function $f_p$ is used to order $Y$
+\item[$\bullet$] Transform into a \emph{chain} of binary classification problems
+\item[$\bullet$] Enrich representations at step $j$ by concatenating each $x_i$ with the confidence of preceeding $1,\dots,j-1$-th classifiers
+\item[$\bullet$] Predict relevant labels for unknown instances by iteratively traversing the classifier chain
+\item[$\bullet$] $\lambda_{\tau(j)}^x \in \{-1,+1\}$ represents the predicted binary assignment
+of $y_{\tau(j)}$, which is recursively obtained
 \end{itemize}
-todo prediction
 
@@
 \begin{itemize}
-\item Higher-order method: exploitation of label correlations to a degree, but in a random manner
-\item Highly sensitive to $f_p$. Ensemble schemes attempt to overcome it.
-\item Iterative operation prevents parallel implementation
+\item[$\bullet$] High-order method: exploitation of label correlations to a degree, but in a random manner
+\item[$\bullet$] Highly sensitive to $f_p$. Ensemble schemes attempt to overcome it.
+\item[$\bullet$] Iterative operation prevents parallel implementation
 \end{itemize}
@@
 \begin{itemize}
-\item Transform into a label pairwise comparison ranking problem
-\item For $q$ labels, generate $q(q-1)/2$ binary classifiers by pairwise comparison
-\item Construct training sets $D_{jk} : \{x_i, Y_i | y_j \in Y_i \oplus y_k \in Y_i\}$
-\item For an unknown instance, all classifiers' votes are aggregated and ranked
-\item A virtual label $y_v$ serves as the artificial splitting point between relevant and irrelevant labels
+\item[$\bullet$] Transform into a label pairwise comparison ranking problem
+\item[$\bullet$] For $q$ labels, generate $q(q-1)/2$ binary classifiers by pairwise comparison and use a binary algorithm $g_{jk}(x)$
+\item[$\bullet$] Construct training sets $D_{jk} : \{x_i, Y_i | y_j \in Y_i \oplus y_k \in Y_i\}$.
+System votes for each example and if $g_{jk}(x) >0$,$x_i$ is associated with $y_j$, otherwise with $y_k$
+\item[$\bullet$] For an unknown instance, all classifiers' votes are aggregated and ranked
+\item[$\bullet$] A virtual label $y_v$ serves as the artificial splitting point between relevant and irrelevant labels
 \end{itemize}
-todo how it works exactly cardum
 Pros \& cons:
 \begin{itemize}
-\item Second-order approach algorithm. One-vs-one scheme.
-\item Smooths out the class-imbalance problem
-\item Number of classifiers s quadratic to $|Y|$ (compared to linear for BR)
+\item[$\bullet$] Second-order approach algorithm. One-vs-one scheme.
+\item[$\bullet$] Smooths out the class-imbalance problem
+\item[$\bullet$] Number of classifiers s quadratic to $|Y|$ (compared to linear for BR)
 \begin{itemize}
-\item Pruning methods to reduce the search space
+\item[$\circ$] Pruning methods to reduce the search space
 \end{itemize}
@@
 \begin{itemize}
-\item Decompose to an ensemble of multi-class classification problems
-\item Each component targets a random subset of $\mathbb{P}(Y)$ that appears in $X$, classified with Label Powerset (LP) techniques:
+\item[$\bullet$] Decompose to an ensemble of multi-class classification problems
+\item[$\bullet$] Each component targets a random subset of $\mathbb{P}(Y)$ that appears in $X$, classified with Label Powerset (LP) techniques:
 \begin{itemize}
-\item Transform to single-label data by treating each distinct labelset as a new class
-\item Each example is reassigned with the new mapped class and classified through regular single-label classification
-\item M-L classify $x$ with $y_i$ when the votes received for $y_i$ from the ensemble exceed half the max possible it can get
+\item[$\circ$] Transform to single-label data by treating each distinct labelset as a new class
+\item[$\circ$] Each example is reassigned with the new mapped class and classified through regular single-label classification
+\item[$\circ$] M-L classify $x$ with $y_i$ when the votes received for $y_i$ from the ensemble exceed half the max possible it can get
 \end{itemize}
@@
 \begin{itemize}
-\item High-order approach algorithm
-\item Data-sensitive: Cannot generalize to labelsets not in the training set, too few examples for some labelsets
-\item Large $Y$ implies high training complexity.
-\item Improve by invoking an ensemble on random k-sized labelsets
+\item[$\bullet$] High-order approach algorithm
+\item[$\bullet$] Data-sensitive: Cannot generalize to labelsets not in the training set, too few examples for some labelsets
+\item[$\bullet$] Large $Y$ implies high training complexity.
+\item[$\bullet$] Improve by invoking an ensemble on random k-sized labelsets
 \end{itemize}
@@
 \begin{itemize}
-\item First-order approach algorithm, adapting knn techniques
-\item Compares the MAP probabilities: $P(K|C_j), K \in \{H_j, \neg H_j\}$
-\item Priors $P(K)$ computed by smoothed frequency estimation in the training data
-\item Likelihoods $P(C_j|K)$ are computed as a function of:
+\item[$\bullet$] First-order approach algorithm, adapting knn techniques
+\item[$\bullet$] Compares the MAP probabilities: $P(K|C_j), K \in \{H_j, \neg H_j\}$
+\item[$\bullet$] Priors $P(K)$ computed by smoothed frequency estimation in the training data
+\item[$\bullet$] Likelihoods $P(C_j|K)$ are computed as a function of:
 \begin{itemize}
-\item $\kappa(r)$, the number of examples labelled $y_j$ with $r$ neighbours labelled $y_j$
-\item $\overset{\sim}{\kappa}(r)$, the number of examples not labelled $y_j$ with $r$ neighbours labelled $y_j$
+\item[$\circ$] $\kappa(r)$, the number of examples labelled $y_j$ with $r$ neighbours labelled $y_j$
+\item[$\circ$] $\overset{\sim}{\kappa}(r)$, the number of examples not labelled $y_j$ with $r$ neighbours labelled $y_j$
 \end{itemize}
@@
 \begin{itemize}
-\item First-order approach, Bayesian and Lazy reasoning
-\item Decision boundary can be modified on-line via new instances
-\item Class imbalance can be overcome by calculating priors
-\item Extensions proposed for label correlation exploitations
+\item[$\bullet$] First-order approach, Bayesian and Lazy reasoning
+\item[$\bullet$] Decision boundary can be modified on-line via new instances
+\item[$\bullet$] Class imbalance can be overcome by calculating priors
+\item[$\bullet$] Extensions proposed for label correlation exploitations
 \end{itemize}
@@
 \begin{itemize}
-\item Multi-label entropy is used to build a decision tree recursively
-\item Split position is such that the information gain criterion is maximized
-\item Dataset $T$ is split on the $l$-th feature valued $\theta$ to branches ($T^-$ \& $T^+$), where $ x_{il} \leq \theta, x_i \in T^-$ and $ x_{il} > \theta, x_i \in T^+$
-\item Recurse on subtrees until a stopping criterion is met (e.g. child size)
-\item Single-label entropy is computed by considering labelsets as new classes
-\item Assumes independence among labels to ensure low computational cost
-\item Unseen instances are assigned the label of the majority of members of the leaf they arrive
+\item[$\bullet$] Multi-label entropy is used to build a decision tree recursively
+\item[$\bullet$] Split position is such that the information gain criterion is maximized
+\item[$\bullet$] Dataset $T$ is split on the $l$-th feature valued $\theta$ to branches ($T^-$ \& $T^+$), where $ x_{il} \leq \theta, x_i \in T^-$ and $ x_{il} > \theta, x_i \in T^+$
+\item[$\bullet$] Recurse on subtrees until a stopping criterion is met (e.g. child size)
+\item[$\bullet$] Single-label entropy is computed by considering labelsets as new classes
+\item[$\bullet$] Assumes independence among labels to ensure low computational cost
+\item[$\bullet$] Unseen instances are assigned the label of the majority of members of the leaf they arrive
 \end{itemize}
@@
 \begin{itemize}
-\item First-order algorithm, highly efficient
-\item Assumes label independence
-\item Pruning and ensemble strategies proposed
+\item[$\bullet$] First-order algorithm, highly efficient
+\item[$\bullet$] Assumes label independence
+\item[$\bullet$] Pruning and ensemble strategies proposed
 \end{itemize}
@@
 \begin{itemize}
-\item $q$ linear classifiers, optimized to minimize the empirical ranking loss
-\item For each relevant-nonrelevant label pair $(y_j,y_k) \in Y_j \times \bar Y$ their decision boundary is marked by $\left<w_j,w_k\right> + b_j - b_k=0$
-\item Retrieves a ranked label pairs list for each unseen instance
-\item Non-linearity achievable through feature mapping and the kernel trick
-\item Convex quadratic programming problem, solved by any QP solver
+\item[$\bullet$] $q$ linear classifiers, optimized to minimize the empirical ranking loss
+\item[$\bullet$] For each relevant-nonrelevant label pair $(y_j,y_k) \in Y_j \times \bar Y$ their decision boundary is marked by $\left<w_j,w_k\right> + b_j - b_k=0$
+\item[$\bullet$] Retrieves a ranked label pairs list for each unseen instance
+\item[$\bullet$] Non-linearity achievable through feature mapping and the kernel trick
+\item[$\bullet$] Convex quadratic programming problem, solved by any QP solver
 \end{itemize}
@@
 \begin{itemize}
-\item Second-order approach, maximum margin strategy
-\item Extensible by kernel learning, kernel selection can be overcome by MKL techniques
-\item Adaptable learning by selecting an appropriate loss function
+\item[$\bullet$] Second-order approach, maximum margin strategy
+\item[$\bullet$] Extensible by kernel learning, kernel selection can be overcome by MKL techniques
+\item[$\bullet$] Adaptable learning by selecting an appropriate loss function
 \end{itemize}
@@
 \begin{itemize}
-\item Labels encoded as binary random vectors into a joint pdf
-\item Label correlations encoded into the joint pdf as constraints
-\item Model via the maximum information entropy criterion: $H_p(x, y)$, subject to constraints $\mathbb{E_p}[f_k(x,y)] = F_k$, where $F_k$ are estimated from the training set
-\item Objective arrived by Lagrange multipliers, solvable by any constraint optimization solver
-\item Unseen instances labelled with $Y = \underset{y}{\text{argmax }} P(y | x)$
+\item[$\bullet$] Labels encoded as binary random vectors into a joint pdf
+\item[$\bullet$] Label correlations encoded into the joint pdf as constraints
+\item[$\bullet$] Model via the maximum information entropy criterion: $H_p(x, y)$, subject to constraints $\mathbb{E_p}[f_k(x,y)] = F_k$, where $F_k$ are estimated from the training set
+\item[$\bullet$] Objective arrived by Lagrange multipliers, solvable by any constraint optimization solver
+\item[$\bullet$] Unseen instances labelled with $Y = \underset{y}{\text{argmax }} P(y | x)$
 \end{itemize}
@@
 \begin{itemize}
-\item Second-order approach. All label pairs considered, not just relevant-irrelevant ones.
-\item Intractable $\text{argmax}$ operation for a large label space without pruning.
+\item[$\bullet$] Second-order approach. All label pairs considered, not just relevant-irrelevant ones.
+\item[$\bullet$] Intractable $\text{argmax}$ operation for a large label space without pruning.
 \end{itemize}
@@
 \begin{itemize}
-\item Multi-instance learning
+\item[$\bullet$] Multi-instance learning
 \begin{itemize}
-\item Each ML example described by a bag of instances while associated with a single/binary label
-\item Models the object’s complex semantics in input space, rather than its output
+\item[$\circ$] Each ML example described by a bag of instances while associated with a single/binary label
+\item[$\circ$] Models the object’s complex semantics in input space, rather than its output
 \end{itemize}
-\item Ordinal classification
+\item[$\bullet$] Ordinal classification
 \begin{itemize}
-\item Model class co-relevance in a (vague) graded membership vector
-\item Transform ML problem to a set of ordinal set of problems
+\item[$\circ$] Model class co-relevance in a (vague) graded membership vector
+\item[$\circ$] Transform ML problem to a set of ordinal set of problems
 \end{itemize}
-\item Multi-task learning
+\item[$\bullet$] Multi-task learning
 \begin{itemize}
-\item Multiple tasks trained in parallel, sharing information
-\item Knowledge from related tasks used as an inductive bias to improve generalization
-\item Shared or different feature space, small task workload
+\item[$\circ$] Multiple tasks trained in parallel, sharing information
+\item[$\circ$] Knowledge from related tasks used as an inductive bias to improve generalization
+\item[$\circ$] Shared or different feature space, small task workload
 \end{itemize}
-\item Data streams classification
+\item[$\bullet$] Data streams classification
 \begin{itemize}
-\item Real-world objects are generated online and processed in  real-time
-\item Concept drift problem
+\item[$\circ$] Real-world objects are generated online and processed in  real-time
+\item[$\circ$] Concept drift problem
 \end{itemize}
@@
 \begin{itemize}
-\item Multi-label learning problem definition
-\item Multi-label learning representative algorithms
-\item Related learning settings
+\item[$\bullet$] Multi-label learning problem definition
+\item[$\bullet$] Multi-label learning representative algorithms
+\item[$\bullet$] Related learning settings
 \end{itemize}
@@
 \begin{itemize}
-\item Holy grail: Formal characterization on the underlying concept / mechanism on the appropriate usage of label correlations, especially on large output spaces
-\item Thorough experimental comparative study to discover pros and cons of different multi-label learning algorithms
+\item[$\bullet$] Holy grail: Formal characterization on the underlying concept / mechanism on the appropriate usage of label correlations, especially on large output spaces
+\item[$\bullet$] Thorough experimental comparative study to discover pros and cons of different multi-label learning algorithms
 \end{itemize}
@@
 \begin{itemize}
-\item Madjarov, Gjorgji, et al. "An extensive experimental comparison of methods for multi-label learning." Pattern Recognition 45.9 (2012): 3084-3104
-\item Tsoumakas, Grigorios, and Ioannis Katakis. "Multi-label classification: An overview." International Journal of Data Warehousing and Mining 3.3 (2006)
-\item Zhang, Min-Ling, and Kun Zhang.
+\item[$\bullet$] Madjarov, Gjorgji, et al. "An extensive experimental comparison of methods for multi-label learning." Pattern Recognition 45.9 (2012): 3084-3104
+\item[$\bullet$] Tsoumakas, Grigorios, and Ioannis Katakis. "Multi-label classification: An overview." International Journal of Data Warehousing and Mining 3.3 (2006)
+\item[$\bullet$] Zhang, Min-Ling, and Kun Zhang.
 "Multi-label learning by exploiting label dependency." Proceedings of the 16th ACM SIGKDD international conference on Knowledge discovery and data mining.

### 2018-01-16 ProjectReport/BigDataHomework.tex
update report

@@
 The project is divided in two parts: in the first part the goals are: to preprocess and clean the given data as well as to visualize five of the given trips in the train\_set file.
-In the second part, the tasks are: to find for all test trips the k nearest neabours from the given train trips, to find the train trips with the longest similar subroute for each test trip, to use 10-fold cross validation to the training data and classify them with three different methods (k-nearest neighbours, logistic regression and random forest) and finally to choose one of the above classification methods and improve the result.
+In the second part, the tasks are: to find for all test trips the k nearest neighbors from the given train trips, to find the train trips with the longest similar sub-route for each test trip, to use 10-fold cross validation to the training data and classify them with three different methods (k-nearest neighbors, logistic regression and random forest), to choose one of the above classification methods and improve the result and finally to test the classifier on some given test data.
 
     In the main.py file, it is required to provide the path to the input folder (containing the training and test files) and the output folder where all results will be stored.
-In this folder an additional directories are created, in order to keep all maps extracted from this application and also to store the charts showing the classification accuracy.
+Afterwards, an object containing all the necessary for the program options is created.
+This object includes options such as the given input and output directories, the train and the required test files, the number of cells in the grid to be constructed, the number of nearest neighbors etc.
 There are two ways in order to run this project, either add these folders in the edit configurations choice from the menu, or use the run.sh in order to provide these args.
@@
 	\subsection{Data Preprocessing (question A)}
-	The requirement in this part is to preprocess the input training data and create a new file with 3 columns, containingthe trip\_id (a number incrementing by 1 for each trip), the journey\_id(excluding trips with null id) and all points and timestamps in the format: [[timestamp1, longitude1, latitude1], [timestamp2, longitude2, latitude2],...].
-The given csv file with the training data was transformed to a dataframe using the relevant pandas library.
-As already mentioned, the data with null journeyPatternId should be removed from the dataframe, therefore the lines are checked for NaN and "null" journey id.
+	The requirement in this part is to preprocess the input training data and create a new file with 3 columns, containing the trip\_id (a number incrementing by 1 for each trip), the journey\_id(excluding trips with null id) and all points and timestamps in the format: [[timestamp1, longitude1, latitude1], [timestamp2, longitude2, latitude2],...].
+The given csv file with the training data was transformed to a dataframe using the pandas library.
+The data with null journeyPatternId should be removed from the dataframe, therefore the lines are checked for NaN and "null" journey id.
 Then, the columns vehicleID and journeyPatternId are groupedby three times in order to create three different dataframes.
-Foe example, the first dataframe contains the columns: vehicleID, journeyPatternId and timpestamp.
+For example, the first dataframe contains the columns: vehicleID, journeyPatternId and timpestamp.
 In a similar way, the two remaining dataframes were formatted.
@@
 	\section{Exercise 2}
-	The files implementing the requirements of this part are in the file question2.py and some functions implementing the logic for this question are in the files: NearestNeighbours.py, NearestSubroutes.py, MapInGridView, JourneyClassification.py.
+	The files implementing the requirements of this part are in the file question2.py and some functions implementing the logic for this question are in the files: NearestNeighbors.py, NearestSubroutes.py, MapInGridView, JourneyClassification.py.
 
-	\subsection{Nearest Neighbours (question A1)}
+	\subsection{Nearest Neighbors (question A1)}
 	For this part a test file is given (test\_set\_a1.csv, included in the input folder) containing additional trips.
-For each of these trips, the k nearest neighbours are targeted (in this exercise k=5) and an image file is produced, showing all six trips (the test trip and the trips from the training file) on six different maps.
-Under each map, the test or neighbour index, the journey\_id, the dynamic time warping and the time needed to process the neighbours of this test trip in millisec.
+For each of these trips, the k nearest neighbors are targeted (in this exercise k=5) and an image file is produced, showing all six trips (the test trip and the trips from the training file) on six different maps.
+Under each map, the test or neighbor index, the journey\_id, the dynamic time warping and the time needed to process the neighbors of this test trip in millisecond.
 
 	In the function question\_a1, two dataframes are created for both the test and the training trips.
-The test dataframe's rows are iterated and for each row the function calculate\_nns from the file NearestNeighbours.py is called in order to get the nearest neighbours, calculating also the time needed for this action.
+The test dataframe's rows are iterated and for each row the function calculate\_nns from the file NearestNeighbors.py is called in order to get the nearest neighbors, calculating also the time needed for this action.
 
@@
 In the function run\_with\_processes an async task is executed for each process and then the result is stored in a list, while in the function run\_with\_threads, the dataframe is splitted in sub dataframes (the number is specified by the number of threads) and then distance is calculated for each subdataframe.
-After thread join, the result is also stored in the nearest\_neighbours list.
+After thread join, the result is also stored in the nearest\_neighbors list.
 
 	In the function calculate\_dists, the 3rd column of the dataframe is provided and it is transformed in tuples with the latitudes and longitudes.
-The list nearest neighbours will keep in tuples for each train trip the trip\_id, the calculated distance (which is calculated using the dynamic time wraping algorithm), the journeyId and finally all the points of the journey.
+The list nearest neighbors will keep in tuples for each train trip the trip\_id, the calculated distance (which is calculated using the dynamic time wraping algorithm), the journeyId and finally all the points of the journey.
 
-	The algorithm dynamic time wraping is implemented with two ways based on the input (diy or diy\_initial). \todo{diy and diyinitial}
+	The algorithm dynamic time warping is implemented with two ways based on the input (diy or diy\_initial). \todo{diy and diyinitial}
 

### 2018-01-16 Paper/PaperReport/PaperReport.tex
update paper report

@@
 \usepackage{todonotes}
+\usepackage{stmaryrd}
+\usepackage{amsfonts}
 \graphicspath{{images/}}
@@
 
-	\section*{Summary}
-	\addcontentsline{toc}{section}{Summary}
+	\section*{Introduction}
+	\addcontentsline{toc}{section}{Introduction}
 
+	\subsection*{Multi-label learning}
+	\addcontentsline{toc}{subsection}{Multi-label learning}
+	The paper "A Review on Multi-Label Learning Algorithms" analyzes the multi-label learning field in supervised classification.
+Supervised classification trains a model $f: X \rightarrow Y$ and for each pair $f(x_i,y_i)$, the model gives as output a number that represents the confidence that label $y_i$ characterizes $x_i$.
+Traditional single-label learning maps an example to only one label, ignoring completely multiple semantics that real-objects usually have.
+On the other hand, multi-label learning allows an example/instance $x_i$ to be mapped to more than one label.
+However, this results to an exponentially growing search space, since search space might include any number of subsets from the powerset of label-set $Y$ ($|\mathbb{P(Y)}|$).
+
+	One straightforward solution, is to exploit label correlations.
+Real-objects are usually characterized by similar concepts.
+For example, if an article is characterized by label "football" it is very likely that it will be also characterized by the term "sports", while label "science" seems to be irrelevant.
+In order to exploit such kind of label correlations, three different algorithm strategies were developed:
+
+	\begin{itemize}
+		\item \textbf{First-order strategies: }These algorithms ignore label correlations and perform a "one-vs-all" strategy.
+They often transform a multi-label problem into multiple single-label problems.
+Algorithms of this category are simple, scalable, have the possibility of parallel implementation, but on the other hand their results are suboptimal.
+		\item \textbf{Second-order strategies: }In this category, relevant algorithms consider pairwise label correlations, having a good trade-off between generalization performance and scalability.
+However, there are still deficiencies in some real-world applications.
+		\item \textbf{High-order strategies: }These algorithms capture even more complicated label correlations and they have strong modeling capabilities.
+But, on the other hand, they are computationally more demanding and less scalable.
+	\end{itemize}
+
+	\subsection*{Evaluation metrics}
+	\addcontentsline{toc}{subsection}{Evaluation metrics}
+	One other issue that worths to be mentioned are the evaluation metrics.
+Since multi-label problems are more complicated, it is necessary to extend existing sigle-label metrics in order to fit the multi-label case.
+There are two categories of metrics, divided in two perspectives:
+	\begin{itemize}
+		\item \textbf{Example-based: }This kind of metrics evaluate multi-labeled performance on each example and then the result is spread to the whole dataset.
+The metrics are divided into two perspectives:
+		\begin{itemize}
+			\item \textbf{Classification:} 
+			\begin{itemize}
+				\item \emph{Subset Accuracy: } $subsetacc(h) = \frac{1}{p} \sum_{i=1}^p \llbracket h(x_i) = Y_i \rrbracket$
+				\item \emph{Hamming Loss: }$hloss(h) = \frac{1}{p} \sum_{i=1}^{p} [h(x_i)\Delta Y_i] $
+				\item \emph{Accuracy:} $Accuracy_{exam}(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i \cup h(x_i)|}$
+				\item \emph{Precision: }$Precision_{exam}(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{h(x_i)|}$
+				\item \emph{Recall: }$Recall_{exam}(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i}$
+				\item \emph{$F^ \beta$ -metric: } $F^ \beta _{exam} (h) = \frac{(1+ \beta ^2) \cdot Precision_{exam}(h) \cdot Recall_{exam}(h)}{\beta ^2 \cdot Precision_{exam}(h) + Recall_{exam}(h)}$
+			\end{itemize}
+			\item \textbf{Ranking:}
+			\begin{itemize}
+				\item \emph{One-error: }$one-error(f) = \frac{1}{p} \llbracket [argmax_{y \in Y}f(x_i,y)] \notin Y_i \rrbracket$
+				\item \emph{Coverage: } $coverage(f) = \frac{1}{p} \sum_{i=1}^{p} max_{y \in Y_i} rank_f(x_i,y) - 1$
+				\item \emph{Ranking loss: }$rloss(f) = \frac{1}{p} \sum_{i=1}^{p} $ \todo{complete it}
+				\item \emph{Average Precision: } \todo{average precision}
+			\end{itemize}
+		\end{itemize}
+		\item \textbf{Label-based: }
+		\begin{itemize}
+			\item \todo{label based metrics}
+		\end{itemize}
+	\end{itemize}
 	%\begin{figure}[H]
@@
 
+	\section*{Multi-label learning algorithms}
+	\addcontentsline{toc}{section}{Multi-label learning algorithms}
+	In this paper, 8 representative multi-label algorithms are presented and analyzed, seleceted with respect to below criteria:
+	\begin{itemize}
+		\item[$\checkmark$] Broad spectrum (unique characteristics)
+		\item[$\checkmark$] Primitive impact (lead to a number follow-up related methods)
+		\item[$\checkmark$] Favorable influence (highly-cited) 
+	\end{itemize}
+
+	The algorithms to be presented are grouped in two categories:
+	\begin{itemize}
+		\item \textbf{Problem Transformation Methods: }Algorithms of this category, in order to handle a multi-label problem, transform it into other well-established learning scenarios ("Fit data to algorithm" philosophy)
+		\item \textbf{Algorithm Adaptation Methods: }Here, algorithms adapt popular learning techniques to deal with multi-label data directly ("Fit algorithm to data" philosophy)
+	\end{itemize}
+
+	\subsection*{Problem Transformation Methods}
+	\addcontentsline{toc}{subsection}{Problem Transformation Methods}
+
+	\subsubsection*{Binary Relevance}
+	\addcontentsline{toc}{subsubsection}{Binary Relevance}
+	This algorithm decomposes the multi-learning problem into $q$ independent binary classification problems, each of them corresponding to a possible label in the label-set.
+First task is to construct a corresponding binary training set and then to use a binary learning algorithm in order to train the relevant binary classifier.
+Each training instance is involved in the learning process of all $q$ classifiers(cross-training strategy), which is regarded as positive instance in case of relevant labels and irrelevant otherwise.
+For new instances, the algorithm predicts the relevant label set by querying each binary classifier and combining the relevant labels.
+
+	This is a first-order approach algorithm, since it ignores completely any label associations.
+However, since the one-vs-rest scheme is used and classifiers are build for each label separately, the algorithm has the possibility of parallel implementation.
+Finally, the algorithm is sensitive to class-imbalance data (true positives are much less than false positives).
 	%-------------------------------------------------------------------------------
@@
 	\begin{thebibliography}{10}
-		\bibitem{intelligent} 
-		Madhu, G., Dr A. Govardhan, and Dr TV Rajinikanth. "Intelligent semantic web search engines: a brief survey." \url{https://arxiv.org/pdf/1102.0831.pdf}
+		\bibitem{comp} 
+		Madjarov, Gjorgji, et al. "An extensive experimental comparison of methods for multi-label learning." Pattern Recognition 45.9 (2012): 3084-3104
+		\bibitem{overview} Tsoumakas, Grigorios, and Ioannis Katakis. "Multi-label classification: An overview." International Journal of Data Warehousing and Mining 3.3 (2006)
+		\bibitem{correlations} Zhang, Min-Ling, and Kun Zhang.
+"Multi-label learning by exploiting label dependency." Proceedings of the 16th ACM SIGKDD international conference on Knowledge discovery and data mining.
+ACM, 2010
 	\end{thebibliography}

### 2018-01-16 Paper/Presentation/mll_review.tex
update presentation
Update presentation

@@
 \begin{itemize}
-\item[$\bullet$] First-order approach algorithm, adapting knn techniques
-\item[$\bullet$] Compares the MAP probabilities: $P(K|C_j), K \in \{H_j, \neg H_j\}$
-\item[$\bullet$] Priors $P(K)$ computed by smoothed frequency estimation in the training data
+	\item[$\bullet$] Let $H_j \equiv (y_j \in Y_i)$, $C_j = \sum_{x_k \in N(x_i)} [\delta(y_j \in Y_k)]$
+\item[$\bullet$] Compares the MAP probabilities: $P(K|C_j), K \in \{H_j, \neg H_j\}$ to decide if to include $y_j$ in the prediction. Compute with the Bayes' theorem.
+\item[$\bullet$] Priors $P(K)$ computed by smoothed frequency counting in the training data
 \item[$\bullet$] Likelihoods $P(C_j|K)$ are computed as a function of:
@@
 \begin{itemize}
-\item[$\bullet$] First-order approach, Bayesian and Lazy reasoning
+\item[$\bullet$] First-order approach, Bayesian reasoning
 \item[$\bullet$] Decision boundary can be modified on-line via new instances
@@
 \item[$\bullet$] Multi-label entropy is used to build a decision tree recursively
-\item[$\bullet$] Split position is such that the information gain criterion is maximized
+\item[$\bullet$] Split $x$ at feature $l$ s.t. the information gain criterion is maximized
 \item[$\bullet$] Dataset $T$ is split on the $l$-th feature valued $\theta$ to branches ($T^-$ \& $T^+$), where $ x_{il} \leq \theta, x_i \in T^-$ and $ x_{il} > \theta, x_i \in T^+$
@@
 \begin{itemize}
-\item[$\bullet$] $q$ linear classifiers, optimized to minimize the empirical ranking loss
-\item[$\bullet$] For each relevant-nonrelevant label pair $(y_j,y_k) \in Y_j \times \bar Y$ their decision boundary is marked by $\left<w_j,w_k\right> + b_j - b_k=0$
-\item[$\bullet$] Retrieves a ranked label pairs list for each unseen instance
+	\item[$\bullet$] $q$ linear classifiers $h_i(x)$, optimized with the empirical ranking loss
+\item[$\bullet$] Relevant-nonrelevant label pairs of $x_i$ $(y_j,y_k) \in Y_i \times \bar Y_i$ are separated by a function of their classifiers $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$
+\item[$\bullet$] Retrieves a ranked list of label pairs for each unseen instance
 \item[$\bullet$] Non-linearity achievable through feature mapping and the kernel trick
@@
 \end{itemize}
+TODO : details of operation ixta
 Pros \& cons:
@@
 \item[$\bullet$] Second-order approach, maximum margin strategy
-\item[$\bullet$] Extensible by kernel learning, kernel selection can be overcome by MKL techniques
+\item[$\bullet$] Extensible by kernel learning: kernel selection can be overcome by Multiple Kernel Learning techniques
 \item[$\bullet$] Adaptable learning by selecting an appropriate loss function
@@
 \end{itemize}
+how the fuck
 Pros \& cons:
@@
 \begin{itemize}
-\item[$\bullet$] Multi-instance learning
+\item[$\bullet$] Multi-instance learning Todoooo
 \begin{itemize}
@@
 \end{itemize}
-\item[$\bullet$] Ordinal classification
+\item[$\bullet$] Ordinal classification Todoooo
 \begin{itemize}
@@
 \begin{frame}
-\frametitle{\insertsection : Online resources}
-\begin{figure}
-\begin{center}
-\includegraphics[scale = 0.47]{images/online.png}
-\caption{Online Resources for Multi-Label Learning}
-\end{center}
-\end{figure}
-\end{frame}
-%------------------------------------------------
-\begin{frame}
 \frametitle{\insertsection}
@@
 
+%------------------------------------------------
+\subsection{Online resources}
+\begin{frame}
+\frametitle{\insertsection : \insertsubsection}
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale = 0.47]{images/online.png}
+		\caption{Online Resources for Multi-Label Learning}
+	\end{center}
+\end{figure}
+\end{frame}
 %------------------------------------------------

### 2018-01-17 Paper/Presentation/mll_review.tex
update presentation and paper report
Update presentation

@@
 \end{itemize}
-%TODO: Include stuff about multi-labeled dataset characterization?
 \end{frame}
@@
 \end{itemize}
-TODO : details of operation ixta
+TODO : details of svm operation
 Pros \& cons:
@@
 \begin{itemize}
-\item[$\bullet$] Multi-instance learning Todoooo
+\item[$\bullet$] Multi-instance learning
+  % https://en.wikipedia.org/wiki/Multiple-instance_learning
 \begin{itemize}
-\item[$\circ$] Each ML example described by a bag of instances while associated with a single/binary label
-\item[$\circ$] Models the object’s complex semantics in input space, rather than its output
+\item[$\circ$] Instead of labeled instances, get binary-labeled \emph{bags of instances}
+\item[$\circ$] Assign positive label to bag if at least one member is positive
+\item[$\circ$] Models complex semantics of $x_i$ in input space, rather than its output
 \end{itemize}
-\item[$\bullet$] Ordinal classification Todoooo
+\item[$\bullet$] Ordinal classification
+  % https://en.wikipedia.org/wiki/Ordinal_regression
 \begin{itemize}
-\item[$\circ$] Model class co-relevance in a (vague) graded membership vector
+\item[$\circ$] Assume labels are related through a natural ordering
+\item[$\circ$] Instead of explicit class relevance, assume an ordering of
+  relevance between labels via a membership vector
 \item[$\circ$] Transform ML problem to a set of ordinal set of problems

### 2018-01-17 Paper/PaperReport/PaperReport.tex

@@
 Finally, the algorithm is sensitive to class-imbalance data (true positives are much less than false positives).
+
+	\subsubsection*{Classifier Chains}
+	\addcontentsline{toc}{subsubsection}{Classifier Chains}
+	The logic of this algorithm is to transform the multi-label learning problem into a chain of binary classification problems, using subsequent binary classifiers in the chain, built upon the predictions of preceding ones and ordered by a permutation function.
+First of all, a binary training set is constructed by enriching each instance with the confidence of preceding classifiers.
+Moreover, a binary algorithm is used to determine whether $y_i$ label is relevant or not for an instance.
+For unknown instances, the algorithm predicts the relevant labels by iteratively traversing the classifier chain ($\lambda _{\tau(j)}^x \in {-1, +1}$ represents the predicted binary assignment of $y_{\tau(j)}$)
+
+	This is a high-order method, that exploits label correlations to a degree, but in a random manner.
+However, it is highly sensitive to the permutation function, since the classifiers' ordering directly affects the result.
+One other disadvantage is that iterative operation prevents parallel implementation.
+Finally, one possible improvement to overcome the affect of classifiers' ordering is to use ensemble schemes.
+
+	\subsubsection*{Claibrated Label Ranking}
+	\addcontentsline{toc}{subsubsection}{Claibrated Label Ranking}
+	This algorithm transforms the learning problem into a label pairwise comparison ranking problem.
+For $q$ labels, $q(q-1)/2$ binary classfiers are generated by pairwise comparison and a binary algorithm $g_{jk}(x)$ is used.
+Firstly, training sets $D_{jk}$ are constructing, where $D_{jk}: \{x_i, Y_i | y_j \in Y_i \oplus y_k \in Y_i\}$.
+The learning system votes for each example and if $g_{jk}(x)>0$, then $x_i$ is associated with label $y_j$, otherwise $y_j$ is irrelevant and $x_i$ is assigned to $y_k$.
+For unknown instances, classifiers' votes are aggregated and ranked.
+In addition, a virtual label $y_v$ is used as an artificial splitting point between relevant and irrelevant labels.
+
+	This is a second-order approach algorithm using a one-vs-one scheme.
+The advantage of this method is that it smooths out the class-imbalance problem (where the number of a class positives is much less than the other's class positives).
+The disadvantage is the number of classifiers is quadratic to $|Y|$, compared to linear for Binary Relevance algorithm.
+One improvement to this issue is to use pruning methods to reduce the search space.
+
+	\subsubsection*{Random k-Labelsets}
+	\addcontentsline{toc}{subsubsection}{Random k-Labelsets}
 	%-------------------------------------------------------------------------------
@@
 	\newpage
-	\addcontentsline{toc}{section}{Bibliography}
+	\addcontentsline{toc}{section}{Related Work}
 	\begin{thebibliography}{10}

### 2018-01-17 Paper/PaperReport/PaperReport.tex
Update paper report

@@
 \usepackage{amsfonts}
+\usepackage{amsmath}
 \graphicspath{{images/}}
@@
 	\end{itemize}
-	%\begin{figure}[H]
-	%	\centering
-	%	\includegraphics[width=0.8\textwidth]{semsearch.png}
-	%	\caption{An overview of the SemSearch architecture}
-	%	\centering
-	%\end{figure}
 
@@
 	\addcontentsline{toc}{subsection}{Problem Transformation Methods}
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.9\textwidth]{pt.png}
+		\caption{Problem Transformation Algorithms}
+		\centering
+	\end{figure}
 
@@
 	\addcontentsline{toc}{subsubsection}{Random k-Labelsets}
+	Random k-labelsets algorithm transforms the multi-label learning problem into an ensemble of multi-class classification problems where each component targets a random subset of $Y$, that appears in $X$ classified with Label Powerset (LP) techniques.
+In more detail, the multi-label dataset is transformed into single-label data by treating each distinct labelset as a new class.
+Each training example is reassigned with the new mapped class and classified through regular single-label classification.
+An instance $x_i$ is assigned to label $y_i$ when the votes received for $y_i$ from the ensemble exceed half the max possible that this label can get.
+For unseen instances $x$, LP predicts its associated label set $Y$ by querying the prediction of multi-class classifier and then mapping it back to the power set of $Y$.
+
+	This is a high-order approach algorithm, utilizing label correlations. However, there are two limitations in terms of practical feasibility:
+
+	\begin{itemize}
+		\item[$\diamond$] \emph{Incompleteness: }The algorithm is data-sensitive and cannot generalize to labelsets that do not appear into the training set
+		\item[$\diamond$] \emph{Inefficiency: }Large $Y$ implies high training complexity and extremely few training examples for some newly mapped classes.
+	\end{itemize}
+
+	To overcome above issues while keeping LP's simplicity, Random k-Labelsets chooses to combine ensemble learning with LP.
+The latter is invoked only on random k-labelsets (each leading to a multi-class classifier $g_{y^k(l_r)}^+$) to guarantee computational efficiency and then ensemble a number of LP classifiers to achieve predictive completeness.
+For new instances $x$, two quantities are calculated: $\tau(\textbf{x},y_j)$, which counts the maximum number of votes that $y_j$ can receive from the ensemble and $\mu(\textbf{x},y_j)$, which counts the actual number of votes that $y_j$ receives from the ensemble.
+Therefore, the predicted labelset is calculated as follows: $Y = \{y_j|\mu(\textbf{x},y_j)/\tau(\textbf{x},y_j)>0.5,  1 \leq j \leq q\}$.
+This means that when the actual number of votes exceeds half of the maximum number of votes, $y_j$ is considered as a relevant label.For an ensemble created by n k-labelsets, the maximum number of votes on each label is $nk/q$ on average (usually: k=3 and n = 2q).
+Finally, another way to improve LP is to prune distinct label set in D appearing less than a pre-specified counting threshold.
+
+	\subsection*{Algorithm Adaptation Methods}
+	\addcontentsline{toc}{subsection}{Algorithm Adaptation Methods}
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.9\textwidth]{aa.png}
+		\caption{Algorithm Adaptation Methods}
+		\centering
+	\end{figure}
+
+	\subsubsection*{Multi-Label k-Nearest Neighbor (ML-kNN)}
+	\addcontentsline{toc}{subsubsection}{Multi-Label k-Nearest Neighbor (ML-kNN)}
+	ML-kNN algorithm adapts k-nearest neighbor techniques utilizing maximum a posteriori (MAP) rule to make prediction by reasoning with the labeling information embodied in the neighbors.
+For new instances $x$, $N(x)$ represents the set of its k nearest neighbors identified in the dataset (similarity among instances is calculated by Euclidean distance).
+Let $H_j \equiv (y_j \in Y_i)$, $C_j = \sum_{x_k \in N(x_i)} [\delta(y_j \in Y_k)]$.
+$C_j$ records the number of x’s neighbors with label $y_j$, while $H_j$ is the event that $\textbf{x}$ has the label $y_j$.
+Moreover, $P(H_j|C_j)$ is the posterior probability that $H_j$ can be applied in the condition that $\textbf{x}$ has exactly $C_j$ neighbors.
+Similarly, $P(\neg H_j|C_j)$ needs to be calculated in order to decide if to include $y_j$ in the prediction.
+This will be computed using the Bayes' theorem:
+
+	\begin{center}
+		$\frac{P(H_j|C_J)}{P(\neg H_j|C_j)} = \frac{P(H_j) \cdot P(C_j|H_j)}{P(\neg H_j) \cdot P(C_j| \neg H_j)}$
+	\end{center}
+
+	Prior probabilities ($P(H_j)$ and $P(\neg H_j)$) are computed via the frequency counting strategy, where a smoothing parameter $s$ is used to control the effect of uniform prior on the estimation (usually $s=1$).
+In addition, likelyhoods $P(C_j|H_j)$ and $P(C_j| \neg H_j)$ are computed using two frequency arrays:
+	\begin{itemize}
+		\item $\kappa(r)$, the number of examples labelled $y_j$ with $r$ neighbours labelled $y_j$
+		\item $\overset{\sim}{\kappa}(r)$, the number of examples not labelled $y_j$ with $r$ neighbours labelled $y_j$
+	\end{itemize}
+
+	This algorithm is a first-order approach using Bayesian reasoning.
+The decision boundary can be modified on-line when new instances appear.
+Moreover, Through prior probabilities calculation class-imbalance issue can be overcome.
+However, since label correlations are ignored, some extensions are proposed that take into account label associations.
+
+	\subsubsection*{Multi-Label Decision Tree (ML-DT)}
+	\addcontentsline{toc}{subsubsection}{Multi-Label Decision Tree (ML-DT)}
+	ML-DT adapts decision tree techniques using multi-label entropy to build a decision tree recursively.
+Let $T = \{(x_i, Y_i) | 1 \leq i \leq n \}$ be the multi-label dataset with n examples.
+The algorithm splits the $x$ vector at the feature $l$ that maximizes the information gain criterion.
+Moreover, dataset $T$ is split on the $l$-th feature valued $\theta$ into branches ($T^-$ \& $T^+$), where $x_{il} \leq \theta, x_i \in T^-$ and $x_{il} > \theta, x_i \in T^+$.
+This process is invoked recursively by treating each time one of the two branches as the new root node and terminates until some stopping criterion $C$ is met (e.g.
+child size).
+Each subset is treated as a new class and single-label entropy is computed.
+Any new instances are assigned the label of the majority of members of the leaf that they arrive.
+
+	In order to ensure low computational cost and high efficiency, the algorithm assumes independence among the labels (first-order approach).
+However, extended strategies are proposes, such as pruning and ensemble strategies, so as to exploit label correlations as well.
+
+	\subsubsection*{Ranking Support Vector Machine (Rank-SVM)}
+	\addcontentsline{toc}{subsubsection}{Ranking Support Vector Machine (Rank-SVM)}
+	Rank-SVM adapts maximum margin strategy, where $q$ linear classifiers are optimized to minimize empirical ranking loss and enabled to handle nonlinear cases with kernel tricks.
+The learning system is composed by those classifiers $W = \{w_j,b_j | 1 \leq j \leq q\}$, where $w_j \in \mathbb{R}^d$ and $b_j \in \mathbb{R}$ are the weight vector and the bias
+	for the j-th class label $y_j$ respectively.
+Rank-SVM defines the system’s margin on $(x_i,Y_i)$ by considering its ranking ability on the relevant and irrelevant label pairs of $x_i$ $(y_j,y_k) \in Y_i \times \bar Y_i$, which are separated by a function of their classifiers $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$.
+When the learning system is capable of properly ranking every relevant-irrelevant label pair for each training example, the classifier will return positive margin.
+For each new instances $x$, the algorithms retrieves a ranked list of label pairs.
+
+	Moreover, non-linear problems can be solved through feature mapping and kernel trick, while convex quadratic programming problems can be solved by any QP solver.
+This algorithm is a second-order approach using pairwise label associations.
+Improvements can be achieved by: replacing ranking with hamming loss which can be cast as a general form of	structured output classification, accomplishing thresholding strategy with techniques other than stacking-style procedure and avoiding the problem of kernel selection by employing multiple kernel learning techniques to learn from multi-label data.
+
+	 \subsubsection*{Collective Multi-Label Classifier}
+	 \addcontentsline{toc}{subsubsection}{Collective Multi-Label Classifier}
+	 This algorithm adapts maximum entropy principle, where correlations among labels are encoded as binary random vectors into a joint probability distribution function(pdf) and label correlations are encoded into the joint pdf as constraints.
+This is a second-order approach, where all label pairs are considered and not just the relevant-irrelevant ones.
+Let $H_p(x, y)$ represent the information entropy of $(x,y)$ given their distribution $p(\cdot , \cdot)$.
+The principle of maximum entropy assumes that the distribution best modeling the current state of knowledge is the one maximizing $H_p(x, y)$ subject to a collection K of given facts: $\mathbb{E_p}[f_k(x,y)] = F_k$, where $F_k$ are estimated from the training set.
+The fact is expressed as constraint on the expectation of  some  function  over.
+Along with the normalization constraint on $p(\cdot , \cdot)$ , the constrained optimization problem can be solved with standard Lagrange Multiplier techniques.
+Any unseen instance $x$ is labeled with $Y = \underset{y}{\text{argmax }} P(y | x)$.
+Exact inference with arg max is only tractable for small label space, otherwise pruning strategies should be applied.
+
+	 \subsection*{Summary of Multi-Label Learning Algorithms}
+	 \addcontentsline{toc}{subsection}{Summary of Multi-Label Learning Algorithms}
+
+	 \begin{figure}[H]
+	 	\centering
+	 	\includegraphics[width=1\textwidth]{summary.png}
+	 	\caption{Summary of Multi-Label Learning Algorithms}
+	 	\centering
+	 \end{figure}
+
+ 	\section*{Related Learning Methods}
+ 	\addcontentsline{toc}{section}{Related Learning Methods}
+ 	Four learning methods are mentioned in this paper, which can be found below:
+ 	\begin{enumerate}
+ 		\item \textbf{Multi-instance learning: }This method studies the problem where each example is described by a bag of instances while associated with a single/binary label.
+Bag is assigned positive label in case of at least one positive member.
+Finally, it models complex semantics of $x_i$ in input space rather than its output.
+ 		\item \textbf{Ordinal classification: }This method assumes that labels are related through a natural ordering and instead of explicit class relevance, it assumes an ordering of relevant between labels via a membership vector.
+Finally, the multi-label problem is transformed into a set of ordinal set of problems.
+ 		\item \textbf{Multi-task learning: }This method studies the problem where multiple tasks are trained in parallel, sharing information, where knowledge from related tasks is used as an inductive bias to help improve the generalization performance of other tasks.
+The  tasks can be in the same feature space or different feature spaces, while in multi-label learning all the examples share the same feature space.
+Finally, in multi-task learning it is not reasonable to have a large number of tasks, while in multi-label learning it is usual to deal with large label space.
+ 		\item \textbf{Data streams classification: }This method studies the problem	where real-world objects are generated online and processed in a real-time manner.
+Key factor is to deal with the concept drift problem.
+Some existing strategies are: classifiers update significantly whenever a new batch of examples arrive, assumption that the influence of past data gradually declines as time evolves or change detector alerting whenever a concept drift is detected.
+ 	\end{enumerate}
+
+	\section*{Conclusion}
+	\addcontentsline{toc}{section}{Conclusion}
+	Summarizing all the above, this paper defines the multi-label learning problem, presents 8 multi-label learning representative algorithms and mentiones some additional related learning methods.
+Some important future goals are the formal characterization on the underlying concept/mechanism on the appropriate usage of label correlations (especially on large output spaces) and also a thorough experimental comparative study to discover the advantages and disadvantages of different multi-label learning algorithms.
+Below, some online resources on multi-label learning are presented:
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=1\textwidth]{online.png}
+		\caption{Online resources for Multi-Label Learning}
+		\centering
+	\end{figure}
+
 	%-------------------------------------------------------------------------------
@@
 	%-------------------------------------------------------------------------------
-	\newpage
 	\addcontentsline{toc}{section}{Related Work}
-	\begin{thebibliography}{10}
+	\renewcommand\bibname{Related Work}
+	\begin{thebibliography}{9}
 		\bibitem{comp} 

### 2018-01-17 Paper/PaperReport/PaperReport.tex
update paper report

@@
 				\item \emph{Coverage: } $coverage(f) = \frac{1}{p} \sum_{i=1}^{p} max_{y \in Y_i} rank_f(x_i,y) - 1$
-				\item \emph{Ranking loss: }$rloss(f) = \frac{1}{p} \sum_{i=1}^{p} $ \todo{complete it}
-				\item \emph{Average Precision: } \todo{average precision}
+				\item \emph{Ranking loss: }$rloss(f) = \frac{1}{p} \sum_{i=1}^{p} \frac{1}{|Y_i| {\overset{\sim}{|Y_i|}}} |\{(y', y'') | f(x_i,y') \leq f(x_i,y''), (y', y'') \in Y_i \times {\overset{\sim}{|Y_i|}} \}| $
+				\item \emph{Average Precision: } $agvprec(f) = \frac{1}{p} \sum_{i=1}^{p} \frac{1}{|Y_i|} \sum_{y \in Y_i} \frac{|\{y' | rank_f(x_i,y') \leq rank_f(x_i,y), y' \in Y_i\}|}{rank_f(x_i,y)}$
 			\end{itemize}
 		\end{itemize}
-		\item \textbf{Label-based: }
+		\item \textbf{Label-based: }Before proceeding with the metrics, true positive(TP), false positive(FP), true negative (TN) and false negative (FN) will be defined:
 		\begin{itemize}
-			\item \todo{label based metrics}
+			\item \emph{True positive: } $TP_j = |\{x_i | y_j \in Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|;$
+			\item \emph{False positive: }$FP_j = |\{x_i | y_j \notin Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|;$
+			\item \emph{True negative: }$TN_J = |\{x_i | y_j \notin Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|;$
+			\item \emph{False negative: }$FN_J = |\{x_i | y_j \in Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|;$
+		\end{itemize}
+		The label-based metrics are also separated in two perspectives:
+		\begin{itemize}
+			\item \textbf{Classification: }Let $B(TP_j, FP_j, TN_J, FN_J)$ represent some specific binary classification metric $(B \in \{Accuracy, Precision, Recall, F^ \beta \})$.
+The metrics from classification perspective are the following:
+			\begin{itemize}
+				\item \emph{Macro-averaging: }$B_{macro}(h) = \frac{1}{q} \sum_{1}^{q} B(TP_j, FP_j, TN_J, FN_J)$
+				\item \emph{Micro-averaging: }$B_{micro}(h) = B(\sum_{1}^{q}TP_j, \sum_{1}^{q}FP_j, \sum_{1}^{q}TN_j, \sum_{1}^{q}FN_j)$
+			\end{itemize}
+			\item \textbf{Ranking:}
+			\begin{itemize}
+				\item \emph{AUC macro: }$AUC_{macro} = \frac{1}{q} \sum_{1}^{q} \frac{|\{(x', x'') | f(x',Y_j) \geq f(x'',Y_j), (x',x'') \in Z_j \times \overset{\sim}{|Z_j|} \}|}{|Z_j|{\overset{\sim}{|Z_j|}}}$
+				\item \emph{AUC micro: }$AUC_{micro} = \frac{|\{(x',x'',y',y'') | f(x',y') \geq f(x'',y''), (x',y') \in S^+, (x'',y'') \in S^- \}|}{|S^+||S^-|}$
+			\end{itemize}
 		\end{itemize} 

### 2018-01-17 Paper/Presentation/mll_review.tex
Update presentation on svm, cml

@@
 \frametitle{\insertsection : \insertsubsection}
-Applied to supervised classification
+Learning $\rightarrow$ build a model from data, to accomplish a task
 \begin{itemize}
-\item[$\bullet$] Given input data and labels, goal is to learn a model $f: X \rightarrow Y$ 
+  \item Supervised: we have both data \emph{and labels}
+\item Applied to classification in this study
+\end{itemize}
+Supervised classification:
+\begin{itemize}
+\item[$\bullet$] Given input data $X$ and labels  $Y$,
+  learn the function $f: X \rightarrow Y$ 
 \item[$\bullet$] $f(x_i, y_i) = r$ where $r\in \mathbb{R}$ is the confidence that $y_i$ characterizes $x_i$
@@
 \item[$\bullet$] Smooths out the class-imbalance problem
-\item[$\bullet$] Number of classifiers s quadratic to $|Y|$ (compared to linear for BR)
+\item[$\bullet$] Number of classifiers is quadratic to $|Y|$ (compared to linear for BR)
 \begin{itemize}
@@
 	\item[$\bullet$] $q$ linear classifiers $h_i(x)$, optimized with the empirical ranking loss
-\item[$\bullet$] Relevant-nonrelevant label pairs of $x_i$ $(y_j,y_k) \in Y_i \times \bar Y_i$ are separated by a function of their classifiers $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$
-\item[$\bullet$] Retrieves a ranked list of label pairs for each unseen instance
+\item[$\bullet$] Relevant-nonrelevant label pairs of $x_i$: $(y_j,y_k) \in Y_i
+  \times \bar Y_i$: decision hyperplane\footnote{w.r.t. singed classifier distance} is: $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$
+\item[$\bullet$] Separation margin is the minimum distance of $x_i$ from every  $Y_i \times \bar Y_i$
+\item[$\bullet$] SVM $\rightarrow$ minimize loss \emph{and} maximize margin.
+  Approximate $max(\cdot)$ and introduce slack variables.
+
+\item[$\bullet$] Ranks labels w.r.t. singled distance
 \item[$\bullet$] Non-linearity achievable through feature mapping and the kernel trick
-\item[$\bullet$] Convex quadratic programming problem, solved by any QP solver
 \end{itemize}
-TODO : details of svm operation
 Pros \& cons:
@@
 \item[$\bullet$] Second-order approach, maximum margin strategy
+\item[$\bullet$] Convex quadratic programming problem, solved by any QP solver
 \item[$\bullet$] Extensible by kernel learning: kernel selection can be overcome by Multiple Kernel Learning techniques
@@
 \begin{itemize}
-\item[$\bullet$] Labels encoded as binary random vectors into a joint pdf
-\item[$\bullet$] Label correlations encoded into the joint pdf as constraints
-\item[$\bullet$] Model via the maximum information entropy criterion: $H_p(x, y)$, subject to constraints $\mathbb{E_p}[f_k(x,y)] = F_k$, where $F_k$ are estimated from the training set
-\item[$\bullet$] Objective arrived by Lagrange multipliers, solvable by any constraint optimization solver
-\item[$\bullet$] Unseen instances labelled with $Y = \underset{y}{\text{argmax }} P(y | x)$
+\item[$\bullet$] Encodes labels as binary random vectors $\bm{y}=\{-1,1\}\}^q$
+\item[$\bullet$] For data $x,\bm{y}$ learn joint probability distribution $p(x,\bm{y})$
+\item[$\bullet$] Learn pdf $p(x,\bm{y})$ via the principle of maximun entropy
+  \begin{itemize}
+    \item Maximize $H_p(x,y)$ subject to $K$ constraints $\mathbb{E}[f_k(x,\bm{y})] = F_k$
+    \item $f_k$ models label correlations with $F_k$ learned from the training data
   \end{itemize}
-how the fuck
+\item[$\bullet$] Optimize via Lagrange multipliers, arrive at $p(\bm{y} | x)$
+\item[$\bullet$] Unseen instances labelled with $Y = \underset{y}{\text{argmax }} p(y | x)$
+\end{itemize}
 Pros \& cons:
@@
 \item[$\bullet$] Second-order approach. All label pairs considered, not just relevant-irrelevant ones.
+\item[$\bullet$] Convex constraint optimization problem $\rightarrow$ solvable by any
+  CP solver
 \item[$\bullet$] Intractable $\text{argmax}$ operation for a large label space without pruning.
@@
 \end{frame}
-
+%------------------------------------------------
+\begin{frame}
+\frametitle{\insertsection : \insertsubsection}
+Useful additional bibliography
+\begin{itemize}
+\item[$\bullet$] Elisseeff, André, and Jason Weston. "A kernel method for multi-labelled classification." Advances in neural information processing systems. 2002.
+\item[$\bullet$]  Ghamrawi, Nadia, and Andrew McCallum.
+"Collective multi-label classification." Proceedings of the 14th ACM international conference on Information and knowledge management.
+ACM, 2005.
+\end{itemize}
+\end{frame}
 %------------------------------------------------

### 2018-01-17 Paper/PaperReport/PaperReport.tex

@@
 	postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
+}
+
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks,
+	citecolor=black,
+	filecolor=black,
+	linkcolor=black,
+	urlcolor=black
 }
@@
 		\begin{itemize}
-			\item \textbf{Classification:} 
+			\item Classification:
 			\begin{itemize}
@@
 			\end{itemize}
-			\item \textbf{Ranking:}
+			\item Ranking:
 			\begin{itemize}
@@
 		\begin{itemize}
-			\item \textbf{Classification: }Let $B(TP_j, FP_j, TN_J, FN_J)$ represent some specific binary classification metric $(B \in \{Accuracy, Precision, Recall, F^ \beta \})$.
+			\item Classification: Let $B(TP_j, FP_j, TN_J, FN_J)$ represent some specific binary classification metric $(B \in \{Accuracy, Precision, Recall, F^ \beta \})$.
 The metrics from classification perspective are the following:

### 2018-01-17 Paper/PaperReport/PaperReport.tex
Update

--- 2018-01-17 Paper/PaperReport/PaperReport.tex
@@
 	\addcontentsline{toc}{subsection}{Multi-label learning}
-	The paper "A Review on Multi-Label Learning Algorithms" analyzes the multi-label learning field in supervised classification.
-Supervised classification trains a model $f: X \rightarrow Y$ and for each pair $f(x_i,y_i)$, the model gives as output a number that represents the confidence that label $y_i$ characterizes $x_i$.
-Traditional single-label learning maps an example to only one label, ignoring completely multiple semantics that real-objects usually have.
-On the other hand, multi-label learning allows an example/instance $x_i$ to be mapped to more than one label.
-However, this results to an exponentially growing search space, since search space might include any number of subsets from the powerset of label-set $Y$ ($|\mathbb{P(Y)}|$).
+	The paper "A Review on Multi-Label Learning Algorithms" analyzes the
+  multi-label learning field in supervised classification. Supervised
+  classification trains a model $f: X \rightarrow Y$ and for each pair
+  $f(x_i,y_i)$, it outputs a real number $r$ that represents the confidence that
+  label $y_i$ characterizes $x_i$. Given $r$ and a thresholding scheme, labels
+  are marked as relevant or not relevant to the input data vector $x_i$ (e.g.
+  for $r$ representing probability estimates, $r \in [0,1]$, a simple
+  thresholding function is $t(x_i) = 0.5$. In other words, labels with outputs
+  greater than $0.5$ are assigned to $x_i$ by our model). Standard, single-label learning maps an
+  example to a single label, completely ignoring multiple semantics that
+  real-objects usually have. On the other hand, multi-label learning allows an
+  example/instance $x_i$ to be mapped to more than one label, i.e. to any label
+  vector which is a subset of $Y$. However, this results to an exponentially growing search space, since
+  the latter depends on the size of the powerset of $Y$, 
+  $|\mathbb{P(Y)}| = 2^{|Y|}$.
 
-	One straightforward solution, is to exploit label correlations.
-Real-objects are usually characterized by similar concepts.
-For example, if an article is characterized by label "football" it is very likely that it will be also characterized by the term "sports", while label "science" seems to be irrelevant.
-In order to exploit such kind of label correlations, three different algorithm strategies were developed:
+	One straightforward solution is to exploit label correlations. Real objects
+  are usually characterized by similar concepts. For example, if an article is
+  characterized by label "football" it follows that it will be also
+  characterized by the term "sports" and it is likely that it will be labeled
+  with ``crowd'' or ``entertainment''. On the other hand, the label "science"
+  will be likely to be irrelevant. In order to exploit such kind of label
+  correlations, three different algorithm strategies are examined:
 
 	\begin{itemize}
-		\item \textbf{First-order strategies: }These algorithms ignore label correlations and perform a "one-vs-all" strategy.
-They often transform a multi-label problem into multiple single-label problems.
-Algorithms of this category are simple, scalable, have the possibility of parallel implementation, but on the other hand their results are suboptimal.
-		\item \textbf{Second-order strategies: }In this category, relevant algorithms consider pairwise label correlations, having a good trade-off between generalization performance and scalability.
-However, there are still deficiencies in some real-world applications.
-		\item \textbf{High-order strategies: }These algorithms capture even more complicated label correlations and they have strong modeling capabilities.
-But, on the other hand, they are computationally more demanding and less scalable.
+		\item \textbf{First-order strategies: }These algorithms ignore label
+      correlations and perform a "one-vs-all" strategy. They often transform a
+      multi-label problem into multiple single-label problems, combining the
+      result in an aggregation-like manner. Algorithms of this category are
+      simple, scalable, have the capability of parallel implementation, but
+      offer suboptimal performance.
+		\item \textbf{Second-order strategies: }Algorithms in this category consider
+      \emph{pairwise} label correlations, i.e. tuples of labels. Having a good
+      trade-off between generalization performance and scalability, they
+      generally outperform first-order strategies but their limited exploration
+      of the label dependency space leads to lacking performance in some real
+      world applications.
+		\item \textbf{High-order strategies: }These algorithms capture a high degree
+      of label correlations, having a large comlexity and strong modeling
+      capabilities. This of course leads to increased performance but renders them computationally demanding and less scalable.
 	\end{itemize}
@@
 	\addcontentsline{toc}{subsection}{Evaluation metrics}
-	One other issue that worths to be mentioned are the evaluation metrics.
-Since multi-label problems are more complicated, it is necessary to extend existing sigle-label metrics in order to fit the multi-label case.
-There are two categories of metrics, divided in two perspectives:
+
+	With respect to performance evaluation, traditional single-label evaluation metrics are extended to
+  incorporate the multi-labeled output. This extension is categorizable into two
+  groups. One with respect to the expansion ``axis'' (across examples or labels)
+  and the task type (classification or ranking).
+
+    Before proceeding with the metrics, true positive(TP), false positive(FP),
+    true negative (TN) and false negative (FN) of label $y_j$ will be defined:
+		\begin{itemize}
+			\item \emph{True positive: } $TP_j = |\{x_i | y_j \in Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|$
+			\item \emph{False positive: }$FP_j = |\{x_i | y_j \notin Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|$
+			\item \emph{True negative: }$TN_J = |\{x_i | y_j \notin Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|$
+			\item \emph{False negative: }$FN_J = |\{x_i | y_j \in Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|$
+		\end{itemize}
+
 	\begin{itemize}
@@
 The metrics are divided into two perspectives:
+      \todos{Form them up into a table}
 		\begin{itemize}
-			\item \textbf{Classification:} 
+			\item Classification:
+        \begin{table}
+          \centering
+          \begin{tabular}{cc} 
+          \toprule
+            one & two \\
+          \midrule
+            one & two \\
+            one & two \\
+          \bottomrule
+          \end{tabular}
+          \caption{Evaluation metrics for multi-label learning}
+          \label{label:evaluation_metrics}
+        \end{table}
+
 			\begin{itemize}
-				\item \emph{Subset Accuracy: } $subsetacc(h) = \frac{1}{p} \sum_{i=1}^p \llbracket h(x_i) = Y_i \rrbracket$
-				\item \emph{Hamming Loss: }$hloss(h) = \frac{1}{p} \sum_{i=1}^{p} [h(x_i)\Delta Y_i] $
-				\item \emph{Accuracy:} $Accuracy_{exam}(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i \cup h(x_i)|}$
-				\item \emph{Precision: }$Precision_{exam}(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{h(x_i)|}$
-				\item \emph{Recall: }$Recall_{exam}(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i}$
-				\item \emph{$F^ \beta$ -metric: } $F^ \beta _{exam} (h) = \frac{(1+ \beta ^2) \cdot Precision_{exam}(h) \cdot Recall_{exam}(h)}{\beta ^2 \cdot Precision_{exam}(h) + Recall_{exam}(h)}$
+				\item Precision: $P(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{h(x_i)|}$
+				\item Recall: $R(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i}$
+        \item Subset Accuracy:  $SA(h) = \frac{1}{p} \sum_{i=1}^p \llbracket h(x_i) = Y_i \rrbracket$
+				\item Hamming Loss: $H(h) = \frac{1}{p} \sum_{i=1}^{p} [h(x_i)\Delta Y_i] $
+				\item Accuracy: $A(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i \cup h(x_i)|}$
+				\item $F^ \beta$ -metric:  $F^ \beta _{exam} (h) = \frac{(1+ \beta ^2) \cdot Precision_{exam}(h) \cdot Recall_{exam}(h)}{\beta ^2 \cdot Precision_{exam}(h) + Recall_{exam}(h)}$
 			\end{itemize}
-			\item \textbf{Ranking:}
+			\item Ranking:
 			\begin{itemize}
-				\item \emph{One-error: }$one-error(f) = \frac{1}{p} \llbracket [argmax_{y \in Y}f(x_i,y)] \notin Y_i \rrbracket$
-				\item \emph{Coverage: } $coverage(f) = \frac{1}{p} \sum_{i=1}^{p} max_{y \in Y_i} rank_f(x_i,y) - 1$
-				\item \emph{Ranking loss: }$rloss(f) = \frac{1}{p} \sum_{i=1}^{p} $ \todo{complete it}
-				\item \emph{Average Precision: } \todo{average precision}
+				\item One-error: $one-error(f) = \frac{1}{p} \llbracket [argmax_{y \in Y}f(x_i,y)] \notin Y_i \rrbracket$
+				\item Coverage:  $coverage(f) = \frac{1}{p} \sum_{i=1}^{p} max_{y \in Y_i} rank_f(x_i,y) - 1$
+				\item Ranking loss: $rloss(f) = \frac{1}{p} \sum_{i=1}^{p} \frac{1}{|Y_i| {\overset{\sim}{|Y_i|}}} |\{(y', y'') | f(x_i,y') \leq f(x_i,y''), (y', y'') \in Y_i \times {\overset{\sim}{|Y_i|}} \}| $ 
+				\item Average Precision:  $agvprec(f) = \frac{1}{p} \sum_{i=1}^{p} \frac{1}{|Y_i|} \sum_{y \in Y_i} \frac{|\{y' | rank_f(x_i,y') \leq rank_f(x_i,y), y' \in Y_i\}|}{rank_f(x_i,y)}$
 			\end{itemize}
@@
   \item \textbf{Label-based: }
+		The label-based metrics are also separated in two perspectives:
 		\begin{itemize}
-			\item \todo{label based metrics}
+			\item Classification: Let $B(TP_j, FP_j, TN_J, FN_J)$ represent some specific binary classification metric $(B \in \{Accuracy, Precision, Recall, F^ \beta \})$.
+The metrics from classification perspective are the following:
+			\begin{itemize}
+				\item Macro-averaging: $B_{macro}(h) = \frac{1}{q} \sum_{1}^{q} B(TP_j, FP_j, TN_J, FN_J)$
+				\item Micro-averaging: $B_{micro}(h) = B(\sum_{1}^{q}TP_j, \sum_{1}^{q}FP_j, \sum_{1}^{q}TN_j, \sum_{1}^{q}FN_j)$
+			\end{itemize}
+			\item Ranking:
+			\begin{itemize}
+				\item \emph{AUC macro: }$AUC_{macro} = \frac{1}{q} \sum_{1}^{q} \frac{|\{(x', x'') | f(x',Y_j) \geq f(x'',Y_j), (x',x'') \in Z_j \times \overset{\sim}{|Z_j|} \}|}{|Z_j|{\overset{\sim}{|Z_j|}}}$
+				\item \emph{AUC micro: }$AUC_{micro} = \frac{|\{(x',x'',y',y'') | f(x',y') \geq f(x'',y''), (x',y') \in S^+, (x'',y'') \in S^- \}|}{|S^+||S^-|}$
+			\end{itemize}
 		\end{itemize} 
@@
 	\addcontentsline{toc}{section}{Multi-label learning algorithms}
-	In this paper, 8 representative multi-label algorithms are presented and analyzed, seleceted with respect to below criteria:
+	In this paper, eight representative multi-label algorithms are presented and analyzed, seleceted with respect to below criteria:
 	\begin{itemize}
-		\item[$\checkmark$] Broad spectrum (unique characteristics)
-		\item[$\checkmark$] Primitive impact (lead to a number follow-up related methods)
-		\item[$\checkmark$] Favorable influence (highly-cited) 
+		\item[$\checkmark$] Broad, noteworthy and/or unique characteristics
+		\item[$\checkmark$] Primitive impact (i.e. leads to a number follow-up related methods)
+		\item[$\checkmark$] Favorable influence (popular and/or highly-cited) 
 	\end{itemize}
@@
 	\begin{itemize}
-		\item \textbf{Problem Transformation Methods: }Algorithms of this category, in order to handle a multi-label problem, transform it into other well-established learning scenarios ("Fit data to algorithm" philosophy)
+		\item \textbf{Problem Transformation Methods: }Algorithms of this category
+      transform the multi-labeled problem into into other well-established
+      single-labeled learning scenarios, fusing the results into a multi-label output (``Fit data to algorithm'' philosophy)
 		\item \textbf{Algorithm Adaptation Methods: }Here, algorithms adapt popular learning techniques to deal with multi-label data directly ("Fit algorithm to data" philosophy)
@@
 	\addcontentsline{toc}{subsubsection}{Binary Relevance}
-	This algorithm decomposes the multi-learning problem into $q$ independent binary classification problems, each of them corresponding to a possible label in the label-set.
-First task is to construct a corresponding binary training set and then to use a binary learning algorithm in order to train the relevant binary classifier.
-Each training instance is involved in the learning process of all $q$ classifiers(cross-training strategy), which is regarded as positive instance in case of relevant labels and irrelevant otherwise.
-For new instances, the algorithm predicts the relevant label set by querying each binary classifier and combining the relevant labels.
+	Binary relevance (BR) decomposes the multi-learning problem into $q$
+  independent binary classification problems, each of them corresponding to a
+  possible label in the label-set $Y$. The first task is to construct a
+  corresponding binary training set for each label, on which a binary learning
+  algorithm is used in order to train the binary classifier that discriminates
+  class $y_j$, i.e. $h_j(\cdot)$. Each training instance is involved in the
+  learning process of all $q$ classifiers(cross-training strategy), which is
+  regarded as positive instance in case of relevant labels and negative
+  otherwise. For new instances, the algorithm predicts the relevant labelset by
+  querying each binary classifier and combining the relevant labels. A label is
+  considered relevant according to the thresholding scheme adopted in the classifier.
 
-	This is a first-order approach algorithm, since it ignores completely any label associations.
-However, since the one-vs-rest scheme is used and classifiers are build for each label separately, the algorithm has the possibility of parallel implementation.
-Finally, the algorithm is sensitive to class-imbalance data (true positives are much less than false positives).
+	This is a first-order approach algorithm, since it ignores completely any
+  label associations. However, since an one-vs-rest scheme is used and
+  classifiers are built for each label separately, the algorithm has the
+  capability of parallel implementation. Finally, this separation renders the
+  algorithm sensitive to class-imbalance in the dataset\footnote{The case where
+    the number of positives and negatives for a class differs significantly}.
 
@@
 	\addcontentsline{toc}{subsubsection}{Classifier Chains}
-	The logic of this algorithm is to transform the multi-label learning problem into a chain of binary classification problems, using subsequent binary classifiers in the chain, built upon the predictions of preceding ones and ordered by a permutation function.
-First of all, a binary training set is constructed by enriching each instance with the confidence of preceding classifiers.
-Moreover, a binary algorithm is used to determine whether $y_i$ label is relevant or not for an instance.
-For unknown instances, the algorithm predicts the relevant labels by iteratively traversing the classifier chain ($\lambda _{\tau(j)}^x \in {-1, +1}$ represents the predicted binary assignment of $y_{\tau(j)}$)
+	The logic of this algorithm is to transform the multi-label learning problem
+  into a chain of binary classification problems, using subsequent binary
+  classifiers in a chain, built upon the predictions of preceding ones and
+  ordered by a permutation function $f_p$. First of all, a binary training set
+  is constructed by enriching each instance with the confidence of preceding
+  classifiers, which are concatenated with the data vector $x_i$. Moreover, a binary algorithm is used in step $i$ to
+  determine whether label $y_i$ label is relevant or not for an instance. For
+  unknown instances the algorithm predicts the relevant labels by iteratively
+  traversing the classifier chain producing thresholded confidence scores ($\lambda _{\tau(j)}^x \in {-1, +1}$ in the
+  same order as within the training process by $f_p$ represents the predicted binary assignment of $y_{\tau(j)}$)
 
@@
 	This algorithm transforms the learning problem into a label pairwise comparison ranking problem.
-For $q$ labels, $q(q-1)/2$ binary classfiers are generated by pairwise comparison and a binary algorithm $g_{jk}(x)$ is used.
+For $q$ labels, $q(q-1)/2$ binary classfiers are generated by pairwise comparison and a binary algorithm $h_{jk}(x)$ is used.
 Firstly, training sets $D_{jk}$ are constructing, where $D_{jk}: \{x_i, Y_i | y_j \in Y_i \oplus y_k \in Y_i\}$.
-The learning system votes for each example and if $g_{jk}(x)>0$, then $x_i$ is associated with label $y_j$, otherwise $y_j$ is irrelevant and $x_i$ is assigned to $y_k$.
+The learning system votes for each example and if $h_{jk}(x)>0$, then $x_i$ is associated with label $y_j$, otherwise $y_j$ is irrelevant and $x_i$ is assigned to $y_k$.
 For unknown instances, classifiers' votes are aggregated and ranked.

### 2018-01-17 Paper/PaperReport/PaperReport.tex
paper report update

--- 2018-01-17 Paper/PaperReport/PaperReport.tex
@@
 
-	\subsubsection*{Claibrated Label Ranking}
-	\addcontentsline{toc}{subsubsection}{Claibrated Label Ranking}
+	\subsubsection*{Calibrated Label Ranking}
+	\addcontentsline{toc}{subsubsection}{Calibrated Label Ranking}
 	This algorithm transforms the learning problem into a label pairwise comparison ranking problem.

### 2018-01-17 Paper/Presentation/mll_review.tex

@@
 Solution: integrate in the learning process potential label correlations.
-This work examines algorithms grouped in three broad categories:
+
+In this work the authors group M-L algorithms in three categories:
 
@@
 \item[$\circ$] Often transform M-L problem to multiple, single-label problems
+  and combine the per-label results
 \item[$\circ$] Simple, scalable, suboptimal
@@
 \item[$\circ$] Capture more complicated label interdependencies
-\item[$\circ$] Strong modeling capabilities
+\item[$\circ$] Strong modeling capabilities that can capture complex relationships
 \item[$\circ$] Computationally demanding, less scalable
@@
 
-Two categories and perspectives:
+Grouped into two categories and perspectives:
 \begin{itemize}
@@
 \item[$\star$] Subset Accuracy, Hamming Loss
-\item[$\star$] Precision, Recall, F-measure, F$^\beta$ -measure
+\item[$\star$] Precision, Recall,  F$^\beta$ -measure
 \end{itemize}
@@
 \begin{itemize}
-\item[$\circ$] Classification perspective: Macro/Micro averaging for single-label example-based, classification-based measures
+\item[$\circ$] Classification perspective: Macro/Micro averaging techniques for single-label example-based, classification-perspective measures
 \item[$\circ$] Ranking perspective: Macro/Micro averaging for AUC
@@
 \end{itemize}
-$\ast$ Classifiers should aim to optimize \emph{multiple} metrics
+$\ast$ Ideally, classifiers should be trained to optimize \emph{multiple} metrics
 \end{frame}
@@
 \begin{itemize}
-\item[$\circ$] Transform the learning problem into other well-established learning scenarios
+\item[$\circ$] Transform the learning problem into other, managable (often
+  single-label) learning
+  problems 
 \item[$\circ$] ``Fit data to algorithm'' philosophy
@@
 
-Include algorithms that:
+Authors include algorithms that:
 \begin{itemize}
 \item[$\checkmark$] Has broad, noteworthy or unique characteristics
-\item[$\checkmark$] Has primitive impact, i.e. leads to a number follow-up related methods
+\item[$\checkmark$] Has important impact, leading to a number follow-up related methods
 \item[$\checkmark$] Is influential and highly-cited in multi-label learning
@@
 \begin{itemize}
-\item[$\bullet$] Decompose multi-label problem to $q$ independent binary classification problems
+\item[$\bullet$] Decompose multi-label problem to $|Y|=q$ independent binary classification problems
 \item[$\bullet$] Construct $q$ binary (one-vs-all) training sets (one for each $y_i$)
-\item[$\bullet$] Cross-train each binary classifier $h_i(x)$ on its respective dataset
+\item[$\bullet$] Independently train each classifier $h_i(x)$ on its respective dataset
 
 \item[$\bullet$] Predict labels of an unseen $x$ by evaluating each classifier $h_i(x)$
-\item[$\bullet$] Assign label $y_i$ according to $h_i(x)$ and the thresholding/ranking scheme 
+\item[$\bullet$] Assign label $y_i$ according to $r = h_i(x)$ and the
+  thresholding setting 
 \end{itemize}
@@
 \begin{itemize}
-\item[$\bullet$] Simple, one-vs-rest scheme, parallelizable
-\item[$\bullet$] Sensitive to class-imbalanced data
+\item[$\bullet$] Simple, one-vs-rest scheme $\rightarrow$ easily parallelizable
+\item[$\bullet$] Sensitive to class-imbalanced data\footnote{Very different
+    number of pos. and neg. examples for a class}
 \item[$\bullet$] Ignores label correlations (first-order method)
@@
 \begin{itemize}
-\item[$\bullet$] A permutation function $f_p$ is used to order $Y$
+\item[$\bullet$] Reorder the label $Y$ set using a permutation function $f_p$ 
 \item[$\bullet$] Transform into a \emph{chain} of binary classification problems
@@
 \item[$\bullet$] Predict relevant labels for unknown instances by iteratively traversing the classifier chain
-\item[$\bullet$] $\lambda_{\tau(j)}^x \in \{-1,+1\}$ represents the predicted binary assignment
-of $y_{\tau(j)}$, which is recursively obtained
 \end{itemize}
@@
 \item[$\bullet$] High-order method: exploitation of label correlations to a degree, but in a random manner
-\item[$\bullet$] Highly sensitive to $f_p$. Ensemble schemes attempt to overcome it.
+\item[$\bullet$] Highly sensitive to $f_p$. Running multiple chain executions in an ensemble
+  attempt to overcome this dependency.
 \item[$\bullet$] Iterative operation prevents parallel implementation
@@
 \item[$\bullet$] Transform into a label pairwise comparison ranking problem
-\item[$\bullet$] For $q$ labels, generate $q(q-1)/2$ binary classifiers by pairwise comparison and use a binary algorithm $g_{jk}(x)$
+\item[$\bullet$] For $q$ labels, generate $q(q-1)/2$ binary classifiers by pairwise comparison and use a binary algorithm $h_{jk}(x)$
 \item[$\bullet$] Construct training sets $D_{jk} : \{x_i, Y_i | y_j \in Y_i \oplus y_k \in Y_i\}$.
-System votes for each example and if $g_{jk}(x) >0$,$x_i$ is associated with $y_j$, otherwise with $y_k$
-\item[$\bullet$] For an unknown instance, all classifiers' votes are aggregated and ranked
-\item[$\bullet$] A virtual label $y_v$ serves as the artificial splitting point between relevant and irrelevant labels
+System votes for each example and if $h_{jk}(x) >0$,$x_i$ is associated with $y_j$, otherwise with $y_k$
+\item[$\bullet$] For an unknown instance, all classifiers' votes are aggregated
+  and resulting labels are ranked according to the total confidence $r$
+\item[$\bullet$] A virtual label $y_v$ is learned as a threshold, serving as the artificial splitting point between relevant and irrelevant labels
 \end{itemize}
@@
 \item[$\bullet$] Second-order approach algorithm. One-vs-one scheme.
-\item[$\bullet$] Smooths out the class-imbalance problem
+\item[$\bullet$] Pairwise comparison smooths out the class-imbalance problem
 \item[$\bullet$] Number of classifiers is quadratic to $|Y|$ (compared to linear for BR)
@@
 \item[$\bullet$] Decompose to an ensemble of multi-class classification problems
-\item[$\bullet$] Each component targets a random subset of $\mathbb{P}(Y)$ that appears in $X$, classified with Label Powerset (LP) techniques:
+\item[$\bullet$] Each component targets a random subset of $\mathbb{P}(Y)$ (that
+  also appears in $X$), classified with Label Powerset (LP) techniques:
 \begin{itemize}
-\item[$\circ$] Transform to single-label data by treating each distinct labelset as a new class
+\item[$\circ$] Transform to single-label data by treating each distinct labelset
+  as a new class $\rightarrow$ multi to single label problem
 \item[$\circ$] Each example is reassigned with the new mapped class and classified through regular single-label classification
@@
 \end{itemize}
-todo how does each LP classify, k-labelsets
 \end{itemize}
@@
 \item[$\bullet$] Compares the MAP probabilities: $P(K|C_j), K \in \{H_j, \neg H_j\}$ to decide if to include $y_j$ in the prediction. Compute with the Bayes' theorem.
-\item[$\bullet$] Priors $P(K)$ computed by smoothed frequency counting in the training data
+\item[$\bullet$] Single-label priors $P(K)$ computed by smoothed frequency counting in the training data
 \item[$\bullet$] Likelihoods $P(C_j|K)$ are computed as a function of:
@@
 \item[$\bullet$] Decision boundary can be modified on-line via new instances
-\item[$\bullet$] Class imbalance can be overcome by calculating priors
+\item[$\bullet$] M-L class imbalance can be overcome by calculating the priors
 \item[$\bullet$] Extensions proposed for label correlation exploitations
@@
 \item[$\bullet$] Multi-label entropy is used to build a decision tree recursively
-\item[$\bullet$] Split $x$ at feature $l$ s.t. the information gain criterion is maximized
+\item[$\bullet$] Split $x$ at the $l$-th feature\footnote{$x$ has to contain distinct features} s.t. the information gain criterion is maximized
 \item[$\bullet$] Dataset $T$ is split on the $l$-th feature valued $\theta$ to branches ($T^-$ \& $T^+$), where $ x_{il} \leq \theta, x_i \in T^-$ and $ x_{il} > \theta, x_i \in T^+$
@@
 	\item[$\bullet$] $q$ linear classifiers $h_i(x)$, optimized with the empirical ranking loss
-\item[$\bullet$] Relevant-nonrelevant label pairs of $x_i$: $(y_j,y_k) \in Y_i
-  \times \bar Y_i$: decision hyperplane\footnote{w.r.t. singed classifier distance} is: $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$
-\item[$\bullet$] Separation margin is the minimum distance of $x_i$ from every  $Y_i \times \bar Y_i$
+\item[$\bullet$] Consider signed distance of $x_i$ from the boundary of every
+  relevant-nonrelevant label pair $(y_j,y_k) \in Y_i \times \bar Y_i$. The
+  pair's decision hyperplane is: $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$
+\item[$\bullet$] The M-L margin is the minimum distance of $x_i$ from every  $Y_i \times \bar Y_i$
 \item[$\bullet$] SVM $\rightarrow$ minimize loss \emph{and} maximize margin.
@@
 
-\item[$\bullet$] Ranks labels w.r.t. singled distance
+\item[$\bullet$] Ranks labels according to the singed distance
 \item[$\bullet$] Non-linearity achievable through feature mapping and the kernel trick
@@
 %------------------------------------------------
-\section{Related Learning Settings}
+\section{Related Learning Methods}
 %------------------------------------------------
@@
 \item[$\bullet$] Multi-label learning representative algorithms
-\item[$\bullet$] Related learning settings
+\item[$\bullet$] Related learning methods
 \end{itemize}
@@
 \begin{itemize}
-\item[$\bullet$] Holy grail: Formal characterization on the underlying concept / mechanism on the appropriate usage of label correlations, especially on large output spaces
+\item[$\star$] Formal characterization on the underlying concept / mechanism on the appropriate usage of label correlations, especially on large output spaces
 \item[$\bullet$] Thorough experimental comparative study to discover pros and cons of different multi-label learning algorithms
@@
 %------------------------------------------------
+
+\subsection{Additional Bibliography}
 \begin{frame}
 \frametitle{\insertsection : \insertsubsection}
-Useful additional bibliography
 \begin{itemize}

### 2018-01-18 Paper/Presentation/mll_review.tex
Update presentation

@@
 \begin{itemize}
-\item[$\star$] Subset Accuracy, Hamming Loss
-\item[$\star$] Precision, Recall,  F$^\beta$ -measure
+\item[$-$] Subset Accuracy, Hamming Loss
+\item[$-$] Precision, Recall,  F$^\beta$ -measure
 \end{itemize}
@@
 \begin{itemize}
-	\item[$\bullet$] Let $H_j \equiv (y_j \in Y_i)$, $C_j = \sum_{x_k \in N(x_i)} [\delta(y_j \in Y_k)]$
+	\item[$\bullet$] Consider events $H_j \equiv (y_j \in Y_i)$, $C_j \equiv (\sum_{x_k \in N(x_i)} [\delta(y_j \in Y_k)])$
 \item[$\bullet$] Compares the MAP probabilities: $P(K|C_j), K \in \{H_j, \neg H_j\}$ to decide if to include $y_j$ in the prediction. Compute with the Bayes' theorem.
@@
 \begin{itemize}
-\item[$\circ$] $\kappa(r)$, the number of examples labelled $y_j$ with $r$ neighbours labelled $y_j$
-\item[$\circ$] $\overset{\sim}{\kappa}(r)$, the number of examples not labelled $y_j$ with $r$ neighbours labelled $y_j$
+\item[$\circ$] $\kappa_j(r)$, the number of examples labelled $y_j$ with $r$ neighbours labelled $y_j$
+\item[$\circ$] $\overset{\sim}{\kappa_j}(r)$, the number of examples not labelled $y_j$ with $r$ neighbours labelled $y_j$
 \end{itemize}
@@
 \item[$\bullet$] Multi-label entropy is used to build a decision tree recursively
-\item[$\bullet$] Split $x$ at the $l$-th feature\footnote{$x$ has to contain distinct features} s.t. the information gain criterion is maximized
-\item[$\bullet$] Dataset $T$ is split on the $l$-th feature valued $\theta$ to branches ($T^-$ \& $T^+$), where $ x_{il} \leq \theta, x_i \in T^-$ and $ x_{il} > \theta, x_i \in T^+$
+\item[$\bullet$] Split node at the $l$-th feature of $x$ such that the information gain
+  criterion is maximized. Node partitions data according to $x_l = \theta$:
+ $T \rightarrow \{T^-,T^+\}$, where $ x_{il} \leq \theta, x_i \in T^-$ and $ x_{il} > \theta, x_i \in T^+$
 \item[$\bullet$] Recurse on subtrees until a stopping criterion is met (e.g. child size)
@@
 	\item[$\bullet$] $q$ linear classifiers $h_i(x)$, optimized with the empirical ranking loss
+	\item[$\bullet$] Combine classifiers to discriminate label \emph{pairs}. Label
+    pair $(j, k)$ decision hyperplane is: $h_{jk} = g(h_j, h_k) = \left<w_j-w_k, x_i \right> + b_j - b_k=0$
 \item[$\bullet$] Consider signed distance of $x_i$ from the boundary of every
-  relevant-nonrelevant label pair $(y_j,y_k) \in Y_i \times \bar Y_i$. The
-  pair's decision hyperplane is: $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$
+  relevant-nonrelevant label pair $(y_j,y_k) \in Y_i \times \bar Y_i$.
+  % The pair's decision hyperplane is: $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$
 \item[$\bullet$] The M-L margin is the minimum distance of $x_i$ from every  $Y_i \times \bar Y_i$
 \item[$\bullet$] SVM $\rightarrow$ minimize loss \emph{and} maximize margin.
-  Approximate $max(\cdot)$ and introduce slack variables.
 
@@
 \item[$\bullet$] Second-order approach, maximum margin strategy
-\item[$\bullet$] Convex quadratic programming problem, solved by any QP solver
-\item[$\bullet$] Extensible by kernel learning: kernel selection can be overcome by Multiple Kernel Learning techniques
+\item[$\bullet$] Optimization is a convex QP problem, solved by any QP solver
+\item[$\bullet$] Kernel SVM, kernel selection can be done with MKL techniques
 \item[$\bullet$] Adaptable learning by selecting an appropriate loss function
@@
 \begin{itemize}
-\item[$\bullet$] Encodes labels as binary random vectors $\bm{y}=\{-1,1\}\}^q$
-\item[$\bullet$] For data $x,\bm{y}$ learn joint probability distribution $p(x,\bm{y})$
-\item[$\bullet$] Learn pdf $p(x,\bm{y})$ via the principle of maximun entropy
+\item[$\bullet$] A Conditional Random Field model for M-L classification
+\item[$\bullet$] Encodes $y$ as binary random vectors $\bm{y}=\{-1,1\}^q$.
+  Learn joint probability distribution $p(x,\bm{y})$
+\item[$\bullet$] Compute conditional $p(\bm{y} | x)$ with the maximum entropy criterion:
   \begin{itemize}
     \item Maximize $H_p(x,y)$ subject to $K$ constraints $\mathbb{E}[f_k(x,\bm{y})] = F_k$
-    \item $f_k$ models label correlations with $F_k$ learned from the training data
+    \item $f_k(\cdot)$ directly model label correlations - the $F_k$ values are learned from the training data
+  \item[$\bullet$] Solve with Lagrange multipliers, assuming gaussian priors
+  \item[$\bullet$] Optimal solution is in the form of a Gibbs probability distribution
   \end{itemize}
-\item[$\bullet$] Optimize via Lagrange multipliers, arrive at $p(\bm{y} | x)$
-\item[$\bullet$] Unseen instances labelled with $Y = \underset{y}{\text{argmax }} p(y | x)$
+%\item[$\bullet$] Unseen instances labelled with $Y = \underset{y}{\text{argmax }} p(y | x)$
+\item[$\bullet$] Unseen instances labelled with $Y = \text{argmax}_y\text{ } p(y | x)$
 \end{itemize}
@@
 \item[$\bullet$] Intractable $\text{argmax}$ operation for a large label space without pruning.
+\item[$\bullet$] Extensible: Different forms of constraints yield various CML variants
 \end{itemize}
@@
 \item[$\bullet$] Multi-instance learning
-  % https://en.wikipedia.org/wiki/Multiple-instance_learning
 \begin{itemize}
@@
 \item[$\bullet$] Ordinal classification
-  % https://en.wikipedia.org/wiki/Ordinal_regression
 \begin{itemize}
-\item[$\circ$] Assume labels are related through a natural ordering
-\item[$\circ$] Instead of explicit class relevance, assume an ordering of
-  relevance between labels via a membership vector
-\item[$\circ$] Transform ML problem to a set of ordinal set of problems
+\item[$\circ$] Assume label relevance is not binary, but soft
+\item[$\circ$] Produce a vector of ordinal graded membership
+\item[$\circ$] Transform M-L problem to a set of ordinal set of problems
 \end{itemize}
@@
 \begin{itemize}
+\item[$\bullet$] Vens, Celine, et al. "Decision trees for hierarchical multi-label classification." Machine Learning 73.2 (2008): 185-214.
+\item[$\bullet$] Tsoumakas, Grigorios, and Ioannis Vlahavas. "Random k-labelsets: An ensemble method for multilabel classification." Machine learning: ECML 2007 (2007): 406-417.
+\item[$\bullet$] Zhang, Min-Ling, and Zhi-Hua Zhou. "ML-KNN: A lazy learning approach to multi-label learning." Pattern recognition 40.7 (2007): 2038-2048.
 \item[$\bullet$] Elisseeff, André, and Jason Weston. "A kernel method for multi-labelled classification." Advances in neural information processing systems. 2002.
@@
 ACM, 2005.
+\item[$\bullet$]  Cheng, Weiwei, Eyke Hüllermeier, and Krzysztof J.
+Dembczynski.
+"Graded multilabel classification: The ordinal case." Proceedings of the 27th international conference on machine learning (ICML-10).
+2010.
 \end{itemize}

### 2018-01-18 Paper/Presentation/mll_review.tex
update presentation

@@
 \frametitle{\insertsection : \insertsubsection}
-Learning $\rightarrow$ build a model from data, to accomplish a task
+\onslide<2-> {Learning $\rightarrow$ build a model from data, to accomplish a task}
 \begin{itemize}
-  \item Supervised: we have both data \emph{and labels}
-\item Applied to classification in this study
+  	\item <3-> Supervised: we have both data \emph{and labels}
+	\item <4-> Applied to classification in this study
 \end{itemize}
-Supervised classification:
+\onslide <5-> {Supervised classification:}
 \begin{itemize}
-\item[$\bullet$] Given input data $X$ and labels  $Y$,
+\item[$\bullet$] <6-> Given input data $X$ and labels  $Y$,
   learn the function $f: X \rightarrow Y$ 
-\item[$\bullet$] $f(x_i, y_i) = r$ where $r\in \mathbb{R}$ is the confidence that $y_i$ characterizes $x_i$
-\item[$\bullet$] Assume that $x_i$ belongs to $y_i$ if $r \ge $ $t(x_i)$
+\item[$\bullet$] <7-> $f(x_i, y_i) = r$ where $r\in \mathbb{R}$ is the confidence that $y_i$ characterizes $x_i$
+\item[$\bullet$] <8-> Assume that $x_i$ belongs to $y_i$ if $r \ge $ $t(x_i)$
 \begin{itemize}%[label={$\checkmark$}]
-\item[$\checkmark$] $t(\cdot)$ can be a predetermined constant function or learned from $X$
+\item[$\checkmark$] <9-> $t(\cdot)$ can be a predetermined constant function or learned from $X$
 \end{itemize}
@@
 
-Single-label learning
+\onslide <10-> {Single-label learning}
 \begin{itemize}
-\item[$\bullet$] Dataset $\{ (x_i, y_i)\}, i = 1, \dots N, x \in X, y \in Y=\{y_1,\dots,y_q\}$
+\item[$\bullet$] <11-> Dataset $\{ (x_i, y_i)\}, i = 1, \dots N, x \in X, y \in Y=\{y_1,\dots,y_q\}$
 \end{itemize}
 
-Multi-label learning
+\onslide <12->{Multi-label learning}
 \begin{itemize}
-\item[$\bullet$] Multi-label dataset: Multiple labels per instance. $\{ (x_i, \bm{y}_i)\}, i = 1, \dots N, x \in X, \bm{y} \in \mathbb{P}(Y)$
+\item[$\bullet$] <13-> Multi-label dataset: Multiple labels per instance. $\{ (x_i, \bm{y}_i)\}, i = 1, \dots N, x \in X, \bm{y} \in \mathbb{P}(Y)$
 % \item Learn a model $f: X \times Y \rightarrow r$
@@
 \frametitle{\insertsection : \insertsubsection}
-Label search space $\mathbb{S_Y}$ grows exponentially as a function of $|Y|=q$
+\onslide <2->{Label search space $\mathbb{S_Y}$ grows exponentially as a function of $|Y|=q$}
 \begin{itemize}
-\item[$\triangleright$] e.g. for $q=20, |\mathbb{S_Y}| = 2 ^ {|\mathbb{P(Y)}|} = 2^{20} \ge 10^6$
+\item[$\triangleright$] <3-> e.g. for $q=20, |\mathbb{S_Y}| = 2 ^ {|\mathbb{P(Y)}|} = 2^{20} \ge 10^6$
 \end{itemize}
-Solution: integrate in the learning process potential label correlations.
+\onslide <4->{Solution: integrate in the learning process potential label correlations.
 
-In this work the authors group M-L algorithms in three categories:
+In this work the authors group M-L algorithms in three categories:}
 
 \begin{enumerate}
-\item First-order strategies
+\item <5->First-order strategies
 \begin{itemize}
-\item[$\circ$] Ignore label correlations
-\item[$\circ$] Often transform M-L problem to multiple, single-label problems
+\item[$\circ$] <6-> Ignore label correlations
+\item[$\circ$] <7-> Often transform M-L problem to multiple, single-label problems
   and combine the per-label results
-\item[$\circ$] Simple, scalable, suboptimal
+\item[$\circ$] <8-> Simple, scalable, suboptimal
 \end{itemize}
-\item Second-order strategies
+\item <9-> Second-order strategies
 \begin{itemize}
-\item[$\circ$] Consider \emph{pairwise} label relations
-\item[$\circ$] Good trade-off between generalization performance and scalability
-\item[$\circ$] Lacking in some real-world applications
+\item[$\circ$] <10-> Consider \emph{pairwise} label relations
+\item[$\circ$] <11-> Good trade-off between generalization performance and scalability
+\item[$\circ$] <12-> Lacking in some real-world applications
 \end{itemize}
-\item Higher-order strategies
+\item <13->Higher-order strategies
 \begin{itemize}
-\item[$\circ$] Capture more complicated label interdependencies
-\item[$\circ$] Strong modeling capabilities that can capture complex relationships
-\item[$\circ$] Computationally demanding, less scalable
+\item[$\circ$] <14-> Capture more complicated label interdependencies
+\item[$\circ$] <15-> Strong modeling capabilities that can capture complex relationships
+\item[$\circ$] <16-> Computationally demanding, less scalable
 \end{itemize}
@@
 \frametitle{\insertsection : \insertsubsection}
-Extension of single-label metrics to the M-L case.
+\onslide <2->{Extension of single-label metrics to the M-L case.
 
-Grouped into two categories and perspectives:
+Grouped into two categories and perspectives:}
 \begin{itemize}
-\item[$\bullet$] \emph{Example-based: }Evaluates multi-labeled performance on each example, extrapolate to whole dataset
+\item[$\bullet$] <3-> \emph{Example-based: }Evaluates multi-labeled performance on each example, extrapolate to whole dataset
 \begin{itemize}
-\item[$\circ$] \emph{Classification perspective:}
+\item[$\circ$] <4-> \emph{Classification perspective:}
 \begin{itemize}
-\item[$-$] Subset Accuracy, Hamming Loss
-\item[$-$] Precision, Recall,  F$^\beta$ -measure
+\item[$-$] <5-> Subset Accuracy, Hamming Loss
+\item[$-$] <6-> Precision, Recall,  F$^\beta$ -measure
 \end{itemize}
-\item[$\circ$] \emph{Ranking perspective:} one-error, coverage, ranking loss, average precision
+\item[$\circ$] <7-> \emph{Ranking perspective:} one-error, coverage, ranking loss, average precision
 \end{itemize}
-\item[$\bullet$] \emph{Label-based: }Evaluates performance on each label separately, extrapolate to whole label set
+\item[$\bullet$] <8-> \emph{Label-based: }Evaluates performance on each label separately, extrapolate to whole label set
 \begin{itemize}
-\item[$\circ$] Classification perspective: Macro/Micro averaging techniques for single-label example-based, classification-perspective measures
-\item[$\circ$] Ranking perspective: Macro/Micro averaging for AUC
+\item[$\circ$] <9-> Classification perspective: Macro/Micro averaging techniques for single-label example-based, classification-perspective measures
+\item[$\circ$] <10-> Ranking perspective: Macro/Micro averaging for AUC
 \end{itemize}
 \end{itemize}
-$\ast$ Ideally, classifiers should be trained to optimize \emph{multiple} metrics
+\onslide <11->{$\ast$ Ideally, classifiers should be trained to optimize \emph{multiple} metrics}
 \end{frame}
@@
 
-Group algorithms in two categories:
+\onslide <2->{Group algorithms in two categories:}
 \begin{itemize}
-\item[$\bullet$] \emph{Problem Transformation Methods:}
+\item[$\bullet$] <3-> \emph{Problem Transformation Methods:}
 
 \begin{itemize}
-\item[$\circ$] Transform the learning problem into other, managable (often
+\item[$\circ$] <4-> Transform the learning problem into other, managable (often
   single-label) learning
   problems 
-\item[$\circ$] ``Fit data to algorithm'' philosophy
+\item[$\circ$] <5-> ``Fit data to algorithm'' philosophy
 \end{itemize}
 
-\item[$\bullet$] \emph{Algorithm Adaptation Methods:}
+\item[$\bullet$] <6-> \emph{Algorithm Adaptation Methods:}
 \begin{itemize}
-\item[$\circ$] Adapt popular learning techniques to deal with multi-label data directly
-\item[$\circ$] ``Fit algorithm to data'' philosophy
+\item[$\circ$] <7-> Adapt popular learning techniques to deal with multi-label data directly
+\item[$\circ$] <8-> ``Fit algorithm to data'' philosophy
 \end{itemize}
@@
 
-Authors include algorithms that:
+\onslide <9->{Authors include algorithms that:}
 \begin{itemize}
-\item[$\checkmark$] Has broad, noteworthy or unique characteristics
-\item[$\checkmark$] Has important impact, leading to a number follow-up related methods
-\item[$\checkmark$] Is influential and highly-cited in multi-label learning
+\item[$\checkmark$] <10-> Has broad, noteworthy or unique characteristics
+\item[$\checkmark$] <11-> Has important impact, leading to a number follow-up related methods
+\item[$\checkmark$] <12-> Is influential and highly-cited in multi-label learning
 \end{itemize}
-
 \end{frame}
@@
 \begin{itemize}
-\item[$\bullet$] Decompose multi-label problem to $|Y|=q$ independent binary classification problems
-\item[$\bullet$] Construct $q$ binary (one-vs-all) training sets (one for each $y_i$)
-\item[$\bullet$] Independently train each classifier $h_i(x)$ on its respective dataset
+\item[$\bullet$] <2-> Decompose multi-label problem to $|Y|=q$ independent binary classification problems
+\item[$\bullet$] <3-> Construct $q$ binary (one-vs-all) training sets (one for each $y_i$)
+\item[$\bullet$] <4-> Independently train each classifier $h_i(x)$ on its respective dataset
 
-\item[$\bullet$] Predict labels of an unseen $x$ by evaluating each classifier $h_i(x)$
-\item[$\bullet$] Assign label $y_i$ according to $r = h_i(x)$ and the
+\item[$\bullet$] <5-> Predict labels of an unseen $x$ by evaluating each classifier $h_i(x)$
+\item[$\bullet$] <6-> Assign label $y_i$ according to $r = h_i(x)$ and the
   thresholding setting 
@@
 
-Pros \& cons:
+\onslide <7->{Pros \& cons:}
 \begin{itemize}
-\item[$\bullet$] Simple, one-vs-rest scheme $\rightarrow$ easily parallelizable
-\item[$\bullet$] Sensitive to class-imbalanced data\footnote{Very different
+\item[$\bullet$] <8-> Simple, one-vs-rest scheme $\rightarrow$ easily parallelizable
+\item[$\bullet$] <9-> Sensitive to class-imbalanced data\footnote{Very different
     number of pos. and neg. examples for a class}
-\item[$\bullet$] Ignores label correlations (first-order method)
+\item[$\bullet$] <10-> Ignores label correlations (first-order method)
 \end{itemize}
@@
 \begin{itemize}
-\item[$\bullet$] Reorder the label $Y$ set using a permutation function $f_p$ 
-\item[$\bullet$] Transform into a \emph{chain} of binary classification problems
-\item[$\bullet$] Enrich representations at step $j$ by concatenating each $x_i$ with the confidence of preceeding $1,\dots,j-1$-th classifiers
-\item[$\bullet$] Predict relevant labels for unknown instances by iteratively traversing the classifier chain
+\item[$\bullet$] <2-> Reorder the label $Y$ set using a permutation function $f_p$ 
+\item[$\bullet$] <3-> Transform into a \emph{chain} of binary classification problems
+\item[$\bullet$] <4-> Enrich representations at step $j$ by concatenating each $x_i$ with the confidence of preceeding $1,\dots,j-1$-th classifiers
+\item[$\bullet$] <5-> Predict relevant labels for unknown instances by iteratively traversing the classifier chain
 \end{itemize}
 
-Pros \& cons:
+\onslide <6->{Pros \& cons:}
 \begin{itemize}
-\item[$\bullet$] High-order method: exploitation of label correlations to a degree, but in a random manner
-\item[$\bullet$] Highly sensitive to $f_p$. Running multiple chain executions in an ensemble
-  attempt to overcome this dependency.
-\item[$\bullet$] Iterative operation prevents parallel implementation
+\item[$\bullet$] <7-> High-order method: exploitation of label correlations to a degree, but in a random manner
+\item[$\bullet$] <8-> Highly sensitive to $f_p$. Running multiple chain executions in an ensemble attempt to overcome this dependency.
+\item[$\bullet$] <9-> Iterative operation prevents parallel implementation
 \end{itemize}
@@
 \begin{itemize}
-\item[$\bullet$] Transform into a label pairwise comparison ranking problem
-\item[$\bullet$] For $q$ labels, generate $q(q-1)/2$ binary classifiers by pairwise comparison and use a binary algorithm $h_{jk}(x)$
-\item[$\bullet$] Construct training sets $D_{jk} : \{x_i, Y_i | y_j \in Y_i \oplus y_k \in Y_i\}$.
+\item[$\bullet$] <2-> Transform into a label pairwise comparison ranking problem
+\item[$\bullet$] <3-> For $q$ labels, generate $q(q-1)/2$ binary classifiers by pairwise comparison and use a binary algorithm $h_{jk}(x)$
+\item[$\bullet$] <4-> Construct training sets $D_{jk} : \{x_i, Y_i | y_j \in Y_i \oplus y_k \in Y_i\}$.
 System votes for each example and if $h_{jk}(x) >0$,$x_i$ is associated with $y_j$, otherwise with $y_k$
-\item[$\bullet$] For an unknown instance, all classifiers' votes are aggregated
+\item[$\bullet$] <5-> For an unknown instance, all classifiers' votes are aggregated
   and resulting labels are ranked according to the total confidence $r$
-\item[$\bullet$] A virtual label $y_v$ is learned as a threshold, serving as the artificial splitting point between relevant and irrelevant labels
+\item[$\bullet$] <6-> A virtual label $y_v$ is learned as a threshold, serving as the artificial splitting point between relevant and irrelevant labels
 \end{itemize}
-Pros \& cons:
+\onslide <7->{Pros \& cons:}
 \begin{itemize}
-\item[$\bullet$] Second-order approach algorithm. One-vs-one scheme.
-\item[$\bullet$] Pairwise comparison smooths out the class-imbalance problem
-\item[$\bullet$] Number of classifiers is quadratic to $|Y|$ (compared to linear for BR)
+\item[$\bullet$] <8-> Second-order approach algorithm. One-vs-one scheme.
+\item[$\bullet$] <9-> Pairwise comparison smooths out the class-imbalance problem
+\item[$\bullet$] <10-> Number of classifiers is quadratic to $|Y|$ (compared to linear for BR)
 \begin{itemize}
-\item[$\circ$] Pruning methods to reduce the search space
+\item[$\circ$] <11-> Pruning methods to reduce the search space
 \end{itemize}
@@
 \begin{itemize}
-\item[$\bullet$] Decompose to an ensemble of multi-class classification problems
-\item[$\bullet$] Each component targets a random subset of $\mathbb{P}(Y)$ (that
+\item[$\bullet$] <2-> Decompose to an ensemble of multi-class classification problems
+\item[$\bullet$] <3-> Each component targets a random subset of $\mathbb{P}(Y)$ (that
   also appears in $X$), classified with Label Powerset (LP) techniques:
 \begin{itemize}
-\item[$\circ$] Transform to single-label data by treating each distinct labelset
+\item[$\circ$] <4-> Transform to single-label data by treating each distinct labelset
   as a new class $\rightarrow$ multi to single label problem
-\item[$\circ$] Each example is reassigned with the new mapped class and classified through regular single-label classification
-\item[$\circ$] M-L classify $x$ with $y_i$ when the votes received for $y_i$ from the ensemble exceed half the max possible it can get
+\item[$\circ$] <5-> Each example is reassigned with the new mapped class and classified through regular single-label classification
+\item[$\circ$] <6-> M-L classify $x$ with $y_i$ when the votes received for $y_i$ from the ensemble exceed half the max possible it can get
 \end{itemize}
 \end{itemize}
-Pros \& cons:
+\onslide <7->{Pros \& cons:}
 \begin{itemize}
-\item[$\bullet$] High-order approach algorithm
-\item[$\bullet$] Data-sensitive: Cannot generalize to labelsets not in the training set, too few examples for some labelsets
-\item[$\bullet$] Large $Y$ implies high training complexity.
-\item[$\bullet$] Improve by invoking an ensemble on random k-sized labelsets
+\item[$\bullet$] <8-> High-order approach algorithm
+\item[$\bullet$] <9-> Data-sensitive: Cannot generalize to labelsets not in the training set, too few examples for some labelsets
+\item[$\bullet$] <10-> Large $Y$ implies high training complexity.
+\item[$\bullet$] <11-> Improve by invoking an ensemble on random k-sized labelsets
 \end{itemize}
@@
 \begin{itemize}
-	\item[$\bullet$] Consider events $H_j \equiv (y_j \in Y_i)$, $C_j \equiv (\sum_{x_k \in N(x_i)} [\delta(y_j \in Y_k)])$
-\item[$\bullet$] Compares the MAP probabilities: $P(K|C_j), K \in \{H_j, \neg H_j\}$ to decide if to include $y_j$ in the prediction. Compute with the Bayes' theorem.
-\item[$\bullet$] Single-label priors $P(K)$ computed by smoothed frequency counting in the training data
-\item[$\bullet$] Likelihoods $P(C_j|K)$ are computed as a function of:
+	\item[$\bullet$] <2-> Consider events $H_j \equiv (y_j \in Y_i)$, $C_j \equiv (\sum_{x_k \in N(x_i)} [\delta(y_j \in Y_k)])$
+\item[$\bullet$] <3-> Compares the MAP probabilities: $P(K|C_j), K \in \{H_j, \neg H_j\}$ to decide if to include $y_j$ in the prediction. Compute with the Bayes' theorem.
+\item[$\bullet$] <4-> Single-label priors $P(K)$ computed by smoothed frequency counting in the training data
+\item[$\bullet$] <5-> Likelihoods $P(C_j|K)$ are computed as a function of:
 \begin{itemize}
-\item[$\circ$] $\kappa_j(r)$, the number of examples labelled $y_j$ with $r$ neighbours labelled $y_j$
-\item[$\circ$] $\overset{\sim}{\kappa_j}(r)$, the number of examples not labelled $y_j$ with $r$ neighbours labelled $y_j$
+\item[$\circ$] <6-> $\kappa_j(r)$, the number of examples labelled $y_j$ with $r$ neighbours labelled $y_j$
+\item[$\circ$] <7-> $\overset{\sim}{\kappa_j}(r)$, the number of examples not labelled $y_j$ with $r$ neighbours labelled $y_j$
 \end{itemize}
@@
 
-Pros \& cons:
+\onslide <8->{Pros \& cons:}
 \begin{itemize}
-\item[$\bullet$] First-order approach, Bayesian reasoning
-\item[$\bullet$] Decision boundary can be modified on-line via new instances
-\item[$\bullet$] M-L class imbalance can be overcome by calculating the priors
-\item[$\bullet$] Extensions proposed for label correlation exploitations
+\item[$\bullet$] <9-> First-order approach, Bayesian reasoning
+\item[$\bullet$] <10-> Decision boundary can be modified on-line via new instances
+\item[$\bullet$] <11-> M-L class imbalance can be overcome by calculating the priors
+\item[$\bullet$] <12-> Extensions proposed for label correlation exploitations
 \end{itemize}
-
 \end{frame}
@@
 \begin{itemize}
-\item[$\bullet$] Multi-label entropy is used to build a decision tree recursively
-\item[$\bullet$] Split node at the $l$-th feature of $x$ such that the information gain
-  criterion is maximized. Node partitions data according to $x_l = \theta$:
+\item[$\bullet$] <2-> Multi-label entropy is used to build a decision tree recursively
+\item[$\bullet$] <3-> Split node at the $l$-th feature of $x$ such that the information gain criterion is maximized. Node partitions data according to $x_l = \theta$:
  $T \rightarrow \{T^-,T^+\}$, where $ x_{il} \leq \theta, x_i \in T^-$ and $ x_{il} > \theta, x_i \in T^+$
-\item[$\bullet$] Recurse on subtrees until a stopping criterion is met (e.g. child size)
-\item[$\bullet$] Single-label entropy is computed by considering labelsets as new classes
-\item[$\bullet$] Assumes independence among labels to ensure low computational cost
-\item[$\bullet$] Unseen instances are assigned the label of the majority of members of the leaf they arrive
+\item[$\bullet$] <4-> Recurse on subtrees until a stopping criterion is met (e.g. child size)
+\item[$\bullet$] <5-> Single-label entropy is computed by considering labelsets as new classes
+\item[$\bullet$] <6-> Assumes independence among labels to ensure low computational cost
+\item[$\bullet$] <7-> Unseen instances are assigned the label of the majority of members of the leaf they arrive
 \end{itemize}
-Pros \& cons:
+\onslide <8->{Pros \& cons:}
 \begin{itemize}
-\item[$\bullet$] First-order algorithm, highly efficient
-\item[$\bullet$] Assumes label independence
-\item[$\bullet$] Pruning and ensemble strategies proposed
+\item[$\bullet$] <9-> First-order algorithm, highly efficient
+\item[$\bullet$] <10-> Assumes label independence
+\item[$\bullet$] <11-> Pruning and ensemble strategies proposed
 \end{itemize}
@@
 \begin{itemize}
-	\item[$\bullet$] $q$ linear classifiers $h_i(x)$, optimized with the empirical ranking loss
-	\item[$\bullet$] Combine classifiers to discriminate label \emph{pairs}. Label
-    pair $(j, k)$ decision hyperplane is: $h_{jk} = g(h_j, h_k) = \left<w_j-w_k, x_i \right> + b_j - b_k=0$
-\item[$\bullet$] Consider signed distance of $x_i$ from the boundary of every
+	\item[$\bullet$] <2-> $q$ linear classifiers $h_i(x)$, optimized with the empirical ranking loss
+	\item[$\bullet$] <3-> Combine classifiers to discriminate label \emph{pairs}. Label pair $(j, k)$ decision hyperplane is: $h_{jk} = g(h_j, h_k) = \left<w_j-w_k, x_i \right> + b_j - b_k=0$
+\item[$\bullet$] <4-> Consider signed distance of $x_i$ from the boundary of every
   relevant-nonrelevant label pair $(y_j,y_k) \in Y_i \times \bar Y_i$.
   % The pair's decision hyperplane is: $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$
-\item[$\bullet$] The M-L margin is the minimum distance of $x_i$ from every  $Y_i \times \bar Y_i$
-\item[$\bullet$] SVM $\rightarrow$ minimize loss \emph{and} maximize margin.
+\item[$\bullet$] <5-> The M-L margin is the minimum distance of $x_i$ from every  $Y_i \times \bar Y_i$
+\item[$\bullet$] <6-> SVM $\rightarrow$ minimize loss \emph{and} maximize margin.
 
-\item[$\bullet$] Ranks labels according to the singed distance
-\item[$\bullet$] Non-linearity achievable through feature mapping and the kernel trick
+\item[$\bullet$] <7-> Ranks labels according to the singed distance
+\item[$\bullet$] <8-> Non-linearity achievable through feature mapping and the kernel trick
 \end{itemize}
-Pros \& cons:
+\onslide <9->{Pros \& cons:}
 \begin{itemize}
-\item[$\bullet$] Second-order approach, maximum margin strategy
-\item[$\bullet$] Optimization is a convex QP problem, solved by any QP solver
-\item[$\bullet$] Kernel SVM, kernel selection can be done with MKL techniques
-\item[$\bullet$] Adaptable learning by selecting an appropriate loss function
+\item[$\bullet$] <10-> Second-order approach, maximum margin strategy
+\item[$\bullet$] <11-> Optimization is a convex QP problem, solved by any QP solver
+\item[$\bullet$] <12-> Kernel SVM, kernel selection can be done with MKL techniques
+\item[$\bullet$] <13-> Adaptable learning by selecting an appropriate loss function
 \end{itemize}
@@
 \begin{itemize}
-\item[$\bullet$] A Conditional Random Field model for M-L classification
-\item[$\bullet$] Encodes $y$ as binary random vectors $\bm{y}=\{-1,1\}^q$.
+\item[$\bullet$] <2-> A Conditional Random Field model for M-L classification
+\item[$\bullet$] <3-> Encodes $y$ as binary random vectors $\bm{y}=\{-1,1\}^q$.
   Learn joint probability distribution $p(x,\bm{y})$
-\item[$\bullet$] Compute conditional $p(\bm{y} | x)$ with the maximum entropy criterion:
+\item[$\bullet$] <4-> Compute conditional $p(\bm{y} | x)$ with the maximum entropy criterion:
   \begin{itemize}
-    \item Maximize $H_p(x,y)$ subject to $K$ constraints $\mathbb{E}[f_k(x,\bm{y})] = F_k$
-    \item $f_k(\cdot)$ directly model label correlations - the $F_k$ values are learned from the training data
-  \item[$\bullet$] Solve with Lagrange multipliers, assuming gaussian priors
-  \item[$\bullet$] Optimal solution is in the form of a Gibbs probability distribution
+    \item <5-> Maximize $H_p(x,y)$ subject to $K$ constraints $\mathbb{E}[f_k(x,\bm{y})] = F_k$
+    \item <6-> $f_k(\cdot)$ directly model label correlations - the $F_k$ values are learned from the training data
+  \item[$\bullet$] <7-> Solve with Lagrange multipliers, assuming gaussian priors
+  \item[$\bullet$] <8-> Optimal solution is in the form of a Gibbs probability distribution
   \end{itemize}
 %\item[$\bullet$] Unseen instances labelled with $Y = \underset{y}{\text{argmax }} p(y | x)$
-\item[$\bullet$] Unseen instances labelled with $Y = \text{argmax}_y\text{ } p(y | x)$
+\item[$\bullet$] <10-> Unseen instances labelled with $Y = \text{argmax}_y\text{ } p(y | x)$
 \end{itemize}
-Pros \& cons:
+\onslide <11->{Pros \& cons:}
 \begin{itemize}
-\item[$\bullet$] Second-order approach. All label pairs considered, not just relevant-irrelevant ones.
-\item[$\bullet$] Convex constraint optimization problem $\rightarrow$ solvable by any
-  CP solver
-\item[$\bullet$] Intractable $\text{argmax}$ operation for a large label space without pruning.
-\item[$\bullet$] Extensible: Different forms of constraints yield various CML variants
+\item[$\bullet$] <12-> Second-order approach. All label pairs considered, not just relevant-irrelevant ones.
+\item[$\bullet$] <13-> Convex constraint optimization problem $\rightarrow$ solvable by any CP solver
+\item[$\bullet$] <14-> Intractable $\text{argmax}$ operation for a large label space without pruning.
+\item[$\bullet$] <15-> Extensible: Different forms of constraints yield various CML variants
 \end{itemize}
@@
 \begin{itemize}
-\item[$\bullet$] Multi-instance learning
+\item[$\bullet$] <2-> Multi-instance learning
 \begin{itemize}
-\item[$\circ$] Instead of labeled instances, get binary-labeled \emph{bags of instances}
-\item[$\circ$] Assign positive label to bag if at least one member is positive
-\item[$\circ$] Models complex semantics of $x_i$ in input space, rather than its output
+\item[$\circ$] <3-> Instead of labeled instances, get binary-labeled \emph{bags of instances}
+\item[$\circ$] <4-> Assign positive label to bag if at least one member is positive
+\item[$\circ$] <5-> Models complex semantics of $x_i$ in input space, rather than its output
 \end{itemize}
-\item[$\bullet$] Ordinal classification
+\item[$\bullet$] <6-> Ordinal classification
 \begin{itemize}
-\item[$\circ$] Assume label relevance is not binary, but soft
-\item[$\circ$] Produce a vector of ordinal graded membership
-\item[$\circ$] Transform M-L problem to a set of ordinal set of problems
+\item[$\circ$] <7-> Assume label relevance is not binary, but soft
+\item[$\circ$] <8-> Produce a vector of ordinal graded membership
+\item[$\circ$] <9-> Transform M-L problem to a set of ordinal set of problems
 \end{itemize}
-\item[$\bullet$] Multi-task learning
+\item[$\bullet$] <10-> Multi-task learning
 \begin{itemize}
-\item[$\circ$] Multiple tasks trained in parallel, sharing information
-\item[$\circ$] Knowledge from related tasks used as an inductive bias to improve generalization
-\item[$\circ$] Shared or different feature space, small task workload
+\item[$\circ$] <11-> Multiple tasks trained in parallel, sharing information
+\item[$\circ$] <12-> Knowledge from related tasks used as an inductive bias to improve generalization
+\item[$\circ$] <13-> Shared or different feature space, small task workload
 \end{itemize}
-\item[$\bullet$] Data streams classification
+\item[$\bullet$] <14-> Data streams classification
 \begin{itemize}
-\item[$\circ$] Real-world objects are generated online and processed in  real-time
-\item[$\circ$] Concept drift problem
+\item[$\circ$] <15-> Real-world objects are generated online and processed in  real-time
+\item[$\circ$] <16-> Concept drift problem
 \end{itemize}
@@
 \frametitle{\insertsection}
-Summary:
+\onslide <2->{Summary:}
 \begin{itemize}
-\item[$\bullet$] Multi-label learning problem definition
-\item[$\bullet$] Multi-label learning representative algorithms
-\item[$\bullet$] Related learning methods
+\item[$\bullet$] <3-> Multi-label learning problem definition
+\item[$\bullet$] <4-> Multi-label learning representative algorithms
+\item[$\bullet$] <5-> Related learning methods
 \end{itemize}
 
-Future goals:
+\onslide <6->{Future goals:}
 \begin{itemize}
-\item[$\star$] Formal characterization on the underlying concept / mechanism on the appropriate usage of label correlations, especially on large output spaces
-\item[$\bullet$] Thorough experimental comparative study to discover pros and cons of different multi-label learning algorithms
+\item[$\bullet$] <7-> Formal characterization on the underlying concept / mechanism on the appropriate usage of label correlations, especially on large output spaces
+\item[$\bullet$] <8-> Thorough experimental comparative study to discover pros and cons of different multi-label learning algorithms
 \end{itemize}

### 2018-01-20 Paper/PaperReport/PaperReport.tex
Update project report

--- 2018-01-17 Paper/PaperReport/PaperReport.tex
@@
 \usepackage{amsmath}
+\usepackage{bm}
 \graphicspath{{images/}}
@@
 The metrics are divided into two perspectives:
-      \todos{Form them up into a table}
+      \todo{Form them up into a table}
 		\begin{itemize}
@@
 	\subsection*{Problem Transformation Methods}
+  We move on to present the first category of methods, which transform the
+  multi-label problem into a set of single label methods. See figure \ref{figure:ptm} for an overview of the approaches examined.
+
 	\addcontentsline{toc}{subsection}{Problem Transformation Methods}
@@
 		\centering
+    \label{figure:ptm}
 	\end{figure}
@@
 	\addcontentsline{toc}{subsubsection}{Claibrated Label Ranking}
-	This algorithm transforms the learning problem into a label pairwise comparison ranking problem.
-For $q$ labels, $q(q-1)/2$ binary classfiers are generated by pairwise comparison and a binary algorithm $h_{jk}(x)$ is used.
-Firstly, training sets $D_{jk}$ are constructing, where $D_{jk}: \{x_i, Y_i | y_j \in Y_i \oplus y_k \in Y_i\}$.
-The learning system votes for each example and if $h_{jk}(x)>0$, then $x_i$ is associated with label $y_j$, otherwise $y_j$ is irrelevant and $x_i$ is assigned to $y_k$.
-For unknown instances, classifiers' votes are aggregated and ranked.
-In addition, a virtual label $y_v$ is used as an artificial splitting point between relevant and irrelevant labels.
+	This algorithm transforms the learning problem into a label pairwise
+  comparison ranking problem. For $q$ labels, $q(q-1)/2$ binary classifiers are
+  constructed to discriminate between each label pair via a binary
+  classification algorithm $h_{jk}(x)$. Firstly, respective pairwise training
+  sets $D_{jk}$ are constructed, where $D_{jk}: \{x_i, Y_i | y_j \in Y_i \oplus
+  y_k \in Y_i\}$, i.e. each $D_{jk}$ contains examples either in class $j$ or
+  $k$. The learning system votes for each example and if $h_{jk}(x)>0$, then
+  $x_i$ is associated with label $y_j$, otherwise $y_j$ is irrelevant and $x_i$
+  is assigned to $y_k$. For unknown instances, all classifiers' votes are
+  aggregated and ranked. In addition, a virtual label $y_v$ is used as an
+  artificial splitting point between relevant and irrelevant labels, serving as
+  a decision threshold for label relevance.
 
-	This is a second-order approach algorithm using a one-vs-one scheme.
-The advantage of this method is that it smooths out the class-imbalance problem (where the number of a class positives is much less than the other's class positives).
-The disadvantage is the number of classifiers is quadratic to $|Y|$, compared to linear for Binary Relevance algorithm.
-One improvement to this issue is to use pruning methods to reduce the search space.
+	This is a second-order approach algorithm using a one-vs-one scheme. The
+  advantage of this method is that it smooths out the class-imbalance problem
+  (where the number of a class positives is much less than the other's class
+  positives) by operating on label pairs. The disadvantage is that the number of
+  classifiers is quadratic to $|Y|$, compared to linear for Binary Relevance or
+  Classifier Chains algorithms. However, the performance can be improved by
+  using pruning methods to reduce the search space.
 
@@
 	\addcontentsline{toc}{subsubsection}{Random k-Labelsets}
-	Random k-labelsets algorithm transforms the multi-label learning problem into an ensemble of multi-class classification problems where each component targets a random subset of $Y$, that appears in $X$ classified with Label Powerset (LP) techniques.
-In more detail, the multi-label dataset is transformed into single-label data by treating each distinct labelset as a new class.
-Each training example is reassigned with the new mapped class and classified through regular single-label classification.
+	The random k-labelsets algorithm transforms the multi-label learning problem
+  into an ensemble of multi-class classification problems. Each such partition
+  targets a random subset of $\mathbb{P}(Y)$, limited to subsets present in the
+  training dataset $X$, classifying it using Label Powerset (LP) techniques. In
+  more detail, the multi-label dataset is transformed into single-label by
+  treating each distinct labelset as a new class. Thus, each training example
+  mapped to a new single class and classified through some single-label
+  classification method.
 An instance $x_i$ is assigned to label $y_i$ when the votes received for $y_i$ from the ensemble exceed half the max possible that this label can get.
@@
 
-	This is a high-order approach algorithm, utilizing label correlations. However, there are two limitations in terms of practical feasibility:
-
+	This is a high-order approach algorithm since it utilizes a variety of label correlations. However, there are two limitations in terms of practical feasibility: 
 	\begin{itemize}
-		\item[$\diamond$] \emph{Incompleteness: }The algorithm is data-sensitive and cannot generalize to labelsets that do not appear into the training set
-		\item[$\diamond$] \emph{Inefficiency: }Large $Y$ implies high training complexity and extremely few training examples for some newly mapped classes.
+		\item[$\diamond$] \emph{Incompleteness: }The algorithm is data-sensitive and cannot generalize to labelsets that do not appear into the training set.
+		\item[$\diamond$] \emph{Inefficiency: }A large label set $Y$ implies high training complexity and extremely few training examples for some newly mapped classes.
 	\end{itemize}
 
-	To overcome above issues while keeping LP's simplicity, Random k-Labelsets chooses to combine ensemble learning with LP.
-The latter is invoked only on random k-labelsets (each leading to a multi-class classifier $g_{y^k(l_r)}^+$) to guarantee computational efficiency and then ensemble a number of LP classifiers to achieve predictive completeness.
-For new instances $x$, two quantities are calculated: $\tau(\textbf{x},y_j)$, which counts the maximum number of votes that $y_j$ can receive from the ensemble and $\mu(\textbf{x},y_j)$, which counts the actual number of votes that $y_j$ receives from the ensemble.
-Therefore, the predicted labelset is calculated as follows: $Y = \{y_j|\mu(\textbf{x},y_j)/\tau(\textbf{x},y_j)>0.5,  1 \leq j \leq q\}$.
-This means that when the actual number of votes exceeds half of the maximum number of votes, $y_j$ is considered as a relevant label.For an ensemble created by n k-labelsets, the maximum number of votes on each label is $nk/q$ on average (usually: k=3 and n = 2q).
-Finally, another way to improve LP is to prune distinct label set in D appearing less than a pre-specified counting threshold.
+	A simple but powerful modification is the Random k-Labelsets variant, which
+  runs an ensemble of executions, each with $k$-sized label sets. The ensemble
+  scheme imporives on the predictive completeness of the system, while the
+  limited labelset size guarantees computational efficiency. Additionally,
+  pruning distinct labelsets which are less frequent than a lower threshold is
+  another strategy to improve the method's performance.
 
 	\subsection*{Algorithm Adaptation Methods}
+  The next category refers to algorithm adaptation methods, i.e. strategies that
+  directly address the multi-label nature of the problem. See figure
+  \ref{figure:aam} for an overview of the approaches examined.
+
 	\addcontentsline{toc}{subsection}{Algorithm Adaptation Methods}
@@
 		\centering
+    \label{figure:aam}
 	\end{figure}
@@
 	\addcontentsline{toc}{subsubsection}{Multi-Label k-Nearest Neighbor (ML-kNN)}
-	ML-kNN algorithm adapts k-nearest neighbor techniques utilizing maximum a posteriori (MAP) rule to make prediction by reasoning with the labeling information embodied in the neighbors.
-For new instances $x$, $N(x)$ represents the set of its k nearest neighbors identified in the dataset (similarity among instances is calculated by Euclidean distance).
-Let $H_j \equiv (y_j \in Y_i)$, $C_j = \sum_{x_k \in N(x_i)} [\delta(y_j \in Y_k)]$.
-$C_j$ records the number of x’s neighbors with label $y_j$, while $H_j$ is the event that $\textbf{x}$ has the label $y_j$.
-Moreover, $P(H_j|C_j)$ is the posterior probability that $H_j$ can be applied in the condition that $\textbf{x}$ has exactly $C_j$ neighbors.
-Similarly, $P(\neg H_j|C_j)$ needs to be calculated in order to decide if to include $y_j$ in the prediction.
-This will be computed using the Bayes' theorem:
+	The ML-kNN algorithm adapts nearest neighbor techniques to the multi-label
+  setting by utilizing the maximum a posteriori (MAP) rule, to make a prediction
+  by reasoning with the labeling information embodied in the neighbors. For new
+  instances $x$, $N(x)$ represents the set of its $k$ nearest neighbors
+  identified in the dataset by a specified similarity or distance measure. Let
+  $H_j \equiv (y_j \in Y_i)$, $C_j = \sum_{x_k \in N(x_i)} [\delta(y_j \in
+  Y_k)]$, where $H_j$ is the event that $\textbf{x}$ has the label $y_j$ and
+  $C_j$ records the number of x’s neighbors with label $y_j$. Moreover,
+  $P(H_j|C_j)$ is the posterior probability that $H_j$ holds in the condition
+  that $\textbf{x}$ has exactly $C_j$ neighbors. The posteriors $P(H_j|C_j)$ are
+  calculated using Bayes' theorem with respect to the priors and the likelihoods.
 
-	\begin{center}
-		$\frac{P(H_j|C_J)}{P(\neg H_j|C_j)} = \frac{P(H_j) \cdot P(C_j|H_j)}{P(\neg H_j) \cdot P(C_j| \neg H_j)}$
-	\end{center}
-
-	Prior probabilities ($P(H_j)$ and $P(\neg H_j)$) are computed via the frequency counting strategy, where a smoothing parameter $s$ is used to control the effect of uniform prior on the estimation (usually $s=1$).
-In addition, likelyhoods $P(C_j|H_j)$ and $P(C_j| \neg H_j)$ are computed using two frequency arrays:
+	The prior probabilities ($P(H_j)$, $P(\neg H_j)$) are computed via a frequency counting strategy, where a smoothing parameter $s$ is used to control the effect of uniform prior on the estimation (usually $s=1$).
+In addition, likelihoods $P(C_j|H_j)$ and $P(C_j| \neg H_j)$ are computed using two frequency arrays:
 	\begin{itemize}
@@
 
-	This algorithm is a first-order approach using Bayesian reasoning.
-The decision boundary can be modified on-line when new instances appear.
-Moreover, Through prior probabilities calculation class-imbalance issue can be overcome.
-However, since label correlations are ignored, some extensions are proposed that take into account label associations.
+	This algorithm is a first-order approach that uses Bayesian reasoning. Being a
+  nearest-neighbour method, the decision boundary can be modified on-line when
+  new instances appear. Moreover, the use of prior probabilities can amend the
+  class imbalance issue to a degree. A number of extensions have been proposed that take into account label associations.
 
@@
 	\addcontentsline{toc}{subsubsection}{Multi-Label Decision Tree (ML-DT)}
-	ML-DT adapts decision tree techniques using multi-label entropy to build a decision tree recursively.
-Let $T = \{(x_i, Y_i) | 1 \leq i \leq n \}$ be the multi-label dataset with n examples.
-The algorithm splits the $x$ vector at the feature $l$ that maximizes the information gain criterion.
-Moreover, dataset $T$ is split on the $l$-th feature valued $\theta$ into branches ($T^-$ \& $T^+$), where $x_{il} \leq \theta, x_i \in T^-$ and $x_{il} > \theta, x_i \in T^+$.
-This process is invoked recursively by treating each time one of the two branches as the new root node and terminates until some stopping criterion $C$ is met (e.g.
-child size).
-Each subset is treated as a new class and single-label entropy is computed.
-Any new instances are assigned the label of the majority of members of the leaf that they arrive.
+	This method adapts decision tree techniques for multi-label classification.
+  Using the multi-label entropy measure, decision tree is built recursively by
+  splitting features that yield the maximum information gain. Let $T = \{(x_i,
+  Y_i) | 1 \leq i \leq n \}$ be the multi-label dataset with n examples. The
+  algorithm splits the $x$ vector at the feature $l$ that maximizes the
+  information gain criterion, where that value in $x$ is  $\theta$. This
+  partitions the dataset $T$ into branches ($T^-$ \& $T^+$), where $x_{il} \leq
+  \theta, x_i \in T^-$ and $x_{il} > \theta, x_i \in T^+$. This process is
+  invoked recursively by treating each time one of the two branches as the new
+  root node and terminates until some stopping criterion $C$ is met (e.g.
+  subtree size). Each label subset is treated as a new class and single-label entropy is computed. Any new instances are assigned the label of the majority of members of the leaf that they arrive.
 
-	In order to ensure low computational cost and high efficiency, the algorithm assumes independence among the labels (first-order approach).
-However, extended strategies are proposes, such as pruning and ensemble strategies, so as to exploit label correlations as well.
+	In order to ensure low computational cost and high efficiency, the algorithm
+  assumes independence among the labels; thus, it is a first-order approach.
+  Extensions are proposed to reduce computational cost and exploit label interrelations, such as pruning and ensemble strategies.
 
@@
 	\addcontentsline{toc}{subsubsection}{Ranking Support Vector Machine (Rank-SVM)}
-	Rank-SVM adapts maximum margin strategy, where $q$ linear classifiers are optimized to minimize empirical ranking loss and enabled to handle nonlinear cases with kernel tricks.
-The learning system is composed by those classifiers $W = \{w_j,b_j | 1 \leq j \leq q\}$, where $w_j \in \mathbb{R}^d$ and $b_j \in \mathbb{R}$ are the weight vector and the bias
-	for the j-th class label $y_j$ respectively.
-Rank-SVM defines the system’s margin on $(x_i,Y_i)$ by considering its ranking ability on the relevant and irrelevant label pairs of $x_i$ $(y_j,y_k) \in Y_i \times \bar Y_i$, which are separated by a function of their classifiers $h_j, h_k$: $\left<w_j-w_k, x_i \right> + b_j - b_k=0$.
-When the learning system is capable of properly ranking every relevant-irrelevant label pair for each training example, the classifier will return positive margin.
-For each new instances $x$, the algorithms retrieves a ranked list of label pairs.
+	Rank-SVM adapts the maximum margin strategy to the multi-label classification
+  problem. $q$ linear classifiers are optimized to minimize the empirical
+  ranking loss. The learning system is composed by classifiers that distinguish
+  between relevant and irrelevant pairs.
+Each classifier is parameterized with $W = \{w_j,b_j | 1 \leq j \leq q\}$, where $w_j \in \mathbb{R}^d$ and $b_j \in \mathbb{R}$ are the weight vector and the bias
+	for the j-th class label $y_j$ respectively. Rank-SVM defines the system’s
+  margin on $(x_i,Y_i)$ by considering its ranking ability on the relevant and
+  irrelevant label pairs of $x_i$ $(y_j,y_k) \in Y_i \times \bar Y_i$, which are
+  separated by a function of their classifiers $h_j, h_k$: $\left<w_j-w_k, x_i
+  \right> + b_j - b_k=0$. The margin is defined as the minimum distance from all
+  related-unrelated pairs and is minimized along with the training loss
+  function. At consistent training, the classifier will return positive margin.
+  For new test instances, the algorithms retrieves a ranked list of label pairs
+  according to the signed distance to the decision hyperplane.
 
-	Moreover, non-linear problems can be solved through feature mapping and kernel trick, while convex quadratic programming problems can be solved by any QP solver.
+	Non-linear problems can be managed by exploiting kernel feature mapping via
+  the kernel trick, which is popular in SVM methods. The optimization method is
+  a convex quadratic programming problems and can thus be solved by any QP
+  solver.
   This algorithm is a second-order approach using pairwise label associations.
-Improvements can be achieved by: replacing ranking with hamming loss which can be cast as a general form of	structured output classification, accomplishing thresholding strategy with techniques other than stacking-style procedure and avoiding the problem of kernel selection by employing multiple kernel learning techniques to learn from multi-label data.
+  It is adaptable to a variety of tasks and data by using alternative loss
+  functions and kernel feature mapping. Kernel selection can be accomplished by
+  multiple kernel learning (MKL) techniques.
 
@@
 	 \addcontentsline{toc}{subsubsection}{Collective Multi-Label Classifier}
-	 This algorithm adapts maximum entropy principle, where correlations among labels are encoded as binary random vectors into a joint probability distribution function(pdf) and label correlations are encoded into the joint pdf as constraints.
-This is a second-order approach, where all label pairs are considered and not just the relevant-irrelevant ones.
-Let $H_p(x, y)$ represent the information entropy of $(x,y)$ given their distribution $p(\cdot , \cdot)$.
-The principle of maximum entropy assumes that the distribution best modeling the current state of knowledge is the one maximizing $H_p(x, y)$ subject to a collection K of given facts: $\mathbb{E_p}[f_k(x,y)] = F_k$, where $F_k$ are estimated from the training set.
-The fact is expressed as constraint on the expectation of  some  function  over.
-Along with the normalization constraint on $p(\cdot , \cdot)$ , the constrained optimization problem can be solved with standard Lagrange Multiplier techniques.
-Any unseen instance $x$ is labeled with $Y = \underset{y}{\text{argmax }} P(y | x)$.
-Exact inference with arg max is only tractable for small label space, otherwise pruning strategies should be applied.
+   The CML method is a multi-label variant of Conditional Random Field methods.
+
+	 This algorithm involves the maximum entropy principle, where correlations
+   among labels are encoded as binary random vectors into a joint probability
+   distribution function and label correlations are encoded into the
+   distribution as constraints.
+   The conditional $p(\bm{y} | x)$ is computed via the maximum entropy
+   criterion, where the information entropy measure $H_p(x,\bm{y})$ is maximized subject to each constraint.
+   The latter are modeled as $ \mathbb{E} [ f_k(x,\bm{y}) ] = F_k  $, where $F_k$ is
+ learned from the traininng data.
+ The optimal solution is arrived at via Lagrange multipliers, assuming gaussian
+ priors for the variables involved, and has the form of a Gibbs probability
+ distribution.
+ Test instances are labelled with respect to $\text{argmax}_y\text{ }p(\bm{y} | x)$.
+
+   This is a second-order approach, where all label pairs are considered and not
+   just the relevant-irrelevant ones which was the Rank-SVM approach. It is a
+   convex constraint optimization problem, and is thus solvable by off the shelf
+   CP solvers. The argmax operation during inference is expensive, which renders
+   the method very computationally inefficient for large label spaces, if
+   pruning is not applied. Varying the expression form of the constraints gives rise to difference
+   version of the CML method.
 
@@
  	\addcontentsline{toc}{section}{Related Learning Methods}
- 	Four learning methods are mentioned in this paper, which can be found below:
+ 	Four related learning methods are mentioned in this paper, which can be found below:
  	\begin{enumerate}
- 		\item \textbf{Multi-instance learning: }This method studies the problem where each example is described by a bag of instances while associated with a single/binary label.
-Bag is assigned positive label in case of at least one positive member.
-Finally, it models complex semantics of $x_i$ in input space rather than its output.
- 		\item \textbf{Ordinal classification: }This method assumes that labels are related through a natural ordering and instead of explicit class relevance, it assumes an ordering of relevant between labels via a membership vector.
-Finally, the multi-label problem is transformed into a set of ordinal set of problems.
- 		\item \textbf{Multi-task learning: }This method studies the problem where multiple tasks are trained in parallel, sharing information, where knowledge from related tasks is used as an inductive bias to help improve the generalization performance of other tasks.
-The  tasks can be in the same feature space or different feature spaces, while in multi-label learning all the examples share the same feature space.
-Finally, in multi-task learning it is not reasonable to have a large number of tasks, while in multi-label learning it is usual to deal with large label space.
- 		\item \textbf{Data streams classification: }This method studies the problem	where real-world objects are generated online and processed in a real-time manner.
-Key factor is to deal with the concept drift problem.
-Some existing strategies are: classifiers update significantly whenever a new batch of examples arrive, assumption that the influence of past data gradually declines as time evolves or change detector alerting whenever a concept drift is detected.
+ 		\item \textbf{Multi-instance learning: }This method studies the problem
+      where each example is described by a bag of instances while associated
+      with a single binary label. The bag is assigned a positive label in the
+      case of at least one positively labelled member. It models complex semantics of $x_i$ in input space rather than its output.
+ 		\item \textbf{Ordinal classification: } This scheme assumes label relevance
+      to an instance is not binary but graded, producing a vector of ordinal graded
+      membership of $x$ to a label. Multi-label classification is induced via a
+      transformation of the original problem into a set of ordinal problems.
+ 		\item \textbf{Multi-task learning: }This method studies the problem where
+      multiple tasks are trained in parallel. Tasks can share information, and
+      knowledge from related tasks is used as an inductive bias to help improve
+      the generalization performance of other tasks. The  tasks can share a
+      common feature space or use different representations, as is common in
+      multi-label learning. In this setting, it is reasonable to have small task
+      workloads, since the execution is of a parallel nature. 
+ 		\item \textbf{Data streams classification: } Here, data is generated in real
+      time, corresponding to the real-time nature of some task. Apart from the
+      tight temporal constraints, the defining challenge here is the concept
+      drift problem, where the concept classified can change over time. A common
+      strategy to this effect is to gradually discount the significance of past
+      data in favor of newly arrived batches.
+
  	\end{enumerate}
@@
 	\addcontentsline{toc}{section}{Conclusion}
-	Summarizing all the above, this paper defines the multi-label learning problem, presents 8 multi-label learning representative algorithms and mentiones some additional related learning methods.
-Some important future goals are the formal characterization on the underlying concept/mechanism on the appropriate usage of label correlations (especially on large output spaces) and also a thorough experimental comparative study to discover the advantages and disadvantages of different multi-label learning algorithms.
-Below, some online resources on multi-label learning are presented:
+	Summarizing, this paper defines the multi-label learning problem, presenting 8
+  multi-label learning representative algorithms, mentions some additional
+  related learning settings. Some important future goals are the formal
+  characterization on the underlying concept and / or mechanism on the
+  appropriate usage of label correlations (especially on large output spaces).
+  In addition, thorough experimental comparative studies are essential to
+  discover the advantages and disadvantages of different multi-label learning
+  algorithms, so as to aid the conquest of the aforementioned theoretical goal.
+  Some online resources on multi-label learning are presented on figure \ref{figure:resources}.
 
@@
 		\caption{Online resources for Multi-Label Learning}
-		\centering
+    \label{figure:resources}
 	\end{figure}
@@
 	%-------------------------------------------------------------------------------
-	\addcontentsline{toc}{section}{Related Work}
-	\renewcommand\bibname{Related Work}
+	\addcontentsline{toc}{section}{Related Work and Bibliography}
+	\renewcommand\bibname{Related Work and Bibliography}
 	\begin{thebibliography}{9}
@@
 ACM, 2010
+
+\bibitem{trees} Vens, Celine, et al. "Decision trees for hierarchical multi-label classification." Machine Learning 73.2 (2008): 185-214.
+\bibitem{rkl} Tsoumakas, Grigorios, and Ioannis Vlahavas. "Random k-labelsets: An ensemble method for multilabel classification." Machine learning: ECML 2007 (2007): 406-417.
+\bibitem{mlknn} Zhang, Min-Ling, and Zhi-Hua Zhou. "ML-KNN: A lazy learning approach to multi-label learning." Pattern recognition 40.7 (2007): 2038-2048.
+\bibitem{rksvm} Elisseeff, André, and Jason Weston. "A kernel method for multi-labelled classification." Advances in neural information processing systems. 2002.
+\bibitem{cml}  Ghamrawi, Nadia, and Andrew McCallum. "Collective multi-label classification." Proceedings of the 14th ACM international conference on Information and knowledge management. ACM, 2005.
+\bibitem{ordinals}  Cheng, Weiwei, Eyke Hüllermeier, and Krzysztof J.
+Dembczynski.
+"Graded multilabel classification: The ordinal case." Proceedings of the 27th international conference on machine learning (ICML-10).
+2010.
+
 	\end{thebibliography}

### 2018-01-21 Paper/PaperReport/PaperReport.tex
Update report

@@
 	\addcontentsline{toc}{subsection}{Multi-label learning}
-	The paper "A Review on Multi-Label Learning Algorithms" analyzes the
-  multi-label learning field in supervised classification. Supervised
+	The paper ``A Review on Multi-Label Learning Algorithms'' by Zhang et al.
+  studies multi-label learning, applied in supervised classification. Supervised
   classification trains a model $f: X \rightarrow Y$ and for each pair
   $f(x_i,y_i)$, it outputs a real number $r$ that represents the confidence that
-  label $y_i$ characterizes $x_i$. Given $r$ and a thresholding scheme, labels
+  the label $y_i$ characterizes $x_i$. Given $r$ and a thresholding scheme, labels
   are marked as relevant or not relevant to the input data vector $x_i$ (e.g.
@@
   thresholding function is $t(x_i) = 0.5$. In other words, labels with outputs
-  greater than $0.5$ are assigned to $x_i$ by our model). Standard, single-label learning maps an
+  greater than $0.5$ are assigned to $x_i$ by the model). Standard, single-label learning maps an
   example to a single label, completely ignoring multiple semantics that
@@
   example/instance $x_i$ to be mapped to more than one label, i.e. to any label
-  vector which is a subset of $Y$. However, this results to an exponentially growing search space, since
-  the latter depends on the size of the powerset of $Y$, 
-  $|\mathbb{P(Y)}| = 2^{|Y|}$.
+  vector which is a member of the powerset of the label set $Y$. This however
+  results to an exponentially growing search space,
+  since the size of the latter is an exponential function of the number of labels, i.e. $|\mathbb{P(Y)}| = 2^{|Y|}$.
 
 	One straightforward solution is to exploit label correlations. Real objects
-  are usually characterized by similar concepts. For example, if an article is
+  are usually characterized by similar and correlated concepts. For example, if an article is
   characterized by label "football" it follows that it will be also
@@
   with ``crowd'' or ``entertainment''. On the other hand, the label "science"
-  will be likely to be irrelevant. In order to exploit such kind of label
-  correlations, three different algorithm strategies are examined:
+  will be likely to be irrelevant. Exploiting this label interdependence gives
+  rise to techniques that perform better and more efficiently that brute force
+  search on the large space of all label sets. The authors categorize approaches
+  to three general strategies, according to the extent of label correlation exploitation:
 
@@
   incorporate the multi-labeled output. This extension is categorizable into two
-  groups. One with respect to the expansion ``axis'' (across examples or labels)
+  groups. One with respect to the expansion ``axis'' (expansion to multiple examples or labels)
   and the task type (classification or ranking).
 
-    Before proceeding with the metrics, true positive(TP), false positive(FP),
-    true negative (TN) and false negative (FN) of label $y_j$ will be defined:
+    Before proceeding with the metrics, we define the concepts of true positive(TP), false positive(FP),
+    true negative (TN) and false negative (FN) of a label $y_j$:
 		\begin{itemize}
-			\item \emph{True positive: } $TP_j = |\{x_i | y_j \in Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|$
-			\item \emph{False positive: }$FP_j = |\{x_i | y_j \notin Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|$
-			\item \emph{True negative: }$TN_J = |\{x_i | y_j \notin Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|$
-			\item \emph{False negative: }$FN_J = |\{x_i | y_j \in Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|$
+			\item True positive:  $TP_j = |\{x_i | y_j \in Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|$
+			\item False positive: $FP_j = |\{x_i | y_j \notin Y_i \wedge y_j \in h(x_i), 1 \leq i \leq p \}|$
+			\item True negative: $TN_j = |\{x_i | y_j \notin Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|$
+			\item False negative: $FN_j = |\{x_i | y_j \in Y_i \wedge y_j \notin h(x_i), 1 \leq i \leq p \}|$
 		\end{itemize}
 
-	\begin{itemize}
-		\item \textbf{Example-based: }This kind of metrics evaluate multi-labeled performance on each example and then the result is spread to the whole dataset.
-The metrics are divided into two perspectives:
-      \todo{Form them up into a table}
-		\begin{itemize}
-			\item Classification:
-        \begin{table}
-          \centering
-          \begin{tabular}{cc} 
-          \toprule
-            one & two \\
-          \midrule
-            one & two \\
-            one & two \\
-          \bottomrule
-          \end{tabular}
-          \caption{Evaluation metrics for multi-label learning}
-          \label{label:evaluation_metrics}
-        \end{table}
+    Many metrics are defined in terms of the above concepts. We proceed to
+    present multi-label evaluation metrics.
 
 	\begin{itemize}
-				\item Precision: $P(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{h(x_i)|}$
-				\item Recall: $R(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i}$
+		\item \textbf{Example-based: } This type of metrics evaluate multi-labeled
+      performance on each example and then the result is spread to the whole
+      dataset. The metrics with respect to classification-based and ranking
+      based approaches are:
+
+		\begin{itemize}
+			\item Classification-wise metrics:
+			\begin{itemize}
+				\item Precision: $P(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|h(x_i)|}$
+
+          Precision measures the proportion of examples correctly classified to $y_i$, out
+          of the total examples assigned to the class by the model.
+				\item Recall: $R(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i|}$
+
+          Recall measures the proportion of examples correctly classified to $y_i$, out
+          of the total examples assigned to the class in the ground truth.
         \item Subset Accuracy:  $SA(h) = \frac{1}{p} \sum_{i=1}^p \llbracket h(x_i) = Y_i \rrbracket$
+
+          Subset accuracy compares the predicted and ground truth sets for equality.
 				\item Hamming Loss: $H(h) = \frac{1}{p} \sum_{i=1}^{p} [h(x_i)\Delta Y_i] $
+
+          The Hamming loss compares the predicted and ground truth sets,
+          measuring the set difference of the two.
 				\item Accuracy: $A(h) = \frac{1}{p} \sum_{i=1}^{p} \frac{|Y_i \cap h(x_i)|}{|Y_i \cup h(x_i)|}$
+
+          Accuracy measures the proportion of examples correctly classified to
+          $y_i$, out of the total examples assigned to that class either by the
+          model or by the ground truth.
 				\item $F^ \beta$ -metric:  $F^ \beta _{exam} (h) = \frac{(1+ \beta ^2) \cdot Precision_{exam}(h) \cdot Recall_{exam}(h)}{\beta ^2 \cdot Precision_{exam}(h) + Recall_{exam}(h)}$
+
+          The F-measure combines the precision and recall metrics via a $beta$-
+          parameterized harmonic mean.
 			\end{itemize}
-			\item Ranking:
+			\item Ranking-wise metrics:
 			\begin{itemize}
 				\item One-error: $one-error(f) = \frac{1}{p} \llbracket [argmax_{y \in Y}f(x_i,y)] \notin Y_i \rrbracket$
+
+          The One-error measure checks whether the top ranked class is within
+          the ground truth label set.
 				\item Coverage:  $coverage(f) = \frac{1}{p} \sum_{i=1}^{p} max_{y \in Y_i} rank_f(x_i,y) - 1$
+
+          The coverage measure counts the steps required to meet all members of
+          the ground truh label set in the ranked list, starting from the top.
 				\item Ranking loss: $rloss(f) = \frac{1}{p} \sum_{i=1}^{p} \frac{1}{|Y_i| {\overset{\sim}{|Y_i|}}} |\{(y', y'') | f(x_i,y') \leq f(x_i,y''), (y', y'') \in Y_i \times {\overset{\sim}{|Y_i|}} \}| $
+
+          The ranking loss measure examines the frequency of reversely ranked
+          label pairs.
 				\item Average Precision:  $agvprec(f) = \frac{1}{p} \sum_{i=1}^{p} \frac{1}{|Y_i|} \sum_{y \in Y_i} \frac{|\{y' | rank_f(x_i,y') \leq rank_f(x_i,y), y' \in Y_i\}|}{rank_f(x_i,y)}$
+
+          The average precision measures the average of labels in the ground
+          truth which are placed higher than $y_i$.
 			\end{itemize}
@@
 		\begin{itemize}
-			\item Classification: Let $B(TP_j, FP_j, TN_J, FN_J)$ represent some specific binary classification metric $(B \in \{Accuracy, Precision, Recall, F^ \beta \})$.
-The metrics from classification perspective are the following:
+			\item Classification: Let $B(TP_j, FP_j, TN_J, FN_J)$ represent some
+        specific binary classification metric $(B \in \{Accuracy, Precision,
+        Recall, F^ \beta \})$. The metrics from a classification perspective are the following:
 			\begin{itemize}
 				\item Macro-averaging: $B_{macro}(h) = \frac{1}{q} \sum_{1}^{q} B(TP_j, FP_j, TN_J, FN_J)$
+
+          Macro-averaging produces the average across all classes.
 				\item Micro-averaging: $B_{micro}(h) = B(\sum_{1}^{q}TP_j, \sum_{1}^{q}FP_j, \sum_{1}^{q}TN_j, \sum_{1}^{q}FN_j)$
+
+          Micro-averaging produces class-wise metrics by considering True /
+          False Positive / Negative elements from all classes.
 			\end{itemize}
@@
 			\begin{itemize}
-				\item \emph{AUC macro: }$AUC_{macro} = \frac{1}{q} \sum_{1}^{q} \frac{|\{(x', x'') | f(x',Y_j) \geq f(x'',Y_j), (x',x'') \in Z_j \times \overset{\sim}{|Z_j|} \}|}{|Z_j|{\overset{\sim}{|Z_j|}}}$
-				\item \emph{AUC micro: }$AUC_{micro} = \frac{|\{(x',x'',y',y'') | f(x',y') \geq f(x'',y''), (x',y') \in S^+, (x'',y'') \in S^- \}|}{|S^+||S^-|}$
+				\item Macro AUC: $AUC_{macro} = \frac{1}{q} \sum_{1}^{q} \frac{|\{(x', x'') | f(x',Y_j) \geq f(x'',Y_j), (x',x'') \in Z_j \times \overset{\sim}{|Z_j|} \}|}{|Z_j|{\overset{\sim}{|Z_j|}}}$
+				\item Micro AUC: $AUC_{micro} = \frac{|\{(x',x'',y',y'') | f(x',y') \geq f(x'',y''), (x',y') \in S^+, (x'',y'') \in S^- \}|}{|S^+||S^-|}$
+
+          The AUC (Area Under Curve) evaluation measures the model performance with respect
+          to the True and False positive rates. The expansion to the multi-label
+          scheme is accomplished by the micro and macro averaging techniques, in
+          the same way as above.
 			\end{itemize}
@@
 	\addcontentsline{toc}{section}{Multi-label learning algorithms}
-	In this paper, eight representative multi-label algorithms are presented and analyzed, seleceted with respect to below criteria:
+	In this paper, eight representative multi-label algorithms are presented and analyzed, selected with respect to below criteria:
 	\begin{itemize}
@@
   independent binary classification problems, each of them corresponding to a
-  possible label in the label-set $Y$. The first task is to construct a
-  corresponding binary training set for each label, on which a binary learning
-  algorithm is used in order to train the binary classifier that discriminates
+  possible label in the label set $Y$. The first task is to construct a
+  binary training set for each label, on which a binary learning
+  algorithm is used, constructing a classifier that discriminates the
   class $y_j$, i.e. $h_j(\cdot)$. Each training instance is involved in the
-  learning process of all $q$ classifiers(cross-training strategy), which is
-  regarded as positive instance in case of relevant labels and negative
-  otherwise. For new instances, the algorithm predicts the relevant labelset by
-  querying each binary classifier and combining the relevant labels. A label is
-  considered relevant according to the thresholding scheme adopted in the classifier.
+  learning process of all $q$ classifiers (cross-training strategy), being 
+  regarded as a positive instance in case of relevant labels and negative
+  otherwise. For new instances, the algorithm predicts the relevant label set by
+  querying each binary classifier and combining the marginal results. A label is
+  considered relevant according to the thresholding scheme adopted in the
+  multi-label process.
 
-	This is a first-order approach algorithm, since it ignores completely any
+	This is a first-order approach algorithm, since it ignores any
   label associations. However, since an one-vs-rest scheme is used and
-  classifiers are built for each label separately, the algorithm has the
-  capability of parallel implementation. Finally, this separation renders the
-  algorithm sensitive to class-imbalance in the dataset\footnote{The case where
+  classifiers are built and trained for each label separately, the algorithm has the
+  advantage of a straightforward parallel implementation. This separation
+  however renders the algorithm highly sensitive to class-imbalance in the dataset\footnote{The case where
     the number of positives and negatives for a class differs significantly}.
@@
 	The logic of this algorithm is to transform the multi-label learning problem
-  into a chain of binary classification problems, using subsequent binary
-  classifiers in a chain, built upon the predictions of preceding ones and
-  ordered by a permutation function $f_p$. First of all, a binary training set
+  into a chain of binary classification problems, using binary
+  classifiers organized in an execution chain. Each classifier is built upon the predictions of
+  preceding ones, while the class order is dictated by a permutation function $f_p$. Firstly, a binary training set
   is constructed by enriching each instance with the confidence of preceding
-  classifiers, which are concatenated with the data vector $x_i$. Moreover, a binary algorithm is used in step $i$ to
-  determine whether label $y_i$ label is relevant or not for an instance. For
-  unknown instances the algorithm predicts the relevant labels by iteratively
+  classifiers, which are iteratively concatenated with the data vector $x_i$.
+  The binary algorithm in step $i$ determines whether label $y_i$ is relevant or not for an instance. For
+  test instances, the algorithm predicts the relevant labels by iteratively
   traversing the classifier chain producing thresholded confidence scores ($\lambda _{\tau(j)}^x \in {-1, +1}$ in the
-  same order as within the training process by $f_p$ represents the predicted binary assignment of $y_{\tau(j)}$)
+  same order as within the training process (i.e., as dictated by $f_p$) represents the predicted binary assignment of $y_{\tau(j)}$).
 
-	This is a high-order method, that exploits label correlations to a degree, but in a random manner.
-However, it is highly sensitive to the permutation function, since the classifiers' ordering directly affects the result.
-One other disadvantage is that iterative operation prevents parallel implementation.
-Finally, one possible improvement to overcome the affect of classifiers' ordering is to use ensemble schemes.
+	This is a high-order method that exploits label correlations to a degree, but
+  in a random manner.
+  It is highly sensitive to the permutation function, since the classifiers'
+  ordering directly affects the result. One other disadvantage is that its
+  iterative operation prevents parallel implementation, since each classifier 
+  has to wait for the results of all previous classifiers in the chain.
+  Ensemble schemes can be used to search for and select favorable permutation functions $f_p$.
 
-	\subsubsection*{Claibrated Label Ranking}
+	\subsubsection*{Calibrated Label Ranking}
 	\addcontentsline{toc}{subsubsection}{Claibrated Label Ranking}
-	This algorithm transforms the learning problem into a label pairwise
+	This algorithm transforms the learning problem into a pairwise label
   comparison ranking problem. For $q$ labels, $q(q-1)/2$ binary classifiers are
@@
   advantage of this method is that it smooths out the class-imbalance problem
-  (where the number of a class positives is much less than the other's class
-  positives) by operating on label pairs. The disadvantage is that the number of
-  classifiers is quadratic to $|Y|$, compared to linear for Binary Relevance or
+   by operating on label pairs. The disadvantage is that the number of
+  classifiers is quadratic to $|Y|$, compared to the linear complexity for Binary Relevance or
   Classifier Chains algorithms. However, the performance can be improved by
-  using pruning methods to reduce the search space.
+  using pruning methods to reduce the labelset search space.
 
-	\subsubsection*{Random k-Labelsets}
-	\addcontentsline{toc}{subsubsection}{Random k-Labelsets}
-	The random k-labelsets algorithm transforms the multi-label learning problem
+	\subsubsection*{Random k-Label Sets}
+	\addcontentsline{toc}{subsubsection}{Random k-Label Sets}
+	The random k-label sets algorithm transforms the multi-label learning problem
   into an ensemble of multi-class classification problems. Each such partition
   targets a random subset of $\mathbb{P}(Y)$, limited to subsets present in the
-  training dataset $X$, classifying it using Label Powerset (LP) techniques. In
-  more detail, the multi-label dataset is transformed into single-label by
-  treating each distinct labelset as a new class. Thus, each training example
-  mapped to a new single class and classified through some single-label
-  classification method.
-An instance $x_i$ is assigned to label $y_i$ when the votes received for $y_i$ from the ensemble exceed half the max possible that this label can get.
-For unseen instances $x$, LP predicts its associated label set $Y$ by querying the prediction of multi-class classifier and then mapping it back to the power set of $Y$.
+  training dataset $X$, classifying it using Label Powerset (LP) techniques.
+  These techniques involve transforming the multi-label dataset into
+  single-label data by treating each distinct label set as a new class. Thus, each training example
+  is mapped to a new, single class and classified through some single-label multiclass
+  classification method. An instance $x_i$ is assigned to a label $y_i$ when the
+  votes received for $y_i$ from the ensemble exceed half the max possible that
+  this label can get.
+  For unseen instances $x_j$, LP predicts its associated label set $Y_j$ by
+  querying the single-label predictions of the LP multi-class classifiers and then mapping these
+  results it back to the labelsets in $\mathbb{P}(Y)$.
 
@@
 	\begin{itemize}
-		\item[$\diamond$] \emph{Incompleteness: }The algorithm is data-sensitive and cannot generalize to labelsets that do not appear into the training set.
+		\item[$\diamond$] \emph{Incompleteness: }The algorithm is data-sensitive and cannot generalize to label sets that do not appear into the training set.
 		\item[$\diamond$] \emph{Inefficiency: }A large label set $Y$ implies high training complexity and extremely few training examples for some newly mapped classes.
@@
 
-	A simple but powerful modification is the Random k-Labelsets variant, which
+	A simple but powerful modification is the Random k-Label Sets variant, which
   runs an ensemble of executions, each with $k$-sized label sets. The ensemble
-  scheme imporives on the predictive completeness of the system, while the
-  limited labelset size guarantees computational efficiency. Additionally,
-  pruning distinct labelsets which are less frequent than a lower threshold is
-  another strategy to improve the method's performance.
+  scheme improves on the predictive completeness of the system, while the
+  limited label set size guarantees computational efficiency. Additionally,
+  pruning distinct label sets which are less frequent than a specified threshold is
+  another common strategy to improve the method's performance.
 
@@
 	The ML-kNN algorithm adapts nearest neighbor techniques to the multi-label
-  setting by utilizing the maximum a posteriori (MAP) rule, to make a prediction
-  by reasoning with the labeling information embodied in the neighbors. For new
-  instances $x$, $N(x)$ represents the set of its $k$ nearest neighbors
+  setting by utilizing the maximum a posteriori (MAP) rule. The method produces
+  predictions by taking into account the labeling information embodied in the neighbors. For new
+  instances $x$, the set $N(x)$ represents its $k$ nearest neighbors
   identified in the dataset by a specified similarity or distance measure. Let
   $H_j \equiv (y_j \in Y_i)$, $C_j = \sum_{x_k \in N(x_i)} [\delta(y_j \in
-  Y_k)]$, where $H_j$ is the event that $\textbf{x}$ has the label $y_j$ and
-  $C_j$ records the number of x’s neighbors with label $y_j$. Moreover,
+  Y_k)]$, i.e. $H_j$ is the event that $\textbf{x}$ has the label $y_j$ and
+  $C_j$ records the number of neighbours of $x$ that have the label $y_j$. Moreover,
   $P(H_j|C_j)$ is the posterior probability that $H_j$ holds in the condition
@@
 
-	The prior probabilities ($P(H_j)$, $P(\neg H_j)$) are computed via a frequency counting strategy, where a smoothing parameter $s$ is used to control the effect of uniform prior on the estimation (usually $s=1$).
-In addition, likelihoods $P(C_j|H_j)$ and $P(C_j| \neg H_j)$ are computed using two frequency arrays:
+	The prior probabilities ($P(H_j)$, $P(\neg H_j)$) are computed via a frequency
+  counting strategy, where a smoothing parameter $s$ is used to control the
+  effect of uniform prior on the estimation (a common value is  $s=1$). In addition, likelihoods $P(C_j|H_j)$ and $P(C_j| \neg H_j)$ are computed using two frequency arrays:
 	\begin{itemize}
@@
 	This method adapts decision tree techniques for multi-label classification.
-  Using the multi-label entropy measure, decision tree is built recursively by
-  splitting features that yield the maximum information gain. Let $T = \{(x_i,
-  Y_i) | 1 \leq i \leq n \}$ be the multi-label dataset with n examples. The
+  Using the multi-label entropy measure, a decision tree is built recursively by
+  splitting features in positions such that a maximum information gain is produced. Let $T = \{(x_i,
+  Y_i) | 1 \leq i \leq n \}$ be the multi-label dataset with $n$ examples. The
   algorithm splits the $x$ vector at the feature $l$ that maximizes the
-  information gain criterion, where that value in $x$ is  $\theta$. This
+  information gain criterion, where $\theta$ is the value in that position. This
   partitions the dataset $T$ into branches ($T^-$ \& $T^+$), where $x_{il} \leq
   \theta, x_i \in T^-$ and $x_{il} > \theta, x_i \in T^+$. This process is
-  invoked recursively by treating each time one of the two branches as the new
+  invoked recursively by treating each new branch as the new
   root node and terminates until some stopping criterion $C$ is met (e.g.
-  subtree size). Each label subset is treated as a new class and single-label entropy is computed. Any new instances are assigned the label of the majority of members of the leaf that they arrive.
+  subtree size). Each label subset is treated as a new class and the single-label entropy is computed. Any new instances are assigned the label of the majority of members of the leaf that they arrive.
 
@@
 	Rank-SVM adapts the maximum margin strategy to the multi-label classification
-  problem. $q$ linear classifiers are optimized to minimize the empirical
+  problem. $q$ linear classifiers are optimized with the empirical
   ranking loss. The learning system is composed by classifiers that distinguish
@@
 Each classifier is parameterized with $W = \{w_j,b_j | 1 \leq j \leq q\}$, where $w_j \in \mathbb{R}^d$ and $b_j \in \mathbb{R}$ are the weight vector and the bias
-	for the j-th class label $y_j$ respectively. Rank-SVM defines the system’s
+	for the j-th class label $y_j$, respectively. Rank-SVM defines the system’s
   margin on $(x_i,Y_i)$ by considering its ranking ability on the relevant and
@@
   related-unrelated pairs and is minimized along with the training loss
-  function. At consistent training, the classifier will return positive margin.
-  For new test instances, the algorithms retrieves a ranked list of label pairs
-  according to the signed distance to the decision hyperplane.
+  function. At consistent training, the classifier will return positive margin
+  for each training example, since the singed distance will be positive with
+  respect to relevant label hyperplanes and negative for irrelevant ones.
+  For new test instances, the algorithms retrieves a ranked list of label pairs, ordered
+  according to the signed distance to the separation hyperplane.
 
@@
   solver.
-  This algorithm is a second-order approach using pairwise label associations.
+  This algorithm is a second-order approach as it utilizes pairwise label associations.
   It is adaptable to a variety of tasks and data by using alternative loss
-  functions and kernel feature mapping. Kernel selection can be accomplished by
+  functions and kernel feature mapping. Kernel selection for the latter can be accomplished by
   multiple kernel learning (MKL) techniques.
@@
 
-	 This algorithm involves the maximum entropy principle, where correlations
-   among labels are encoded as binary random vectors into a joint probability
-   distribution function and label correlations are encoded into the
+	 This algorithm involves computing a joint probability distribution on $x, Y$
+   where label sets are encoded as binary random vectors and label correlations are encoded into the
    distribution as constraints.
@@
    The latter are modeled as $ \mathbb{E} [ f_k(x,\bm{y}) ] = F_k  $, where $F_k$ is
- learned from the traininng data.
- The optimal solution is arrived at via Lagrange multipliers, assuming gaussian
+ learned from the training data.
+ The optimal solution is arrived at via Lagrange multipliers, assuming Gaussian
  priors for the variables involved, and has the form of a Gibbs probability
@@
    This is a second-order approach, where all label pairs are considered and not
-   just the relevant-irrelevant ones which was the Rank-SVM approach. It is a
+   just the relevant-irrelevant ones which was the case with the Rank-SVM approach. It is a
    convex constraint optimization problem, and is thus solvable by off the shelf
@@
    the method very computationally inefficient for large label spaces, if
-   pruning is not applied. Varying the expression form of the constraints gives rise to difference
+   pruning is not applied. Finally, varying the expression form of the constraints gives rise to difference
    version of the CML method.
@@
  	\addcontentsline{toc}{section}{Related Learning Methods}
- 	Four related learning methods are mentioned in this paper, which can be found below:
+ 	Four related learning methods are mentioned in this paper, which are summarized below:
  	\begin{enumerate}
@@
       with a single binary label. The bag is assigned a positive label in the
-      case of at least one positively labelled member. It models complex semantics of $x_i$ in input space rather than its output.
+      case of at least one positively labelled member. It models complex
+      semantics of $x_i$ in \emph{input space} (i.e. in the variability of each
+      bag of instances) rather than its output (i.e. the different labels in the instance's
+      label set).
  		\item \textbf{Ordinal classification: } This scheme assumes label relevance
@@
 	\addcontentsline{toc}{section}{Conclusion}
-	Summarizing, this paper defines the multi-label learning problem, presenting 8
-  multi-label learning representative algorithms, mentions some additional
-  related learning settings. Some important future goals are the formal
-  characterization on the underlying concept and / or mechanism on the
-  appropriate usage of label correlations (especially on large output spaces).
+	Summarizing, this paper defines the multi-label learning problem, presenting
+  an outline of evaluation methods and describing 8 multi-label learning
+  representative algorithms. It closes by mentioning some learning scenarios
+  related to multi-label learning.
+
+  Some important future goals in the field are the formal description of the
+  underlying concept behind label correlations, and of the mechanism by which interdependencies ought to
+  be handled and modelled in learning problems - especially in large output spaces. 
   In addition, thorough experimental comparative studies are essential to
@@
   algorithms, so as to aid the conquest of the aforementioned theoretical goal.
+
   Some online resources on multi-label learning are presented on figure \ref{figure:resources}.
@@
 \bibitem{trees} Vens, Celine, et al. "Decision trees for hierarchical multi-label classification." Machine Learning 73.2 (2008): 185-214.
-\bibitem{rkl} Tsoumakas, Grigorios, and Ioannis Vlahavas. "Random k-labelsets: An ensemble method for multilabel classification." Machine learning: ECML 2007 (2007): 406-417.
+\bibitem{rkl} Tsoumakas, Grigorios, and Ioannis Vlahavas. "Random k-label sets: An ensemble method for multilabel classification." Machine learning: ECML 2007 (2007): 406-417.
 \bibitem{mlknn} Zhang, Min-Ling, and Zhi-Hua Zhou. "ML-KNN: A lazy learning approach to multi-label learning." Pattern recognition 40.7 (2007): 2038-2048.

### 2018-01-28 Project/main.py
Minor changes

@@ -24,10 +24,10 @@ def question_2(opts):
     print("=====================")
     # Question 2
     print("\n>>> Running question 2a1 - Nearest neighbours computation")
-    # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"])
+    qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"])
 
     print("\n>>> Running question 2a2 - Nearest subroutes computation")
-    # qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],opts["conseq_lcss"], opts["k"], opts["paropts"])
+    qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],opts["conseq_lcss"], opts["k"], opts["paropts"])
 
     print("\n>>> Running question 2b - Cell grid quantization")
     cellgrid = opts["grid"]
@@ -67,9 +67,9 @@ if __name__ == '__main__':
     np.random.seed(options["seed"])
     random.seed(options["seed"])
 
-    # paropts = ("processes", 10)
+    paropts = ("processes", 10)
     # paropts = ("threads", 10)
-    paropts = None
+    #paropts = None
     options["paropts"] = paropts
 
     # question 1
@@ -90,7 +90,7 @@ if __name__ == '__main__':
     os.makedirs(options["classifdir"], exist_ok=True)
 
     # run
-    # question_1(options)
+    question_1(options)
 
     # question 2
     ############

### 2018-01-28 Project/question2.py

@@ -9,9 +9,9 @@ import pickle
 
 
 def question_a1(output_folder, clean_file, test_file, paropts, k):
-    print("Extracting %d nearest neighbours for each test trip" % k)
     test_df = pd.read_csv(test_file, delimiter="\n")
     train_df = pd.read_csv(clean_file)
+    print("Extracting %d nearest neighbours out of %d cleaned train data, for each test trip" % (k,len(train_df)))
     for index, row in test_df.iterrows():
         print("Examining test element %d / %d" % (index + 1, len(test_df)))
         outfile_name = os.path.join(output_folder, "nn_%d_" % (index + 1))

### 2018-01-28 ProjectReport/BigDataHomework.tex

@@
 
-    In the main.py file, it is required to provide the path to the input folder (containing the training and test files) and the output folder where all results will be stored.
+    In the \texttt{main.py} file, it is required to provide the path to the input folder (containing the training and test files) and the output folder where all results will be stored.
 Afterwards, an object containing all the necessary for the program options is created.
@@
 	In this section, further details are provided related to the preprocessing and cleaning of the training data as well as to the trip's visualization.
-All files implementing the requirements of this part are in the file question1.py.
+All files implementing the requirements of this part are in the file \texttt{question1.py}.
 
@@
 The data with null journeyPatternId should be removed from the dataframe, therefore the lines are checked for NaN and "null" journey id.
-Then, the columns vehicleID and journeyPatternId are groupedby three times in order to create three different dataframes.
+Then, the columns vehicleID and journeyPatternId are grouped by three times in order to create three different dataframes.
 For example, the first dataframe contains the columns: vehicleID, journeyPatternId and timpestamp.
@@
 
-	The above logic is implemented in function filter\_trips.
-The two lists (trips\_too\_small and trips\_too\_big) will keep the excluded trips due to either not valid total distance or to invalid max distance between successive points.
-For each trip in the training dataset the total distance is calculated, using the function calculate\_lonlat\_distance, which returns the distance between two successive points using the haversine formula.
+	The above logic is implemented in function \texttt{filter\_trips}.
+The two lists (\texttt{trips\_too\_small and trips\_too\_big}) will keep the excluded trips due to either not valid total distance or to invalid max distance between successive points.
+For each trip in the training dataset the total distance is calculated, using the function \texttt{calculate\_lonlat\_distance}, which returns the distance between two successive points using the haversine formula.
 Following the same logic, the max distance between successive points of a trip is calculated and a distance is found more than 2km, the trip is excluded.
 
-	Finally, the distances are checked and rows with invalid trips are removed from the dataframe, which is, afterwards, exported in a new file (trips\_clean.csv).
+	Finally, the distances are checked and rows with invalid trips are removed from the dataframe, which is, afterwards, exported in a new file (\texttt{trips\_clean.csv}).
 
 	\subsection{Trip Visualization (question C)}
-	This task is implemented in the function visualize\_trips and the created output is stored in the output folder in the gmplots directory.
+	This task is implemented in the function \texttt{visualize\_trips} and the created output is stored in the output folder in the gmplots directory.
 The requirement is to visualize five of the given train trips.
-For this purporse, the trips list is shuffled in order to visualize random trips and then iterated only for 5 trips.
-Then, a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).
+For this purpose, the trips list is shuffled in order to visualize random trips and then iterated only for 5 trips.
+Then, a points list is created having two tuples with all the longitudes and latitudes(function \texttt{idx\_to\_lonlat} in the \texttt{utils.py}).
 
-	Finally, the function write\_group\_gml from the utils.py file is called.
+	Finally, the function \texttt{write\_group\_gml} from the \texttt{utils.py} file is called.
 In this function, it is calculated the mean for both longitudes and latitudes and then gmplot is used so as to visualize the selected trips.
@@
 	\section{Exercise 2}
-	The files implementing the requirements of this part are in the file question2.py and some functions implementing the logic for this question are in the files: NearestNeighbors.py, NearestSubroutes.py, MapInGridView, JourneyClassification.py.
+	The files implementing the requirements of this part are in the file question2.py and some functions implementing the logic for this question are in the files: \texttt{NearestNeighbors.py, NearestSubroutes.py, MapInGridView, JourneyClassification.py}.
 
@@
 
-	In the function question\_a1, two dataframes are created for both the test and the training trips.
-The test dataframe's rows are iterated and for each row the function calculate\_nns from the file NearestNeighbors.py is called in order to get the nearest neighbors, calculating also the time needed for this action.
+	In the function \texttt{question\_a1}, two dataframes are created for both the test and the training trips.
+The test dataframe's rows are iterated and for each row the function \texttt{calculate\_nns} from the file NearestNeighbors.py is called in order to get the nearest neighbors, calculating also the time needed for this action.
 
@@
 This information is provided as input giving the kind of execution and the number of processes/threads as well.
-In the function run\_with\_processes an async task is executed for each process and then the result is stored in a list, while in the function run\_with\_threads, the dataframe is splitted in sub dataframes (the number is specified by the number of threads) and then distance is calculated for each subdataframe.
-After thread join, the result is also stored in the nearest\_neighbors list.
+In the function \texttt{run\_with\_processes} an async task is executed for each process and then the result is stored in a list, while in the function \texttt{run\_with\_threads}, the dataframe is splitted in sub dataframes (the number is specified by the number of threads) and then distance is calculated for each subdataframe.
+After thread join, the result is also stored in the \texttt{nearest\_neighbors} list.
 
-	In the function calculate\_dists, the 3rd column of the dataframe is provided and it is transformed in tuples with the latitudes and longitudes.
+	In the function \texttt{calculate\_dists}, the 3rd column of the dataframe is provided and it is transformed in tuples with the latitudes and longitudes.
 The list nearest neighbors will keep in tuples for each train trip the trip\_id, the calculated distance (which is calculated using the dynamic time wraping algorithm), the journeyId and finally all the points of the journey.
@@
 	The final list is sorted and only the 5 nearest neighbours are kept for visualization.
-However, before the visualization, the preprocessing\_for\_visualization function is used, where three list are created: a list with all the points (longitudes and latitudes) of the test trip and its neighbours, a list with the labels and a list with the colors (in this question the only color is the blue).
-All this data is provided to the function visualize\_paths in the utils.py file in order to export the trips as map images.
+However, before the visualization, the \texttt{preprocessing\_for\_visualization function} is used, where three list are created: a list with all the points (longitudes and latitudes) of the test trip and its neighbours, a list with the labels and a list with the colors (in this question the only color is the blue).
+All this data is provided to the function \texttt{visualize\_paths} in the \texttt{utils.py} file in order to export the trips as map images.
 
-	In order to be able to create an image containing all the six maps each time, the rasterize.js file is required as well as phantomjs.
+	In order to be able to create an image containing all the six maps each time, the \texttt{rasterize.js} file is required as well as \texttt{phantomjs}.
 In this function, both .html and .jpg files are created.
-First of all, the function write\_group\_gml is used so as to create the html files (this function is described above in question c).
-These html files are converted to images using the function html\_to\_png.
+First of all, the function \texttt{write\_group\_gml} is used so as to create the html files (this function is described above in question c).
+These html files are converted to images using the function \texttt{html\_to\_png}.
 File names are kept in a list and when the final image file is created (containing all six maps) these files are deleted.
@@
 	\subsection{Nearest Sub-routes (question A2)}
-	Following the same logic as above, in the function question\_a2, two dataframes are created for the test and the training dataset using pandas library.
+	Following the same logic as above, in the function \texttt{question\_a2}, two dataframes are created for the test and the training dataset using pandas library.
 The test dataframe is iterated row by row, in order to find the 5 trips from the training dataset which have the longest common subsequence with the specific test trip.
 
-	Similarly to the above question, the points are stored as tuples and the required time is calculated using the relevant methods from the utils.py file.
-To begin with, the method calc\_lcss is called.
+	Similarly to the above question, the points are stored as tuples and the required time is calculated using the relevant methods from the \texttt{utils.py} file.
+To begin with, the method \texttt{calc\_lcss} is called.
 In this function, the list of lonlat coordinate tuples are given as input.
 We keep all similar point values and indexes in two lists while iterating the two given lists.
-The distance between two points is calculated using the method calculate\_lonlat\_distance in the question1.py file.
+The distance between two points is calculated using the method \texttt{calculate\_lonlat\_distance} in the \texttt{question1.py} file.
 Based on the exercise requirements, two points are considered as equals if their distance is less than 200 metres.
@@
 
-	Having now found the longest subsequences, we sort the indexes based on the length of the subsequences and we call the update\_current\_maxsubseq in order to update our result comparing previous subsequences.
+	Having now found the longest subsequences, we sort the indexes based on the length of the subsequences and we call the \texttt{update\_current\_maxsubseq} in order to update our result comparing previous subsequences.
 First of all, the new sequences found are iterated, checking if the length of the subsequence list has reached the given number (in our case 5).
@@
 This time the points are separated so as to show which of them belong to the nearest sub-route and color them with red color.
-The function for the visualization is the same that was used for the question 2a1 in the utils.py file.
+The function for the visualization is the same that was used for the \texttt{question 2a1} in the \texttt{utils.py} file.
 
@@
 
-	In order to be able to draw the grid, it was necessary to find the min and max longitudes and latitudes (function: find\_min\_max\_latlong, file:MapInGridView.py).
-Having now this information, the next function called is the create\_grid giving as input the number\_of\_cells and the result of the previous method.
+	In order to be able to draw the grid, it was necessary to find the min and max longitudes and latitudes (function: \texttt{find\_min\_max\_latlong}, file: \texttt{MapInGridView.py}).
+Having now this information, the next function called is the create\_grid giving as input the \texttt{number\_of\_cells} and the result of the previous method.
 The distance per cell in the grid is calculated as the difference between the max and the min lat/long divided by the total number of cells.
@@
 
-	Afterwards, all points in the trip dataframe should be replaced by the cell names that each point corresponds to.
-This logic is implemented in the method replace\_points.
-For each row in the dataframe, all coordinates are transformed in tuples format and using the find\_index function a name is mapped to this point.
+	For the features extraction, we have used two methods.
+In the first method, all points in the trip dataframe should be replaced by the cell names that each point corresponds to.
+This logic is implemented in the methods \texttt{map\_to\_features} and \texttt{map\_to\_features\_pointwise}.
+For each row in the dataframe, all coordinates are transformed in tuples format and using the \texttt{find\_index} function a name is mapped to this point.
 After having replaced all points with their new names, the dataframe is exported in .csv format.
-In order to be able to easily check the created grid, the method visualize\_grid is implemented so as to plot the grid.
+In order to be able to easily check the created grid, the method \texttt{visualize\_grid} is implemented so as to plot the grid.
+
+	The second method is implemented in the function \texttt{map\_to\_features\_bow}.
+The difference here is that we create a vector for each trip with size equal to the number of cells in the grid.
+The vector is populated with the number of points that belong to the respective cell.
+All these vectors are added in a list which is exported in a csv file.
 
 	\subsection{Classification (question C)}
-	In the method question\_c, a new dataframe is created from the features file, containing the grid cell names.
-All functions used in this question are in the file JourneyClassification.py.
-Before starting the classification, the method preprocess\_data was used in order to get the journey\_id of all trips as numbers and also to remove the letter C from the cell names.
+	In the method \texttt{question\_c}, a new dataframe is created from the features file, containing the grid cell names.
+All functions used in this question are in the file \texttt{JourneyClassification.py}.
+Before starting the classification, the method \texttt{preprocess\_train\_data} was used in order to get the journey\_id of all trips as numbers and also to remove the letter C from the cell names.
 
-	In the requirements of this question it is mentioned to use 10-fold cross validation, therefore the KFold class is used providing the required folds number and then the training dataset is splitted in training and test folds.
-It worths to be mentioned here, that method asarrays from numpy was used to the, necessary for this question, lists.
+	Afterwards, the \texttt{train} function we use 10-fold cross validation, therefore the \texttt{KFold} class is used providing the required folds number and then the training dataset is splitted in training and test folds.
+It worths to be mentioned here, that method \texttt{asarrays} from \texttt{numpy} was used to the, necessary for this question, lists.
 In addition, a list with the names of the three classifiers is created and for each of them, the relevant function is called in order to train the classifier and then count the predictions' accuracy on the test dataset.
 
-	In this question, the scikit learn library was used.
-Firstly, for the knn classifier the class KNeighborsClassifier was used, and more specifically the fit and predict methods.
-In order to calculate the accuracy of the classifiers we have used the accuracy\_score method (this result is return from each classification function in the JourneyClassification.py), but also the classification\_report in order to print the precision and recall values for each prediction of the classifier.
+	In this question, the \texttt{scikit learn} library was used.
+Firstly, for the knn classifier the class \texttt{KNeighborsClassifier} was used, and more specifically the fit and predict methods.
+In order to calculate the accuracy of the classifiers we have used the \texttt{accuracy\_score} method (this result is return from each classification function in the \texttt{JourneyClassification.py}), but also the \texttt{classification\_report} in order to print the precision and recall values for each prediction of the classifier.
 The same logic is followed for the other two classifiers(logistic regression and random forest).
-What should be mentioned, though, is that the LogisticRegression class implements the predict\_proba method, which returns a list with probabilities and argmax from numpy was used in order to determine the result with the max probability.
+What should be mentioned, though, is that the \texttt{LogisticRegression} class implements the \texttt{predict\_proba} method, which returns a list with probabilities and \texttt{argmax} from \texttt{numpy} was used in order to determine the result with the max probability.
 
-	\subsubsection{Beat ​ ​the ​ ​Benchmark}
+	\subsubsection{Improve classification}
+	\subsection{Test improved classifier}
 

### 2018-02-11 ProjectReport/BigDataHomework.tex
Update report, minor changes

--- 2018-01-16 ProjectReport/BigDataHomework.tex
@@
 
-\title{Spatiotemporal Data Classification}   % Title
+\title{Project report}   % Title
 \author{ Nikiforos Pittaras, M1422 \\ 
@@
 
+  Some unclear points in the project instructions where dealt with respective
+  variables and multiple implementations. These points where:
+  \begin{itemize}
+  \item Unique journey ids in subroutes (\texttt{unique\_jids} parameter,
+    default true.)
+    \item Consequtive common subsequences (\texttt{conseq\_lcss} parameter,
+      default true.)
+    \end{itemize}
+
 	\section{Exercise 1}
 	In this section, further details are provided related to the preprocessing and cleaning of the training data as well as to the trip's visualization.
-All files implementing the requirements of this part are in the file question1.py.
+All files implementing the requirements of this part are in the file \texttt{question1.py}.
 
 	\subsection{Data Preprocessing (question A)}
-	The requirement in this part is to preprocess the input training data and create a new file with 3 columns, containing the trip\_id (a number incrementing by 1 for each trip), the journey\_id(excluding trips with null id) and all points and timestamps in the format: [[timestamp1, longitude1, latitude1], [timestamp2, longitude2, latitude2],...].
-The given csv file with the training data was transformed to a dataframe using the pandas library.
-The data with null journeyPatternId should be removed from the dataframe, therefore the lines are checked for NaN and "null" journey id.
-Then, the columns vehicleID and journeyPatternId are groupedby three times in order to create three different dataframes.
-For example, the first dataframe contains the columns: vehicleID, journeyPatternId and timpestamp.
-In a similar way, the two remaining dataframes were formatted.
-By concatenating them, the final dataframe contains five columns, where the three last columns contain a list of timestamps, a list of longitudes and a list of latitudes respectively.
+	The goal in this part is to preprocess the input training data and create a
+  new file with 3 columns:
+  \begin{itemize}
+  \item  the trip\_id (a number incrementing by 1 for each trip)
+  \item the journey\_id(excluding trips with null id)
+  \item all points and timestamps in the format: [[timestamp1, longitude1,
+    latitude1], [timestamp2, longitude2, latitude2],...].
+  \end{itemize}
+  The given csv file with the training data was transformed to a dataframe using
+  the pandas library. Data with null journey ids are removed from the dataframe,
+  therefore lines with NaN or "null" journey ids are discarded. All data is
+  subsequently grouped by the vehicleID and the journeyPatternId.
+The final dataframe contains five columns, where the three last columns contain a list of timestamps, a list of longitudes and a list of latitudes respectively.
 
-	By iterating to each dataframe's row, the three lists are now concatenated in one column and therefore the other two are dropped from the dataframe.
+	To produce the required final format [timestamp, longitude, latitude], we
+  iterate each row, producing a column with the appropriate list data.
+  % the three lists are now concatenated in one column and therefore the other two are dropped from the dataframe.
 Then, the column vehicleID is replaced by the tripId column which is the index of this dataframe, starting from 0 and incrementing by 1.
@@
 	\subsection{Clean Data (question B)}
-	The second goal was to clean the given data, based on two criteria: total trip distance should not be less than 2km while max distance between two successive points should not exceed 2km.
+	We clean the given data based on two criteria:
+  \begin{enumerate}
+  \item the total trip distance should not be less than 2km
+  \item the max distance between two successive points should not exceed 2km.
+    \end{enumerate}
 
-	The above logic is implemented in function filter\_trips.
-The two lists (trips\_too\_small and trips\_too\_big) will keep the excluded trips due to either not valid total distance or to invalid max distance between successive points.
-For each trip in the training dataset the total distance is calculated, using the function calculate\_lonlat\_distance, which returns the distance between two successive points using the haversine formula.
-Following the same logic, the max distance between successive points of a trip is calculated and a distance is found more than 2km, the trip is excluded.
+	The above logic is implemented in function \texttt{filter\_trips}. The two
+  lists (\texttt{trips\_too\_small and trips\_too\_big}) will keep the trips
+  discarded due to the aforementioned criteria. For each trip in the training
+  dataset the total distance is calculated using the function
+  \texttt{calculate\_total\_distance}, which uses the
+  \texttt{calculate\_lonlat\_distance} function that returns the distance
+  between two points using the haversine formula. Using these functions we
+  compute the required distances and discard the trips that match the exclusion
+  criteria.
 
-	Finally, the distances are checked and rows with invalid trips are removed from the dataframe, which is, afterwards, exported in a new file (trips\_clean.csv).
+	The resulting filtered trips are stored in the file \texttt{trips\_clean.csv}.
 
 	\subsection{Trip Visualization (question C)}
-	This task is implemented in the function visualize\_trips and the created output is stored in the output folder in the gmplots directory.
-The requirement is to visualize five of the given train trips.
-For this purporse, the trips list is shuffled in order to visualize random trips and then iterated only for 5 trips.
-Then, a points list is created having two tuples with all the longitudes and latitudes(function idx\_to\_lonlat in the utils.py).
-
-	Finally, the function write\_group\_gml from the utils.py file is called.
-In this function, it is calculated the mean for both longitudes and latitudes and then gmplot is used so as to visualize the selected trips.
-Below, there is an example of a trip shown on map is presented:
+	This trip visualization task is implemented in the function
+  \texttt{visualize\_trips} and the created output is stored in the output
+  folder in the \texttt{gmplot} directory. The trips list is shuffled in order
+  to show random items, the first 5 of which are fetched for
+  visualization. For each trip, we generate a tuple of longitudes and latitudes
+  with the function \texttt{utils.idx\_to\_lonlat} which we pass on to the
+  plotting function \texttt{utils.write\_group\_gml}. There, the longitude and
+  latitude means are calculated as a visualization centroid for the produced map, and the gmplot
+  library is used so as to visualize the respective point sequence. See the
+  following figure for a visualization example.
 
@@
 			\includegraphics [scale = 0.75] {questionCexample.jpg}
-			\caption{Example of trip visualization}
+			\caption{Example of a trip visualization.}
 		\end{center}
+    \label{gmlplot_example}
 	\end{figure}
@@
 	\section{Exercise 2}
-	The files implementing the requirements of this part are in the file question2.py and some functions implementing the logic for this question are in the files: NearestNeighbors.py, NearestSubroutes.py, MapInGridView, JourneyClassification.py.
+	This exercise is implemented in the file question2.py and some functions implementing the logic for this question are in the files: \texttt{NearestNeighbors.py, NearestSubroutes.py, MapInGridView, JourneyClassification.py}.
 
 	\subsection{Nearest Neighbors (question A1)}
-	For this part a test file is given (test\_set\_a1.csv, included in the input folder) containing additional trips.
-For each of these trips, the k nearest neighbors are targeted (in this exercise k=5) and an image file is produced, showing all six trips (the test trip and the trips from the training file) on six different maps.
+	For this part, a test file is given (test\_set\_a1.csv, included in the input
+  folder) containing additional test trip data. For each test trip, the k
+  nearest neighbors are extracted (in this exercise k=5) and an image file is
+  produced, showing all six trips (the test trip and its 5 neighbour trips from the training file) on six different maps.
 Under each map, the test or neighbor index, the journey\_id, the dynamic time warping and the time needed to process the neighbors of this test trip in millisecond.
 
-	In the function question\_a1, two dataframes are created for both the test and the training trips.
-The test dataframe's rows are iterated and for each row the function calculate\_nns from the file NearestNeighbors.py is called in order to get the nearest neighbors, calculating also the time needed for this action.
+	In the function \texttt{question\_a1}, two dataframes are created for both the
+  test and the training trips. For each test trip, \texttt{calculate\_nns} from
+  the file NearestNeighbors.py is called in order to get the nearest neighbors
+  with respect to all training trips, calculating also the time needed for this calculation.
 
-	There are three ways to execute this task: using processes, threads or serial execution.
-This information is provided as input giving the kind of execution and the number of processes/threads as well.
-In the function run\_with\_processes an async task is executed for each process and then the result is stored in a list, while in the function run\_with\_threads, the dataframe is splitted in sub dataframes (the number is specified by the number of threads) and then distance is calculated for each subdataframe.
-After thread join, the result is also stored in the nearest\_neighbors list.
+	In \texttt{calculate\_nns}, we explored parallelization options to compute the neighbours,
+  due to the computational intensivenes of the task.
+  \begin{enumerate}
+  \item using multiple processes (\texttt{run\_with\_processes} function) with the
+    \texttt{multiprocessing} module, where an async task is executed for each process.
+  \item multiple threads (\texttt{run\_with\_threads} function) with the \texttt{threading} module, where the task is
+    broken to multiple threads with separate result containers.
+  \item standard serial execution, where each training datum is examined iteratively.
+  \end{enumerate}
+  This parallelization type and parameter (i.e. number of processes or threads) is contained in tie \texttt{paropts} variable in main.
 
-	In the function calculate\_dists, the 3rd column of the dataframe is provided and it is transformed in tuples with the latitudes and longitudes.
-The list nearest neighbors will keep in tuples for each train trip the trip\_id, the calculated distance (which is calculated using the dynamic time wraping algorithm), the journeyId and finally all the points of the journey.
+	In the function \texttt{calculate\_dists}, the 3rd column of the dataframe is
+  provided and
+  In the function, each training trip's time series data is transformed to a
+  list of (longitude, latitude) tuples, which is passed to the \texttt{calculate\_dynamic\_time\_warping}
+  function to calculate the DTW distance. The latter is computed using the
+  harvesine distance calculation function and the DTW wikipedia lemma. The
+  resulting list contains the training instance information along with its distance to
+  the test trip. Sorting the list to increasing distance and keeping its k first
+  elements provides the k nearest neighbours.
 
-	The algorithm dynamic time warping is implemented with two ways based on the input (diy or diy\_initial). \todo{diy and diyinitial}
-
-	The final list is sorted and only the 5 nearest neighbours are kept for visualization.
-However, before the visualization, the preprocessing\_for\_visualization function is used, where three list are created: a list with all the points (longitudes and latitudes) of the test trip and its neighbours, a list with the labels and a list with the colors (in this question the only color is the blue).
-All this data is provided to the function visualize\_paths in the utils.py file in order to export the trips as map images.
-
-	In order to be able to create an image containing all the six maps each time, the rasterize.js file is required as well as phantomjs.
-In this function, both .html and .jpg files are created.
-First of all, the function write\_group\_gml is used so as to create the html files (this function is described above in question c).
-These html files are converted to images using the function html\_to\_png.
-File names are kept in a list and when the final image file is created (containing all six maps) these files are deleted.
-Afterwards, images as displayed as collection of plots and exported in the .jpg format.
-
+  To visualize the neighbours, the \texttt{visualize\_nns}
+  function is used, where three lists are created:
+  \begin{enumerate}
+    \item a list with all the route points (longitudes and latitudes) of the
+      test trip and its neighbours
+    \item a list with the journey labels and the nn computation data
+      \item a list with the colors with which to draw (all blue in this
+        question).
+      \end{enumerate}
+      All this data is provided to the
+      \texttt{utils.visualize\_point\_sequences} function. This function is
+      supplied with point sequences with corresponding color sequences, and
+      produces colored gml plot maps with the \texttt{write\_group\_gml}
+      function. Each html output map is converted to a png image using a conversion tool in the \texttt{html\_to\_png} function, in
+order to automate image creation from the generated maps (the tool is the
+\texttt{rasterize.js} utility from the \texttt{phantomjs} module).
+However we found the gml plot zoom difficult to automatically configure correctly so that
+all points are visible, and thus the images in the report were fetched with a printscreen.
+The converted images are used to create 6 subplots (1 test trip + k=5 nearest
+neighbours). All plotting is done with the pylab package.
+l 
 	An example of this question image output can be found below:
@@
 		\begin{center}
-			\includegraphics [scale = 0.50] {question2a1example.jpg}
+			\includegraphics [scale = 0.40] {question2a1example.jpg}
 			\caption{Example of nearest neighbours trips visualization}
@@
 	\subsection{Nearest Sub-routes (question A2)}
-	Following the same logic as above, in the function question\_a2, two dataframes are created for the test and the training dataset using pandas library.
-The test dataframe is iterated row by row, in order to find the 5 trips from the training dataset which have the longest common subsequence with the specific test trip.
+	Following the same logic as above, in the function \texttt{question\_a2} we
+  compute the nearest subroutes of each test trip with the following key
+  differences:
+  \begin{itemize}
+    \item Instead of DTW, we use the LCSS algorithm in the \texttt{calc\_lcss}
+      method, using the harvesine formula for pointwise distance as before. Give
+      a test and train point sequence, the LCSS function returns a list of all
+      common subsequence points and indexes of the two sequence inputs. The
+      parameter \texttt{unique\_jids} controls wether more than one subsequence
+      is allowed per training trip (default false). Based on the exercise
+      requirements, two points are considered as equals if their distance is less than 200 metres.
 
-	Similarly to the above question, the points are stored as tuples and the required time is calculated using the relevant methods from the utils.py file.
-To begin with, the method calc\_lcss is called.
-In this function, the list of lonlat coordinate tuples are given as input.
-We keep all similar point values and indexes in two lists while iterating the two given lists.
-The distance between two points is calculated using the method calculate\_lonlat\_distance in the question1.py file.
-Based on the exercise requirements, two points are considered as equals if their distance is less than 200 metres.
-Therefore, if the points are equal, and the indexes are the first point of one of the given lists the cell is marked to unit cost and a new similar sequence is started.
-Otherwise, we continue to an existing similar sequence, increasing by 1 its length, keeping the current cost in the z variable.
-On the other hand, if points are not equal, then a new subsequence should start, and therefore L[i][j] is set to zero.
+      \item The collection of longest subsequences is updated on-line in the
+        \\ \texttt{update\_current\_maxsubseq} function. Each new subsequences list
+        from a training trip is sorted length-wise, and the current max k
+        subsequences list is updated accordingly (i.e. if there is enough space
+        or if thew new entries are longer than existing ones). Special care is
+        taken to keep unique journey ids, if the \texttt{unique\_jids} is activated.
+      \end{itemize}
 
-	Having now found the longest subsequences, we sort the indexes based on the length of the subsequences and we call the update\_current\_maxsubseq in order to update our result comparing previous subsequences.
-First of all, the new sequences found are iterated, checking if the length of the subsequence list has reached the given number (in our case 5).
-If there is still space to add the subsequence, then it is added, otherwise we check if the new subsequence is longer from those kept in the list and therefore it should replace it.
+	Finally, a similar data preparation is followed as with the nearest neighbour
+  case to visualize the data. For each training trip, the points belonging to
+  the subroute are colored red, while the others are plotted in blue, by storing
+  corresponding point lists and colors to lists. The required metadata are
+  supplied as labels to the \texttt{utils.visualize\_point\_sequences} function,
+  as before.
 
-	Finally, a similar data preprocessing is followed before visualize the data.
-This time the points are separated so as to show which of them belong to the nearest sub-route and color them with red color.
-The function for the visualization is the same that was used for the question 2a1 in the utils.py file.
-
-	Below, there is an example of the 6 produced maps, showing the nearest sub-routes related to a test trip:
+	Below there is an example of the 6 produced maps, showing the nearest sub-routes related to a test trip:
 
@@
 
-	\subsection{Features extraction (question B)}
-	The purpose of this task was to export features for classification for the training dataset.
-In short, a grid should be created and the points of each trip trace should be replaced by a specific square of the grid.
-The output file where the exported features are stored is the tripFeatures.csv.
+	\subsection{Feature extraction (question B)}
+	The purpose of this task was to extract features for classification from the
+  training dataset.
+In short, a quantization grid should be created and the points of each trip trace should be replaced by a specific square of the grid.
+The output file where the exported features are stored is \texttt{tripFeatures.csv}.
 
-	In order to be able to draw the grid, it was necessary to find the min and max longitudes and latitudes (function: find\_min\_max\_latlong, file:MapInGridView.py).
-Having now this information, the next function called is the create\_grid giving as input the number\_of\_cells and the result of the previous method.
-The distance per cell in the grid is calculated as the difference between the max and the min lat/long divided by the total number of cells.
-The next step is to create the lines of the rows and the columns and then to provide a name for each cell.
+The grid is assumed to be rectangular, with dimensions specified in the options
+dictionary. In order to be able to draw the grid, we find the min and max
+longitudes and latitudes (function: \texttt{find\_min\_max\_latlong}, file:
+\texttt{MapInGridView.py}). The next function called is the
+\texttt{create\_grid}, parameterized with the discovered bounds and the grid
+dimensions. The next step is to create the lines of the rows and the columns and
+then to provide a name for each cell. The training points and grid structure is
+visualized with the \texttt{visualize\_grid} function, and illustrated in the
+figures below.
 
-	Afterwards, all points in the trip dataframe should be replaced by the cell names that each point corresponds to.
-This logic is implemented in the method replace\_points.
-For each row in the dataframe, all coordinates are transformed in tuples format and using the find\_index function a name is mapped to this point.
-After having replaced all points with their new names, the dataframe is exported in .csv format.
-In order to be able to easily check the created grid, the method visualize\_grid is implemented so as to plot the grid.
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.70] {data_minmax_grid_extent.png}
+			\caption{Training data for grid creation.}
+		\end{center}
+	\end{figure} 
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.70] {grid.png}
+			\caption{Example rectangular quantization grid of 5 cells, resulting in 4
+        row and column lines.}
+		\end{center}
+	\end{figure}
+
+  Feature construction uses the bag-of-words technique with the \texttt{map\_to\_features\_bow} function. A histogram vector is
+  initialized to zeros for each trip, and each coordinate point in the trip is 
+  mapped to a grid cell as a function of the grid rows, grid columns and the
+  point's coordinates (funtion \texttt{find\_cell\_index}). This assignment increases the accumulation on that
+  cell in the histogram vector. The resulting feature fectors are saved in a csv file.
 
 	\subsection{Classification (question C)}
-	In the method question\_c, a new dataframe is created from the features file, containing the grid cell names.
-All functions used in this question are in the file JourneyClassification.py.
-Before starting the classification, the method preprocess\_data was used in order to get the journey\_id of all trips as numbers and also to remove the letter C from the cell names.
+ All functions used in this question are in the file \texttt{JourneyClassification.py}. 
+	In the method \texttt{question\_c}, a new dataframe is created from the
+  features file, containing the quantized features.
+  Data is prepared for classification in the method
+  \texttt{preprocess\_train\_data}. There, data is shuffled and the target
+  journey ids are uniquely mapped to integers to serve as target classes.
 
-	In the requirements of this question it is mentioned to use 10-fold cross validation, therefore the KFold class is used providing the required folds number and then the training dataset is splitted in training and test folds.
-It worths to be mentioned here, that method asarrays from numpy was used to the, necessary for this question, lists.
-In addition, a list with the names of the three classifiers is created and for each of them, the relevant function is called in order to train the classifier and then count the predictions' accuracy on the test dataset.
+  The \texttt{train} function uses the numpy scikit-learn libraries to perform training.
+  Data are converted to numpy arrays, partitioned to K=10 cross-validation folds
+  and training is performed for each fold using the appropriate scikit-learn
+  implementations for the Logistic Regression, Random Forests K-Nearest Neighbours (K=5)
+  classifiers and the cross-validation itself. In summary, each classifier is
+  implemented as an object, is trained and evaluated with the \texttt{fit} and
+  \texttt{predict} functions respectively. Prediction produces the predicted
+  classes in all classifiers but the Logistic Regression, which outputs class
+  probabilities - there, the predicted class index is extracted by considering
+  the argmax of the output. Training and validation accuracy
+  is collected per fold and classifier with the \texttt{accuracy\_score} function
+  of scikit-learn, both illustrated in a barchart, constructed using
+  pylab and matplotlib. The mean accuracies accross folds is returned and printed.
 
-	In this question, the scikit learn library was used.
-Firstly, for the knn classifier the class KNeighborsClassifier was used, and more specifically the fit and predict methods.
-In order to calculate the accuracy of the classifiers we have used the accuracy\_score method (this result is return from each classification function in the JourneyClassification.py), but also the classification\_report in order to print the precision and recall values for each prediction of the classifier.
-The same logic is followed for the other two classifiers(logistic regression and random forest).
-What should be mentioned, though, is that the LogisticRegression class implements the predict\_proba method, which returns a list with probabilities and argmax from numpy was used in order to determine the result with the max probability.
+  For beating the benchmark we have selected the Logistic Regression classifier,
+  and attempted to improve its performance by the following techniques:
+  \begin{itemize}
+    \item Feature normalization, where we reduce the variance in the data by
+      applying feature normalization techniques, using the $L_1, L_2$ and $max$ norms.
+      \item Feature scaling, where the features are scaled to zero mean and unit variance.
+    \item Varying the regularization term ($C$) in the Logistic Regression hyperplane
+      to attempt to find a better fit in the data. We tested values from $0.2$ to $2$.
+     \item Replace the bag-of-words feature with the VLAD (Vector of Locally
+       Aggregated Descriptors) feature. In VLAD,
+       instead of considering frequency histograms, we accumulate difference
+       from the representative elements in the quantization vector. For example,
+       if a point $p$ is assigned to cell $c_n$, in bag-of-words the $n$-th
+       coordinate in the quantization vector is incremented by $1$. In VLAD, the
+       increment equals $d(p,c_n)$, where $d(\cdot,\cdot)$ is a distance function.
+       We used the cell centroid as the cell representative point and the
+       harvesine distance.
 
-	\subsubsection{Beat ​ ​the ​ ​Benchmark}
+  \end{itemize}
+  Results for all methods are in the table below.
+
+  \begin{table}[H]
+    \centering
+    \begin{tabular}{|c|c|c|}
+    \hline
+      method & train & val \\ \hline
+      logreg & 0.123 & 0.0123 \\ \hline
+
+      % ...
+\hline
+      improvements & & \\ \hline
+       impr.method1 & 123123 & 3412312 \\ \hline
+    \end{tabular}
+    \end{table}
+	\subsection{Test improved classifier}
+  We tested the improved classifier in the test data. The results are attached
+  in the specified format to the file \texttt{testSet\_journeyPatternIDs.csv}.
 

### 2018-02-11 Project/JourneyClassification.py

@@ -168,6 +168,15 @@ def improve_classification(clean_trips_file, grid_file, bow_features_file, num_f
         mean_acc = train(proc_features, targets, num_folds, classifier, output_folder, tag, classifier_obj=classifier_obj)
         mean_accuracies[tag] = (mean_acc, classifier_obj)
 
+    # try feature scaling
+    tag = "scaling"
+    print("\nTrying strategy: %s" % tag, end='')
+    classifier_obj = LogisticRegression()
+    features_scaled = sklearn.preprocessing.scale(features)
+    mean_acc = train(features_scaled, targets, num_folds, classifier, output_folder, tag, classifier_obj=classifier_obj)
+    mean_accuracies[tag] = (mean_acc, classifier_obj)
+
+
     # try various regularization strengths
     Cs = [0.2, 0.5 , 1.5, 2.0]
     for c in Cs:
@@ -187,6 +196,7 @@ def improve_classification(clean_trips_file, grid_file, bow_features_file, num_f
     mean_accuracies[tag] = (mean_acc, classifier_obj)
 
 
+
     return mean_accuracies
 
 
@@ -199,6 +209,7 @@ def test(logreg_object, classifier_name, best_technique, test_file, grid_file, j
         grid = pickle.load(f)
     print("Transforming test data to features...")
     features = gvp.map_to_features_bow(test_data_df, grid, None)
+    # mean subtraction
 
     jids = [j for j in jid_mapping]
     numeric_ids = [jid_mapping[j] for j in jids]

### 2018-02-11 Project/MapInGridView.py

@@ -131,8 +131,8 @@ def map_to_features_bow(data_df, grid, output_file):
         for i,lonlat in enumerate(train_lonlats):
             lon = lonlat[0]  # for columns
             lat = lonlat[1]  # for rows
-            row_idx = find_index(rows, lat)
-            col_idx = find_index(columns, lon)
+            row_idx = find_cell_index(rows, lat)
+            col_idx = find_cell_index(columns, lon)
             linear_idx = row_idx * len(columns) + col_idx
             bow_vector[linear_idx] += 1
 
@@ -191,8 +191,8 @@ def map_to_features_pointwise(data_df, grid):
         for i,lonlat in enumerate(train_lonlats):
             lon = lonlat[0]  # for columns
             lat = lonlat[1]  # for rows
-            row_idx = find_index(rows, lat)
-            col_idx = find_index(columns, lon)
+            row_idx = find_cell_index(rows, lat)
+            col_idx = find_cell_index(columns, lon)
             cell_name = 'C'+cell_names[row_idx][col_idx]
             # visualize_grid(rows,columns,None,None,[[lon],[lat]])
             grid_hist[cell_name] += 1
@@ -209,7 +209,7 @@ def map_to_features_pointwise(data_df, grid):
         ssum += grid_hist[name]
     return features, timestamps
 
-def find_index(points_list, point):
+def find_cell_index(points_list, point):
     count = 0
     for p in points_list:
         if point < p:

### 2018-02-11 Project/NearestNeighbours.py

@@ -94,7 +94,7 @@ def calculate_dists(test_lonlat, train_df, ret_container = None, paropts = None)
 
 def calculate_dynamic_time_warping(latlons1, latlons2, paropts = None, impl = "diy_initial"):
     '''
-    Calculate the DTW of two points, used in order find the nearest neighbours
+    Calculate the DTW of two sequences, used in order find the nearest neighbours
     :param latlons1:
     :param latlons2:
     :return:
@@ -155,7 +155,7 @@ def compute_dists(points_list, paropts):
     return reslist
 
 
-def preprocessing_for_visualization(test_points, nns_ids_distances, outfile_name, elapsed, i):
+def visualize_nns(test_points, nns_ids_distances, outfile_name, elapsed, i):
     '''
     :param test_trip: the given test trip from the test file
     :param nearest_neighbours: the 5 nearest neighbours of this test trip, in format id, distance

### 2018-02-11 Project/main.py

@@ -27,8 +27,8 @@ def question_2(opts):
     # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"])
 
     print("\n>>> Running question 2a2 - Nearest subroutes computation")
-    qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],\
-                    opts["conseq_lcss"], opts["k"], opts["paropts"], opts["verbosity"], options["unique_subroute_jids"])
+    # qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],\
+    #                opts["conseq_lcss"], opts["k"], opts["paropts"], opts["verbosity"], options["unique_subroute_jids"])
 
     print("\n>>> Running question 2b - Cell grid quantization")
     cellgrid = opts["grid"]

### 2018-02-11 Project/question2.py

@@ -24,11 +24,12 @@ def question_a1(output_folder, clean_file, test_file, paropts, k):
         # get time elapsed
         elapsed = utils.tictoc(millis_start)
         # visualize
-        nn.preprocessing_for_visualization(test_points, nns_ids_distances, outfile_name, elapsed, index)
+        nn.visualize_nns(test_points, nns_ids_distances, outfile_name, elapsed, index)
 
 
 def question_a2(output_folder, test_file, train_file, conseq_lcss, k, paropts, verbosity, unique_jids = True):
-    print("Extracting %d subroutes for each test trip" % k)
+    lcss_type = "consequtive" if conseq_lcss else "non-consequtive"
+    print("Extracting %d %s subroutes for each test trip" % (k, lcss_type))
     test_df = pd.read_csv(test_file, delimiter="\n")
     train_df = pd.read_csv(train_file)
     for index, row in test_df.iterrows():

### 2018-02-14 ProjectReport/BigDataHomework.tex
Update report, add images, etc

@@
 	\listoffigures
-	\newpage
-	\listoftodos
 	\newpage
@@
 		\begin{center}
-			\includegraphics [scale = 0.75] {questionCexample.jpg}
-			\caption{Example of a trip visualization.}
+			\includegraphics [scale = 0.55] {questionC1.png}
+			\caption{First trip visualization.}
+		\end{center}
+    \label{gmlplot_example}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.60] {questionC2.png}
+			\caption{Second trip visualization.}
+		\end{center}
+		\label{gmlplot_example}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.75] {questionC3.png}
+			\caption{Third trip visualization.}
+		\end{center}
+		\label{gmlplot_example}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.75] {questionC4.png}
+			\caption{Third trip visualization.}
+		\end{center}
+		\label{gmlplot_example}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.75] {questionC5.png}
+			\caption{Third trip visualization.}
 		\end{center}
@@
 neighbours). All plotting is done with the pylab package.
-l 
-	An example of this question image output can be found below:
+
+	Below we present the results from this question in the requested format:
 
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {question2a1example.jpg}
-			\caption{Example of nearest neighbours trips visualization}
+			\includegraphics [scale = 0.40] {nn1.jpg}
+			\caption{Nearest neighbours for the first test trip}
 		\end{center}
 	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.40] {nn2.jpg}
+			\caption{Nearest neighbours for the second test trip}
+		\end{center}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.40] {nn3.jpg}
+			\caption{Nearest neighbours for the third test trip}
+		\end{center}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.40] {nn4.jpg}
+			\caption{Nearest neighbours for the fourth test trip}
+		\end{center}
+	\end{figure}
+
+	In \ref{nnresults}, we show the total results for the Nearest Neighbors.
+
+	\begin{table}[H]
+		\centering
+		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
+			\hline
+			\textbf{}  & \textbf{Time} & \textbf{id1} & \textbf{d1} & \textbf{id2} & \textbf{d2} & \textbf{id3} & \textbf{d3} & \textbf{id4} & \textbf{d4} & \textbf{id5} & \textbf{d5} \\ \hline
+			\textbf{1} & 02:129:716    & 309          & 5.2           & 1599         & 6.6           & 1575         & 6.7           & 608          & 6.9           & 1477         & 8.2           \\ \hline
+			\textbf{2} & 01:66:739     & 929          & 4.1           & 955          & 65.5          & 582          & 85.3          & 2291         & 105.6         & 1970         & 106.7         \\ \hline
+			\textbf{3} & 02:127:758    & 2060         & 4.8           & 567          & 5.8           & 901          & 6.0           & 554          & 7.8           & 571          & 16.0          \\ \hline
+			\textbf{4} & 01:83:246     & 2415         & 2.5           & 603          & 3.5           & 2407         & 3.5           & 786          & 64.4          & 753          & 65.7          \\ \hline
+			\textbf{5} & 01:80:492     & 740          & 6.6           & 1304         & 32.4          & 2686         & 46.1          & 3138         & 72.3          & 1435         & 145.1         \\ \hline
+		\end{tabular}
+	\caption{Nearest Neighbors results}
+	\label{nnresults}
+	\end{table}
 
@@
 
-	Below there is an example of the 6 produced maps, showing the nearest sub-routes related to a test trip:
+  Below, we present the nearest subroutes for each given test trip:
 
@@
 		\begin{center}
-			\includegraphics [scale = 0.50] {question2a2example.jpg}
-			\caption{Example of nearest sub-routes trips visualization}
+			\includegraphics [scale = 0.40] {subroutes1.jpg}
+			\caption{Nearest subroutes for the first test trip}
 		\end{center}
 	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.40] {subroutes2.jpg}
+			\caption{Nearest subroutes for the second test trip}
+		\end{center}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.40] {subroutes3.jpg}
+			\caption{Nearest subroutes for the third test trip}
+		\end{center}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.40] {subroutes4.jpg}
+			\caption{Nearest subroutes for the fourth test trip}
+		\end{center}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.40] {subroutes5.jpg}
+			\caption{Nearest subroutes for the fifth test trip}
+		\end{center}
+	\end{figure}
+
+	As in the Nearest Neighbours case, we have also included the html files in the gmplot folder. in \ref{subres} we show the total results for Nearest Subroutes.
+
+	\begin{table}[H]
+		\centering
+		\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|}
+			\hline
+			\textbf{}  & \textbf{Time} & \textbf{id1} & \textbf{pts1} & \textbf{id2} & \textbf{pts2} & \textbf{id3} & \textbf{pts3} & \textbf{id4} & \textbf{pts4} & \textbf{id5} & \textbf{pts5} \\ \hline
+			\textbf{1} & 01:62:303     & 491          & 31            & 667          & 31            & 1366         & 22            & 503          & 21            & 426          & 19            \\ \hline
+			\textbf{2} & 01:63:755     & 2620         & 42            & 510          & 37            & 1093         & 34            & 426          & 31            & 363          & 30            \\ \hline
+			\textbf{3} & 00:32:319     & 200          & 40            & 1885         & 40            & 2763         & 40            & 142          & 39            & 135          & 35            \\ \hline
+			\textbf{4} & 00:46:297     & 566          & 32            & 1269         & 21            & 227          & 20            & 2561         & 17            & 252          & 15            \\ \hline
+			\textbf{5} & 00:57:303     & 940          & 33            & 1932         & 22            & 1384         & 21            & 1606         & 21            & 1822         & 20            \\ \hline
+		\end{tabular}
+	\caption{Nearest subroutes results}
+	\label{subres}
+	\end{table}
 
@@
  		\hline
-      method & train & val \\ \hline
-      logreg & 0.123 & 0.0123 \\ \hline
+ 		\textbf{Strategy}            & \textbf{Train Accuracy} & \textbf{Validation Accuracy} \\ \hline
+ 		\textbf{KNN}                 & 0.446                   & 0.221                        \\ \hline
+ 		\textbf{Logistic Regression} & 0.269                   & 0.170                        \\ \hline
+ 		\textbf{Random Forest}       & 0.969                   & 0.281                        \\ \hline
+ 	\end{tabular}
+ \caption{Classification using KNN, logreg and random forest}
+ \label{class}
+ \end{table}
 
-      % ...
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|l|c|c|}
 		\hline
-      improvements & & \\ \hline
-       impr.method1 & 123123 & 3412312 \\ \hline
+		strategy & validaton accuracy & change  \\ \hline
+		norm\_l2 &  0.063  &  -62.84\%  \\ \hline
+		norm\_l1 &  0.048  &  -71.56\%  \\ \hline
+		norm\_max &  0.072  &  -57.34\%  \\ \hline
+		scaling &  0.131  &  -22.94\%  \\ \hline
+		C\_0.200 &  0.15  &  -11.93\%  \\ \hline
+		C\_0.500 &  0.157  &  -7.80\%  \\ \hline
+		C\_1.500 &  0.175  &  2.75\%  \\ \hline
+		C\_2.000 &  0.178  &  5.05\%  \\ \hline
+		vlad &  0.003  &  -97.71\%  \\ \hline
+
 	\end{tabular}
+	\label{impr}
+	\caption{Classification improvements using logistic regression}
 \end{table}
+
 	\subsection{Test improved classifier}
+
   We tested the improved classifier in the test data. The results are attached

### 2018-02-14 Project/NearestSubroutes.py

@@ -49,7 +49,7 @@ def serial_execution(df, test_lonlat, k, conseq_lcss, verbosity = False, unique_
             max_subseqs = update_current_maxsubseq(max_subseqs, subseqs_idx, k, elapsed, row, unique_jids = unique_jids)
             # print("Updated max subseqs, len now:",len(max_subseqs))
     if verbosity:
-        print("Got %d subseqs:" % len(max_subseqs), [ (x,y,z["journeyId"]) for (x,y,z) in max_subseqs])
+        print("Got %d subseqs:" % len(max_subseqs), [ (x,y,z["tripId"]) for (x,y,z) in max_subseqs])
     return max_subseqs
 
 

### 2018-02-14 Project/main.py

@@ -24,11 +24,11 @@ def question_2(opts):
     print("=====================")
     # Question 2
     print("\n>>> Running question 2a1 - Nearest neighbours computation")
-    # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"])
+    qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"])
 
     print("\n>>> Running question 2a2 - Nearest subroutes computation")
-    # qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],\
-    #                opts["conseq_lcss"], opts["k"], opts["paropts"], opts["verbosity"], options["unique_subroute_jids"])
+    qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],\
+                    opts["conseq_lcss"], opts["k"], opts["paropts"], opts["verbosity"], options["unique_subroute_jids"])
 
     print("\n>>> Running question 2b - Cell grid quantization")
     cellgrid = opts["grid"]
@@ -69,7 +69,7 @@ if __name__ == '__main__':
     np.random.seed(options["seed"])
     random.seed(options["seed"])
 
-    #paropts = ("processes", 10)
+    # paropts = ("processes", 10)
     # paropts = ("threads", 10)
     paropts = None
     options["paropts"] = paropts
@@ -84,7 +84,7 @@ if __name__ == '__main__':
     options["mapsdir"]    = join(options["outdir"], "gmplots")
     options["classifdir"] = join(options["outdir"], "classification_charts")
     options["classiffile"] = join(options["outdir"],"​testSet_JourneyPatternIDs.csv")
-    options["folds"] = 2
+    options["folds"] = 10
     options["grid"] = (5,5)
 
     os.makedirs(options["outdir"], exist_ok=True)
@@ -92,7 +92,7 @@ if __name__ == '__main__':
     os.makedirs(options["classifdir"], exist_ok=True)
 
     # run
-    # question_1(options)
+    question_1(options)
 
     # question 2
     ############

### 2018-02-14 Project/NearestNeighbours.py
final updates

@@ -5,7 +5,7 @@ from dtw import dtw as libdtw
 import question1 as qp1
 
 
-def calculate_nns(test_points, train_df, paropts=None, k=5):
+def calculate_nns(test_points, train_df, paropts=None, k=5, unique_jids = False):
     # parallelization type
     if paropts:
         print("Parallelizing with", paropts)
@@ -28,6 +28,19 @@ def calculate_nns(test_points, train_df, paropts=None, k=5):
         nearest_neighbours = calculate_dists(test_lonlat, train_df)
     # sort the list to increasing distance
     nearest_neighbours = sorted(nearest_neighbours, key=lambda k: k[1])
+    # keep unique jids, if needed
+    if unique_jids:
+        print("Restricting to single neighbour per jid")
+        keep = [0 for _ in range(len(nearest_neighbours))]
+        already_encountered = []
+        for i, nn in enumerate(nearest_neighbours):
+            jid = nn[2]
+            if jid not in already_encountered:
+                already_encountered.append(jid)
+                keep[i] = True
+                continue
+        nearest_neighbours = [ nearest_neighbours[i] for i in range(len(nearest_neighbours)) if keep[i]]
+
     # return the top 5
     nearest_neighbours = nearest_neighbours[:k]
     print("Neighbours:",[n[0] for n in nearest_neighbours])

### 2018-02-14 Project/main.py

@@ -24,7 +24,7 @@ def question_2(opts):
     print("=====================")
     # Question 2
     print("\n>>> Running question 2a1 - Nearest neighbours computation")
-    qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"])
+    qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"], options["unique_subroute_jids"])
 
     print("\n>>> Running question 2a2 - Nearest subroutes computation")
     qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],\

### 2018-02-14 Project/question2.py

@@ -8,7 +8,7 @@ import os
 import pickle
 
 
-def question_a1(output_folder, clean_file, test_file, paropts, k):
+def question_a1(output_folder, clean_file, test_file, paropts, k, unique_jids = False):
     test_df = pd.read_csv(test_file, delimiter="\n")
     train_df = pd.read_csv(clean_file)
     print("Extracting %d nearest neighbours out of %d cleaned train data, for each test trip" % (k,len(train_df)))
@@ -20,7 +20,7 @@ def question_a1(output_folder, clean_file, test_file, paropts, k):
         millis_start = utils.tic()
         # compute nearest neighbours
         test_points = eval(row["Trajectory"])
-        nns_ids_distances = nn.calculate_nns(test_points, train_df, paropts=paropts)
+        nns_ids_distances = nn.calculate_nns(test_points, train_df, paropts=paropts, unique_jids = unique_jids)
         # get time elapsed
         elapsed = utils.tictoc(millis_start)
         # visualize

### 2018-02-14 Project/test.sh

@@ -1,0 +1,22 @@
+ls -lh 	Project/NearestSubroutes.py
+ls -lh 	Project/main.py
+ls -lh 	Project/runlog.txt
+ls -lh 	ProjectReport/BigDataHomework.tex
+ls -lh 	ProjectReport/images/nn1.jpg
+ls -lh 	ProjectReport/images/nn2.jpg
+ls -lh 	ProjectReport/images/nn3.jpg
+ls -lh 	ProjectReport/images/nn4.jpg
+ls -lh 	ProjectReport/images/nn5.jpg
+ls -lh 	ProjectReport/images/question2a1example.jpg
+ls -lh 	ProjectReport/images/question2a2example.jpg
+ls -lh 	ProjectReport/images/questionC1.png
+ls -lh 	ProjectReport/images/questionC2.png
+ls -lh 	ProjectReport/images/questionC3.png
+ls -lh 	ProjectReport/images/questionC4.png
+ls -lh 	ProjectReport/images/questionC5.png
+ls -lh 	ProjectReport/images/questionCexample.jpg
+ls -lh 	ProjectReport/images/subroutes1.jpg
+ls -lh 	ProjectReport/images/subroutes2.jpg
+ls -lh 	ProjectReport/images/subroutes3.jpg
+ls -lh 	ProjectReport/images/subroutes4.jpg
+ls -lh 	ProjectReport/images/subroutes5.jpg

### 2018-02-15 ProjectReport/BigDataHomework.tex
final updates
update report with new data, problem with subroutes results

@@
 	To produce the required final format [timestamp, longitude, latitude], we
-  iterate each row, producing a column with the appropriate list data.
+  iterate each row, producing a column with the appropriate list data. The total number of trips extracted to the trips.csv file is 6651.
   % the three lists are now concatenated in one column and therefore the other two are dropped from the dataframe.
@@
 	The resulting filtered trips are stored in the file \texttt{trips\_clean.csv}.
+The total number of filtered trips is 6020 (192 are deleted due to having total distance less than 2km while 439 are excluded due having to max distance between two points more than 2km).
 
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {nn1.jpg}
+			\includegraphics [scale = 0.60] {nn1.png}
 			\caption{Nearest neighbours for the first test trip}
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {nn2.jpg}
+			\includegraphics [scale = 0.60] {nn2.png}
 			\caption{Nearest neighbours for the second test trip}
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {nn3.jpg}
+			\includegraphics [scale = 0.60] {nn3.png}
 			\caption{Nearest neighbours for the third test trip}
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {nn4.jpg}
+			\includegraphics [scale = 0.60] {nn4.png}
 			\caption{Nearest neighbours for the fourth test trip}
+		\end{center}
+	\end{figure}
+
+	\begin{figure} [H]
+		\begin{center}
+			\includegraphics [scale = 0.60] {nn5.png}
+			\caption{Nearest neighbours for the fifth test trip}
 		\end{center}
@@
 			\textbf{}  & \textbf{Time} & \textbf{id1} & \textbf{d1} & \textbf{id2} & \textbf{d2} & \textbf{id3} & \textbf{d3} & \textbf{id4} & \textbf{d4} & \textbf{id5} & \textbf{d5} \\ \hline
-			\textbf{1} & 02:129:716    & 309          & 5.2           & 1599         & 6.6           & 1575         & 6.7           & 608          & 6.9           & 1477         & 8.2           \\ \hline
-			\textbf{2} & 01:66:739     & 929          & 4.1           & 955          & 65.5          & 582          & 85.3          & 2291         & 105.6         & 1970         & 106.7         \\ \hline
-			\textbf{3} & 02:127:758    & 2060         & 4.8           & 567          & 5.8           & 901          & 6.0           & 554          & 7.8           & 571          & 16.0          \\ \hline
-			\textbf{4} & 01:83:246     & 2415         & 2.5           & 603          & 3.5           & 2407         & 3.5           & 786          & 64.4          & 753          & 65.7          \\ \hline
-			\textbf{5} & 01:80:492     & 740          & 6.6           & 1304         & 32.4          & 2686         & 46.1          & 3138         & 72.3          & 1435         & 145.1         \\ \hline
+			\textbf{1} & 11:694:365    & 4139          & 0.0           & 1756         & 3.5           & 584         & 3.8           & 3119          & 4.0           & 3428         & 4.1           \\ \hline
+			\textbf{2} & 05:358:252     & 1554          & 0.0           & 6461          & 2.8          & 1985          & 3.4          & 611         & 3.4         & 1434         & 3.5         \\ \hline
+			\textbf{3} & 12:732:353    & 3317         & 0.0           & 1939          & 4.7           & 3200          & 4.8           & 5374          & 4.8           & 1508          & 4.8          \\ \hline
+			\textbf{4} & 07:475:129     & 603         & 0.0           & 559          & 2.5           & 951         & 2.8           & 1118          & 3.2          & 684          & 3.5          \\ \hline
+			\textbf{5} & 07:444:862     & 1497          & 0.0           & 1688         & 4.3          & 927         & 4.5          & 2578         & 4.6          & 6252         & 4.7         \\ \hline
 		\end{tabular}
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {subroutes1.jpg}
+			\includegraphics [scale = 0.80] {subroutes1.png}
 			\caption{Nearest subroutes for the first test trip}
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {subroutes2.jpg}
+			\includegraphics [scale = 0.80] {subroutes2.png}
 			\caption{Nearest subroutes for the second test trip}
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {subroutes3.jpg}
+			\includegraphics [scale = 0.80] {subroutes3.png}
 			\caption{Nearest subroutes for the third test trip}
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {subroutes4.jpg}
+			\includegraphics [scale = 0.80] {subroutes4.png}
 			\caption{Nearest subroutes for the fourth test trip}
@@
 		\begin{center}
-			\includegraphics [scale = 0.40] {subroutes5.jpg}
+			\includegraphics [scale = 0.80] {subroutes5.png}
 			\caption{Nearest subroutes for the fifth test trip}
@@
 			\textbf{}  & \textbf{Time} & \textbf{id1} & \textbf{pts1} & \textbf{id2} & \textbf{pts2} & \textbf{id3} & \textbf{pts3} & \textbf{id4} & \textbf{pts4} & \textbf{id5} & \textbf{pts5} \\ \hline
-			\textbf{1} & 01:62:303     & 491          & 31            & 667          & 31            & 1366         & 22            & 503          & 21            & 426          & 19            \\ \hline
+			\textbf{1} & 05:346:265     & 211          & 31            & 728          & 31            & 3590         & 22            & 4985          & 21            & 843          & 19            \\ \hline
 			\textbf{2} & 01:63:755     & 2620         & 42            & 510          & 37            & 1093         & 34            & 426          & 31            & 363          & 30            \\ \hline

### 2018-02-15 Project/NearestSubroutes.py

@@ -222,13 +222,13 @@ def update_current_maxsubseq(current, new_seqs, k, elapsed, row, unique_jids = T
     :return:
     """
     if unique_jids:
-        new_jid = row['journeyId']
-        current_jids = [r['journeyId'] for (_,_,r) in current]
+        new_jid = row['tripId']
+        current_jids = [r['tripId'] for (_,_,r) in current]
         if new_jid in current_jids:
             # replace only if new sequence length is larger
-            existing_item = [item for item in current if item[2]['journeyId'] == new_jid]
+            existing_item = [item for item in current if item[2]['tripId'] == new_jid]
             if len(existing_item) > 1:
-                print("Found more than one item with jid",new_jid,"impossible!")
+                print("Found more than one item with tripId",new_jid,"impossible!")
                 exit(1)
             # get the item, its position and its sequence length
             existing_item = existing_item[0]

### 2018-02-15 Project/main.py

@@ -26,8 +26,8 @@ def question_2(opts):
     print("=====================")
     # Question 2
     print("\n>>> Running question 2a1 - Nearest neighbours computation")
-    qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"], options["unique_subroute_jids"])
-
+    # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"], options["unique_subroute_jids"])
+    # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"], False)
     print("\n>>> Running question 2a2 - Nearest subroutes computation")
     qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],\
                     opts["conseq_lcss"], opts["k"], opts["paropts"], opts["verbosity"], options["unique_subroute_jids"])
@@ -103,7 +103,7 @@ if __name__ == '__main__':
     test_files = [join(options["indir"], "test_set_a%d.csv" % t) for t in [1,2]] + ["test_set.csv"]
     test_files = [join(options["indir"],t) for t in test_files]
     options["testfiles"] = test_files
-    options["unique_subroute_jids"] = False
+    options["unique_subroute_jids"] = True
     options["k"] = 5
 
     # run

### 2018-02-17 ProjectReport/BigDataHomework.tex
Do data read updates

@@
   \end{itemize}
-  The given csv file with the training data was transformed to a dataframe using
-  the pandas library. Data with null journey ids are removed from the dataframe,
-  therefore lines with NaN or "null" journey ids are discarded. All data is
-  subsequently grouped by the vehicleID and the journeyPatternId.
-The final dataframe contains five columns, where the three last columns contain a list of timestamps, a list of longitudes and a list of latitudes respectively.
+  The given csv file with the training data was transformed to a dataframe using the pandas library.
+Data with null journey ids are removed from the dataframe, therefore lines with NaN or "null" journey ids are discarded.
+Then, we iterate in dataframes rows and we store the data into a dictionary having as key the vehicleId and as value the data in the requested format i.e.
+[journeyId, [timestamp, longitude, latitude]].
+In each iteration we check if the vehicleId is in the dictionary (otherwise we add it) and if the current journeyId is the one stored in the dictionary for this vehicleId.
+If this is true, we append the data list otherwise we create a new trip in the trips list and we delete those data from the dictionary, starting a new trip for the specific vehicle.
 
-	To produce the required final format [timestamp, longitude, latitude], we
-  iterate each row, producing a column with the appropriate list data. The total number of trips extracted to the trips.csv file is 6651.
-  % the three lists are now concatenated in one column and therefore the other two are dropped from the dataframe.
-Then, the column vehicleID is replaced by the tripId column which is the index of this dataframe, starting from 0 and incrementing by 1.
-Finally, the column points, which now contains the list of lists with the timestamps and the points, is sorted based on the timestamp of each list item.
-The final result is exported in a file with the requested format.
+  After having collected all the trips, we have created three different lists (tripIds, jids and points) storing the trip ids, the journey ids and the points respectively for each trip.
+Then we have creating three different dataframes from each list and by concatenating them the final dataframe was creating, which was exported in the file \texttt{trips.csv}.
 
@@
 			\includegraphics [scale = 0.70] {grid.png}
-			\caption{Example rectangular quantization grid of 5 cells, resulting in 4
-        row and column lines.}
+			\caption{Example rectangular quantization grid of 30x30 cells}
 		\end{center}
@@
   mapped to a grid cell as a function of the grid rows, grid columns and the
-  point's coordinates (funtion \texttt{find\_cell\_index}). This assignment increases the accumulation on that
-  cell in the histogram vector. The resulting feature fectors are saved in a csv file.
+  point's coordinates (funtion \texttt{find\_cell\_index}).
+This assignment increases the accumulation on that cell in the histogram vector.
+In addition, in order to keep the direction as feature (to characterize trips with different journeyId and direction but with similar points) we check the first and last point of each trip.
+If the first point has greater longitude we store all features as negative numbers(in case that the longitude is the same, we check the latitude).
+The resulting feature fectors are saved in a csv file.
 
@@
 
-  For beating the benchmark we have selected the Logistic Regression classifier,
-  and attempted to improve its performance by the following techniques:
+  For beating the benchmark we have selected the Random Forest classifier, and attempted to improve its performance by the following techniques:
+	\begin{itemize}
+		\item Experimenting with the number of trees in the forest by changing the estimator number(5,20,25) - the default number was 10.
+		\item Changing the max features number to be considered when looking for the best split.
+The tested stategies are: sqrt, log2, None while the default strategy tested in baseline was the auto strategy.
+		\item Experimenting with the max depth strategy of the tree, with values 10,50 and 100 while the default value was 2.
+	\end{itemize}
+
+  Results for all methods are in the table below.
+
+ \begin{table}[H]
+ 	\centering
+ 	\begin{tabular}{|c|c|c|}
+ 		\hline
+ 		\textbf{Strategy}            & \textbf{Train Accuracy} & \textbf{Validation Accuracy} \\ \hline
+ 		\textbf{KNN}                 & 0.904                   & 0.861                        \\ \hline
+ 		\textbf{Logistic Regression} & 0.874                   & 0.831                        \\ \hline
+ 		\textbf{Random Forest}       & 0.995                   & 0.883                        \\ \hline
+ 	\end{tabular}
+ \caption{Classification using KNN, logreg and random forest}
+ \label{class}
+ \end{table}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{|c|c|c|}
+		\hline
+		\textbf{Strategy} & \textbf{Validation Accuracy} & \textbf{Change} \\ \hline
+		Estimator=5       & 0.886                        & 0.40\%          \\ \hline
+		Estimator=20      & 0.887                        & 0.45\%          \\ \hline
+		Estimator = 25    & 0.886                        & 0.36\%          \\ \hline
+		Max feature sqrt  & 0.887                        & 0.47\%          \\ \hline
+		Max feature log2  & 0.887                        & 0.49\%          \\ \hline
+		Max feature None  & 0.891                        & 0.92\%          \\ \hline
+		Max depth=10      & 0.887                        & 0.45\%          \\ \hline
+		Max depth=50      & 0.888                        & 0.56\%          \\ \hline
+		Max depth=100     & 0.888                        & 0.60\%          \\ \hline
+	\end{tabular}
+	\label{impr1}
+	\caption{Classification improvements using random forest}
+\end{table}
+
+Since we had small improvement with Random Forest classifier, we decided to improve also Logistic Regression classifier. In order to improve its performance we have used the following techniques:
+
 \begin{itemize}
@@
 	harvesine distance. 
+\end{itemize}
 
-  \end{itemize}
-  Results for all methods are in the table below.
-
- \begin{table}[H]
- 	\centering
- 	\begin{tabular}{|c|c|c|}
- 		\hline
- 		\textbf{Strategy}            & \textbf{Train Accuracy} & \textbf{Validation Accuracy} \\ \hline
- 		\textbf{KNN}                 & 0.446                   & 0.221                        \\ \hline
- 		\textbf{Logistic Regression} & 0.269                   & 0.170                        \\ \hline
- 		\textbf{Random Forest}       & 0.969                   & 0.281                        \\ \hline
- 	\end{tabular}
- \caption{Classification using KNN, logreg and random forest}
- \label{class}
- \end{table}
+Below we show the results from the techniques used for the Logistic Regression classifier:
 
@@
 		\hline
-		strategy & validaton accuracy & change  \\ \hline
-		norm\_l2 &  0.063  &  -62.84\%  \\ \hline
-		norm\_l1 &  0.048  &  -71.56\%  \\ \hline
-		norm\_max &  0.072  &  -57.34\%  \\ \hline
-		scaling &  0.131  &  -22.94\%  \\ \hline
-		C\_0.200 &  0.15  &  -11.93\%  \\ \hline
-		C\_0.500 &  0.157  &  -7.80\%  \\ \hline
-		C\_1.500 &  0.175  &  2.75\%  \\ \hline
-		C\_2.000 &  0.178  &  5.05\%  \\ \hline
-		vlad &  0.003  &  -97.71\%  \\ \hline
-
+		\textbf{Strategy} & \textbf{Validation Accuracy} & \textbf{Change} \\ \hline
+		norm\_l2 &  0.794  &  -4.56\%  \\ \hline
+		norm\_l1 &  0.616  &  -25.93\%  \\ \hline
+		norm\_max &  0.790  &  -4.98\%  \\ \hline
+		scaling &  0.832  &  0.12\%  \\ \hline
+		C\_0.200 &  0.834  &  0.34\%  \\ \hline
+		C\_0.500 &  0.833  &  0.18\%  \\ \hline
+		C\_1.500 &  0.831  &  -0.02\%  \\ \hline
+		C\_2.000 &  0.831  &  -0.02\%  \\ \hline
 	\end{tabular}
-	\label{impr}
+	\label{impr2}
 	\caption{Classification improvements using logistic regression}
 \end{table}
+
+All above techniques were implemented using a 30x30 grid.
+We have also tried to increase the size of grid to 50x50 but with no improvement to the baseline.
+Finally, we have also tried to decrease the size of the grid to 20x20 and we observed that the results were slightly worse, therefore we have decided to keep the grid of 30x30.
+All these results are included in the runlog file.
 

### 2018-02-17 Project/JourneyClassification.py

@@ -14,51 +14,6 @@ import utils
 import os
 
 
-def train(features, targets, num_folds, classifiers, output_folder, filename_tag ="", classifier_obj = None):
-    kf = KFold(n_splits = num_folds)
-    folds_idxs = list(kf.split(features))
-    trips_array = np.asarray(features)
-    targets = np.asarray(targets)
-
-    if type(classifiers) != list:
-        classifiers = [classifiers]
-    # train/val accuracies
-    accuracies = {}
-    mean_accuracies = {}
-    # classify
-    for classifier in classifiers:
-        classif_start = utils.tic()
-        accuracies[classifier] = []
-        print("\nTesting classifier [%s]" % classifier)
-        # train & test each classifier
-        # for each fold
-        for i, (train_idx, val_idx) in enumerate(folds_idxs):
-            print("\tClassifying fold %d/%d" % (i + 1, len(folds_idxs)), end=" ")
-            train = (trips_array[train_idx], targets[train_idx])
-            val = (trips_array[val_idx], targets[val_idx])
-            if classifier == "knn":
-                k = 5
-                accTrain, accVal = knn_classification(train, val, k)
-            elif classifier == "logreg":
-                accTrain, accVal = logreg_classification(train, val, lr_classifier=classifier_obj)
-            elif classifier == "randfor":
-                accTrain, accVal = randfor_classification(train, val)
-            accuracies[classifier].append((accTrain, accVal))
-            print("- accuracies train/val:",accuracies[classifier][-1])
-        elapsed = utils.tictoc(classif_start)
-        print("Done in:", elapsed)
-
-        # accuracy across all folds
-        mean_accuracies[classifier] = [np.mean([x[0] for x in accuracies[classifier]]), \
-                                       np.mean([x[1] for x in accuracies[classifier]])]
-        titlestr = "%s, overall accuracy train/val: %s" % (classifier, str(mean_accuracies[classifier]))
-        chart_filename = os.path.join(output_folder, classifier + "_" + filename_tag + "_chart")
-        utils.barchart(list(range(1, num_folds + 1)), accuracies[classifier], title=titlestr, ylabel="accuracy", legend=["train","val"],
-                    save=chart_filename)
-
-    return mean_accuracies
-
-
 def preprocess_train_data(feature_df, seed):
     print("Preprocessing training data")
     targets = []
[…]

### 2018-02-17 Project/MapInGridView.py

@@ -2,6 +2,7 @@ import os
 import matplotlib.pyplot as plt
 import utils
 import numpy as np
+from itertools import combinations
 import math
 from collections import OrderedDict
 
@@ -67,6 +68,115 @@ def create_cell_names(number_of_cells):
     return cell_names
 
 
+def map_to_features(data_df, grid, output_file):
+    rows, columns, cell_names = grid
+    raw_features, timestamps = map_to_features_pointwise(data_df, grid)
+    headername = "points" if "points" in data_df else "Trajectory"
+    maxlen = -1
+    for i,(featlist,ts) in enumerate(zip(raw_features,timestamps)):
+        # squeeze duplicate points
+        sq_feats = []
+        for f,t in zip(featlist,ts):
+            if (not sq_feats) or (f not in [ v[-1] for v in sq_feats]): # alternate check
+            # if (not sq_feats) or (f != sq_feats[-1][-1]): # alternate check
+                sq_feats.append([t, f])
+        if len(sq_feats) > maxlen:
+            maxlen = len(sq_feats)
+        data_df.at[i,headername] = sq_feats
+    print("Max squeezed feature length:",maxlen)
+    if output_file is not None:
+        data_df.to_csv(output_file)
+    else:
+        return data_df
+
+
+def map_to_features_pointwise(data_df, grid):
+    rows, columns, cell_names = grid
+    # measure some statistics
+    grid_hist = {}
+    total_points = 0
+    numcells = (len(rows) + 1) * (len(columns) + 1)
+    for cc in cell_names:
+        for c in cc:
+            grid_hist['C' + str(c)] = 0
+
+    points_header = "points" if "points" in data_df else "Trajectory"
+    features, timestamps = [], []
+    for index,row in data_df.iterrows():
+        train_points = row[points_header]
+        train_points = eval(train_points)
+
+        ts = [p[0] for p in train_points]
+        timestamps.append(ts)
+        train_lonlats = utils.idx_to_lonlat(train_points, format="tuples")
+        feature_list = []
+        for i,lonlat in enumerate(train_lonlats):
+            lon = lonlat[0]  # for columns
+            lat = lonlat[1]  # for rows
+            row_idx = find_cell_index(rows, lat)
+            col_idx = find_cell_index(columns, lon)
+            cell_name = 'C'+cell_names[row_idx][col_idx]
+            # visualize_grid(rows,columns,None,None,[[lon],[lat]])
[…]

### 2018-02-17 Project/main.py

@@ -2,7 +2,6 @@ import sys, os, pkgutil
 from os.path import join, abspath
 import question1 as qp1
 import question2 as qp2
-import question1trial as qp1t
 import random
 import numpy as np
 
@@ -12,10 +11,9 @@ def question_1(opts):
     print("\nRunning question #1")
     print("=====================")
     print("\n>>> Running question 1a - parsing the training data")
-    # trips_list, df = qp1.create_trips_file(opts["trainfile"], opts["tripsfile"])
-    trips_list, df = qp1t.read_file(opts["trainfile"], opts["tripsfile"])
+    df = qp1.create_trips_file(opts["trainfile"], opts["tripsfile"])
     print("\n>>> Running question 1b - cleaning the training data")
-    trips_list, df = qp1.filter_trips(opts["cleanfile"], df)
+    df = qp1.filter_trips(opts["cleanfile"], df)
     print("\n>>> Running question 1c - visualizing the training data")
     qp1.visualize_trips(opts["mapsdir"], df)
     print("\nFinished question1!")
@@ -29,9 +27,8 @@ def question_2(opts):
     # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"], options["unique_subroute_jids"])
     # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"], False)
     print("\n>>> Running question 2a2 - Nearest subroutes computation")
-    qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],\
-                    opts["conseq_lcss"], opts["k"], opts["paropts"], opts["verbosity"], options["unique_subroute_jids"])
-
+    # qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"],\
+                   # opts["conseq_lcss"], opts["k"], opts["paropts"], opts["verbosity"], options["unique_subroute_jids"])
     print("\n>>> Running question 2b - Cell grid quantization")
     cellgrid = opts["grid"]
     print("Using cell grid with dimensions", cellgrid)
@@ -87,7 +84,7 @@ if __name__ == '__main__':
     options["classifdir"] = join(options["outdir"], "classification_charts")
     options["classiffile"] = join(options["outdir"],"​testSet_JourneyPatternIDs.csv")
     options["folds"] = 10
[…]

### 2018-02-17 Project/question1.py

@@ -9,44 +9,46 @@ import utils
 #################
 
 def create_trips_file(input_file, output_file):
+    df = pd.read_csv(input_file)
     print("Reading training file", input_file, " - will ignore null journey ids")
     print("Parsing csv...")
-    df = pd.read_csv(input_file)
-    # remove potential NaNs and "null" jids
     print("Removing nulls and NaNs ...")
     df = df[pd.notnull(df['journeyPatternId'])]
     df = df[df['journeyPatternId'] != "null"].reindex()
-    timestamps = df.groupby(["vehicleID", "journeyPatternId"])["timestamp"].apply(list)
-    lons = df.groupby(["vehicleID", "journeyPatternId"])["longitude"].apply(list)
-    lats = df.groupby(["vehicleID", "journeyPatternId"])["latitude"].apply(list)
-    df = pd.concat([timestamps, lons, lats], axis=1, ignore_index=False)
-    print("Done reading training file!")
+    vids = {}
+    trips = []
     print("Transforming data to the required format")
     for index, row in df.iterrows():
-        tslist, lonslist, latslist = [], [], []
-        tslist = row["timestamp"]
-        lonslist = row["longitude"]
-        latslist = row["latitude"]
-        data_list = []
-        for i,timestamp in enumerate(tslist):
-            new_list = []
-            new_list.append(timestamp)
-            new_list.append(lonslist[i])
-            new_list.append(latslist[i])
-            data_list.append(new_list)
-        row["timestamp"] = data_list
-    new_df = create_new_dataframe()
-    new_df = pd.concat([new_df, df["timestamp"]], axis=1, ignore_index=False)
-    new_df.to_csv(output_file)
-    df = pd.read_csv(output_file)
-    df = df.drop(["vehicleID", "tripId"], axis=1)
-    df.index.name = "tripId"
-    df.columns = ["journeyId", "points"]
-    df = sort_timestamps(df)
+        vid = row["vehicleID"]
+        jid = row["journeyPatternId"]
+        data = [row["timestamp"], row["longitude"], row["latitude"]]
+        if vid not in vids:
+            vids[vid] = []
+        if not vids[vid]:
+            vids[vid] = [jid, [data]]
+        else:
+            if vids[vid][0] == row["journeyPatternId"]:
+                vids[vid][1].append(data)
+            else:
+                trips.append(vids[vid])
+                vids[vid] = []
[…]

### 2018-02-17 Project/question1trial.py

@@ -1,45 +1,0 @@
-import pandas as pd
-
-
-def read_file(input_file, output_file):
-    df = pd.read_csv(input_file)
-    print("Reading training file", input_file, " - will ignore null journey ids")
-    print("Parsing csv...")
-    print("Removing nulls and NaNs ...")
-    df = df[pd.notnull(df['journeyPatternId'])]
-    df = df[df['journeyPatternId'] != "null"].reindex()
-    vids={}
-    trips=[]
-    print("Transforming data to the required format")
-    for index, row in df.iterrows():
-        vid = row["vehicleID"]
-        jid = row["journeyPatternId"]
-        data = [row["timestamp"], row["longitude"], row["latitude"]]
-        if vid not in vids:
-            vids[vid] = []
-        if not vids[vid]:
-            vids[vid] = [jid, [data]]
-        else:
-            if vids[vid][0] == row["journeyPatternId"]:
-                vids[vid][1].append(data)
-            else:
-                trips.append(vids[vid])
-                vids[vid] = []
-                vids[vid] = [jid, [data]]
-        print(vids[vid])
-    points = []
-    jids=[]
-    tripids = []
-    print("Producing new dataframe")
-    for i,trip in enumerate(trips):
-        points.append(trip[1])
-        jids.append(trip[0])
-        tripids.append(i)
-    df1 = pd.DataFrame({'tripId': tripids})
-    df2 = pd.DataFrame({'journeyId': jids})
-    df3 = pd.DataFrame({'points': points})
-    df = pd.concat([df1, df2, df3], axis=1, ignore_index=False)
-    df.to_csv(output_file)
-    trips_list = df.to_dict(orient='dict')
-    print("Done transforming data!")
-    return trips_list, df

### 2018-02-17 Project/question2.py

@@ -51,17 +51,20 @@ def question_b(train_file, number_of_cells, output_folder):
     # save grid and transform data
     with open(grid_file, "wb") as f:
         pickle.dump(grid, f)
-    # gvp.map_to_features(train_df, grid, feature_file)
+
+    feats_start = utils.tic()
     gvp.map_to_features_bow(train_df, grid, feature_file)
+    print("Generated features in",utils.tictoc(feats_start))
     return feature_file, grid_file
+
 
 
 def question_c(clean_file, features_file, grid_file, test_file, output_folder, seed, classif_file, num_folds):
     total_start = utils.tic()
     df_features = pd.read_csv(features_file)
     features, jid_mapping, targets = jcp.preprocess_train_data(df_features, seed)
-    classifiers = ["knn", "logreg", "randfor"]
-    # classifiers = ["logreg"]
+    classifiers = ["knn", "logreg","randfor"]
+    classifiers = ["randfor"]
     mean_accuracies = jcp.train(features, targets, num_folds, classifiers, output_folder)
 
     # print mean accuracy per classifier
@@ -69,39 +72,14 @@ def question_c(clean_file, features_file, grid_file, test_file, output_folder, s
     for classifier in mean_accuracies:
         print(classifier, "accuracy train/val:", mean_accuracies[classifier])
 
-    # select the logistic regression algorithm to beat the benchmark
-    impr_classifier_name = "logreg"
+    # select the random forest algorithm to beat the benchmark
+    impr_classifier_name = "randfor"
     baseline_accuracy = mean_accuracies[impr_classifier_name][-1]
 
-    best_accuracy, best_classifier, best_technique = -1, None, None
     print()
     print("Improving classification for classifier", impr_classifier_name)
-    mean_accuracies = jcp.improve_classification(clean_file, grid_file, features_file, num_folds, output_folder, impr_classifier_name, seed)
-
-    print()
-    print("Performance comparison:")
-    # print updated accuracies
-    for technique in mean_accuracies:
-        accuracy, classifier = mean_accuracies[technique]
-        # get validation accuracy
-        accuracy = accuracy['logreg'][-1]
-        print(impr_classifier_name, ", technique",technique,", validation accuracy :", accuracy,
-              "change over baseline: %2.2f%%" % ((accuracy-baseline_accuracy) / baseline_accuracy * 100))
-        if best_accuracy < accuracy:
-            best_classifier = classifier
-            best_technique = technique
[…]

### 2018-02-17 ProjectReport/BigDataHomework.tex
small changes in classification

@@
 		\textbf{Strategy} & \textbf{Validation Accuracy} & \textbf{Change} \\ \hline
-		Estimator=5       & 0.886                        & 0.40\%          \\ \hline
-		Estimator=20      & 0.887                        & 0.45\%          \\ \hline
-		Estimator = 25    & 0.886                        & 0.36\%          \\ \hline
-		Max feature sqrt  & 0.887                        & 0.47\%          \\ \hline
-		Max feature log2  & 0.887                        & 0.49\%          \\ \hline
-		Max feature None  & 0.891                        & 0.92\%          \\ \hline
-		Max depth=10      & 0.887                        & 0.45\%          \\ \hline
-		Max depth=50      & 0.888                        & 0.56\%          \\ \hline
-		Max depth=100     & 0.888                        & 0.60\%          \\ \hline
+		Estimator=5       & 0.865                        & -1.92\%         \\ \hline
+		Estimator=20      & 0.893                        & 1.19\%          \\ \hline
+		Estimator = 25    & 0.894                        & 1.30\%          \\ \hline
+		Max feature sqrt  & 0.882                        & 0.00\%          \\ \hline
+		Max feature log2  & 0.884                        & 0.23\%          \\ \hline
+		Max feature None  & 0.882                        & 0.00\%          \\ \hline
+		Max depth=10      & 0.661                        & -25.04\%        \\ \hline
+		Max depth=50      & 0.879                        & -0.34\%         \\ \hline
+		Max depth=100     & 0.885                        & 0.34\%          \\ \hline
+		Min samples=4     & 0.883                        & 0.08\%          \\ \hline
+		Min samples=6     & 0.888                        & 0.58\%          \\ \hline
+		Min samples=10    & 0.884                        & 0.21\%          \\ \hline
 	\end{tabular}
@@
 
-All above techniques were implemented using a 30x30 grid.
-We have also tried to increase the size of grid to 50x50 but with no improvement to the baseline.
-Finally, we have also tried to decrease the size of the grid to 20x20 and we observed that the results were slightly worse, therefore we have decided to keep the grid of 30x30.
-All these results are included in the runlog file.
+All above techniques were implemented using a 20x20 grid.
+We have also tried to increase the size of grid to 50x50 and 30x30 but the results were worse.
+Therefore we have decided to keep the 20x20 grid.
+All the results are included in the runlog file.
+In addition, in the folder classification\_charts, we have included all the histograms  showing the performance of all tested classifiers.
 

### 2018-02-17 Project/JourneyClassification.py

@@ -47,7 +47,7 @@ def preprocess_train_data(feature_df, seed):
 
 
 def train(features, targets, num_folds, classifiers, output_folder, seed=None, filename_tag ="", classifier_obj = None):
-    kf = KFold(n_splits = num_folds)
+    kf = KFold(n_splits = num_folds, random_state=seed)
     folds_idxs = list(kf.split(features))
     trips_array = np.asarray(features)
     targets = np.asarray(targets)
@@ -148,7 +148,7 @@ def improve_randfor(baseline_accuracy, bow_features_file, num_folds, output_fold
         tag = "estimator_%s" % estimator
         print("\nTrying strategy: %s" % tag, end='')
         classifier_obj = RandomForestClassifier(n_estimators=estimator, random_state=seed)
-        mean_acc = train(features, targets, num_folds, classifier, output_folder, tag,
+        mean_acc = train(features, targets, num_folds, classifier, output_folder, seed=seed, filename_tag=tag,
                          classifier_obj=classifier_obj)
         mean_accuracies[tag] = (mean_acc, classifier_obj)
 
@@ -158,7 +158,7 @@ def improve_randfor(baseline_accuracy, bow_features_file, num_folds, output_fold
         tag = "max_feature_strategy_%s" % max
         print("\nTrying strategy: %s" % tag, end='')
         classifier_obj = RandomForestClassifier(max_features=max, random_state=seed)
-        mean_acc = train(features, targets, num_folds, classifier, output_folder, tag,
+        mean_acc = train(features, targets, num_folds, classifier, output_folder, seed=seed, filename_tag=tag,
                             classifier_obj=classifier_obj)
         mean_accuracies[tag] = (mean_acc, classifier_obj)
     # test max depth strategy
@@ -167,7 +167,7 @@ def improve_randfor(baseline_accuracy, bow_features_file, num_folds, output_fold
         tag = "max_depth_strategy_%s" % max
         print("\nTrying strategy: %s" % tag, end='')
         classifier_obj = RandomForestClassifier(max_depth=max, random_state=seed)
-        mean_acc = train(features, targets, num_folds, classifier, output_folder, tag,
+        mean_acc = train(features, targets, num_folds, classifier, output_folder, seed=seed, filename_tag=tag,
                             classifier_obj=classifier_obj)
         mean_accuracies[tag] = (mean_acc, classifier_obj)
     min_samples = [4,6,10]
[…]

### 2018-02-17 Project/main.py

@@ -84,7 +84,7 @@ if __name__ == '__main__':
     options["classifdir"] = join(options["outdir"], "classification_charts")
     options["classiffile"] = join(options["outdir"],"​testSet_JourneyPatternIDs.csv")
     options["folds"] = 10
-    options["grid"] = (15,15)
+    options["grid"] = (20,20)
 
     os.makedirs(options["outdir"], exist_ok=True)
     os.makedirs(options["mapsdir"], exist_ok=True)

### 2018-02-17 Project/question2.py

@@ -64,7 +64,7 @@ def question_c(features_file, grid_file, test_file, output_folder, seed, classif
     df_features = pd.read_csv(features_file)
     features, jid_mapping, targets = jcp.preprocess_train_data(df_features, seed)
     classifiers = ["knn", "logreg","randfor"]
-    classifiers = ["randfor"]
+    # classifiers = ["randfor"]
     mean_accuracies = jcp.train(features, targets, num_folds, classifiers, output_folder, seed=seed)
 
     # print mean accuracy per classifier

### 2018-02-18 Project/NearestNeighbours.py
update report

@@ -103,7 +103,7 @@ def calculate_dists(test_lonlat, train_df, ret_container = None, paropts = None)
         #distance = calculate_dynamic_time_warping(test_lonlat, trip_lonlat, paropts, impl = "libdtw")
         # print("Calculated distance: %.2f for trip: %d/%d : %s" % (distance, index+1, len(train_df.index), str(row["journeyId"])))
         dists[index] = (int(row["tripId"]), distance, jid, train_points)
-        print("Computed distance for train trip %d/%d" % (index+1, len(train_df)))
+        # print("Computed distance for train trip %d/%d" % (index+1, len(train_df)))
     return dists
 
 

### 2018-02-18 Project/NearestSubroutes.py

@@ -18,9 +18,9 @@ def find_similar_subroutes_per_test_trip(test_points, train_df, k, paropts=None,
     if partype:
         # num threads or processes
         if partype == "processes":
-            max_subseqs = exec_with_processes(train_df, numpar, test_lonlat, k, conseq_lcss, unique_trip)
+            max_subseqs = exec_with_processes(train_df, numpar, test_lonlat, unique_trip, k)
         elif partype == "threads":
-            max_subseqs = exec_with_threads(train_df, numpar, test_lonlat, k, conseq_lcss, unique_trip)
+            max_subseqs = exec_with_threads(train_df, numpar, test_lonlat, unique_trip, k)
     else:
         max_subseqs = serial_execution(train_df, test_lonlat, k, conseq_lcss, verbosity = verbosity, unique_trip = unique_trip)
     if len(max_subseqs) != k:
@@ -303,7 +303,7 @@ def update_current_maxsubseq(current, new_seqs, k, elapsed, row, unique_trip = T
             # replace only if new sequence length is larger
             existing_item = [item for item in current if item[2]['tripId'] == new_trip]
             if len(existing_item) > 1:
-                print("Found more than one item with tripId",new_jid,"impossible!")
+                print("Found more than one item with tripId",new_trip,"impossible!")
                 exit(1)
             # get the item, its position and its sequence length
             existing_item = existing_item[0]

### 2018-02-18 Project/main.py

@@ -25,10 +25,9 @@ def question_2(opts):
     # Question 2
     print("\n>>> Running question 2a1 - Nearest neighbours computation")
     # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"], options["unique_subroute_jids"])
-    # qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"], False)
+    qp2.question_a1(opts["mapsdir"], opts["cleanfile"], opts["testfiles"][0], opts["paropts"], opts["k"], False)
     print("\n>>> Running question 2a2 - Nearest subroutes computation")
     qp2.question_a2(opts["mapsdir"], opts["testfiles"][1], opts["cleanfile"], opts["conseq_lcss"], opts["k"], opts["paropts"], opts["verbosity"], options["unique_subroute_ids"])
-    return
     print("\n>>> Running question 2b - Cell grid quantization")
     cellgrid = opts["grid"]
     print("Using cell grid with dimensions", cellgrid)

### 2018-02-18 ProjectReport/BigDataHomework.tex
update report
update readme

@@
   latitude means are calculated as a visualization centroid for the produced map, and the gmplot
-  library is used so as to visualize the respective point sequence. See the
-  following figure for a visualization example.
+  library is used so as to visualize the respective point sequence. Below, we present the resulting plots for the five selected trips.
 
@@
       \item a list with the colors with which to draw (all blue in this
-        question).
+        question)
       \end{enumerate}
@@
 \texttt{rasterize.js} utility from the \texttt{phantomjs} module).
-However we found the gml plot zoom difficult to automatically configure correctly so that
-all points are visible, and thus the images in the report were fetched with a printscreen.
-The converted images are used to create 6 subplots (1 test trip + k=5 nearest
-neighbours). All plotting is done with the pylab package.
+However we found the gml plot zoom difficult to automatically configure correctly so that all points are visible, and thus the images in the report were fetched with a printscreen.
+The converted images are used to create 6 subplots (1 test trip + k=5 nearest neighbours).
+All plotting is done with the pylab package.
+
+In order to improve time performance, we have used a cache strategy exploiting the commutative property of the distance between two points.
+Therefore, we have used a dictionary storing a mapping of the point pair indices $(i,j)$ as well as $(j,i)$.
 
@@
  		\hline
- 		\textbf{Strategy}            & \textbf{Train Accuracy} & \textbf{Validation Accuracy} \\ \hline
+ 		\textbf{Classifier}            & \textbf{Train Accuracy} & \textbf{Validation Accuracy} \\ \hline
  		\textbf{KNN}                 & 0.904                   & 0.861                        \\ \hline

### 2018-02-21 ProjectReport/BigDataHomework.tex
update report

@@
 
-	As in the Nearest Neighbours case, we have also included the html files in the gmplot folder. in \ref{subres} we show the total results for Nearest Subroutes.
+	As in the Nearest Neighbours case, we have also included the html files in the gmplot folder.
+The automated way with phantomjs does not work always correctly and in above images the red points are not included in the images.
+However, in html files, all the common points are presented.
+In \ref{subres} we show the total results for Nearest Subroutes.
 
@@
 			\textbf{}  & \textbf{Time} & \textbf{id1} & \textbf{pts1} & \textbf{id2} & \textbf{pts2} & \textbf{id3} & \textbf{pts3} & \textbf{id4} & \textbf{pts4} & \textbf{id5} & \textbf{pts5} \\ \hline
-			\textbf{1} & 05:346:265     & 211          & 31            & 728          & 31            & 3590         & 22            & 4985          & 21            & 843          & 19            \\ \hline
-			\textbf{2} & 01:63:755     & 2620         & 42            & 510          & 37            & 1093         & 34            & 426          & 31            & 363          & 30            \\ \hline
-			\textbf{3} & 00:32:319     & 200          & 40            & 1885         & 40            & 2763         & 40            & 142          & 39            & 135          & 35            \\ \hline
-			\textbf{4} & 00:46:297     & 566          & 32            & 1269         & 21            & 227          & 20            & 2561         & 17            & 252          & 15            \\ \hline
-			\textbf{5} & 00:57:303     & 940          & 33            & 1932         & 22            & 1384         & 21            & 1606         & 21            & 1822         & 20            \\ \hline
+			\textbf{1} & 07:427:941     & 5152          & 125            & 5683          & 124            & 4621         & 118            & 593          & 116            & 278          & 111            \\ \hline
+			\textbf{2} & 06:408:627     & 620         & 175            & 1078          & 136            & 4422         & 129            & 5079          & 126            & 2927          & 119            \\ \hline
+			\textbf{3} & 03:226:050     & 3822          & 202            & 6401         & 136            & 1229         & 120            & 3028          & 111            & 2308          & 110            \\ \hline
+			\textbf{4} & 05:305:995     & 6401          & 180            & 1249         & 111            & 1740          & 111            & 586         & 108            & 5858          & 105            \\ \hline
+			\textbf{5} & 06:367:621     & 5010          & 144            & 723         & 129            & 1870         & 124            & 2792         & 115            & 4759         & 115            \\ \hline
 		\end{tabular}

### 2018-02-21 Project/NearestSubroutes.py

@@ -33,8 +33,6 @@ def serial_execution(df, test_lonlat, k, verbosity = False):
     max_subseqs = []
     # for each trip in the training data
     for index, row in df.iterrows():
-        if index > 500:
-            break
         train_points = row["points"]
         train_points = eval(train_points)
         train_lonlat = utils.idx_to_lonlat(train_points, format="tuples")
@@ -49,9 +47,9 @@ def serial_execution(df, test_lonlat, k, verbosity = False):
         # update the list of the longest subsequences
         if subseqs_idx:
             max_subseqs = update_current_maxsubseq(max_subseqs, subseqs_idx, k, elapsed, row)
-            print("Max subseq length:",len(max_subseqs))
-            print([x[0] for x in max_subseqs])
-            print("Updated max subseqs, lens now:",[len(x[0]) for x in max_subseqs])
+            # print("Max subseq length:",len(max_subseqs))
+            #print([x[0] for x in max_subseqs])
+            # print("Updated max subseqs, lens now:",[len(x[0]) for x in max_subseqs])
     if verbosity:
         print("Got %d subseqs:" % len(max_subseqs), [ (x,y,z["tripId"]) for (x,y,z) in max_subseqs])
 
@@ -184,7 +182,7 @@ def calc_lcss(t1, t2, subseqs=None, subseqs_idx=None):
                         seqs.append(new_seq)
                         idxs.append(new_seq_idxs)
                     # else:
-                    #     print("Curr len at",i,j,":",p1,p2,"not gr8r than curr_len:",curr_len," -> not adding to the seqs")
+                    # print("Curr len at",i,j,":",p1,p2,"not gr8r than curr_len:",curr_len," -> not adding to the seqs")
             else:
                 L[i][j] = 0
 
@@ -266,7 +264,7 @@ def preprocessing_for_visualisation(test_points, max_subseqs, file_name, index):
         # label
         str = ["neighbour %d" % j, "jid: %s" % jid, "Matching pts: %d" % num_points, "Delta-t: %s " % sseq[1]]
         labels.append("\n".join(str))
-        print("seq first/last idxs:",[(s[0],s[-1]) for s in subseq_idxs])
+        # print("seq first/last idxs:",[(s[0],s[-1]) for s in subseq_idxs])
 
         # get the points data from the pandas dataframe to lonlat tuples
         train_points = sseq[2]["points"]
[…]
